### 2022-08-03 how_to.qmd
testing add apps

---
title: "How to"
editor: visual
---

To create a folder of shinylive apps deployed to netlify within your repo...

1.  Get netlify CLI up an running - see here: [https://docs.netlify.com/cli/get-started](https://docs.netlify.com/cli/get-started/)
2.  Run `netlify init` in the command line (from your local repo).
    1.  Follow the options.

    2.  Provide a domain name

    3.  For the build command, use `hugo build`

    4.  For the directory to deploy, use the 'site' folder associated with the folder of apps. For me this is `./site_apps`

    5.  Accept default functions folder

    6.  Upon prompting ask to create new `netlify.toml` file

### 2022-08-03 test_pages/test_with_shinylive.qmd
testing add apps

---
title: "A test notebook with Shinylive"
---

``` {r}
library(tidyverse)
library(here)
```

``` {r}
a<- c(1:10)
b<- a * 2 + 1

df<- data.frame(a, b)

fp<- paste0(here(),"/shiny_apps/shiny_app1/data.csv")

write.csv(df, fp)

```

### 2022-08-03 publish.sh

@@ -1,0 +1,32 @@
+#!/bin/zsh
+
+# For documentation of how to create and publish shinylive apps, see https://shiny.rstudio.com/py/docs/shinylive.html#sharing-shinylive-applications
+
+# To create a new app run: shiny create <PATH>, where for me PATH = shiny_apps/<APP_NAME>
+
+# Create the dirstribution for each 
+
+folder_prefix="./"
+app_prefix="./shiny_apps/"
+
+suffix="/"
+
+for d in ./shiny_apps/*/; do
+    
+    folder=${d#"$folder_prefix"}
+    folder=${folder%"$suffix"}
+
+    app=${d#"$app_prefix"}
+    app=${app%"$suffix"}
+
+    echo "$d"
+    echo "$folder"
+    echo "$app"
+
+    shiny static $folder site_apps --subdir $app 
+
+done
+
+#shiny static shiny_apps/app1 site_apps --subdir app1   
+
+#quarto publish netlify

### 2022-08-03 new_app.sh
Add publish and new app shell scripts

@@ -1,0 +1,14 @@
+#!/bin/zsh
+
+
+apps_folder_name="shiny_apps"
+site_folder="site_apps"
+
+
+echo "What shall we call the new folder in which the app will be saved?"
+read app_name
+
+
+mkdir $apps_folder_name"/"$app_name
+
+shiny create $apps_folder_name"/"$app_name"/script"

### 2022-08-03 publish.sh

@@ -6,27 +6,27 @@
 
 # Create the dirstribution for each 
 
-folder_prefix="./"
-app_prefix="./shiny_apps/"
+apps_folder_name="shiny_apps"
+site_folder="site_apps"
 
-suffix="/"
-
-for d in ./shiny_apps/*/; do
+for d in ./$apps_folder_name/*/; do
     
-    folder=${d#"$folder_prefix"}
-    folder=${folder%"$suffix"}
 
-    app=${d#"$app_prefix"}
-    app=${app%"$suffix"}
+    folder=${d#"./"}
+    folder=${folder%"/"}
+    app=${folder#$apps_folder_name"/"}
 
     echo "$d"
     echo "$folder"
     echo "$app"
 
-    shiny static $folder site_apps --subdir $app 
+    #shiny static $folder"/script" $folder"/site"
+    shiny static $folder $site_folder --subdir $app
 
 done
 
-#shiny static shiny_apps/app1 site_apps --subdir app1   
 
-#quarto publish netlify
+netlify deploy --prod    
+
+
+quarto publish netlify

### 2022-08-03 shiny_apps/app1/app.py

@@ -1,17 +1,26 @@
+import matplotlib.pyplot as plt
+import pandas as pd
 from shiny import App, render, ui
+from pathlib import Path
+
+
+infile = Path(__file__).parent / 'data.csv'
+df = pd.read_csv(infile)
 
 app_ui = ui.page_fluid(
     ui.h2("Hello Shiny!"),
-    ui.input_slider("n", "N", 0, 100, 20),
-    ui.output_text_verbatim("txt"),
+    ui.input_slider("max_a", "Slider input", min=0, max=10, value=5),
+    ui.output_plot("plt1"),
 )
 
 
 def server(input, output, session):
     @output
-    @render.text
-    def txt():
-        return f"n*2 is {input.n() * 2}"
+    @render.plot
+    def plt1():
+        df_filter = df[df['a']<=input.max_a()]
+        fig = plt.scatter(df_filter.a, df_filter.b)
+        return fig
 
 
 app = App(app_ui, server)

### 2022-08-03 shiny_apps/app1/test.py

@@ -1,0 +1,13 @@
+#%%
+import matplotlib.pyplot as plt
+
+import pandas as pd
+
+df = pd.read_csv('data.csv')
+
+
+# %%
+
+plt.scatter(df.a, df.b)
+
+# %%

### 2022-10-07 PhD_pages/paper_pages/overview.qmd
Add in PhD

---
title: "Overview of papers"
---

```{r, echo=FALSE, message = FALSE, warning=FALSE}

library(tidyverse)
library(trackdown)

```

The revival of research on top incomes since 2000 (refs) has come to have a huge impact on our understanding of economic inequality and its evolution around the world, across academia, policy and civil society alike.

#### Two sets of methods

The top incomes approach is separated by a range of different methodological differences, the key ones being:

-   **the data sources relied upon:**

The use of administrative tax data, national accounts aggregates, or rich lists in place of, or in addition to, household survey data; - **its conception of inequality:**

The approach has an emphasis on the share of income captured by the rich, in contrast to measures of overall inequality such as the Gini coefficient; - **the welfare concept used:**

Fiscal income, or more recently pre-tax national income as defined within the DINA methodology. This contrasts with household disposable income (Canberra Group 2011, chapter 2) or consumption.

Initially at least, these methodological choices by necessity came somewhat bundled together. The choice to use tax data imposed both restrictions on the income concept (taxable income) and a focus on the top (given the profile of tax filers). Conversely, concerns surrounding the ability of survey data to reliably capture top incomes has always been a main motivation for turning to tax data, less affected by issues of undersampling, underreporting or nonresponse.

Methodological innovations within the top incomes literature, including the production of Distributional National Accounts (Alvaredo et al. 2016), have to some extent relaxed this constraint that the data source, conception of inequality and welfare concept must come as a package. Within DINA, survey data, national accounting aggregates, and administrative data are combined to produce 'synthetic microfiles' that can be used to model the whole distribution for a range of welfare concepts.

Nonetheless a primary focus on top income shares of pre-tax income continues to be a defining feature of research efforts in this space, marking it out from conventional approaches focussing on disposable income or consumption, often measured with Gini coefficient.

#### Two views of the world

This distinction between in terms of methods is reflected different hypothesis: two concerning the empirical reality of inequality around the world, one concerning the drivers of these trends.

##### 1) Accounting for 'missing' top incomes, inequality within countries around the world is higher and has risen faster than we thought

Much of the 'action' in the income distribution in recent decades has been towards a greater concentration at the top. This is a phenomenon that standard approaches to inequality measurement -- based on survey data and often using the Gini coefficient as the main summary measure -- are ill-equipped to monitor.

Compared to World Bank assessment -- Inequality falling in a majority of countries. \##### 2) Inequality within countries are considerably greater than inequalities between countries Because within-country inequality is much higher than suggested by survey data, so too is its contribution to global inequality. Working with DINA estimates from the World Inequality Database, Chancel and Piketty (2021) disaggregate global income inequality into between- and within-country components. They find that the majority of global inequality relates to within-country income differences, inverting the results of previous studies anchored more closely to household survey data (Lakner and Milanovic 2016; van Zanden et al. 2014; Bourguignon and Morrisson 2002).

##### 3) The accumulation and concentration of wealth is a key driver of these trends, and an inherent feature of capitalism

A key theme in the top incomes literature -- brought to wider prominence by Piketty's *Capital in the 21st Century* (Piketty 2014) -- is that the rise in inequality in recent decades to a large extent reflects 'fundamental dynamics of capitalism'. The framework set out in Piketty (2014) connects the rise in inequality to rising wealth-to-income ratios. This trend is interpreted within a standard neoclassical aggregate growth model, whereby the stock of capital is growing faster than total income towards a high steady-state capital-to-income ratio. This increases the share of total income received by owners of that capital, which is in turn a key driver of rising inequality, since the vast majority of capital is owned by those earning top incomes. The return to factor incomes as a , stands in contrast with conventional frameworks -- the returns to education, skills premia, labour market interventions, or the progressivity of tax and benefit regime.

This is a simplifification -- Obviously many people thinking about top incomes in terms of adjustment of survey data. We can think of these as 'stylized hypotheses' s: the themes that connect and delineate a given literature and have driven much of the development of its methods -- even if the boundaries are fuzzy.

But taken together ends up in two distinct views of the world:

-   An 'orthodox' view of the world where inequalities within countries are mainly a story of taxes and benefits, skills and wages; where trends in within country inequality -- at least over the period where a global view is possible (roughly 1990 onwards) are very mixed -- reflecting the importance and scope for national policy.; and where global inequalities reflect productivies;

-   And a 'radical' view of the world in which the locus of global inequality is 'the 1%' -- a global elite, driven by wealth accumulation, somewhat inevitable without radical rethink of policy implemented at the global level.

#### Connecting up two worlds of inequality research

Which world we live in matters a great deal for how we prioritise our efforts to reduce global inequality -- national policy making, international policy, in international development and aid, in philanthropy. It affects our view of the consequences of globalisation, good and bad.

And it matters which world we live in: for national and international policy, including on coordinated taxation, international development and aid, philanthropy.

But crucially, these worlds are a set of methods and questions that come bundled together. It is makes the assessment of what world we live in impossible.

The goal of the DPhil is to connect up these two worlds. It does so in two ways.

Firstly on the empirical side. Can we triangulate a clearer view of the world by bringing the methods from both 'worlds'. This is the subject of the first two papers:

Paper 1: With country inequality

Paper 2: Global inequality

Secondly, by addressing the causal hypothesis driving -- namely factor shares and wealth accumulation are driving the trends in income inequality. This is the subject of the second two papers:

Paper 3: Housing and factor shares

Paper 4: Explaining difference between wealth and income inequality in survey data.

### 2022-10-07 PhD_pages/paper_pages/paper1.qmd
Add in PhD

---
title: "Paper 1: Triangulating trends in inequality around the world from secondary databases"
---

```{r, echo=FALSE, message = FALSE, warning=FALSE}

library(tidyverse)
library(trackdown)

```

What has happened to within-country inequality globally?

Is it that there has been a secular rise - suggesting global forces.

Or is it mixed - suggesting much scope for national policy to influence?

Different headline - WID suggests general rise (find a quote). PIP much more mixed... even a fall.

Untangle which methods, definitions.

# Methods and data

## Data sources

### PIP

### WID

### LIS

### Other sources

Should we use Solt? As another check on results?

What about GCIP? For trying to flip between income and consumption...

## Coverage

## Lining up methods

Method 1: Reference years

Method 2: Fully extrapolating/interpolating

The equivalent of the reference year approach where there is no max threshold gap.

Method 3: For the time trend, use a regression with country fixed effects to allow for entry and exit from the sample.

## Comparing distributions

### Comparing welfare concepts

Getting a *global* view of inequalities within countries faces a Catch-22 situation: If you want broad coverage you get very heterogeneous data; if you want comparable data, or to be able to consider different welfare concepts consistently, the coverage you get is much smaller.

Here I'll take two avenues.

Firstly, what we're going for with the 'triangulation' approach is dominance - if different welfare metrics agree

Second we will cross reference to attempts to 'standardise' across these differences: using Solt and GCIP. How valid that is I don't know. But taking the two approaches together we get as close as we can to robust observations.

### A range of inequality metrics

### Lorenz curves

Dominance implies any (well-behaved) summary measure of inequality will agree on the ranking.

Operationalize this by taking 99 points on the Lorenze curve: the percentiles.

This leaves out the *very* top. But this would be quite a warped conception of inequality. I.e. if you have two distributions, A and B, and the share of income received by every whole percentile threshold in B is higher than A, but the Lorenze curves crossed within the top percentile - to claim that inequality is higher in B you'd need a very extreme conception of inequality.

# Results

### Outliers

Russia

# Conclusion

### 2022-10-07 how_to.qmd
Add publish and new app shell scripts
Add in PhD

@@
 ---
-title: "How to"
+title: "How I built this site"
 editor: visual
@@
 
-To create a folder of shinylive apps deployed to netlify within your repo...
+This site is built on three main ingredients:
 
-1.  Get netlify CLI up an running - see here: [https://docs.netlify.com/cli/get-started](https://docs.netlify.com/cli/get-started/)
-2.  Run `netlify init` in the command line (from your local repo).
-    1.  Follow the options.
+-   Quarto notebooks, gathered together into a Quarto website
 
-    2.  Provide a domain name
+-   These are published to the web via Netlify
 
-    3.  For the build command, use `hugo build`
+-   Text editing is done using Google Docs, pulled in to the Quarto notebooks using the R **trackdown** package
 
-    4.  For the directory to deploy, use the 'site' folder associated with the folder of apps. For me this is `./site_apps`
+<!-- To create a folder of shinylive apps deployed to netlify within your repo... -->
 
-    5.  Accept default functions folder
+<!-- 1.  Get netlify CLI up an running - see here: [https://docs.netlify.com/cli/get-started](https://docs.netlify.com/cli/get-started/) -->
 
-    6.  Upon prompting ask to create new `netlify.toml` file
+<!-- 2.  Run `netlify init` in the command line (from your local repo). -->
+
+<!--     1.  Follow the options. -->
+
+<!--     2.  Provide a domain name -->
+
+<!--     3.  For the build command, use `hugo build` -->
+
+<!--     4.  For the directory to deploy, use the 'site' folder associated with the folder of apps. For me this is `./site_apps` -->
+
+<!--     5.  Accept default functions folder -->
+
+<!--     6.  Upon prompting ask to create new `netlify.toml` file -->
+
+<!-- To add a new zsh shell script... -->
+
+<!-- 1.  dsf -->
+
+<!-- 2.  Sort out permissions by running: `chmod u+x <FILNAME>.sh` -->

### 2022-10-07 cms/test_with_trackdown/pull_from_gdocs.R

@@ -1,0 +1,15 @@
+
+
+library(trackdown)
+library(tidyverse)
+
+
+
+fp<- "PhD_pages"
+
+rmd_filedirs <- dir_ls(path = paste0(fp,"/rmarkdown"), glob = "*.Rmd")
+
+# rmd_filedirs <- gsub(".Rmd","-markdown.Rmd", group$group)
+
+
+download_file(gpath = paste0("joehasell.netlify.app/",fp), file = rmd_filedirs)

### 2022-10-07 cms/test_with_trackdown/push_to_gdocs.R

@@ -1,0 +1,12 @@
+
+library(trackdown)
+
+
+fp<- "PhD_pages"
+
+rmd_filedirs <- dir_ls(path = paste0(fp,"/rmarkdown"), glob = "*.Rmd")
+
+
+update_file(gpath = paste0("joehasell.netlify.app/",fp), file = rmd_filedirs, hide_code = TRUE)
+
+

### 2022-10-07 cms/test_with_trackdown/rmarkdown_to_quarto.R

@@ -1,0 +1,21 @@
+library(fs)
+library(stringr)
+
+fp<- "PhD_pages"
+
+rmd_filedirs <- dir_ls(path = paste0(fp,"/rmarkdown"), glob = "*.Rmd")
+
+qmd_filedirs <- str_replace(string = rmd_filedirs,
+                         pattern = "Rmd",
+                         replacement = "qmd")
+
+
+qmd_filedirs <- str_replace(string = qmd_filedirs,
+                            pattern = "/rmarkdown",
+                            replacement = "/quarto")
+
+file_copy(path = rmd_filedirs,
+          new_path = qmd_filedirs,
+          overwrite = TRUE)
+
+

### 2022-10-07 sync_gdocs.R

@@ -1,0 +1,11 @@
+
+
+library(trackdown)
+
+
+upload_file(gpath = "joehasell.netlify.app/PhD_pages", file = "PhD_pages/overview.Rmd", hide_code = TRUE)
+
+
+
+
+download_file(gpath = "joehasell.netlify.app/PhD_pages", file = "PhD_pages/overview.Rmd")

### 2022-10-11 PhD_pages/paper_pages/paper1.qmd
testing out PhD workflow

@@
 
-Is it that there has been a secular rise - suggesting global forces.
-
-Or is it mixed - suggesting much scope for national policy to influence?
-
 Different headline - WID suggests general rise (find a quote). PIP much more mixed... even a fall.
 
-Untangle which methods, definitions.
+Which world we live in matters for how we think we should tackle inequality.
+
+A secular, general rise - suggesting global forces, perhaps something intrinsic to capitalism (Piketty, 2014).
+
+A mixed picture of inequality around the world suggests reflect specific circumstances, with accordingly much scope for national policy to influence.
+
+Both are reasonable methods and have strengths and weaknesses.
+
+I want to:
+
+a\) Triangulate the key stylized facts. Which observations are robust to across available datasets with their different methods, and which are contingent?
+
+b\) Untangle which methods, definitions affect the different conclusions arrived at.
 
@@
 
-## Coverage
+### Coverage
 
-## Lining up methods
+s
+
+## Methods: Comparing databases
+
+### Lining up years
 
@@
 Method 3: For the time trend, use a regression with country fixed effects to allow for entry and exit from the sample.
-
-## Comparing distributions
 

### 2022-10-11 PhD_pages/scripts/PIP/prepare_pip_percentiles.qmd
testing out PhD workflow

```{r}
library(tidyverse)

source('PhD_pages/scripts/functions.r')

```

```{r}
#These variables uniquely id a set of survey results
pip_id_vars<- c('Entity', 'Year', 'reporting_level', 'welfare_type')

```

```{r}
pip_perc_orig<- read.csv("PhD_pages/data/original/pip_percentiles.csv")

```

```{r}
pip_main_vars_orig<- read.csv("PhD_pages/data/original/pip_main_vars.csv")

```

#### Prep 'percentiles' data file

```{r}
pip_perc<- pip_perc_orig
```

```{r}
# Drop duplicate observations of 'headcount' - The PIP API returns the . Given the discrete nature of the survey data, it sometimes returns a headcount ratio that are fairly far from the target percentile queried and the same headcount ratio can be returned for multiple percentile queries. Here we keep only one.

pip_perc<- drop_dupes(pip_perc,
                       id_vars = pip_id_vars,
                        duplicate_data_var = 'headcount')

```

```{r}

# Browse for nonmonotonic rows
# browse_nonmonotonic<- monotonicity_check(
#   df = pip_perc, 
#   id_vars = pip_id_vars,
#   order_var = 'headcount',
#   data_var = 'percentile_value', 
#   action = 'browse')

#There seems to be an issue with China in 1993,  driven by rural China.
# br<- pip_perc %>%
#   filter(Entity == "China",
#          reporting_level == 'national',
#          Year == 1993)

# Let's drop these rows
pip_perc<- monotonicity_check(
  df = pip_perc, 
  id_vars = pip_id_vars,
  order_var = 'headcount',
  data_var = 'percentile_value', 
  action = 'drop')

```

```{r}
#Drop world regions from percentile data

# These are entities with no reporting level - as veried below
# pip_perc %>%
#   filter(reporting_level=="") %>% 
#   select(Entity) %>%
#   unique() %>%
#   print()

pip_perc<- pip_perc %>%
    filter(reporting_level!="")

```

#### Prep 'main vars' data file

```{r}

pip_main_vars<- pip_main_vars_orig %>%
  rename(Entity = country_name,
         Year = reporting_year) %>%
  select(Entity, Year, reporting_level, welfare_type, mean, reporting_pop, reporting_gdp, gini, mld)
```

#### Merge the percentile and main data files

```{r}
pip_perc<- left_join(pip_perc, pip_main_vars)

```

#### Additional cleaning

```{r}
#Keep only 'National' estimates, except for country-years that don't have national estimates.

# Count how many reporting levels there are for each estimate
count_reporting_levels<- pip_main_vars %>%
  select(Entity, Year, reporting_level) %>%
  unique() %>%
  count(Entity, Year) %>%
  rename(number_of_reporting_levels = n)

# We see there only 1 and 3 reporting levels - i.e. either national, rural or uban alone, or else all three are present.
count_reporting_levels %>% select(number_of_reporting_levels) %>% unique()

# Merge the counts back into dataframe
pip_main_vars<- left_join(pip_main_vars, count_reporting_levels)

# Filter to see if there are any cases where only non-national estimates are available
pip_main_vars %>%
filter(number_of_reporting_levels<3 & reporting_level!='national') %>%
select(Entity) %>%
unique()

# Keep only national estimates except for these cases

pip_main_vars<- pip_main_vars %>%
filter(reporting_level=='national' | number_of_reporting_levels<3)

# Check again if there is more than one reporting level
pip_main_vars %>%
  select(Entity, Year, reporting_level) %>%
  unique() %>%
  count(Entity, Year) %>%
  rename(number_of_reporting_levels = n) %>%
    filter(number_of_reporting_levels>1)

```

### Prepare reference year data

```{r}
# I'll prepare the observations from the main_vars file, and then merge in the percentiles

target_years<- c(1990, 2000, 2010, 2018)

df_pip_ref_years<- keep_closest(
  df= pip_main_vars,
  id_vars = c("Entity", "reporting_level"),
  distance_varname = "Year",
  targets = target_years,
  tie_break_dist_up_or_down = "up",
  tie_break_varname = "welfare_type",
  tie_break_value = "income")

# We see that if we impose that a max distance, country coverage falls.
df_count<- df_pip_ref_years %>%
  count(Entity)

df_pip_ref_years_filter<- df_pip_ref_years %>%
  filter(abs_distance<3)

df_count<- df_pip_ref_years_filter %>%
  count(Entity)

```

### 2022-10-11 PhD_pages/scripts/functions.r

@@ -1,0 +1,145 @@
+
+library(tidyverse)
+
+# A function to drop duplicates. Because the API sometimes returns the same headcount share for different requested percentiles we get duplicate headcount/poverty line observations. Later we use this function to drop these... 
+
+drop_dupes<- function(df, id_vars, duplicate_data_var){
+    
+    dedupe_vars<- c(id_vars, duplicate_data_var)
+
+    df<- df %>%
+        distinct(across(all_of(dedupe_vars)), .keep_all = TRUE)
+    
+return(df)
+
+}
+
+
+monotonicity_check<- function(df, id_vars, order_var, data_var, action){
+  
+  
+  # Group, order and lag
+  lag_data_var<- paste0("lag_", data_var)
+  
+  data_var <- sym(data_var)
+  
+  # df<- df %>%
+  #    group_by_at(id_vars) %>% 
+  #    arrange_at(order_var, .by_group = TRUE) %>%
+  #    mutate({{lag_data_var}} := dplyr::lag(!!data_var, n = 1, default = NA)) %>%
+  #    arrange_at(c(id_vars,order_var))
+  
+  df<- df %>%
+    group_by_at(id_vars) %>% 
+    arrange_at(order_var, .by_group = TRUE) %>%
+    mutate({{lag_data_var}} := dplyr::lag(!!data_var, n = 1, default = NA)) %>%
+    arrange_at(c(id_vars,order_var))
+  
+  lag_data_var <- sym(lag_data_var)
+  
+  
+  if(action == "browse"){
+    df<- df %>% filter(!!lag_data_var>=!!data_var)
+  } else if (action == "drop"){
+    df<- df %>% filter(!!lag_data_var<!!data_var)  
+  } else {
+    print("Action parameter must be set to 'browse' or 'drop'.")
+  }
+  
+  df<- df %>% ungroup()
+  return(df)
+  
+}
+
+
+
+keep_closest<- function(df, id_vars, distance_varname, targets, 
+                        tie_break_dist_up_or_down, tie_break_varname = NULL, 
+                        tie_break_value = NULL){
+  
+#The function assumes that distance_var is unique within the id_vars or, if there
+  #is a tie_break_var uniques within id_vars and tie_break_var taken together.
+  
+  # Example values
+  # df<- pip_main_vars
+  # targets<- c(1990, 2010)
+  # t<- 2010
+  # id_vars<- c('Entity', 'reporting_level')
+  # distance_varname<- 'Year'
+  # tie_break_dist_up_or_down<- "down"
[…]

### 2022-10-12 PhD_pages/paper_pages/overview.qmd
Adding to overview and papers 1 and 3

@@
 title: "Overview of papers"
+bibliography: references.bib
 ---
@@
 library(tidyverse)
-library(trackdown)
 
@@
 
-The revival of research on top incomes since 2000 (refs) has come to have a huge impact on our understanding of economic inequality and its evolution around the world, across academia, policy and civil society alike.
+Thanks to the efforts of a number of key teams involved in the production of harmonized inequality datasets, we are in a far better position than ever to describe and monitor trends in income inequality across the world as a whole.
+Recent years have seen considerable improvements, either in terms of spatial or temporal coverage of these datasets, the degree of comparability achieved across data points, or both.
 
-#### Two sets of methods
+However, the key takeaways of this substantial body of empirical work remain surprisingly, and unhelpfully fractured.
 
-The top incomes approach is separated by a range of different methodological differences, the key ones being:
+For example, available inequality estimates based on Distributional National Accounts - as collated and summarise by the World Inequality Lab - depict a near universal rise inequality in recent decades.
 
--   **the data sources relied upon:**
+> Income and wealth inequalities have been on the rise nearly everywhere since the 1980s, following a series of deregulation and liberalization programs which took different forms in different countries.
+[@chancel2021]
 
-The use of administrative tax data, national accounts aggregates, or rich lists in place of, or in addition to, household survey data; - **its conception of inequality:**
+In stark contrast, national survey data collected and harmonized by the World Bank show a far more mixed picture, with overall a decline in average inequality more recently.
 
-The approach has an emphasis on the share of income captured by the rich, in contrast to measures of overall inequality such as the Gini coefficient; - **the welfare concept used:**
+> Within-country inequality as a whole---that is, considering together the trends of all countries on which evidence is available---\[...\] started to narrow in the last decade, after peaking in the 1990s.
+[@WBpoverty2016]
+
+Alongside these differing views of within-country inequality around the world are starkly contrasting views concerning the locus of global inequality: with inverse findings concerning the respective contributions made by within- and between country inequalities.
+[@chancel2021, @BourgMorrisons2002, @laknerMilanovic2016]
+
+Despite the far richer, wider range of distributional data at our disposal, basic questions concerning the nature and direction of inequality around the world remain, effectively, unanswered.
+
+One explanation for this disconnect - between the greatly improved data and the ambiguous understanding of inequality globally that this has yielded - lies in extent to which the available data has emerged from the efforts of different networks of researchers, coalescing around different institutions, most notably the World Bank, the World Inequality Lab and the Luxembourg Income Study.
+These approaches to the measurement of inequality around the world are separated by a range of methodological differences woven together, somewhat as a package.
+Related to choice of methods, the approaches are also separated by different emphasis on hypotheses regarding the nature of inequality and its drivers - in particular the role played by factor shares in driving income inequality, particularly at the top.
+They are also separated technologically - presented on different platforms each with fairly high start-up costs required for proficient use.
+
+A causal observer would be forgiven for thinking that the different efforts occupy almost different 'worlds' of inequality research, where it is difficult to compare results, to understand where approaches agree, where they differ and why.
+Since the answers arrived at within each 'world' of inequality measurement can be strikingly different, very basic questions concerning inequality across the world remain, in effect, unanswered.
+
+In the public and policy debate, this leads very different narratives that tends to focus, on the one hand, on a runaway global elite - 'the 1%' - or on the other hand on 'convergence' and 'shared prosperity'.
+Within these different narratives sit divergent views concerning a variety of important and urgent questions: the outcomes, merits and demerits of global economic development and globalisation; and the scope and proper role of international development aid, philanthropy, and taxation - at national and international levels - to shape a fairer, more equal world.
+
+Which of these 'worlds' of inequality we live in matters a great deal - but the way the data supporting these two views of the world has been presented and compared to date leaves us to a large extent in the dark as to which world we actually live in.
+By not 'speaking to each other', the value to society of the enormous efforts collectively made by hundreds of researchers and statisticians around the world in the production of these datasets is greatly diminished.
+
+#### Distinct methods
+
+\[Breakdown the bundled methods: top incomes approach vs 'conventional' approaches\]
+
+-   **The data sources relied upon:**
+
+The use of administrative tax data, national accounts aggregates, or rich lists in place of, or in addition to, household survey data
+
+-   **The conception of inequality adopted:**
+
+The approach has an emphasis on the share of income captured by the rich, in contrast to measures of overall inequality such as the Gini coefficient;
+
+-   **The welfare concept typically measured:**
 
@@
 
-#### Two views of the world
+#### Distinct views of inequality and its drivers
 
-This distinction between in terms of methods is reflected different hypothesis: two concerning the empirical reality of inequality around the world, one concerning the drivers of these trends.
+This distinction between in terms of methods is reflected different hypothesis concerning the empirical reality of inequality around the world, one concerning the drivers of these trends.
 
@@
 Compared to World Bank assessment -- Inequality falling in a majority of countries.
-\##### 2) Inequality within countries are considerably greater than inequalities between countries Because within-country inequality is much higher than suggested by survey data, so too is its contribution to global inequality.
+
+##### 2) Inequality within countries are considerably greater than inequalities between countries
+
+Because within-country inequality is much higher than suggested by survey data, so too is its contribution to global inequality.
 Working with DINA estimates from the World Inequality Database, Chancel and Piketty (2021) disaggregate global income inequality into between- and within-country components.
@@
 
-#### Connecting up two worlds of inequality research
+#### Distinct platforms
 
-Which world we live in matters a great deal for how we prioritise our efforts to reduce global inequality -- national policy making, international policy, in international development and aid, in philanthropy.
-It affects our view of the consequences of globalisation, good and bad.
+\[I'm not sure whether to include this, but it seems to me that the root of many of these problems is essentially technological: the infrastructure and practices we have to compare and 'switch' between different methodological choices in our analysis of global inequality is very poor.
+Whilst the code behind the production of inequality estimates is become more available, in practice the barriers to this kind of joined-up analysis remain prohibitively high.\]
 
-And it matters which world we live in: for national and international policy, including on coordinated taxation, international development and aid, philanthropy.
+#### Connecting up the worlds of inequality research
 
-But crucially, these worlds are a set of methods and questions that come bundled together. It is makes the assessment of what world we live in impossible.
-
-The goal of the DPhil is to connect up these two worlds. It does so in two ways.
+The goal of the DPhil is to connect up these worlds. It does so in two ways.
 
@@
 
-Paper 4: Explaining difference between wealth and income inequality in survey data.
+Paper 4: Explaining the poor correlation between wealth and income inequality.

### 2022-10-12 PhD_pages/paper_pages/paper3.qmd
Adding to overview and papers 1 and 3

---
title: "Paper 3: Housing and the capital share of income"
editor: visual
---

## The housing share of income

What share of aggregate income is made up by rents on housing assets and how has this evolved in recent decades? The national accounts offers three avenues for answering this question: operating surplus within the household sector ('Household sector' approach); operating surplus within the real estate industry ('Real estate industry' approach); and household final consumption of housing services ('Final consumption' approach).

Figure 1 shows the housing capital share of factor cost value added calculated according to three approaches.

-   **Capital income of the real estate industry\
    **(Give brief description)

-   **Household final consumption of rent\
    **(Give brief description)

-   **Gross operating surplus of the household sector\
    **(Give brief description)

\

### Trends in the housing share

Comparing the housing share series plotted in Figure 1, we see the figures for the industry approach are quite consistent across the two data sources: the early vintage KLEMS data (red line) and OECD data (green line). The series adopting the second approach of final household consumption of rent (blue line) also tracks both these series closely.

As expected, the third approach of gross operating surplus in the household sector (purple line) is generally an outlier. How far this series falls from the others varies considerably across countries, reflecting the varying accounting conventions and tenure structures just discussed.

A notable exception to the above characterisation is the UK, in which the more recent data from the OECD (beginning in 1995) for all three approaches falls substantially above the share calculated from the earlier KLEMS industry data. The reason for this large discrepancy is not entirely clear, but may relate to recent revisions to ONS practices for calculating imputed rent. (FIND REFERENCE).

Before turning to summary statistics, an inspection of Figure 1shows that there have been substantial rises in a majority of the twenty countries. Many countries -- including France, Germany, Italy and Spain -- have seen an increase in the housing share roughly five percentage points across five decades, or one point per decade. To give a sense of scale, it serves to compare this with the changes in the factor distribution found in the existing literature. Karabarbounis and Neiman [(2014)](https://paperpile.com/c/C530Q2/vTOu/?noauthor=1) find a 1.3 points-per-decade fall in the overall 'global' labour share between 1975 and 2012; Rodriguez and Jayadev [(2010)](https://paperpile.com/c/ztky5j/NxrB/?noauthor=1) find xxxx, and Piketty finds xxxx across X countries.

A number of countries, however have seen only modest rises or none at all -- Ireland, Norway and Sweden all falling in the latter group. It is only in Portugal that we see a substantial fall over the period, (Presumably the revolution/nationalizations? But a) why would this affect the housing share so much, b) the fall seems to begin in 1977 -- some time after the revolution an economic downturn).

To facilitate comparisons I construct a single housing share series for each country. For most countries, the series is centred around the OECD industry-level data which is then extended back using the KLEMS industry-level data where available for earlier years. In the case of Canada, United States and Spain, for which recent OECD industry data is not available, the rent consumption share is used. Earlier and later series are linked on a proportionate basis: the earlier data is adjusted by the ratio of the two series for the earliest overlapping year.

(The series based on gross operating surplus in the household sector (approach 3) are excluded entirely due to the comparability issues discussed. For each individual country the available series are 'averaged' by plotting the year fixed effects from a linear regression across all series, normalised so as to coincide with the earliest series. See Appendix **X** for details and regression results.)

Figure 2 plots the 'average' of the resulting housing share series across all twenty countries. This represents the year fixed effects from a linear regression of the synthetic housing share on country and year dummies, normalised so as to coincide with the mean across all earliest series. This is used in order to account for entry and exit of countries from the sample.

On average x point rise.

## Housing and non-housing capital shares

To see how this increase in housing income has contributed to changes in the overall functional distribution, Figure 1 shows the capital share of income across 18 high income countries from roughly 2017 separating out housing (dotted lines) from non-housing (solid lines) capital income. The data exclude net taxes on production, such that capital and labour income sum to 100% of value added (i.e. at factor cost). On this basis, a rise in the capital share of 1 point implies a fall in the labour share of 1 point.

In order to increase coverage, trends are shown from three sources as follows.

\

#### 1) EU-KLEMS -- industry-level data

The EU-KLEMS database collates industry-level data for most European countries as well some larger OECD countries. This includes industry figures for value added, compensation of employees and gross operating surplus (the residual balance after subtracting employee compensation from value added). In addition, labour income attributable to the activity of the self-employed is imputed and separated out from gross operating surplus, resulting in a 'capital' and 'labour' income series for each industry which together sum to total value added. The imputation is made using employment data, based on the assumption that the hourly compensation of the employed and self-employed is equal on average within each industry.

In the series presented below, we use the capital income reported within the real estate industry as a proxy for housing capital income. Since activities other than housing services are included in this industry (such as the rent of offices, or activity involved in the sale or marketing of property) this leads to an overestimate of the housing capital share per se.

Comparison with national accounts data on households' consumption expenditure on rent (both actual and imputed) shows that in many countries the real estate industry has grown somewhat faster than housing services (Appendix X). Arguably, it makes sense to include these additional activities along with housing since their growth is likely to derive in part from the same underlying force -- namely, rising demand for urban space against an inelastic supply. As shown in Appendix X however, where the boundary of 'housing' is drawn -- whether the real estate industry as a whole or only including dwellings -- does not substantively affect the resulting trends in the housing and non-housing capital shares in any case.

Country and year coverage varies across different vintages of the EU-KLEMS data, in part due to changes in industry classification over time. In order to maximise coverage in earlier years, I rely on the 2009 release from which long-run series, generally running from 1970 to 2007, can be obtained for 18 high income countries.

#### 2) OECD -- industry-level data

The OECD also provide national accounts data at the industry-level similar to the EU-KLEMS data used, though based on a more recent industry classification. Country coverage is wider, and observations generally run up to 2017 or later, but in most cases only begin from around 1995.

In the series presented below, capital and labour components of value added (at factor cost) are obtained following the same imputation method for self-employment activity as used in the EU-KLEMS dataset, based on industry employment data also from the OECD (see Appendix X for further discussion).

\

#### 3) OECD -- institutional sector data

A third series is based on national accounts data broken down by institutional sector (corporations, households, government and non-profit institutions serving households) rather than by industry.

This is added in order to further increase coverage, and to connect with the existing factor shares literature which commonly relies on this approach (see for instance Piketty and Zucman [(2014)](https://paperpile.com/c/ztky5j/VuzZ/?noauthor=1), Karabarbounis and Neiman [(2014)](https://paperpile.com/c/ztky5j/qk46/?noauthor=1) and Rognlie [(2016)](https://paperpile.com/c/ztky5j/CmLb/?noauthor=1)). However, it suffers from two important drawbacks.

Firstly, the question of how to treat the mixed income arising from self-employment activity is particularly problematic in the context of sector-level data. In the series presented below, I adopt a commonly used approach whereby mixed income -- reported as a separate line in the household sector accounts -- is split into capital and labour components based on the assumption that the respective shares are equal to those observed in the corporate sector. In addition to the uncertainty surrounding this assumption, however, Pionnier and Guidetti [(2015)](https://paperpile.com/c/ztky5j/Jgmu/?noauthor=1) point to a significant limitation of this approach on account of the varying treatment of unincorporated enterprises across countries. In some countries, substantial self-employment activity may be recorded fully as corporate gross operating surplus, due to the inclusion of some unincorporated enterprises within the corporate sector as quasi-corporations. In the case of Italy and Germany they demonstrate that failing to account for the labour income implicit within such activity biases the labour share downward substantially, and also affect trends as the composition of employment changes over time.

Secondly, sector-level data also presents difficulties when seeking to isolate the capital income relating specifically to housing. Previous studies have tended to take the gross operating surplus of the household sector as the relevant measure [(Piketty and Zucman 2014; Rognlie 2016)](https://paperpile.com/c/ztky5j/VuzZ+CmLb). According to national accounting conventions, this is where the imputed rent attributable to owner-occupied housing is recorded, which for most countries represents the large majority of dwellings.

There are two issues with this approach however. Firstly, whilst the corporate sector involvement in the housing market is often small, it is by no means insignificant in all cases. In addition to houses and apartments owned in the for-profit sector, this includes social housing operated by public or non-profit corporations. The latter comprises, for instance, 18% of the UK housing stock in recent years -- down from roughly 30% through the 1970s.

Secondly, there appears to be inconsistency across national accounting practices with respect to private rentals on the part of unincorporated individuals, which often represent the bulk of the private rented sector. In some countries this is recorded as household gross operating surplus, but in others it is included as mixed income (see Appendix X for further discussion). Consistent with these two observations, household gross operating surplus is often substantially lower than two other more comprehensive measures of housing income available in the national accounts (see Figure A.1).

For this reason, in the series below we use households' final consumption expenditure on rent (both actual and imputed) in order to separate out housing from non-housing capital income within the sector data. These data exclude any costs relating to maintenance, services or utilities, but are, nevertheless, probably a slight overestimate of housing capital income since they may include some labour income (such as the management activity of landlords) and production taxes (since it is measured at market prices).

Given these problems relating to the institutional sector data, this third series is generally not included in the calculation of average trends (section X below). The exceptions are the US and Spain, which lack OECD industry-level data for recent years and are not affected by the aforementioned issue relating to quasi-corporations.

\

### Country trends in the housing and non-housing capital share

The three series are plotted in Figure 1. (Further discussion)

Table 1 indicates the average annual points change observed in each series, multiplied by 10. As a rough indicator of the centrality of housing to changes in the overall housing share, positive changes in the housing capital share greater than the corresponding change seen in the non-housing capital share are highlighted in red (including where the non-housing capital share fell). We see that in the earlier KLEMS data, the majority of countries saw rises in the housing capital share greater than those in the non-housing capital share. In the two later OECD series we see a more even split between countries with rising and falling housing shares. (Further discussion).

\

#### Average trends

To consider average trends, the various series are clustered into an 'early' grouping and a 'later' grouping. The early grouping is made up of observations from 1970 to 2007 which for almost all countries is taken from the KLEMS industry data. The later grouping is made up of observations from 1995 onwards which for almost all countries is taken from the OECD industry data. The exceptions are Denmark and Norway for which a complete series in the OECD industry data is available. In these cases the OECD data is used twice over, for both the 'early' and 'late' groupings. For Japan, only the KLEMS industry data is available, and again this is included in both the late and early groupings. In the case of the United States and Spain, for which no OECD industry-level data is available, observations for the later grouping are taken from the sector accounts and housing rent consumption data (see discussion in Section X).

'Average' housing and non-housing capital shares are then obtained within each respective grouping by taking the year fixed effects from a linear model controlling for country fixed effects to account for the (few instances of) entry or exit from the sample. These are then normalised such that the value in the initial year corresponds to the mean across countries for that year. These average series are plotted for the early and late groupings in figure 2.

\
The fact that the gradients of the average housing and non-housing capital share trend lines are, broadly speaking, similar gives us an overall indication of the importance of housing to movements in the overall capital share: its contribution alone was comparable to the aggregate across all other industries.

This is confirmed in Table 1, which shows the change in the averages, broken down into sub-periods to facilitate comparison between the early and late series. The first four columns split the trends into two roughly equal subperiods: 1970-1995 and 1995-2017. 1995 marks the year in which the later series for the average capital shares begins (the earliest year for which the OECD industry data is available for most countries).

We see from the earlier (primarily KLEMS) data that the quarter-century from 1970 to 1995 saw a substantial rise in the overall capital share. The average capital share rose by 4.6 points across the 18 high-income countries in our sample, and 3.5 points where weighted by GDP at PPPs. But around two-thirds of this increase owed to rising housing shares.

The later series shows a much more modest rise in the overall capital share between 1995 and 2017, with housing nevertheless playing an important role. Rising housing shares drove almost all of the 1 point rise in the unweighted average for this period, and about half of the 2.5 point rise in the weighted average.

The four columns to the right of Table 2 provide a further breakdown of the later period, split at 2007 which marks the end of the early KLEMS data. As visible from Figure 2, where the early and later series overlap, there is some disagreement between them concerning the non-housing capital share. Between 1995 and 2007 the early series average rose by around 1-1.5 points more than the later series, with no apparent slow-down.

Referring back to Figure 1, we see that whilst the more recent OECD industry data does generally track the earlier KLEMS data very closely, for a small number of countries -- Greece, Italy and to a lesser extent the United Kingdom -- there is a large initial discrepancy beginning in 1995 that decreases over time. Similarly, for the United States the later series based on institutional sectors rises more slowly than indicated in the KLEMS data. \[Can we say anything about the reasons for these discrepancies?\]

Taken together, however, the overall view of the past half century afforded by the available data sources is not much affected by the differences visible across them. Table 3 sums the points changes seen in the averages across the subperiods, as a summary across the full period.

We see the rise in the overall capital share varies modestly depending on whether we make a break to the later series from where it begins in 1995 (left two columns) or from where the earlier series ends in 2007 (right two columns). Similarly, the average rise also depends somewhat on whether countries are weighted by GDP, with some larger economies seeing greater rises (see Table 1). But any combination of the series results in a rise in the overall capital share in the region of 6-7 points between 1970 and 2017.

Likewise, whilst there are differences across specifications, the centrality of housing to changes in the overall capital share over the past five decades is clear. On a weighted basis, rising housing shares contributed to around half of the rise in the overall capital share on average. Taking the unweighted average, the contribution was larger still, at around two-thirds of the overall rise.

\

\

### 2022-10-12 PhD_pages/scripts/PIP/prepare_pip_percentiles.qmd
Adding to overview and papers 1 and 3

@@
 
-target_years<- c(1990, 2000, 2010, 2018)
-
-df_pip_ref_years<- keep_closest(
-  df= pip_main_vars,
-  id_vars = c("Entity", "reporting_level"),
-  distance_varname = "Year",
-  targets = target_years,
-  tie_break_dist_up_or_down = "up",
-  tie_break_varname = "welfare_type",
-  tie_break_value = "income")
-
-# We see that if we impose that a max distance, country coverage falls.
-df_count<- df_pip_ref_years %>%
-  count(Entity)
-
-df_pip_ref_years_filter<- df_pip_ref_years %>%
-  filter(abs_distance<3)
-
-df_count<- df_pip_ref_years_filter %>%
-  count(Entity)
+# target_years<- c(1990, 2000, 2010, 2018)
+# 
+# df_pip_ref_years<- keep_closest(
+#   df= pip_main_vars,
+#   id_vars = c("Entity", "reporting_level"),
+#   distance_varname = "Year",
+#   targets = target_years,
+#   tie_break_dist_up_or_down = "up",
+#   tie_break_varname = "welfare_type",
+#   tie_break_value = "income")
+# 
+# # We see that if we impose that a max distance, country coverage falls.
+# df_count<- df_pip_ref_years %>%
+#   count(Entity)
+# 
+# df_pip_ref_years_filter<- df_pip_ref_years %>%
+#   filter(abs_distance<3)
+# 
+# df_count<- df_pip_ref_years_filter %>%
+#   count(Entity)
 

### 2022-10-31 PhD_pages/data_appendices/3_A_prepare_KLEMS.qmd
Adding data appendix for paper 3

---
title: "Appendix 3.A: Preparing EU KLEMS industry data"
format: html
---

```{r}
library(tidyverse)
```

## Introduction

## Raw data files

Make a list of filenames (both to reference the correct file in the url and as the filename to save locally).
```{r}

files<- c("32_countries",
          "Canada",
          "US_alt")

# The urls where the files can be retrieved
download_urls<- c("http://www.euklems.net/data/09i/all_countries_09I.txt",
               "https://www.worldklems.net/data/basic/CAN_WK_07_2012.xlsx",
               "https://www.worldklems.net/data/basic/usa_wk_mar_2017.xlsx"
)

names(download_urls)<- files

# uses the last part of the download urls as the filenames
filenames<- sub(".*\\/", "", download_urls)

names(filenames)<- files

# Add directory to specify filepaths in which to save a local copy of each file
filepaths<- paste0("PhD_pages/data/original/KLEMS/", filenames)

names(filepaths)<- files

```

Save files locally.
```{r}

for (file in files){
  print(file)
download.file(download_urls[file], filepaths[file])
}

```

## Preparing dataframes

Importing the main data file - with data for 32 countries.
```{r}

df <- read.csv(filepaths["32_countries"], quote="")

# convert to long format and rename variables
df<- df %>%
  pivot_longer(cols = starts_with("X_"), names_to = "year",
    names_prefix = "X_", values_to ="value",
    names_transform = list(year=as.integer),
    values_transform = list(value=as.character)) %>%
  rename(country = X.country,
         transaction = var)

# Remove quotations (These come from not properly specifying the file structure in the read.csv command above. But I was unable to find the right format spec.)
df <- df %>% 
  mutate(country = str_replace(country, '"', ''),
         value = as.numeric(str_replace(value, '"', '')))

```

## Calculatiing derived variables

### 2022-10-31 PhD_pages/data_appendices/3_A_prepare_KLEMS.qmd
prep OECD data

@@
 library(tidyverse)
+library(readxl) # For handling xls source data files
+
 ```
@@
           "Canada",
-          "US_alt")
+          "US_alt",
+          "Japan")
 
@@
                "https://www.worldklems.net/data/basic/CAN_WK_07_2012.xlsx",
-               "https://www.worldklems.net/data/basic/usa_wk_mar_2017.xlsx"
+               "https://www.worldklems.net/data/basic/usa_wk_apr_2013.xlsx",
+               "https://www.worldklems.net/data/basic/JPN_wk_may_2013.xlsx"
 )
@@
 
+```{r}
+df_list<- list()
+```
+
 Importing the main data file - with data for 32 countries.
@@
 
-df <- read.csv(filepaths["32_countries"], quote="")
+df_list[["32_countries"]]<- read.csv(filepaths["32_countries"], quote="")
 
 # convert to long format and rename variables
-df<- df %>%
+df_list[["32_countries"]]<- df_list[["32_countries"]] %>%
   pivot_longer(cols = starts_with("X_"), names_to = "year",
@@
 # Remove quotations (These come from not properly specifying the file structure in the read.csv command above. But I was unable to find the right format spec.)
-df <- df %>% 
+df_list[["32_countries"]]<- df_list[["32_countries"]] %>%
   mutate(country = str_replace(country, '"', ''),
@@
 
-## Calculatiing derived variables
+Data for Canada, Japan and an alternative (longer) series for the US.
+```{r}
+countries<- c("Canada", "US_alt", "Japan")
+
+for (country in countries){
+df_list[[country]]<- as.data.frame(read_excel(filepaths[country],
+                                                         sheet = 'DATA'))
+
+# convert to long format and rename variables
+df_list[[country]]<- df_list[[country]] %>%
+  pivot_longer(cols = starts_with("_"), names_to = "year",
+    names_prefix = "_", values_to ="value",
+    names_transform = list(year=as.integer)) %>%
+  rename(transaction = Variable) %>%
+  mutate(country = country)
+
+}
+
+```
+
+Tidying Canada data
+```{r}
+# Drop some NA values (- data for 2009 and 2010 is given as missing)
+df_list[["Canada"]]<- df_list[["Canada"]] %>%
+  filter(!is.na(value))
+
+# For Canada we need to calculate some industry aggregates to be comparable with other countries
+agg_industries<- list()
+
+agg_industries[['D']]<- c(
+  "15t16",
+  "17t19",
+  "20",
+  "21t22",
+  "23",
+  "24",
+  "25",
+  "26",
+  "27t28",
+  "29",
+  "30t33",
+  "34t35",
+  "36t37")
+
+agg_industries[['G']]<- c(
+  "50",
+  "52")
+
+agg_industries[['I']]<- c(
+  "60t63",
+  "64")
+
+agg_industries[['K']]<- c(
+  "70",
+  "71t74")
+
+for (agg_ind in names(agg_industries)){
+  print(agg_ind)
+
+  df_totals<- df_list[["Canada"]] %>%
+  filter(code %in% agg_industries[[agg_ind]]) %>%
+    group_by(transaction, year) %>%
+    summarise(value = sum(value)) %>%
+    ungroup() %>%
+    mutate(code = agg_ind) %>%
+    mutate(country = "Canada")
+
+  df_list[["Canada"]]<- bind_rows(df_list[["Canada"]], df_totals)
+
+}
+
+```
+Tidying Japan data
+```{r}
+# Drop NA values
+df_list[["Japan"]]<- df_list[["Japan"]] %>% filter(is.na(value))
+
+# Rename the capital compensation variable, to agree with other files
+df_list[["Japan"]]<- df_list[["Japan"]] %>%
+  mutate(transaction = replace(transaction,
+                               transaction == "CAP_GFCF",
+                               "CAP"))
+
+```
+
+Join the different dataframes together into a single dataframe.
+```{r}
+df<- bind_rows(
+  df_list
+)
+
+```
+
+Make a dictionary of industry codes and there descriptions from the combined data sources.
+```{r}
+
+df_industries<- df %>%
+  select(code, desc) %>%
+  drop_na() %>%
+  unique()
+
+# There are a couple of different spellings etc. from the different datasets. Here we just take select one (arbitrarily)
+df_industries<- df_industries %>%
+  distinct(code, .keep_all = TRUE)
+
+# Add in the uniform industry code descriptions
+df<- df %>%
+  select(-desc)
+
+df<- left_join(df, df_industries)
+
+```
+
+Some more tidying
+```{r}
+df<- df %>%
+  filter(transaction %in% c("VA", "CAP", "LAB", "COMP"))
+
+df<- df %>%
+  pivot_wider(names_from = "transaction",
+              values_from = "value")
+
+# Any industry where VA = CAP but LAB is missing, set LAB = 0
+  # This is generally only the case for sector Q, where VA=CAP=0.
+df<- df %>%
+  mutate(LAB = replace(LAB,
+                       VA==CAP & is.na(LAB),
+                       0))
+
+# Any industry where VA = LAB but CAP is missing, set CAP = 0
+  # This is only for sector P for US (NAICS data) and Korea (and Japan where P VA = 0)
+df<- df %>%
+  mutate(CAP = replace(CAP,
+                       VA==LAB & is.na(CAP),
+                       0))
+
+# Drop any country-year-industry obs that do not have data for all 4 transactions
+  # To-do check if COMP is needed, and calculate for US_alt and Japan if so)
+
+df<- df %>% filter(!(is.na(VA) & is.na(CAP) & is.na(LAB)) )
+
+```
+
+Drop USA-NAICS data (in favour of of SIC data)
+```{r}
+df<- df %>%
+  filter(country!="USA-NAICS")
+```
+
+Standardize country names
+```{r}
+
+# Mapping file
+fp<- "PhD_pages/data/original/KLEMS/KLEMS_country_name_mapping.csv"
+
+KLEMS_country_names<- read.csv(fp)
+
+KLEMS_country_names<- KLEMS_country_names %>%
+  rename(country = Country)
+
+# aggregate shares
+df<- merge(df, KLEMS_country_names)
+
+df <- df %>%
+  select(-country) %>%
+  rename(country = Our.World.In.Data.Name)
+
+```
+
+## Complete dataset for aggregate industries
+
+```{r}
+
+# This list gives the codes of the main industries
+main_industry_codes<- c("AtB",
+                        "C",
+                        "D",
+                        "E",
+                        "F",
+                        "G",
+                        "H",
+                        "I",
+                        "J",
+                        "K",
+                        "L",
+                        "M",
+                        "N",
+                        "O",
+                        "P",
+                        "Q")
+
+# filter for these codes sub-industries
+industry_level_values<- df %>%
+  filter(code %in% main_industry_codes)
+
+# Store reported totals separately
+
+reported_total_values<- df %>%
+  filter(code == "TOT") %>%
+  select(-code) %>%
+  gather(transaction, reported_total, -c(country, year)) %>%
+  arrange(country, year, transaction)
+
+```

### 2022-10-31 PhD_pages/data_appendices/3_B_prepare_OECD.qmd
prep OECD data

---
title: "Appendix 3.B: Preparing OECD data"
format: html
---

```{r}
library(tidyverse)
library(rsdmx)
```

## Queries of the OECD API

```{r}
start_year<- 1950
end_year<- 2020

```

### Industry-level data on value added and its components

```{r}

select_table<- "SNA_TABLE6A_ARCHIVE" # Value added and its components by activity, ISIC rev4

countries<- "" # leave blank for all entities
transactions<- "" # i.e. output/GOS/CoE etc. leave blank for all
activities<- "" # ISIC rev 4 industries. leave blank for all industries and sub-industries
measures<- "C+VOB" # Prices. C = constant; VOB = indexed to OECD base year

key_query<- paste(countries,transactions,activities,measures, sep = ".")

industry_accounts <- readSDMX(providerId = "OECD",
                              resource = "data",
                              flowRef = select_table,
                              key = key_query,
                              start = start_year,
                              end = end_year,
                              dsd = TRUE)

industry_accounts<- as.data.frame(industry_accounts)

# Save
save(industry_accounts, file="PhD_pages/data/original/OECD/industry_accounts.Rda")

```

```{r}

# GDP AT PPPs ----

GDP_PPP<- readSDMX(providerId = "OECD",
                       resource = "data",
                      flowRef = "SNA_TABLE1",
                      key = ".B1_GE.VPVOB",
                       start = 1950,
                      end = 2020,
                      dsd = TRUE)

GDP_PPP<- as.data.frame(GDP_PPP)

# Save
save(GDP_PPP, file="PhD_pages/data/original/OECD/GDP_PPP_raw.Rda")

# HOUSING CONSUMPTION ------
rental_consumption <- readSDMX(providerId = "OECD",
                            resource = "data",
                            flowRef = "SNA_TABLE5",
                            key = ".P31CP041+P31CP042.C+VOB",
                            start = 1950,
                            end = 2020,
                            dsd = TRUE)

rental_consumption<- as.data.frame(rental_consumption)

# Save
save(rental_consumption, file="PhD_pages/data/original/OECD/rental_consumption.Rda")

# RENT PRICE -----
# This is the CPI component for actual rentals or else the aggrgeate
  # of actual and imputed rentals and maintenance. See paper for details.

rent_price_index <- readSDMX(providerId = "OECD",
                               resource = "data",
                               flowRef = "HOUSE_PRICES",
                               key = ".RPI",
                               start = 1950,
                               end = 2020,
                               dsd = TRUE)

rent_price_index<- as.data.frame(rent_price_index)

# Save
save(rent_price_index, file="PhD_pages/data/original/OECD/rent_price_index.Rda")

# CPI ------

CPI <- readSDMX(providerId = "OECD",
                             resource = "data",
                             flowRef = "PRICES_CPI",
                             key = ".CPALTT01.IXOB.A",
                             start = 1950,
                             end = 2020,
                             dsd = TRUE)

CPI<- as.data.frame(CPI)

  # the single subject-measure CPALTT01 - IXOB refers to all items
    # CPI index. 2015=100

# Save
save(CPI, file="PhD_pages/data/original/OECD/CPI.Rda")

# GDP deflator ------

GDP_deflator <- readSDMX(providerId = "OECD",
                resource = "data",
                flowRef = "EO",
                key = ".PGDP.A",
                start = 1950,
                end = 2020,
                dsd = TRUE)

GDP_deflator<- as.data.frame(GDP_deflator)

# the single subject-measure CPALTT01 - IXOB refers to all items
# CPI index. 2015=100

# Save
save(GDP_deflator, file="PhD_pages/data/original/OECD/GDP_deflator.Rda")

# SECTOR ACCOUNTS ------

# Grab full dataset

sector_accounts <- readSDMX(providerId = "OECD",
                        resource = "data",
                        flowRef = "SNA_TABLE14A",
                        key = "..S1+S2+S11+S12+S13+S14_S15+S14+S15.C",
                        start = 1950,
                        end = 2020,
                        dsd = TRUE)

sector_accounts<- as.data.frame(sector_accounts)

# Save
save(sector_accounts, file="PhD_pages/data/original/OECD/sector_accounts.Rda")

# INDUSTRY LABOUR INPUTS DATA ----

# set paramaters for API query

select_table<- "SNA_TABLE7A_ARCHIVE" # Labour intput by activity, ISIC rev4

start_year<- 1950
end_year<- 2020

countries<- "" # leave blank for all entities
transactions<- "" # i.e. total employment/employed/self-employed. leave blank for all
activities<- "" # ISIC rev 4 industries. leave blank for all industries and sub-industries
measures<- "HRS" # i.e. hours/jobs/fte. HRS = hours in millions

key_query<- paste(countries,transactions,activities,measures, sep = ".")

employment_by_industry <- readSDMX(providerId = "OECD",
                              resource = "data",
                              flowRef = select_table,
                              key = key_query,
                              start = start_year,
                              end = end_year,
                              dsd = TRUE)

employment_by_industry<- as.data.frame(employment_by_industry)

# Save
save(employment_by_industry, file="PhD_pages/data/original/OECD/employment_by_industry.Rda")

```

### 2022-10-31 PhD_pages/paper_pages/paper3.qmd
prep OECD data

@@
 ---
+
+## Introduction
+
+A substantial literature has documented a long-run shift in the distribution of national income between wages and returns to capital.
+
+Much of the interest in this trend is due to it lying at the junction of two prominent topics within the social sciences: 1) the decline of labour bargaining power, and 2) the rise in income inequality.
+
+(Explain (1) and connection to (2)).
+
+\
 

### 2022-11-30 PhD_pages/data_appendices/3_A_prepare_KLEMS.qmd
Updating papers and data appendices

--- 2022-10-31 PhD_pages/data_appendices/3_A_prepare_KLEMS.qmd
@@
 
+## Processing steps
+
+The titles in each of the boxes below summarise the steps taken to prepare the KLEMS industry data. In each case you can expand the box to see the code used.
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Load libraries and functions
+
 ```{r}
 library(tidyverse)
+library(readxl) # For handling xls source data files
+library(knitr)
+
+source("PhD_pages/scripts/general_functions.R")
 ```
+:::
 
-## Introduction
+::: {.callout-note collapse="true" appearance="minimal"}
+### Specify the location of four original data files
 
-## Raw data files
+We obtain data from four KLEMS datasets:
 
-Make a list of filenames (both to reference the correct file in the url and as the filename to save locally).
+-   EU KLEMS version ...
+-   Canada...
+-   An additional series for the US (different industry classification, longer series..)
+-   Japan
+
+Here we specify corresponding lists of URLs where the original files are located, short names for the files, and local filepaths where a copy of the original data files will be saved.
+
 ```{r}
@@
 files<- c("32_countries",
-          "Canada",
-          "US_alt")
+          "CAN",
+          "USA")
 
@@
                "https://www.worldklems.net/data/basic/CAN_WK_07_2012.xlsx",
-               "https://www.worldklems.net/data/basic/usa_wk_mar_2017.xlsx"
+               "https://www.worldklems.net/data/basic/usa_wk_apr_2013.xlsx"
 )
+
+# JAPAN OLD: "https://www.worldklems.net/data/basic/JPN_wk_may_2013.xlsx"
 
@@
 
-# Add directory to specify filepaths in which to save a local copy of each file
-filepaths<- paste0("PhD_pages/data/original/KLEMS/", filenames)
+# The directory where the copies will be saved
+fp<- "PhD_pages/data/original/KLEMS/"
+
+# Append this directory to the file names to specify file paths
+filepaths<- paste0(fp, filenames)
 
@@
 ```
+:::
 
-Save files locally.
+::: {.callout-note collapse="true" appearance="minimal"}
+### Save a local copy of the original data files
+
+If a new copy of the original data is required, the `grab_original` variable can be set to `TRUE`. Otherwise this step is skipped.
+
 ```{r}
+grab_original<- FALSE
 
+if (grab_original){
   for (file in files){
-  print(file)
+    print(paste0("Downloading ", file))
     download.file(download_urls[file], filepaths[file])
+  }
 }
@@
 ```
+:::
 
-## Preparing dataframes
+::: {.callout-note collapse="true" appearance="minimal"}
+### Load locally saved copies of original data as data frames
+
+```{r}
+df_list<- list()
+```
 
 Importing the main data file - with data for 32 countries.
+
 ```{r}
 
-df <- read.csv(filepaths["32_countries"], quote="")
+df_list[["32_countries"]]<- read.csv(filepaths["32_countries"], quote="")
 
 # convert to long format and rename variables
-df<- df %>%
+df_list[["32_countries"]]<- df_list[["32_countries"]] %>%
   pivot_longer(cols = starts_with("X_"), names_to = "year",
@@
 
-# Remove quotations (These come from not properly specifying the file structure in the read.csv command above. But I was unable to find the right format spec.)
-df <- df %>% 
+# Remove quotations (Ideally this would be handled by specifying the right format in the read.csv command above. But I was unable to find the right format specification)
+df_list[["32_countries"]]<- df_list[["32_countries"]] %>%
   mutate(country = str_replace(country, '"', ''),
@@
 
+# Recode the Entity name for US: 'USA-NAICS' -> 'USA'
+df_list[["32_countries"]]<- df_list[["32_countries"]] %>%
+  mutate(country = if_else(country == "USA-NAICS",
+                           "USA",
+                           country))
+
 ```
 
-## Calculatiing derived variables
+Data for Canada and the alternative series for the US.
+
+```{r}
+countries<- c("CAN", "USA")
+
+for (country in countries){
+df_list[[country]]<- as.data.frame(read_excel(filepaths[country],
+                                                         sheet = 'DATA'))
+
+# convert to long format and rename variables
+df_list[[country]]<- df_list[[country]] %>%
+  pivot_longer(cols = starts_with("_"), names_to = "year",
+    names_prefix = "_", values_to ="value",
+    names_transform = list(year=as.integer)) %>%
+  rename(transaction = Variable) %>%
+  mutate(country = country)
+
+}
+
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Tidying Canada data
+
+```{r}
+# Drop some NA values (- data for 2009 and 2010 is given as missing)
+df_list[["CAN"]]<- df_list[["CAN"]] %>%
+  filter(!is.na(value))
+
+# For Canada we need to calculate some industry aggregates to be comparable with other countries
+agg_industries<- list()
+
+agg_industries[['D']]<- c(
+  "15t16",
+  "17t19",
+  "20",
+  "21t22",
+  "23",
+  "24",
+  "25",
+  "26",
+  "27t28",
+  "29",
+  "30t33",
+  "34t35",
+  "36t37")
+
+agg_industries[['G']]<- c(
+  "50",
+  "52")
+
+agg_industries[['I']]<- c(
+  "60t63",
+  "64")
+
+agg_industries[['K']]<- c(
+  "70",
+  "71t74")
+
+for (agg_ind in names(agg_industries)){
+  print(paste0("Adding aggregate data for industry: ", agg_ind))
+
+  df_totals<- df_list[["CAN"]] %>%
+  filter(code %in% agg_industries[[agg_ind]]) %>%
+    group_by(transaction, year) %>%
+    summarise(value = sum(value)) %>%
+    ungroup() %>%
+    mutate(code = agg_ind) %>%
+    mutate(country = "CAN")
+
+  df_list[["CAN"]]<- bind_rows(df_list[["CAN"]], df_totals)
+
+}
+
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Tidying Japan data - OLD, TO DELETE
+
+```{r}
+# Drop NA values
+# df_list[["JPN"]]<- df_list[["JPN"]] %>% filter(is.na(value))
+# 
+# # Rename the capital compensation variable, to agree with other files
+# df_list[["JPN"]]<- df_list[["JPN"]] %>%
+#   mutate(transaction = replace(transaction,
+#                                transaction == "CAP_GFCF",
+#                                "CAP"))
+
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Join the different dataframes together into a single dataframe
+
+```{r}
+df<- bind_rows(
+  df_list, .id = "source"
+)
+
+# Make a copy of the original merged dataframe, for use later on.
+df_orig<- df
+```
+
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Drop duplicate series for the US
+
+Data for the US is included in the main data file using the NAICS industry classification.
+
+A separate source provides a much longer series using SIC classification.
+
+Here we show the data for aggregate Value Added from the two datasets.
+```{r}
+
+df %>%
+  filter(country %in% c("USA"),
+         code == "TOT",
+ transaction == "VA") %>%
+  ggplot(aes(x = year, y = value, shape = source, colour = source, alpha=0.5)) +
+  geom_point()
+
+```
+
+We drop the shorter NAICS data in favour of the SIC data.
+```{r}
+df<- df %>%
+  filter(!(country == "USA" & source == "32_countries"))
+
+```
+
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Standardize country names
+
+Here is the mapping of country names we use to standardize across data sources.
+```{r}
+
+KLEMS_country_mapping<- read.csv(paste0(fp, "KLEMS_country_name_mapping", ".csv"))
+
+kable(KLEMS_country_mapping)
+
+```
+
+Here we map the joined KLEMS data sets to these standardized names.
+```{r}
+df<- map_names(
+  df = df, 
+  df_mapping = KLEMS_country_mapping, 
+  old_country_varname_df = "country", 
+  old_country_varname_df_mapping = "Country_KLEMS",
+  new_country_varname_df_mapping = "Our_World_In_Data_country_name",
+  new_country_varname_final = "country"
+)
+
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Select relevant transactions and convert to wide format
+
+We keep only the transactions we need: Value Added, Capital income, Labour income and Compensation of employees. And we spread the data to a wider format.
+
+```{r}
+df<- df %>%
+  filter(transaction %in% c("VA", "CAP", "LAB", "COMP"))
+
+df<- df %>%
+  pivot_wider(id_cols = c(country, year, code),
+              names_from = "transaction",
+              values_from = "value")
+
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Replace some NA values with 0
+
+There are many (`r df %>% filter(VA == CAP & is.na(LAB)) %>% count() %>% as.numeric()`) country-year-industry observations where where CAP is 100% of VA, but LAB is missing.
+
+Most of these (`r df %>% filter(VA == CAP & is.na(LAB) & VA == 0) %>% count() %>% as.numeric()`) are where both capital income and value added equal zero.
+Such observations appear only for the following industries:
+
+```{r}
+kable(
+  df %>% filter(VA == CAP & is.na(LAB) & VA == 0) %>% count(code) 
+)
+```
+
+There are also a small number of such observations where VA>0, listed here:
+
+```{r}
+kable(
+  df %>% filter(VA == CAP & is.na(LAB) & VA>0)
+)
+
+```
+
+For both cases (i.e. where LAB is NA and either VA = CAP = 0 or VA = CAP > 0) we set LAB = 0.
+
+```{r}
+df<- df %>%
+  mutate(LAB = replace(LAB,
+                       VA==CAP & is.na(LAB),
+                       0))
+```
+
+There is a similar set of country-year-industry observations (n = `r df %>% filter(VA == LAB & is.na(CAP)) %>% count() %>% as.numeric()`) where VA = LAB but CAP is NA. These we list here.
+
+```{r}
+kable(
+  df %>% filter(VA == LAB & is.na(CAP) & VA>0)
+)
+
+```
+
+For these cases (i.e. where CAP is NA and VA = LAB) we set CAP = 0.
+
+```{r}
+df<- df %>%
+  mutate(CAP = replace(CAP,
+                       VA==LAB & is.na(CAP),
+                       0))
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Filter and store data for main industries and the aggregate economy separately
+
+```{r}
+
+# This list gives the codes of the main industries
+main_industry_codes<- c("AtB",
+                        "C",
+                        "D",
+                        "E",
+                        "F",
+                        "G",
+                        "H",
+                        "I",
+                        "J",
+                        "K",
+                        "L",
+                        "M",
+                        "N",
+                        "O",
+                        "P",
+                        "Q")
+
+# filter for these codes
+main_industry_accounts<- df %>%
+  filter(code %in% main_industry_codes)
+
+# Store reported totals separately
+
+reported_totals<- df %>%
+  filter(code == "TOT") %>%
+  select(-code)
+
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Drop country-year-industry observations that do not have all variables needed
+
+There are some observations (n=`r main_industry_accounts %>% filter(is.na(VA) & is.na(CAP) & is.na(LAB)) %>% count() %>% as.numeric()`) without .... . We group these by country and industry.
+
+```{r}
+kable(
+  main_industry_accounts %>% filter(is.na(VA) | is.na(CAP) | is.na(LAB)) %>% count(country, code)
+)
+
+```
+
+We filter these out.
+
+```{r}
+
+  # TODO check if COMP is needed, and add in check here and/or calculate it where missing? if so)
+
+main_industry_accounts<- main_industry_accounts %>% filter(!(is.na(VA) & is.na(CAP) & is.na(LAB)) )
+
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Add industry code descriptions
+
+Here are the industry descriptions provided in the data.
+```{r}
+
+industry_descriptions<- df_orig %>%
+  filter(code %in% main_industry_codes) %>%
+  select(code, desc) %>%
+  drop_na() %>%
+  unique() %>%
+  arrange(code)
+
+kable(industry_descriptions)
+
+```
+
+We join these back into the clean dataframe.
+```{r}
+
+main_industry_accounts<- left_join(main_industry_accounts, industry_descriptions)
+
+```
+:::
+
+## Sense checking
+
+### How does the sum across all industries compare to the reported economy-wide totals?
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Calculate aggregate across industries and merge with reported economy-wide totals
+
+```{r}
+agg_industries<- main_industry_accounts %>%
+    group_by(country, year) %>%
+    summarise(VA = sum(VA, na.rm = TRUE),
+              CAP = sum(CAP, na.rm = TRUE),
+              COMP = sum(COMP, na.rm = TRUE),
+              LAB = sum(LAB, na.rm = TRUE)
+              )
+
+reported_totals <- reported_totals %>%
+  rename(VA_reported = VA,
+         CAP_reported = CAP,
+         COMP_reported = COMP,
+         LAB_reported = LAB)
+
+check_totals<- left_join(agg_industries, reported_totals) %>%
+  mutate(VA_ratio = VA/VA_reported,
+         CAP_ratio = CAP/CAP_reported,
+         COMP_ratio = COMP/COMP_reported,
+         LAB_ratio = LAB/LAB_reported)
+
+tolerance<- 0.001 # tenth of a percent tolerance
+
+```
+:::
+
+In general the sum across industries is very close to to the reported economy-wide aggregate.
+In the table below we show all observations where the ratio deviates from 100% by more than `r tolerance*100`%.
+
+Note that in Canada - which is one of the countries we will go on to use this data for - the sum across available industries falls short of the reported aggregate (for value added, Capital income, Labour income and Compensation of employees) by around 5%, with no trend over time.
+
+::: {.panel-tabset}
+## Value added
+```{r}
+
+kable(
+  check_totals %>%
+  filter(VA_ratio<(1-tolerance) | VA_ratio>(1+tolerance)) %>%
+    select(country, year, VA_ratio, VA, VA_reported)
+)
+```
+## Capital income
+```{r}
+kable(
+  check_totals %>%
+  filter(CAP_ratio<(1-tolerance) | CAP_ratio>(1+tolerance)) %>%
+    select(country, year, CAP_ratio, CAP, CAP_reported)
+)
+
+```
+
+## Labour income
+
+```{r}
+kable(
+  check_totals %>%
+  filter(LAB_ratio<(1-tolerance) | LAB_ratio>(1+tolerance)) %>%
+    select(country, year, LAB_ratio, LAB, LAB_reported)
+)
+
+```
+
+## Comp. of employees
+```{r}
+kable(
+  check_totals %>%
+  filter(COMP_ratio<(1-tolerance) | COMP_ratio>(1+tolerance)) %>%
+    select(country, year, COMP_ratio, COMP, COMP_reported)
+)
+
+```
+
+:::
+
+## Save cleaned data
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Write cleaned data to csv
+```{r}
+fp<- "PhD_pages/data/clean/"
+
+write.csv(main_industry_accounts, paste0(fp, "KLEMS_gen_income_by_industry", ".csv"))
+
+```
+
+:::

### 2022-11-30 PhD_pages/data_appendices/3_B_Query_OECD_API.qmd
Updating papers and data appendices

---
title: "Appendix 3.B: Preparing OECD data"
format: html
---

```{r}
library(tidyverse)
library(rsdmx)
library(knitr)
```

## Queries of the OECD API

Here is where we will save the outputs of the API without any processing.
```{r}

fp<- "PhD_pages/data/original/OECD/"
```

Set a year range for the queries below.
```{r}
query_API<- FALSE

start_year<- 1950
end_year<- 2020

```

```{r}
#| code-fold: true
#| code-summary: "Industry-level data on value added and its components"

if (query_API){
select_table<- "SNA_TABLE6A_ARCHIVE" # Value added and its components by activity, ISIC rev4

countries<- "" # leave blank for all entities
transactions<- "" # i.e. output/GOS/CoE etc. leave blank for all
activities<- "" # ISIC rev 4 industries. leave blank for all industries and sub-industries
measures<- "C+VOB" # Prices. C = constant; VOB = indexed to OECD base year

key_query<- paste(countries,transactions,activities,measures, sep = ".")

industry_accounts <- readSDMX(providerId = "OECD",
                              resource = "data",
                              flowRef = select_table,
                              key = key_query,
                              start = start_year,
                              end = end_year,
                              dsd = TRUE)

industry_accounts<- as.data.frame(industry_accounts)

# Save
save(industry_accounts, file= paste0(fp, "industry_accounts", ".Rda"))

}
```

```{r}
#| code-fold: true
#| code-summary: "Industry-level data on employment"

if (query_API){

# INDUSTRY LABOUR INPUTS DATA ----

# set paramaters for API query

select_table<- "SNA_TABLE7A_ARCHIVE" # Labour intput by activity, ISIC rev4

countries<- "" # leave blank for all entities
transactions<- "" # i.e. total employment/employed/self-employed. leave blank for all
activities<- "" # ISIC rev 4 industries. leave blank for all industries and sub-industries
measures<- "HRS" # i.e. hours/jobs/fte. HRS = hours in millions

key_query<- paste(countries,transactions,activities,measures, sep = ".")

employment_by_industry <- readSDMX(providerId = "OECD",
                              resource = "data",
                              flowRef = select_table,
                              key = key_query,
                              start = start_year,
                              end = end_year,
                              dsd = TRUE)

employment_by_industry<- as.data.frame(employment_by_industry)

# Save
save(employment_by_industry, file=paste0(fp, "employment_by_industry", ".Rda"))
}

```

```{r}
#| code-fold: true
#| code-summary: "Sector-level accounts (Households, Corporations, Govt etc.)"

if (query_API){

# Grab full dataset

sector_accounts <- readSDMX(providerId = "OECD",
                        resource = "data",
                        flowRef = "SNA_TABLE14A",
                        key = "..S1+S2+S11+S12+S13+S14_S15+S14+S15.C",
                        start = start_year,
                        end = end_year,
                        dsd = TRUE)

sector_accounts<- as.data.frame(sector_accounts)

# Save
save(sector_accounts,
     file= paste0(fp, "sector_accounts", ".Rda"))

}
```

```{r}

#| code-fold: true
#| code-summary: "GDP measured in international-$"

if (query_API){
# GDP AT PPPs ----

GDP_PPP<- readSDMX(providerId = "OECD",
                       resource = "data",
                      flowRef = "SNA_TABLE1",
                      key = ".B1_GE.VPVOB",
                       start = start_year,
                      end = end_year,
                      dsd = TRUE)

GDP_PPP<- as.data.frame(GDP_PPP)

# Save
save(GDP_PPP, file= paste0(fp, "GDP_PPP", ".Rda"))

}

```

```{r}
#| code-fold: true
#| code-summary: "Consumption of rent"

if (query_API){
# HOUSING CONSUMPTION ------
rental_consumption <- readSDMX(providerId = "OECD",
                            resource = "data",
                            flowRef = "SNA_TABLE5",
                            key = ".P31CP041+P31CP042.C+VOB",
                            start = start_year,
                            end = end_year,
                            dsd = TRUE)

rental_consumption<- as.data.frame(rental_consumption)

# Save
save(rental_consumption, file= paste0(fp, "rental_consumption", ".Rda"))
}
```

```{r}
#| code-fold: true
#| code-summary: "Rental price level"

if (query_API){
# RENT PRICE -----
# This is the CPI component for actual rentals or else the aggrgeate
  # of actual and imputed rentals and maintenance. See paper for details.

rent_price_index <- readSDMX(providerId = "OECD",
                               resource = "data",
                               flowRef = "HOUSE_PRICES",
                               key = ".RPI",
                               start = start_year,
                               end = end_year,
                               dsd = TRUE)

rent_price_index<- as.data.frame(rent_price_index)

# Save
save(rent_price_index, file= paste0(fp, "rent_price_index", ".Rda"))

}
```

```{r}
#| code-fold: true
#| code-summary: "CPI"

if (query_API){
# CPI ------

CPI <- readSDMX(providerId = "OECD",
                             resource = "data",
                             flowRef = "PRICES_CPI",
                             key = ".CPALTT01.IXOB.A",
                             start = start_year,
                             end = end_year,
                             dsd = TRUE)

CPI<- as.data.frame(CPI)

  # the single subject-measure CPALTT01 - IXOB refers to all items
    # CPI index. 2015=100

# Save
save(CPI, file= paste0(fp, "CPI", ".Rda"))
}
```

```{r}
#| code-fold: true
#| code-summary: "GDP deflator"

if (query_API){

# GDP deflator ------

GDP_deflator <- readSDMX(providerId = "OECD",
                resource = "data",
                flowRef = "EO",
                key = ".PGDP.A",
                start = start_year,
                end = end_year,
                dsd = TRUE)

GDP_deflator<- as.data.frame(GDP_deflator)

# the single subject-measure CPALTT01 - IXOB refers to all items
# CPI index. 2015=100

# Save
save(GDP_deflator, file= paste0(fp, "GDP_deflator", ".Rda"))

}
```

### 2022-11-30 PhD_pages/data_appendices/3_B_prepare_OECD.qmd
Updating papers and data appendices

@@
 ---
-title: "Appendix 3.B: Preparing OECD data"
+title: "Appendix 3.B: Preparing OECD industry data"
 format: html
 ---
+
+## Processing steps
+
+The titles in each of the boxes below summarise the steps taken to prepare the OECD industry data. In each case you can expand the box to see the code used.
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Load libraries and functions
 
@@
 library(rsdmx)
+library(knitr)
+
+source("PhD_pages/scripts/general_functions.R")
 ```
+:::
 
-## Queries of the OECD API
+::: {.callout-note collapse="true" appearance="minimal"}
+### Load industry generation of income and employment data
 
 ```{r}
-start_year<- 1950
-end_year<- 2020
+fp<- "PhD_pages/data/original/OECD/"
+
+load(paste0(fp, "industry_accounts", ".Rda"))
+load(paste0(fp, "employment_by_industry", ".Rda"))
+
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Standardize country names
+
+Here is the mapping of country names we use to standardize across data sources.
+```{r}
+
+OECD_country_mapping<- read.csv(paste0(fp, "OECD_country_name_mapping", ".csv"))
+
+kable(OECD_country_mapping)
 
@@
 
-### Industry-level data on value added and its components
+Here we map the two data sets to these standardized names.
+```{r}
+
+industry_accounts<- map_names(
+  df = industry_accounts, 
+  df_mapping = OECD_country_mapping, 
+  old_country_varname_df = "LOCATION", 
+  old_country_varname_df_mapping = "OECD_country_name",
+  new_country_varname_df_mapping = "Our_World_In_Data_country_name",
+  new_country_varname_final = "country"
+)
+
+employment_by_industry<- map_names(
+  df = employment_by_industry, 
+  df_mapping = OECD_country_mapping, 
+  old_country_varname_df = "LOCATION", 
+  old_country_varname_df_mapping = "OECD_country_name",
+  new_country_varname_df_mapping = "Our_World_In_Data_country_name",
+  new_country_varname_final = "country"
+)
+
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Clean year variable in both files
+
+```{r}
+clean_year<- function(df){
+  df<- df %>%
+    mutate(year = as.numeric(obsTime))
+}
+
+industry_accounts<- clean_year(industry_accounts)
+employment_by_industry<- clean_year(employment_by_industry)
+
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Further tidying of industry generation of income data
 
@@
 
-select_table<- "SNA_TABLE6A_ARCHIVE" # Value added and its components by activity, ISIC rev4
+# Filter for only current price data and the needed transactions
+industry_accounts<- industry_accounts %>%
+  filter(MEASURE == "C" & 
+           TRANSACT %in% c("B1GA", "D1A","B2G_B3GA", "B2N_B3NA", "D29_D39A"))
 
-countries<- "" # leave blank for all entities
-transactions<- "" # i.e. output/GOS/CoE etc. leave blank for all
-activities<- "" # ISIC rev 4 industries. leave blank for all industries and sub-industries
-measures<- "C+VOB" # Prices. C = constant; VOB = indexed to OECD base year
+# Recode transactions from national accounts codes to something more legible
+recoding_transactions <- c(
+  B1GA = "VA", # (Gross) Value added
+  D1A = "COMP", # Compensation of employees
+  B2G_B3GA = "GOS", # Gross operating surplus
+  B2N_B3NA = "NOS", # Net operating surplus (net of depreciation)
+  D29_D39A = "TOP") # Taxes on production
 
-key_query<- paste(countries,transactions,activities,measures, sep = ".")
+industry_accounts<- industry_accounts %>%
+  mutate(TRANSACT = recode(TRANSACT, !!!recoding_transactions))
 
-industry_accounts <- readSDMX(providerId = "OECD",
-                              resource = "data",
-                              flowRef = select_table,
-                              key = key_query,
-                              start = start_year,
-                              end = end_year,
-                              dsd = TRUE)
+# flip to wide format
+industry_accounts<- industry_accounts %>%
+  pivot_wider(names_from = TRANSACT, values_from = obsValue)
 
-industry_accounts<- as.data.frame(industry_accounts)
+# Select and rename needed vars
+industry_accounts<- industry_accounts %>%
+  rename(ind_code = ACTIVITY) %>%
+  select(country, year, ind_code, VA, COMP, GOS, NOS, TOP)
 
-# Save
-save(industry_accounts, file="PhD_pages/data/original/OECD/industry_accounts.Rda")
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Filter and store generation of income data for main industries and the aggregate economy separately
+
+```{r}
+
+# Main aggregate industries
+
+main_industry_codes<- c("VA0", "VB", "VC", "VD", "VE", "VF", "VG", "VH", "VI",
+                        "VJ", "VK", "VL", "VM", "VN", "VO", "VP", "VQ", "VR",
+                        "VS", "VT", "VU")
+
+# Filter industry data for main industry codes only
+main_industry_accounts<- industry_accounts %>%
+  filter(ind_code %in% main_industry_codes)
+
+# Store economy-wide aggregates separately
+reported_totals <- industry_accounts %>%
+  filter(ind_code == 'VTOT')
+
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Further tidying of industry employment data.
+
+```{r}
+# Select self-employed and employed hours and flip to wide format
+employment_by_industry<- employment_by_industry %>%
+  filter(TRANSACT %in% c("ESEA", "EEMA")) %>%  # codes for self-employed and employed
+  pivot_wider(id_cols = c(country, year, ACTIVITY), names_from = TRANSACT, values_from = obsValue)
+
+# Select and rename needed vars
+employment_by_industry<- employment_by_industry %>%
+  rename(self_employed_hrs = ESEA, 
+         employed_hrs = EEMA,
+         ind_code = ACTIVITY)
+
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Join generation of income and employment data
+
+```{r}
+main_industry_accounts<- left_join(main_industry_accounts, employment_by_industry)
+```
+
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Filter out observations where the implied wage is infinite
+
+In the joined data, there are a small number of rows where employed hours = 0 but compensation of employees \> 0 --- implying an infinite wage.
+
+Here we show these illogical rows, grouped by country and industry.
+
+```{r}
+kable(
+  main_industry_accounts %>%
+  filter((COMP>0 & employed_hrs == 0))
+)
+```
+
+We drop these.
+
+```{r}
+main_industry_accounts<- main_industry_accounts %>%
+  filter(!(COMP>0 & employed_hrs == 0))
+
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Filter out country-year-industry observations without all the variables we need
+
+There are some rows where we do not have all the variables we will need for the calculations: VA, COMP, GOS, and employed and self-employed hours.
+
+```{r}
+kable(
+  main_industry_accounts %>%
+    filter(is.na(VA) | is.na(COMP) | is.na(GOS) | is.na(employed_hrs) | is.na(self_employed_hrs)) %>%
+    count(country, ind_code)
+)
+```
+
+We drop these.
+
+```{r}
+main_industry_accounts<-  main_industry_accounts %>%
+    filter(!(is.na(VA) | is.na(COMP) | is.na(GOS) | is.na(employed_hrs) | is.na(self_employed_hrs)))
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+#### Filter out country-year observations with partial industry coverage
+
+The objective is to have complete coverage of the economy. There are some country years where coverage across industries looks decidedly incomplete.
+
+For instance, out of `r length(main_industry_codes)` aggregate industry codes, these country years have data for fewer than 18:
+
+```{r}
+kable( 
+ to_drop<- main_industry_accounts %>%
+   count(country, year) %>%
+   filter(n<18)
+)
 
@@
 
+We drop these country-years with apparently incomplete data.
+
+```{r}
+main_industry_accounts<- left_join(main_industry_accounts, to_drop) %>%
+  filter(is.na(n)) %>%
+  select(-n)
+
+```
+
+But because of differences in national accounting methods, missing data for a given industry does not imply the accounts are incomplete.
+
+In particular, the dataset contains only `r main_industry_accounts %>% filter(ind_code == "VU")%>%count()` country-year observations for `VU` (as compared to `r main_industry_accounts %>% filter(ind_code == "VA0")%>%count()` for `VA0`) and these are in general only NA or 0 values.
+
+```{r}
+summary(main_industry_accounts %>% filter(ind_code == "VU"))
+```
+
+In the table below we show all remaining country years in which at least one industry --- other than `VU` --- is missing.
+
+We assume that these apparent gaps do not indicate incompleteness of the data, rather just differences in national accounting practices.
+
 ```{r}
 
-# GDP AT PPPs ----
-
-GDP_PPP<- readSDMX(providerId = "OECD",
-                       resource = "data",
-                      flowRef = "SNA_TABLE1",
-                      key = ".B1_GE.VPVOB",
-                       start = 1950,
-                      end = 2020,
-                      dsd = TRUE)
-
-GDP_PPP<- as.data.frame(GDP_PPP)
-
-# Save
-save(GDP_PPP, file="PhD_pages/data/original/OECD/GDP_PPP_raw.Rda")
-
-# HOUSING CONSUMPTION ------
-rental_consumption <- readSDMX(providerId = "OECD",
-                            resource = "data",
-                            flowRef = "SNA_TABLE5",
-                            key = ".P31CP041+P31CP042.C+VOB",
-                            start = 1950,
-                            end = 2020,
-                            dsd = TRUE)
-
-rental_consumption<- as.data.frame(rental_consumption)
-
-# Save
-save(rental_consumption, file="PhD_pages/data/original/OECD/rental_consumption.Rda")
-
-# RENT PRICE -----
-# This is the CPI component for actual rentals or else the aggrgeate
-  # of actual and imputed rentals and maintenance. See paper for details.
-
-rent_price_index <- readSDMX(providerId = "OECD",
-                               resource = "data",
-                               flowRef = "HOUSE_PRICES",
-                               key = ".RPI",
-                               start = 1950,
-                               end = 2020,
-                               dsd = TRUE)
-
-rent_price_index<- as.data.frame(rent_price_index)
-
-# Save
-save(rent_price_index, file="PhD_pages/data/original/OECD/rent_price_index.Rda")
-
-# CPI ------
-
-CPI <- readSDMX(providerId = "OECD",
-                             resource = "data",
-                             flowRef = "PRICES_CPI",
-                             key = ".CPALTT01.IXOB.A",
-                             start = 1950,
-                             end = 2020,
-                             dsd = TRUE)
-
-CPI<- as.data.frame(CPI)
-
-  # the single subject-measure CPALTT01 - IXOB refers to all items
-    # CPI index. 2015=100
-
-# Save
-save(CPI, file="PhD_pages/data/original/OECD/CPI.Rda")
-
-# GDP deflator ------
-
-GDP_deflator <- readSDMX(providerId = "OECD",
-                resource = "data",
-                flowRef = "EO",
-                key = ".PGDP.A",
-                start = 1950,
-                end = 2020,
-                dsd = TRUE)
-
-GDP_deflator<- as.data.frame(GDP_deflator)
-
-# the single subject-measure CPALTT01 - IXOB refers to all items
-# CPI index. 2015=100
-
-# Save
-save(GDP_deflator, file="PhD_pages/data/original/OECD/GDP_deflator.Rda")
-
-# SECTOR ACCOUNTS ------
-
-# Grab full dataset
-
-sector_accounts <- readSDMX(providerId = "OECD",
-                        resource = "data",
-                        flowRef = "SNA_TABLE14A",
-                        key = "..S1+S2+S11+S12+S13+S14_S15+S14+S15.C",
-                        start = 1950,
-                        end = 2020,
-                        dsd = TRUE)
-
-sector_accounts<- as.data.frame(sector_accounts)
-
-# Save
-save(sector_accounts, file="PhD_pages/data/original/OECD/sector_accounts.Rda")
-
-# INDUSTRY LABOUR INPUTS DATA ----
-
-# set paramaters for API query
-
-select_table<- "SNA_TABLE7A_ARCHIVE" # Labour intput by activity, ISIC rev4
-
-start_year<- 1950
-end_year<- 2020
-
-countries<- "" # leave blank for all entities
-transactions<- "" # i.e. total employment/employed/self-employed. leave blank for all
-activities<- "" # ISIC rev 4 industries. leave blank for all industries and sub-industries
-measures<- "HRS" # i.e. hours/jobs/fte. HRS = hours in millions
-
-key_query<- paste(countries,transactions,activities,measures, sep = ".")
-
-employment_by_industry <- readSDMX(providerId = "OECD",
-                              resource = "data",
-                              flowRef = select_table,
-                              key = key_query,
-                              start = start_year,
-                              end = end_year,
-                              dsd = TRUE)
-
-employment_by_industry<- as.data.frame(employment_by_industry)
-
-# Save
-save(employment_by_industry, file="PhD_pages/data/original/OECD/employment_by_industry.Rda")
+kable( main_industry_accounts %>%
+  select(year, country, ind_code, VA) %>%
+  pivot_wider(names_from = ind_code, values_from = VA) %>%
+  filter(if_any(VA0:VT, ~ is.na(.)))
+)
 
 ```
+:::
+
+## Sense checking
+
+### How does the sum across all industries compare to the reported economy-wide totals?
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Calculate aggregate across industries and merge with reported economy-wide totals
+
+```{r}
+agg_industries<- main_industry_accounts %>%
+    group_by(country, year) %>%
+    summarise(VA = sum(VA, na.rm = TRUE),
+              COMP = sum(COMP, na.rm = TRUE),
+              GOS = sum(GOS, na.rm = TRUE),
+              NOS = sum(NOS, na.rm = TRUE)
+              )
+
+reported_totals <- reported_totals %>%
+  rename(VA_reported = VA,
+         COMP_reported = COMP,
+         GOS_reported = GOS,
+         NOS_reported = NOS) %>%
+  select(-ind_code)
+
+check_totals<- left_join(agg_industries, reported_totals) %>%
+  mutate(VA_ratio = VA/VA_reported,
+         COMP_ratio = COMP/COMP_reported,
+         GOS_ratio = GOS/GOS_reported,
+         NOS_ratio = NOS/NOS_reported)
+
+tolerance<- 0.005 # half a percent tolerance
+
+```
+:::
+
+In general the sum across industries is very close to to the reported economy-wide aggregate.
+In the table below we show all observations where the ratio deviates from 100% by more than `r tolerance*100`%.
+
+Note that in Finland - which is one of the countries we will go on to use this data for - the sum across available industries falls short of the reported aggregate (for value added, compensation of employees and gross operating surplus) by around 5%, with no trend over time.
+
+::: {.panel-tabset}
+## Value added
+```{r}
+
+kable(
+  check_totals %>%
+  filter(VA_ratio<(1-tolerance) | VA_ratio>(1+tolerance)) %>%
+    select(country, year, VA_ratio, VA, VA_reported)
+)
+```
+## Comp. of employees
+```{r}
+kable(
+  check_totals %>%
+  filter(COMP_ratio<(1-tolerance) | COMP_ratio>(1+tolerance)) %>%
+    select(country, year, COMP_ratio, COMP, COMP_reported)
+)
+
+```
+## Gross operating surplus
+```{r}
+kable(
+  check_totals %>%
+  filter(GOS_ratio<(1-tolerance) | GOS_ratio>(1+tolerance)) %>%
+    select(country, year, GOS_ratio, GOS, GOS_reported)
+)
+
+```
+
+:::
+
+## Save cleaned data
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Write cleaned data to csv
+```{r}
+fp<- "PhD_pages/data/clean/"
+
+write.csv(main_industry_accounts, paste0(fp, "OECD_gen_income_and_employment_by_industry", ".csv"))
+
+```
+:::

### 2022-11-30 PhD_pages/data_appendices/3_C_calc_factor_shares.qmd
Updating papers and data appendices

---
title: "Calculating housing and factor shares from industry data"
editor: visual
---

## Summary

We take the prepared KLEMS and OECD generation of income by industry data and calculate the labour and capital share of income, separating the housing share from the latter.

(Possibly public sector too?).

The key elements are this:

-   **We use factor income as the denominator\
    **(Explain... Jump to section)
-   **We impute capital and labour components for the self-employed**\
    (Explain... Jump to section))
-   **We prepare two alternative specifications relating to the above imputation, to handle where labour income is greater than 100% of value added**\
    (Explain... Jump to section)) \

## Factor income as the denominator

National accounting practices separates compensation of employees from value added (as well as taxes on production). Gross operating surplus is the residual.

## Impute labour and capital components for self employed

For the purposes of calculating factor shares, this residual includes mixed income - the income of the self-employed which is generated by both their labour and capital inputs.

Within the KLEMS data, this is already separated out - using employment data... (explain).

We do the same process for OECD data.

## Problem of \>100% labour income

## Two approaches

## Calculate factor shares (inc. housing)

TO DO: This was split out from the OECD data prep. I want to process the KLEMS data at the same time.

### Factor shares from industry accounts

#### Imputing self-employed labour and capital income

We have data on the number of employed and self-employed hours are worked in each industry. We also have data on the compensation of the *employed*. The task is to impute the *labour* compensation of the self-employed and then separate this out from *Gross Operating Surplus* in order to arrive at a split between capital and labour income.

The basic strategy is to apply the same hourly rate of compensation for the employed (within each industry) to the self-employed.

```{r}
# main_industry_accounts_inspect<- main_industry_accounts %>%
#   mutate(COMP_per_hour = COMP/employed_hrs,
#          COMP_SE = COMP_per_hour * self_employed_hrs,
#          LAB = COMP + COMP_SE)
```

This however creates a problem: in some cases, the total labour income implied (employed + self employed) is greater than 100% of value added. Here we show these rows grouped by industry.

```{r}
# kable(
#  br<-  main_industry_accounts_inspect %>%
#   mutate(LAB_share = LAB/VA,
#          LAB_share_group = cut(LAB_share,
#                        breaks=c(-100, 1, 1.1, 1.5, 2, 10, 1000000000000),
#                        labels=c('<100%', '100-110%', '110-150%', '150-200%', '200-1,000%', '1,000+'))) %>%
#     count(ind_code, LAB_share_group) %>%
#     pivot_wider(names_from = LAB_share_group, values_from = n)
# )

```

We take two approaches to this problem.

##### Model 1 - imputation with no ceiling

The most simple approach is just to ignore the problem and allow labour income to be greater than value added. This will mean allowing capital income to be negative.

```{r}
# main_industry_accounts_mod1<- main_industry_accounts %>%
#    mutate(COMP_per_hour = COMP/employed_hrs,
#          COMP_SE = COMP_per_hour * self_employed_hrs)
```

##### Model 2 - imputation with ceiling

An alternative approach is to cap self-employed labour income at the full value of *Gross Operating Surplus*. This implies that total labour incomes cannot exceed 100% of value added and total capital incomes cannot be negative.

```{r}
# main_industry_accounts_mod2<- main_industry_accounts %>%
#   mutate(COMP_per_hour = COMP/employed_hrs,
#          COMP_SE = pmin(COMP_per_hour * self_employed_hrs, GOS))
```

##### Calculate labour and capital incomes

On each model, total labour income is equal to the sum of compensation of employees and the self-employed. Total gross (net) capital income is equal to *Gross (Net) Operating Surplus* minus the labour income of the self-employed.

```{r}
# calc_LAB_CAP<- function(df){
#   df<- df %>%
#     mutate(LAB = COMP + COMP_SE,
#            CAP = GOS - COMP_SE,
#            NCAP = NOS - COMP_SE)
# }
# 
# main_industry_accounts_mod1<- calc_LAB_CAP(main_industry_accounts_mod1)
# 
# main_industry_accounts_mod2<- calc_LAB_CAP(main_industry_accounts_mod2)

```

#### Create economy aggregates

For the calculation of factor shares, we will use the sum across industries as the denominator - rather than the reported economy-wide aggregate.

the reason for this. The implicit assumption is that whatever is 'missing' from the aggregate has a factor split equal to that found in the available industry-level data.

Here we calculate the sum across industries.

```{r}

# agg_across_industries<- function(df){
#   
#   df<- df %>%
#     group_by(country, year) %>%
#     summarise(VA = sum(VA, na.rm = TRUE),
#               COMP = sum(COMP, na.rm = TRUE),
#               GOS = sum(GOS, na.rm = TRUE),
#               COMP_SE = sum(COMP_SE, na.rm = TRUE),
#               LAB = sum(LAB, na.rm = TRUE),
#               CAP = sum(CAP, na.rm = TRUE),
#               NCAP = sum(NCAP, na.rm = TRUE)
#               )
#  
# }
# 
# agg_industry_mod1<- agg_across_industries(main_industry_accounts_mod1)
# agg_industry_mod2<- agg_across_industries(main_industry_accounts_mod2)

```

### 2022-11-30 PhD_pages/paper_pages/paper3.qmd
Updating papers and data appendices

--- 2022-10-12 PhD_pages/paper_pages/paper3.qmd
@@
 
+## Introduction
+
+A substantial literature has documented a long-run shift in the distribution of national income between wages and returns to capital.
+
+Much of the interest in this empirical trend owes to its lying at the junction of two prominent topics within the social sciences: 1) the decline of labour bargaining power, and 2) the rise in income inequality.
+
+(Explain (1) and connection to (2)).
+
+The hypothesis underpins much of the motivation for the Top incomes approach.
+
+## Data
+
+###  Background on the structure of national accounts
+
+### Approaches to calculating factor shares
+
+### Available sources
+
 ## The housing share of income
 
-What share of aggregate income is made up by rents on housing assets and how has this evolved in recent decades?
-The national accounts offers three avenues for answering this question: operating surplus within the household sector ('Household sector' approach); operating surplus within the real estate industry ('Real estate industry' approach); and household final consumption of housing services ('Final consumption' approach).
+The unusual nature of housing ... in the national accounts.
 
-Figure 1 shows the housing capital share of factor cost value added calculated according to three approaches.
+As a *stock*, housing is a non-financial asset included in the balance sheet.
+Housing also features as a *flow*, included in each of the three approaches to calculating GDP: output, income and expenditure.
+
+As output, housing services are included in Real Estate activity (SIC....). Sub Sector.
+
+This flow of housing services generates a stream of income, which, in addition to small amounts of labour income (property management, agents' fees), is the yield received by owners of housing assets.[^1] This private or public rented sector , and owner occupied, to which an equivalent rent is imputed.
+includes rent imputed to owner occupiers, as well as actual rents received by private landlords, whether individuals or corporations (both privately and publicly owned).
+
+[^1]: ADD NOTE ABOUT LABOUR INCOME IN ACTUAL RENTS
+
+This income as an expenditure by households, again both actual and imputed rents 'paid'.
+
+In line with national accounting principles, the value of the flow of housing is equal, whether conceived of as a stream of services, income, or expenditure.
+As such, these offer different entry points from which aggregate housing income can be calculated and isolated from non-housing capital income.
+
+Figure 1 shows these housing flows as a share of total income.
+
+In each case the denoninator is gross domestic value added, valued at factor cost.
+This aggregate is similar to GDP, with the exception that taxes and subsidies on production are excluded.
+This results in an aggregate that capital and labour components: *gross operating surplus* and *compensation of employees.*
 

### 2022-11-30 PhD_pages/scripts/general_functions.R

@@ -1,0 +1,26 @@
+library(tidyverse)
+
+
+# Country name standardization
+map_names <- function(df, 
+                      df_mapping, 
+                      old_country_varname_df, 
+                      old_country_varname_df_mapping,
+                      new_country_varname_df_mapping,
+                      new_country_varname_final) {  
+  
+  df<- left_join(df, df_mapping, 
+                 by=setNames(old_country_varname_df_mapping,  old_country_varname_df))
+  
+
+  df<- df %>%
+    select(-c(all_of(old_country_varname_df)))
+  
+  names(df)[names(df) == new_country_varname_df_mapping] <- new_country_varname_final
+  
+  return(df)
+}
+
+
+
+

### 2023-01-08 PhD_pages/paper_pages/paper1.qmd
add test shinylive app that pulls in csv

@@
 
-```{r, echo=FALSE, message = FALSE, warning=FALSE}
+::: {.callout-note collapse="true" appearance="minimal"}
+### Load packages
 
+```{r}
 library(tidyverse)
-library(trackdown)
+library(knitr)
+library(data.table)
 
 ```
+:::
 
@@
 
-Different headline - WID suggests general rise (find a quote). PIP much more mixed... even a fall.
+Different headlines - WID suggests general rise (find a quote). PIP much more mixed... even a fall.
 
@@
 
-A mixed picture of inequality around the world suggests reflect specific circumstances, with accordingly much scope for national policy to influence.
+A mixed picture of inequality around the world suggests reflect specific circumstances, with accordingly much more scope for national policy to influence.
 
@@
 
+::: {.callout-note collapse="true" appearance="minimal"}
+### Load PIP data
+
+```{r}
+# df_pip<- read.csv("https://raw.githubusercontent.com/owid/poverty-data/main/datasets/pip_dataset.csv")
+
+df_pip<- read.csv("PhD_pages/data/pip_dataset (2).csv")
+
+df_pip<- df_pip %>%
+  filter(ppp_version == 2017) %>%
+  rename("Gini" = "gini",
+         "Top_10_share" = "decile10_share",
+         "Entity" = "country",
+         "Year" = "year")
+
+# drop subnational data where national data is available
+df_pip<- df_pip %>%
+  filter((!Entity %in% c("China", "India", "Indonesia")) | reporting_level == "national")
+
+# drop income if both income and consumption are available
+df_pip<- df_pip %>%
+  group_by(Entity, reporting_level, Year) %>%
+  mutate(n = n()) %>%
+  filter(n<2 | welfare_type == "consumption") %>%
+  ungroup()
+
+ #drop regional aggregates
+drop_aggs<- c(
+  "East Asia and Pacific",
+  "Europe and Central Asia",
+  "High income countries",
+  "Latin America and the Caribbean",
+  "Middle East and North Africa",
+  "Sub-Saharan Africa",
+  "World"
+)
+
+df_pip<- df_pip %>%
+  filter(!Entity %in% drop_aggs)
+
+```
+:::
+
 ### WID
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Load WID data
+
+```{r}
+# df_wid<- read.csv("https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/PabloArriagada/WID/data/final/wid_pretax.csv")
+
+df_wid<- read.csv("PhD_pages/data/wid.csv")
+
+df_wid<- df_wid %>%
+  rename("Gini" = "Gini.coefficient",
+         "Top_10_share" = "P90.P100...share.of.the.top.10.")
+
+drop_entities<- c(
+"China - rural",
+"China - urban",
+"East Germany",
+"Other Russia and Central Asia",
+"Other East Asia",
+"Other Western Europe",
+"Other Latin America",
+"Other MENA",
+"Other South & South-East Asia",
+"Other Sub-Saharan Africa",
+"Africa",
+"Asia",
+"Europe",
+"Oceania",
+"Central Asia",
+"East Africa",
+"East Asia",
+"Eastern Europe",
+"Middle Africa",
+"North Africa",
+"North America",
+"South-East Asia",
+"South Africa region",
+"West Africa",
+"West Asia",
+"Western Europe",
+"European Union",
+"World",
+"Asia (excluding Middle East)",
+"North America and Oceania",
+"Sub-Saharan Africa",
+"Latin America",
+"Middle East",
+"MENA",
+"Russia and Central Asia",
+"South & South-East Asia"
+)
+
+df_wid<- df_wid %>%
+  filter(!Entity %in% drop_entities)
+```
+
+:::
 
@@
 
-s
+Give an overview of coverage in each dataset.
 
-## Methods: Comparing databases
+::: {.callout-note collapse="true" appearance="minimal"}
+### Standardize coutry names across datasets
+```{r}
 
-### Lining up years
+pip_countries<- df_pip %>%
+  select(Entity) %>%
+  unique() %>%
+  mutate(Entity_pip = Entity)
 
-Method 1: Reference years
+wid_countries<- df_wid %>%
+  select(Entity) %>%
+  unique() %>%
+  mutate(Entity_wid = Entity)
 
-Method 2: Fully extrapolating/interpolating
+countries<- full_join(pip_countries, wid_countries, keep = TRUE) %>%
+  select(Entity_pip, Entity_wid)
 
-The equivalent of the reference year approach where there is no max threshold gap.
+```
 
-Method 3: For the time trend, use a regression with country fixed effects to allow for entry and exit from the sample.
+Countries that are in the PIP data but not the WID data:
+```{r}
+kable(
+  countries %>%
+       filter(is.na(Entity_wid))
+)
+```
+Countries that are in the WID data but not the PIP data:
+(TO-DO: drop regions and sub-national data from the WID data at an eariler step.)
+```{r}
+kable(
+  countries %>%
+       filter(is.na(Entity_pip))
+)
+```
+:::
 
-### Comparing welfare concepts
+## Comparing databases
 
-Getting a *global* view of inequalities within countries faces a Catch-22 situation: If you want broad coverage you get very heterogeneous data; if you want comparable data, or to be able to consider different welfare concepts consistently, the coverage you get is much smaller.
+#### Method 1: Reference years, with a certain tolerance
 
-Here I'll take two avenues.
+::: {.callout-note collapse="true" appearance="minimal"} 
+## A function for selecting data around a set of reference years, within a certain tolerance
 
-Firstly, what we're going for with the 'triangulation' approach is dominance - if different welfare metrics agree
+```{r}
+## This has been superseded by the function below
+# select_ref_years<- function(df, target_varname, group_by_varname, ref_varname, ref_values,tolerance,tie_break_round_up){
+#   
+#   ## Example arguments for testing
+#   # df_in<- df_wid
+#   # target_varname<- "Gini.coefficient"
+#   # group_by_varname<- "Entity"
+#   # ref_varname<- "Year"
+#   # ref_values<- c(1990, 2020)
+#   # tolerance<- 5
+#   # tie_break_round_up<- TRUE
+#   
+#     # rename vars toß make nonstandard evaluation easier
+#    df_in<- df %>%
+#     rename(target_var = !!target_varname,
+#            group_by_var = !!group_by_varname,
+#            ref_var = !!ref_varname)
+#    
+#    out_list<- list()
+#    
+#    for(ref in ref_values){
+#      
+#      df_ref<- df_in %>%
+#        select(target_var, group_by_var, ref_var) %>%
+#        drop_na() %>%
+#        mutate(distance = ref_var - ref) %>%
+#        mutate(abs_distance = abs(distance)) %>%
+#        group_by(group_by_var) %>%
+#        mutate(min_distance = min(abs_distance)) %>%
+#        filter(abs_distance == min_distance) %>%
+#        group_by(group_by_var, min_distance) %>%
+#        mutate(n = n()) %>%
+#        ungroup()
+#      
+#     # address tie-breaks. If 'round_up' is true, the higher value wins out - i.e. the distance to the reference value must positive. And visa versa.
+#      if(tie_break_round_up){
+#        df_ref<- df_ref %>%
+#          filter(n==1 | distance>0)
+#      } else {
+#        df_ref<- df_ref %>%
+#          filter(n==1 | distance<0)
+#      }
+#      
+#      out_list[[as.character(ref)]]<- df_ref
+#      
+#      
+#    }
+#  
+#   
+#   df_out<- bind_rows(out_list, .id = paste0("reference_", ref_varname))
+#   
+#   # If a tolerance has been set, filter out matches beyond tolerance
+#   if(!is.na(tolerance)){
+#       df_out<- df_out %>%
+#          filter(abs_distance <= tolerance)
+#      }
+#   
+#   # names back to original and drop not needed vars
+#    df_out<- df_out %>%
+#     rename(!!target_varname := target_var,
+#            !!group_by_varname := group_by_var,
+#            !!ref_varname := ref_var) %>%
+#      select(-c(distance, min_distance, n))
+#    
+# }
 
-Second we will cross reference to attempts to 'standardise' across these differences: using Solt and GCIP.
-How valid that is I don't know.
-But taking the two approaches together we get as close as we can to robust observations.
+```
 
-### A range of inequality metrics
+```{r}
+# Next, we define a function that takes a dataframe, a pair of years (year1 < year2), a distance, a list of variable names, and a list of ID columns as arguments
+# # calculate_change <- function(data, year1, year2, distance, vars, id_cols) {
 
-### Lorenz curves
+# #   calc_distances<- list()
 
-Dominance implies any (well-behaved) summary measure of inequality will agree on the ranking.
+# #   for(ref in c(year1, year2)){
 
-Operationalize this by taking 99 points on the Lorenze curve: the percentiles.
+# #   # We add a column stating the reference year and group the data by the ID columns
+# #   data_subset <- data %>%
+# #     mutate(ref_year = ref) %>%
+# #     group_by(across(all_of(id_cols)))
 
-This leaves out the *very* top.
-But this would be quite a warped conception of inequality.
-I.e.
-if you have two distributions, A and B, and the share of income received by every whole percentile threshold in B is higher than A, but the Lorenze curves crossed within the top percentile - to claim that inequality is higher in B you'd need a very extreme conception of inequality.
+# #     # We calculate the difference and absolute difference from the reference year
+# #   data_subset$diff <- data_subset$Year - ref
+# #   data_subset$abs_diff <- abs(data_subset$diff)
 
-# Results
+# #   # Store in list
+# #   calc_distances[[ref]]<- data_subset %>% ungroup()
+# #   }
 
-### Outliers
+# #   data_long<- bind_rows(calc_distances)
 
-Russia
+# #   # A list where we will store the matched observations and changes for each variable
+# #   return_matches<- list()
+
+# #   # na_treatment<- "obs_for_all_vars"
+# #   vars_plus<- c(vars, "obs_for_all_vars_case")
+
+# #     #For each var...
+# #     for(var in vars_plus){
+
+# #       # Filter out observations with missing values in var of interest and calculate the min absolute distance from the two reference years
+# #     if(var == "obs_for_all_vars_case"){
+# #       matches <- data_long %>%
+# #         filter_at(vars,all_vars(!is.na(.)))
+# #     } else {
+# #       matches <- data_long %>%
+# #       filter(!is.na(!!sym(var)))
+# #     }
+
+# #     matches <- matches %>%
+# #       group_by(across(all_of(c(id_cols, "ref_year")))) %>% 
+# #       mutate(min_diff = min(abs_diff)) %>%
+# #       filter(abs_diff == min_diff)
+
+# #      # If there is a tie-break (with two matches, one above and one below the reference year), settle the tie-break so as to maximise the gap between the two observations.
+That means for the early year, below matches win; for the later year, above matches win.
+
+# #     matches <- matches %>%
+# #       group_by(across(all_of(c(id_cols, "ref_year", "min_diff")))) %>%
+# #       mutate(n = n()) %>% # count number of matches (n>1 if there is a tie-break) %>%
+# #     ungroup()
+
+# #     matches <- matches %>% 
+# #       group_by(across(all_of(c(id_cols, "ref_year")))) %>% 
+# #       filter(n==1 | (ref_year == year1 & diff<0) | (ref_year == year2 & diff>0))
+
+# #     # If the same year is matched for the two reference years, drop the observations
+# #      matches <- matches %>% 
+# #       group_by(across(all_of(c(id_cols, "Year")))) %>% 
+# #       mutate(n = n()) %>% # count number of times a year is matched
+# #       filter(!n>1) # Drop if same year is matched
+
+# #     if(var == "obs_for_all_vars_case"){
+# #       keep_vars<- c(id_cols, "Year", vars, "ref_year" )
+# #     } else {
+# #       keep_vars<- c(id_cols, "Year", var, "ref_year" )
+# #     }
+
+# #      matches <- matches %>%
+# #        select_at(keep_vars)
+
+# #      # Make a list to store the matched observations, and also the calculated distances
+# #     return_matches[[var]]<- matches
+
+# #     }
+
+# #   # Finally, we return the ...
+# #   return(return_matches)
+# # }
+
+# # # To test the function, we can create a sample dataframe
+# # data <- data.frame(Year = c(1999, 2000, 2001, 2002, 2003, 2004, 2003),
+# #                    Entity = c("A","A", "A", "A", "B", "B", "C"),
+# #                    Var1 = c(40, NA, 20, 30, 40, 50, 60),
+# #                    Var2 = c(30, 100, 200, 300, 400, 500, 400))
+
+# # year1<- 2000
+# # year2<- 2005
+# # distance<- 2
+# # vars<- c("Var1","Var2")
+# # id_cols<- c("Entity")
+
+# # # Then, we can call the function with a pair of years, a distance, a list of variable names, and a list of ID columns
+# # calculate_change(data, 2000, 2004, 2, c("Var1", "Var2"), c("Entity"))
+
+# ```
+
+# ```{r}
+# compare_ref_years<- function(df_list, target_varname, group_by_varname, ref_varname, ref_values,tolerance,tie_break_round_up, rise_fall_threshold) {
+
+#   # Example arguments for testing
+#   df_list<- list("pip" = df_pip, "wid" = df_wid)
+#   target_varname<- "Gini"
+#   group_by_varname<- "Entity"
+#   ref_varname<- "Year"
+#   ref_values<- c(1990, 2020)
+#   tolerance<- 5
+#   tie_break_round_up<- TRUE
+#   rise_fall_threshold<- 0.01
+
+#   # Store the names of the two dataframes as a vector
+#   df_names<- names(df_list)
+
+#   # Initiate a list to store the processed data for each dataframe
+#   df_out_list<- list()
+
+#   # loop through both data frame names
+#   for(i in 1:2){
+#     df_name<- df_names[i]
+
+#     # rename vars to avoid nonstandard evaluation later on
+#     df_in<- df_list[[df_name]] %>%
+#       rename(target_var = !!target_varname,
+#            group_by_var = !!group_by_varname,
+#            ref_var = !!ref_varname)
+
+#     # initiate a list to store the filtered data for each reference (e.g. reference year)
+#     ref_list_list<- list()
+
+#     # Loop through the references
+#     for(ref in ref_values){
+
+#       # filter for the closest observation, by group (e.g. by country)
+#       df_ref<- df_in %>%
+#         select(target_var, group_by_var, ref_var) %>%
+#         drop_na() %>%
+#         mutate(distance = ref_var - ref) %>%
+#         mutate(abs_distance = abs(distance)) %>%
+#         group_by(group_by_var) %>%
+#         mutate(min_distance = min(abs_distance)) %>%
+#         filter(abs_distance == min_distance) %>%
+#         group_by(group_by_var, min_distance) %>%
+#         mutate(n = n()) %>% # count number of matches (n>1 if there is a tie-break)
+#         ungroup()
+
+#       # Settle tie-breaks. If 'round_up' is true, the higher value wins out - i.e. the distance to the reference value must positive. And visa versa.
+#       if(tie_break_round_up){
+#         df_ref<- df_ref %>%
+#          filter(n==1 | distance>0)
+#       } else {
+#         df_ref<- df_ref %>%
+#          filter(n==1 | distance<0)
+#       }
+
+#       ref_list_list[[as.character(ref)]]<- df_ref
+
+#     }
+
+#     # Collate the matches for each reference observation
+#     df_out<- bind_rows(ref_list_list, .id = "reference")
+
+#     # If a tolerance has been set, filter out matches beyond tolerance
+#     if(!is.na(tolerance)){
+#         df_out<- df_out %>%
+#          filter(abs_distance <= tolerance)
+#      }
+
+#     # drop not needed vars
+#     df_out<- df_out %>%
+#       select(-c(distance, min_distance, n))
+
+#     # code rise, fall, stable
+#     df_out<- df_out %>%
+#       drop_na() %>%
+#       pivot_longer(cols = c(target_var, ref_var, abs_distance), names_to="metric", values_to = "value" ) %>%
+#       pivot_wider(id_cols = group_by_var, names_from = c(metric,reference), values_from = value) %>%
+#   drop_na() %>%
+#   mutate(change = !!sym(paste0("target_var_", max(ref_values))) - !!sym(paste0("target_var_", min(ref_values)))) %>%
+#   mutate(rise_fall = "stable") %>%
+#   mutate(rise_fall = if_else(change< -(rise_fall_threshold),
+#                              "fall",
+#                              rise_fall)) %>%
+#   mutate(rise_fall = if_else(change > rise_fall_threshold,
+#                              "rise",
+#                              rise_fall))
+
+#     # Store the results in a list
+#     df_out_list[[df_name]]<- df_out
+
+#   }
+
+#   # Join the results gathered from each input dataframe
+#   df_out_list[["joint"]]<- full_join(
+#     df_out_list[[df_names[1]]], df_out_list[[df_names[2]]],
+#     by = c("group_by_var"),
+#     suffix = c(paste0("_",df_names[1]), paste0("_",df_names[2])),
+#     keep = TRUE
+#     )
+
+#   # Make a list of fall/rise summaries
+
+#     # Initiate the list
+#     rise_fall_list<- list()
+
+#     #Loop through the two sources
+#     for(df_name in df_names){
+
+#        # For the individual source observations
+#         rise_fall_list[[paste0(df_name, "_allObs")]]<- df_out_list[[df_name]] %>%
+#           count(rise_fall)
+
+#         # for the observations that are in both sources
+#         rise_fall_list[[paste0(df_name, "_sharedObs")]]<- df_out_list[["joint"]] %>% 
+#       drop_na() %>%
+#       count(across(paste0("rise_fall_", df_name)))
+#     }
+
+#   # Make a list of the averages in the two reference periods
+
+#     # Initiate the list
+#     average_list<- list()
+
+#     #Loop through the two sources
+#     for(df_name in df_names){
+
+#        # For the individual source observations
+#         average_list[[paste0(df_name, "_allObs")]]<- df_out_list[[df_name]] %>%
+#           summarise(!!sym(paste0("avg_", ref_values[1])) := mean(!!sym(paste0("target_var_", ref_values[1]))),
+#                     !!sym(paste0("avg_", ref_values[2])) := mean(!!sym(paste0("target_var_", ref_values[2]))))
+
+#         # for the observations that are in both sources
+#         average_list[[paste0(df_name, "_sharedObs")]]<- df_out_list[["joint"]] %>% 
+#       drop_na() %>%
+#       summarise(!!sym(paste0("avg_", ref_values[1])) := mean(!!sym(paste("target_var", ref_values[1], df_name, sep= "_"))),
+#                     !!sym(paste0("avg_", ref_values[2])) := mean(!!sym(paste("target_var", ref_values[1], df_name, sep= "_"))))
+
+#     }
+
+#   # Undo the renaming (used early to avoid nonstandard evaluation problems)
+#   for(df_name in names(df_out_list)){
+
+#   names(df_out_list[[df_name]]) <- gsub(x = names(df_out_list[[df_name]]), pattern = "target_var", replacement = target_varname)
+
+#   names(df_out_list[[df_name]]) <- gsub(x = names(df_out_list[[df_name]]), pattern = "group_by_var", replacement = group_by_varname)
+
+#   names(df_out_list[[df_name]]) <- gsub(x = names(df_out_list[[df_name]]), pattern = "ref_var", replacement = ref_varname)
+# }
+
+#  outputs<- list("dataframes" = df_out_list,
+#                 "summaries" = summary_list)
+
+# return(outputs)
+# }
+
+```
+
+```{r}
+
+# compare_pip_wid<- list()
+
+# for(var in c("Gini", "Top_10_share")){
+
+#   if(var == "Gini"){rise_fall_threshold<- 0.1}
+#   if(var == "Top_10_share"){rise_fall_threshold<- 1}
+
+# compare_for_var<- compare_ref_years(
+#     df_list = list("pip" = df_pip, "wid" = df_wid),
+#     target_varname = var,
+#     group_by_varname = "Entity",
+#     ref_varname = "Year",
+#     ref_values = c(1990, 2020),
+#     tolerance = 5,
+#     tie_break_round_up = FALSE,
+#     rise_fall_threshold = rise_fall_threshold
+#     )
+
+# compare_pip_wid[[var]]<- compare_for_var 
+# }
+
+```
+
+A scatter chart to compare
+```{r}
+
+```
+
+A table to compare rises and falls
+```{r}
+
+# tab_compare_rise_and_falls<- function(summaries_list){
+
+#   summaries_list<- compare_pip_wid[["Gini"]][["summaries"]]
+#   tab<- data.frame(rise_fall = c("rise", "stable", "fall"))
+
+#   source_names<- names(summaries_list)
+
+#   for(source_name in source_names) {
+#   tab<- left_join(tab, summaries_list[[source_name]]) %>%
+#     rename(!!(paste0(source_name, "_all")) := n)
+#   }
+
+#   # Adding data for shared observations
+
+# }
+
+```
+
+Averages
+```{r}
+
+```
+
+:::
+
+#### Method 2: Use a regression to obtain the time trend
+
+### Results for a range of inequality metrics
+
+::: {.callout-note collapse="true" appearance="minimal"} 
+## Calculations for comparisons
+
+Calculations for Gini
+```{r}
+
+# group_by_varname<- "Entity"
+# ref_varname<- "Year"
+# ref_values<- c(1990, 2020)
+# tolerance<- 5
+# tie_break_round_up<- FALSE
+
+# target_varname<- "Gini"
+
+# compare_gini<- compare_pip_and_wid_ref_years(target_varname = "Gini")
+
+# # Spread wide and calculate change over time
+
+# # PIP only
+# compare_gini_wide_pip<- compare_gini
+
+# names(compare_gini_wide_pip) <- gsub(x = names(compare_gini_wide_pip), pattern = "_pip", replacement = "")
+
+# compare_gini_wide_pip<- compare_gini_wide_pip %>%
+#   select(Entity, reference_Year, Gini) %>%
+#   drop_na() %>%
+#   pivot_wider(id_cols = Entity, names_from = reference_Year, values_from = Gini) %>%
+#   drop_na() %>%
+#   mutate(change = `2020` - `1990`) %>%
+#   mutate(rise_fall = "stable") %>%
+#   mutate(rise_fall = if_else(change< -0.01,
+#                              "fall",
+#                              rise_fall)) %>%
+#   mutate(rise_fall = if_else(change > 0.01,
+#                              "rise",
+#                              rise_fall))
+
+# compare_gini_pip_summary<- compare_gini_wide_pip %>%
+#   count(rise_fall)
+
+# # WID only
+# compare_gini_wide_wid<- compare_gini
+
+# names(compare_gini_wide_wid) <- gsub(x = names(compare_gini_wide_wid), pattern = "_wid", replacement = "")
+
+# compare_gini_wide_wid<- compare_gini_wide_wid %>%
+#   select(Entity, reference_Year, Gini) %>%
+#   drop_na() %>%
+#   pivot_wider(id_cols = Entity, names_from = reference_Year, values_from = Gini) %>%
+#   drop_na() %>%
+#   mutate(change = `2020` - `1990`) %>%
+#   mutate(rise_fall = "stable") %>%
+#   mutate(rise_fall = if_else(change< -0.01,
+#                              "fall",
+#                              rise_fall)) %>%
+#   mutate(rise_fall = if_else(change > 0.01,
+#                              "rise",
+#                              rise_fall))
+
+# compare_gini_wid_summary<- compare_gini_wide_wid %>%
+#   count(rise_fall)
+
+# # In both PIP and WID
+
+# compare_gini_pip_and_wid<- full_join(compare_gini_wide_pip, compare_gini_wide_wid, by= c("Entity"), keep = TRUE) %>%
+#   drop_na() %>%
+#   rename(rise_fall_pip = rise_fall.x,
+#          rise_fall_wid = rise_fall.y)
+
+# compare_gini_pip_wid_summary_pip<- compare_gini_pip_and_wid %>%
+#   count(rise_fall_pip)
+
+# compare_gini_pip_wid_summary_wid<- compare_gini_pip_and_wid %>%
+#   count(rise_fall_wid)
+
+# # Make a table
+
+# df_gini_table<- data.frame(
+#   rise_fall = c("rise", "fall", "stable")
+# )
+
+# df_gini_table<- left_join(df_gini_table, compare_gini_pip_summary) %>%
+#   rename(pip_all = n)
+
+# df_gini_table<- left_join(df_gini_table, compare_gini_pip_wid_summary_pip, by = c("rise_fall" = "rise_fall_pip")) %>%
+#   rename(pip_shared = n)
+
+# df_gini_table<- left_join(df_gini_table, compare_gini_wid_summary) %>%
+#   rename(wid_all = n)
+
+# df_gini_table<- left_join(df_gini_table, compare_gini_pip_wid_summary_wid, by = c("rise_fall" = "rise_fall_wid")) %>%
+#   rename(wid_shared = n)
+
+# # country by country table for rise and fall
+# df_gini_country_rise_fall_table<- full_join(compare_gini_wide_pip, compare_gini_wide_wid, by= c("Entity"), keep = TRUE) %>%
+#   filter(rise_fall.x != rise_fall.y) %>%
+#   mutate(Entity = Entity.x) %>%
+#   mutate(Entity = if_else(!is.na(Entity.y),
+#                           Entity.y,
+#                           Entity.x)) %>%
+#   select(Entity, rise_fall.x, rise_fall.y) %>%
+#   rename(pip = rise_fall.x,
+#          wid = rise_fall.y)
+
+```
+
+Calculations for Top 10% share
+```{r}
+
+# compare_top_10<- compare_pip_and_wid_ref_years(target_varname = "Top_10_share")
+
+# # Spread wide and calculate change over time
+
+# # PIP only
+# compare_top_10_wide_pip<- compare_top_10
+
+# names(compare_top_10_wide_pip) <- gsub(x = names(compare_top_10_wide_pip), pattern = "_pip", replacement = "")
+
+# compare_top_10_wide_pip<- compare_top_10_wide_pip %>%
+#   select(Entity, reference_Year, Top_10_share) %>%
+#   drop_na() %>%
+#   pivot_wider(id_cols = Entity, names_from = reference_Year, values_from = Top_10_share) %>%
+#   drop_na() %>%
+#   mutate(change = `2020` - `1990`) %>%
+#   mutate(rise_fall = "stable") %>%
+#   mutate(rise_fall = if_else(change< -1,
+#                              "fall",
+#                              rise_fall)) %>%
+#   mutate(rise_fall = if_else(change > 1,
+#                              "rise",
+#                              rise_fall))
+
+# compare_top_10_pip_summary<- compare_top_10_wide_pip %>%
+#   count(rise_fall)
+
+# # WID only
+# compare_top_10_wide_wid<- compare_top_10
+
+# names(compare_top_10_wide_wid) <- gsub(x = names(compare_top_10_wide_wid), pattern = "_wid", replacement = "")
+
+# compare_top_10_wide_wid<- compare_top_10_wide_wid %>%
+#   select(Entity, reference_Year, Top_10_share) %>%
+#   drop_na() %>%
+#   pivot_wider(id_cols = Entity, names_from = reference_Year, values_from = Top_10_share) %>%
+#   drop_na() %>%
+#   mutate(change = `2020` - `1990`) %>%
+#   mutate(rise_fall = "stable") %>%
+#   mutate(rise_fall = if_else(change< -1,
+#                              "fall",
+#                              rise_fall)) %>%
+#   mutate(rise_fall = if_else(change > 1,
+#                              "rise",
+#                              rise_fall))
+
+# compare_top_10_wid_summary<- compare_top_10_wide_wid %>%
+#   count(rise_fall)
+
+# # In both PIP and WID
+
+# compare_top_10_pip_and_wid<- full_join(compare_top_10_wide_pip, compare_top_10_wide_wid, by= c("Entity"), keep = TRUE) %>%
+#   drop_na() %>%
+#   rename(rise_fall_pip = rise_fall.x,
+#          rise_fall_wid = rise_fall.y)
+
+# compare_top_10_pip_wid_summary_pip<- compare_top_10_pip_and_wid %>%
+#   count(rise_fall_pip)
+
+# compare_top_10_pip_wid_summary_wid<- compare_top_10_pip_and_wid %>%
+#   count(rise_fall_wid)
+
+# # Make a table
+
+# df_top_10_table<- data.frame(
+#   rise_fall = c("rise", "fall", "stable")
+# )
+
+# df_top_10_table<- left_join(df_top_10_table, compare_top_10_pip_summary) %>%
+#   rename(pip_all = n)
+
+# df_top_10_table<- left_join(df_top_10_table, compare_top_10_pip_wid_summary_pip, by = c("rise_fall" = "rise_fall_pip")) %>%
+#   rename(pip_shared = n)
+
+# df_top_10_table<- left_join(df_top_10_table, compare_top_10_wid_summary) %>%
+#   rename(wid_all = n)
+
+# df_top_10_table<- left_join(df_top_10_table, compare_top_10_pip_wid_summary_wid, by = c("rise_fall" = "rise_fall_wid")) %>%
+#   rename(wid_shared = n)
+
+# # country by country table for rise and fall
+# df_top_10_country_rise_fall_table<- full_join(compare_top_10_wide_pip, compare_top_10_wide_wid, by= c("Entity"), keep = TRUE) %>%
+#   filter(rise_fall.x != rise_fall.y) %>%
+#   mutate(Entity = Entity.x) %>%
+#   mutate(Entity = if_else(!is.na(Entity.y),
+#                           Entity.y,
+#                           Entity.x)) %>%
+#   select(Entity, rise_fall.x, rise_fall.y) %>%
+#   rename(pip = rise_fall.x,
+#          wid = rise_fall.y)
+
+```
+
+:::
+
+#### Comparison between 1990 and 2020 (+/- 5 years))
+::: {.panel-tabset}
+### Gini
+```{r}
+# kable(df_gini_table)
+
+# kable(df_gini_country_rise_fall_table)
+
+```
+
+### Top 10% share
+```{r}
+# kable(df_top_10_table)
+
+# kable(df_top_10_country_rise_fall_table)
+
+```
+
+### MLD
+:::
+
+## Explaining differences between PIP and WID data
+
+### For a subset of countries we can bridge between the PIP and WID data
+
+Besides the underlying data sources, there are many differences between what is being measured in the PIP and WID data - notably the welfare concept and unit of analysis/equivalization.
+
+In order to get a better understanding of where differences between the datasets arise, for a subset of countries we can bridge between the two somewhat. For example:
+
+-   WID has post-tax income series for some countries
+-   We can use LIS to look at how changing the welfare concept, unit of analysis, equivalisation affects things.
+(For equivalisation, we could also possibly use the newly available PIP harmonized microdata).
+
+### More in-depth discussion of particular outliers
+
+-   Russia is a very notable case where the two datasets tell very different stories.
 

### 2023-01-08 test_pages/test_with_shinylive.qmd
Add publish and new app shell scripts
add test shinylive app that pulls in csv

@@
 ---
-title: "A test notebook with Shinylive"
+title: Shinylive in Quarto example
+format: html
+filters:
+  - shinylive
 ---
 
-``` {r}
-library(tidyverse)
-library(here)
+This is a Shinylive application embedded in a Quarto doc.
+
+These are Shiny for Python applications which run completely in the browser, using Pyodide (Python compiled to WebAssembly).
+
+See the [Shilylive quarto extension](https://github.com/quarto-ext/shinylive) for further documentation.
+
+And the documentation on [Shiny for Python](https://shiny.rstudio.com/py/docs/putting-it-together.html).
+
+```{shinylive-python}
+#| standalone: true
+
+from shiny import *
+
+app_ui = ui.page_fluid(
+    ui.input_slider("n", "N", 0, 100, 40),
+    ui.output_text_verbatim("txt"),
+)
+
+def server(input, output, session):
+    @output
+    @render.text
+    def txt():
+        return f"The value of n*2 is {input.n() * 2}"
+
+app = App(app_ui, server)
+
 ```
 
-``` {r}
-a<- c(1:10)
-b<- a * 2 + 1
+Here is a second app I am testing:
 
-df<- data.frame(a, b)
+```{shinylive-python}
+#| standalone: true
+#| viewerHeight: 420
 
-fp<- paste0(here(),"/shiny_apps/shiny_app1/data.csv")
+from shiny import *
+from plotnine import *
+from pyodide.http import open_url
+import pandas as pd
 
-write.csv(df, fp)
+app_ui = ui.page_fluid(
+    ui.output_plot("example_plot"),
+)
+
+def server(input, output, session):
+    @output
+    @render.plot
+    def example_plot():
+
+        url = 'https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/JoeHasell/PIP/data/ppp_2017/final/PIP_data_public_download/full_dataset/cons_only/poverty_cons_only.csv'
+
+        df = pd.read_csv(open_url(url))
+
+        plot = (ggplot(df, aes('Year', 'headcount_ratio_365', color = 'Entity'))
+        + geom_line())
+
+        # d = {'col1': [1, 2], 'col2': [3, 4]}
+        # df = pd.DataFrame(data=d)
+        # plot = (ggplot(df, aes('col1', 'col2'))
+        # + geom_point())
+
+        return plot
+
+app = App(app_ui, server)
 

### 2023-01-08 PhD_pages/data_appendices/1_A_prepare_PIP_and_WID.qmd
Shifting data prep for Paper 1 to appendix

---
title: "Appendix 1.A: Preparing PIP and WID inequality data"
format: html
warning: false
---

# Processing steps

The titles in each of the boxes below summarise the steps taken to prepare the PIP and WID data. In each case you can expand the box to see the code used.

::: {.callout-note collapse="true" appearance="minimal"}
### Load packages

```{r}
library(tidyverse)
library(knitr)

```
:::

## PIP data

::: {.callout-note collapse="true" appearance="minimal"}
### Load PIP data
We load the data from a file prepared from the PIP API by Our World in Data.
```{r}
# df_pip<- read.csv("https://raw.githubusercontent.com/owid/poverty-data/main/datasets/pip_dataset.csv")

#  A local version, in case internet isn't available 
df_pip<- read.csv("PhD_pages/data/pip_dataset (2).csv")

```

The dataframe looks like this:
```{r}
kable(

    head(df_pip, 50)

)
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Drop sub-national data

The PIP data includes observations for national, urban and rural populations (`reporting_level`):

```{r}
kable(
    df_pip %>%
        count(reporting_level)
)
```

Here we add a count of the number of reporting levels per country-year.
```{r}

count_reporting_levels<- df_pip %>%
    group_by(country, year) %>%
    count(reporting_level) %>%
    group_by(country, year) %>%
    summarise(n_reporting_levels = n())

df_pip<- left_join(df_pip, count_reporting_levels) %>%
    relocate(n_reporting_levels, .after=reporting_level)

kable(

    head(df_pip, 20)

)
```

We only want one reporting level for each country-year.

That is already true for most country-years. But some include all three reporting levels.
```{r}
kable(

    df_pip %>%
        count(n_reporting_levels)

)

```

Here are the countries which include all three reporting levels:
```{r}
multi_reporting_level_countries<- 
    df_pip %>%
        filter(n_reporting_levels == 3) %>%
        select(country) %>%
        unique() %>%
        pull(country)

print(multi_reporting_level_countries)

```

For the country-years with 3 reporting levels, we keep only the national estimates. 
```{r}
df_pip<- df_pip %>%
    filter(n_reporting_levels != 3 | reporting_level == 'national' )

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Drop regional data

The dataset includes aggregated estiates for world regions.

Here we define a list of these aggregate entities and drop them from the data.
```{r}
drop_aggs<- c(
  "East Asia and Pacific",
  "Europe and Central Asia",
  "High income countries",
  "Latin America and the Caribbean",
  "Middle East and North Africa",
  "Sub-Saharan Africa",
  "World"
)

df_pip<- df_pip %>%
  filter(!country %in% drop_aggs)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Select 2017 PPPs

The data is provided by the World Bank in both 2011 and 2017 PPPs.

We will be using 2017 PPPs for Paper 1.

```{r}

df_pip<- df_pip %>%
  filter(ppp_version == 2017)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Rename and select needed columns
```{r}

df_pip<- df_pip %>%
  rename("Gini" = "gini",
         "Top_10_share" = "decile10_share",
         "MLD" = "mld",
         "P90_P10_ratio" = "p90_p10_ratio",
         "Entity" = "country",
         "Year" = "year")

df_pip<- df_pip %>%
    select(
        Year,
        Entity,
        Gini,
        Top_10_share,
        MLD,
        P90_P10_ratio,
        welfare_type)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### See the prepared PIP data

The first 100 rows of the dataframe looks like this:
```{r}
kable(

    head(df_pip, 100)

)
```
:::

## WID data

::: {.callout-note collapse="true" appearance="minimal"}
### Load WID data
We load the data from a file prepared from the WID Stata package by Our World in Data.

```{r}
# df_wid<- read.csv("https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/PabloArriagada/WID/data/final/wid_pretax.csv")

# A local version, in case internet isn't available 
df_wid<- read.csv("PhD_pages/data/wid.csv")

```

The dataframe looks like this:
```{r}
kable(

    head(df_wid, 50)

)
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Drop subnational, regional and historical geographical entities

Here we create a list of entities in the WID data that do not correspond to countries in existence today. These we drop from the data.
```{r}
drop_entities<- c(
"China - rural",
"China - urban",
"East Germany",
"Other Russia and Central Asia",
"Other East Asia",
"Other Western Europe",
"Other Latin America",
"Other MENA",
"Other South & South-East Asia",
"Other Sub-Saharan Africa",
"Africa",
"Asia",
"Europe",
"Oceania",
"Central Asia",
"East Africa",
"East Asia",
"Eastern Europe",
"Middle Africa",
"North Africa",
"North America",
"South-East Asia",
"South Africa region",
"West Africa",
"West Asia",
"Western Europe",
"European Union",
"World",
"Asia (excluding Middle East)",
"North America and Oceania",
"Sub-Saharan Africa",
"Latin America",
"Middle East",
"MENA",
"Russia and Central Asia",
"South & South-East Asia"
)

df_wid<- df_wid %>%
  filter(!Entity %in% drop_entities)
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Calculate some additional inequality metrics

```{r}

#  Calculate P90_P10_ratio
df_wid<- df_wid %>%
    mutate(P90_P10_ratio = `P90...income.threshold.of.the.top.10.`/P10...income.threshold)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Rename and select needed variables

```{r}
df_wid<- df_wid %>%
    rename("Gini" = "Gini.coefficient",
         "Top_10_share" = "P90.P100...share.of.the.top.10.") %>%
    select(
        Year,
        Entity,
        Gini,
        Top_10_share,
        # MLD,
        P90_P10_ratio)

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### See the prepared WID data

The first 100 rows of the dataframe looks like this:
```{r}
kable(

    head(df_wid, 100)

)
```
:::

## Important notes

### Income and consumption data in PIP
The PIP data includes a mix of income and consumption observations – including both for some country-years.

### Coverage

::: {.callout-note collapse="true" appearance="minimal"}
### Compare countries available across datasets
```{r}

pip_countries<- df_pip %>%
  select(Entity) %>%
  unique() %>%
  mutate(Entity_pip = Entity)

wid_countries<- df_wid %>%
  select(Entity) %>%
  unique() %>%
  mutate(Entity_wid = Entity)

countries<- full_join(pip_countries, wid_countries, keep = TRUE) %>%
  select(Entity_pip, Entity_wid)

```

Countries that are in the PIP data but not the WID data:
```{r}
kable(
  countries %>%
       filter(is.na(Entity_wid))
)
```
Countries that are in the WID data but not the PIP data:
```{r}
kable(
  countries %>%
       filter(is.na(Entity_pip))
)
```
:::

### 2023-01-08 PhD_pages/data_appendices/1_B_within_country_trends.qmd
Shifting data prep for Paper 1 to appendix

--- 2023-01-08 test_pages/test_with_shinylive.qmd
@@
 ---
-title: Shinylive in Quarto example
+title: "Appendix 1.B: Analysing global trends in within-country inequality"
 format: html
+warning: false
 filters:
@@
 
-This is a Shinylive application embedded in a Quarto doc.
+A key difficulty of comparing trends is that the data is incomplete coverage.
+The paper uses two methods to calculate and compare aggregate within-country trends across incomplete data: using reference years and using year fixed effects in a regression.
 
-These are Shiny for Python applications which run completely in the browser, using Pyodide (Python compiled to WebAssembly).
+## Method 1: Reference years
 
-See the [Shilylive quarto extension](https://github.com/quarto-ext/shinylive) for further documentation.
+I need to allow for the income/consumption issue (currently the data includes both – i.e. multi observations per country year).
 
-And the documentation on [Shiny for Python](https://shiny.rstudio.com/py/docs/putting-it-together.html).
+## Method 2: regression analysis
 
-```{shinylive-python}
-#| standalone: true
+## Explore the date
 
-from shiny import *
+I plan to build a Shiny app to help compare trends across datasets (using Shinylive – built on Shiny for Python - ).
 
-app_ui = ui.page_fluid(
-    ui.input_slider("n", "N", 0, 100, 40),
-    ui.output_text_verbatim("txt"),
-)
-
-def server(input, output, session):
-    @output
-    @render.text
-    def txt():
-        return f"The value of n*2 is {input.n() * 2}"
-
-app = App(app_ui, server)
-
-```
-
-Here is a second app I am testing:
+Here is a test app just so I can test the wiring of how such an app works.
 

### 2023-01-08 PhD_pages/paper_pages/paper1.qmd
Shifting data prep for Paper 1 to appendix

@@
 title: "Paper 1: Triangulating trends in inequality around the world from secondary databases"
+format: html
 ---
@@
 
-::: {.callout-note collapse="true" appearance="minimal"}
-### Load PIP data
-
-```{r}
-# df_pip<- read.csv("https://raw.githubusercontent.com/owid/poverty-data/main/datasets/pip_dataset.csv")
-
-df_pip<- read.csv("PhD_pages/data/pip_dataset (2).csv")
-
-df_pip<- df_pip %>%
-  filter(ppp_version == 2017) %>%
-  rename("Gini" = "gini",
-         "Top_10_share" = "decile10_share",
-         "Entity" = "country",
-         "Year" = "year")
-
-# drop subnational data where national data is available
-df_pip<- df_pip %>%
-  filter((!Entity %in% c("China", "India", "Indonesia")) | reporting_level == "national")
-
-# drop income if both income and consumption are available
-df_pip<- df_pip %>%
-  group_by(Entity, reporting_level, Year) %>%
-  mutate(n = n()) %>%
-  filter(n<2 | welfare_type == "consumption") %>%
-  ungroup()
-
- #drop regional aggregates
-drop_aggs<- c(
-  "East Asia and Pacific",
-  "Europe and Central Asia",
-  "High income countries",
-  "Latin America and the Caribbean",
-  "Middle East and North Africa",
-  "Sub-Saharan Africa",
-  "World"
-)
-
-df_pip<- df_pip %>%
-  filter(!Entity %in% drop_aggs)
-
-```
-:::
-
 ### WID
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Load WID data
-
-```{r}
-# df_wid<- read.csv("https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/PabloArriagada/WID/data/final/wid_pretax.csv")
-
-df_wid<- read.csv("PhD_pages/data/wid.csv")
-
-df_wid<- df_wid %>%
-  rename("Gini" = "Gini.coefficient",
-         "Top_10_share" = "P90.P100...share.of.the.top.10.")
-
-drop_entities<- c(
-"China - rural",
-"China - urban",
-"East Germany",
-"Other Russia and Central Asia",
-"Other East Asia",
-"Other Western Europe",
-"Other Latin America",
-"Other MENA",
-"Other South & South-East Asia",
-"Other Sub-Saharan Africa",
-"Africa",
-"Asia",
-"Europe",
-"Oceania",
-"Central Asia",
-"East Africa",
-"East Asia",
-"Eastern Europe",
-"Middle Africa",
-"North Africa",
-"North America",
-"South-East Asia",
-"South Africa region",
-"West Africa",
-"West Asia",
-"Western Europe",
-"European Union",
-"World",
-"Asia (excluding Middle East)",
-"North America and Oceania",
-"Sub-Saharan Africa",
-"Latin America",
-"Middle East",
-"MENA",
-"Russia and Central Asia",
-"South & South-East Asia"
-)
-
-df_wid<- df_wid %>%
-  filter(!Entity %in% drop_entities)
-```
-
-:::
 
@@
 Give an overview of coverage in each dataset.
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Standardize coutry names across datasets
-```{r}
-
-pip_countries<- df_pip %>%
-  select(Entity) %>%
-  unique() %>%
-  mutate(Entity_pip = Entity)
-
-wid_countries<- df_wid %>%
-  select(Entity) %>%
-  unique() %>%
-  mutate(Entity_wid = Entity)
-
-countries<- full_join(pip_countries, wid_countries, keep = TRUE) %>%
-  select(Entity_pip, Entity_wid)
-
-```
-
-Countries that are in the PIP data but not the WID data:
-```{r}
-kable(
-  countries %>%
-       filter(is.na(Entity_wid))
-)
-```
-Countries that are in the WID data but not the PIP data:
-(TO-DO: drop regions and sub-national data from the WID data at an eariler step.)
-```{r}
-kable(
-  countries %>%
-       filter(is.na(Entity_pip))
-)
-```
-:::
 

### 2023-01-18 PhD_pages/data_appendices/1_A_prepare_PIP_and_WID.qmd
restructuring and data work on paper 1

@@
 ```
+
+```{python}
+# For some reason I can't get the owid package working when rendering... only when running in the interactive window in vs code
+# from owid import catalog
+import pandas as pd
+```
 :::
@@
 #  A local version, in case internet isn't available 
-df_pip<- read.csv("PhD_pages/data/pip_dataset (2).csv")
+df_pip<- read.csv("PhD_pages/data_appendices/data/original/pip_dataset.csv")
 
@@
   "East Asia and Pacific",
+  "South Asia",
   "Europe and Central Asia",
@@
 ::: {.callout-note collapse="true" appearance="minimal"}
+### Drop extremely low P10 values
+
+There are some very low values of P10. These lead to exaggerated values of the P90/P10 ratio.
+
+```{r}
+
+kable(
+    df_pip %>% filter(decile1_thr<0.02)
+
+  )
+
+```
+
+We set these values of P10 to NA, and thereby precluding a value for P90/P10 ratio when it is calculated below.
+
+```{r}
+
+df_pip<- df_pip %>%
+  mutate(decile1_thr = if_else(decile1_thr<0.02,
+                              as.numeric(NA),
+                              decile1_thr))
+
+kable(
+    df_pip %>% filter(is.na(decile1_thr))
+
+  )
+
+```
+
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Calculate additional variables
+
+```{r}
+
+df_pip<- df_pip %>%
+  mutate(Bottom_50_share = decile1_share + decile2_share + decile3_share + decile4_share + decile5_share,
+          P90_P10_ratio = decile9_thr/decile1_thr)
+
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
 ### Rename and select needed columns
@@
          "MLD" = "mld",
-         "P90_P10_ratio" = "p90_p10_ratio",
          "Entity" = "country",
@@
         Top_10_share,
+        Bottom_50_share,
         MLD,
@@
 ```
+
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Save the data
+I save the prepared data locally:
+```{r}
+
+write.csv(df_pip, 
+      "PhD_pages/data_appendices/data/clean/pip.csv", row.names=FALSE)
+
+```
 :::
@@
 # A local version, in case internet isn't available 
-df_wid<- read.csv("PhD_pages/data/wid.csv")
+df_wid<- read.csv("PhD_pages/data_appendices/data/original/wid.csv")
 
@@
 "East Asia",
+"South Asia",
 "Eastern Europe",
@@
 ::: {.callout-note collapse="true" appearance="minimal"}
+### Drop extremely low P10 values
+
+There are some very low values of P10. These lead to exaggerated values of the P90/P10 ratio. Here are all values less than $1 a day (using pretax national income)
+
+```{r}
+
+kable(
+    df_wid %>% filter(P10...income.threshold/365<1)
+
+  )
+
+```
+
+We set these values of P10 to NA, and thereby precluding a value for P90/P10 ratio when it is calculated below.
+
+```{r}
+
+df_wid<- df_wid %>%
+  mutate(P10...income.threshold = if_else(P10...income.threshold/365<1,
+                              as.numeric(NA),
+                              P10...income.threshold))
+
+                              kable(
+    df_wid %>% filter(is.na(P10...income.threshold))
+
+  )
+
+```
+
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
 ### Calculate some additional inequality metrics
@@
     rename("Gini" = "Gini.coefficient",
-         "Top_10_share" = "P90.P100...share.of.the.top.10.") %>%
+         "Top_10_share" = "P90.P100...share.of.the.top.10.",
+         "Bottom_50_share"= "P0.P50...share.of.the.bottom.50.",
+         "P90" = "P90...income.threshold.of.the.top.10.",
+         "P10" = "P10...income.threshold") %>%
     select(
@@
         # MLD,
-        P90_P10_ratio)
+        P90_P10_ratio,
+        Bottom_50_share,
+        P90,
+        P10)
 
@@
 )
+```
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Save the data
+I save the prepared data locally:
+```{r}
+
+write.csv(df_wid, 
+      "PhD_pages/data_appendices/data/clean/wid.csv", row.names=FALSE)
+
+```
+:::
+:::
+
+## Additional variables
+::: {.callout-note collapse="true" appearance="minimal"}
+### Region classfications
+For regional classifications I take WID definitions, as specified on their website [here](https://wid.world/codes-dictionary/#country-code).
+
+I read in a csv prepared from the table in that webpage, which has been mapped to the Our World in Data standarized country names used throughout this project.
+```{r}
+
+df_regions<- read.csv("PhD_pages/data_appendices/data/original/WID_regions_country_standardized.csv")
+
+```
+
+I create a clean region mapping table using two regional definitions.
+One is the 'standard' definitions used by WID.
+In an alternative set I split Western and Eastern Europe, ccombining the latter with Russia & Central Asia to make a joint Eastern Europe and Central Asia region.
+```{r}
+
+df_regions<- df_regions %>%
+  select(Our.World.In.Data.Name, `region..standard.`, sub.region) %>%
+  rename(Entity = Our.World.In.Data.Name,
+        WID_region = `region..standard.`) %>%
+  mutate(region_alt = if_else(sub.region == 'Western Europe',
+  'Western Europe',
+  WID_region)) %>%
+  mutate(region_alt = if_else((sub.region == 'Eastern Europe' | WID_region == "Russia & Central Asia"),
+  'Eastern Europe & Central Asia',
+  WID_region)) %>%
+  select(Entity, WID_region, region_alt)
+
+kable(
+  df_regions
+)
+```
+
+Check that no countries are not matched for either dataset.
+```{r}
+# Merge in the regions dataset
+df_pip<- left_join(df_pip, df_regions)
+df_wid<- left_join(df_wid, df_regions)
+
+# PIP check
+kable(
+  df_pip %>% 
+    select(Entity, WID_region, region_alt) %>%
+    unique() %>%
+    filter(is.na(WID_region))
+)
+
+# WID check
+kable(
+  df_wid %>% 
+    select(Entity, WID_region, region_alt) %>%
+    unique() %>%
+    filter(is.na(WID_region))
+)
+
+```
+
+Save the region mapping
+```{r}
+write.csv(df_regions, 
+      "PhD_pages/data_appendices/data/clean/WID_region_mapping.csv", row.names=FALSE)
+
+```
+
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Region classfications
+
+```{python}
+# For some reason I can't get the owid package working when rendering... only when running in the interactive window in vs code
+# results = catalog.find('population')
+# owid_pop = results[results['path'] == 'garden/owid/latest/key_indicators/population']
+
+# df_pop = owid_pop.iloc[0].load().reset_index().rename(columns={'year':'Year', 'country': 'Entity'})
+
+# df_pop.head()
+
+```
+
+Save the population data
+```{python}
+# df_pop.to_csv('PhD_pages/data_appendices/data/clean/population.csv')
+
 ```
@@
 :::
+
+### Outliers
+::: {.callout-note collapse="true" appearance="minimal"}
+### P90/P10 ratio
+
+For the P90/P10 ratio, a few observations have extreme values. In the PIP data
+```{r}
+kable(
+
+  df_pip %>% arrange(-P90_P10_ratio) %>% head()
+)
+
+```
+
+And the WID data:
+
+```{r}
+kable(
+
+  df_wid %>% arrange(-P90_P10_ratio) %>% head(100)
+)
+
+```
+
+:::

### 2023-01-18 PhD_pages/data_appendices/1_B_within_country_trends.qmd
restructuring and data work on paper 1

@@
 ---
+```{python}
+
+import pandas as pd
+import numpy as np
+from plotnine import *
+import plotly.graph_objects as go
+```
+
+Load in the data
+```{python}
+
+df_wid = pd.read_csv("data/clean/wid.csv")
+df_pip = pd.read_csv("data/clean/pip.csv")
+df_regions = pd.read_csv("data/clean/WID_region_mapping.csv")
+
+```
+
+```{python}
+
+# df_wid = pd.read_csv("PhD_pages/data_appendices/data/clean/wid.csv")
+# df_pip = pd.read_csv("PhD_pages/data_appendices/data/clean/pip.csv")
+# df_regions = pd.read_csv("PhD_pages/data_appendices/data/clean/WID_region_mapping.csv")
+
+```
 
@@
 I need to allow for the income/consumption issue (currently the data includes both – i.e. multi observations per country year).
+
+### Mapping data to a reference year
+This function grabs the observation, by a grouping variable (e.g. by country), for which a reference variable (e.g. year) is closest to a reference value (e.g. 2002).
+You can specify a maximum distance from the reference value, beyond which no match will be returned.
+Because there may be tie-breaks – matches equally distant above or below the reference value – there is an argmuent to specify how these tie-breaks should resolved.
+```{python}
+
+#  Function get matching for ref years
+def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):
+
+    df = df.loc[:, [reference_col, group_by_col, value_col]]
+
+    # Drop NAs
+    df = df.dropna()
+
+    # Calculate absolute distance from reference value
+    df['ref_diff'] = abs(df[reference_col] - reference_val)
+
+    # Drop any rows with a distance beyond threshold
+    if not pd.isna(max_dist_from_ref):
+      df = df.loc[df['ref_diff'] <= max_dist_from_ref]
+
+    # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
+    df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)
+
+    # Settle tie-breaks
+    if tie_break == 'below':
+      df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)
+
+    elif tie_break == 'above':
+      df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)
+
+    df = df.drop('ref_diff', axis=1)
+
+    return df
+
+```
+
+I test this function here:
+```{python}
+
+test = closest_to_reference(
+  df = df_wid, 
+  reference_val = 2002,
+  max_dist_from_ref = 5,
+  reference_col = 'Year',
+  group_by_col = 'Entity',
+  value_col = 'Gini',
+  tie_break = 'below'
+  )
+
+test.head()
+
+```
+
+### Obtaining a pair of observations
+
+This function runs the `closest_to_reference` function twice, over a pair of reference values.
+Tie-breaks are settled so as to maximise the gap between the two reference values.
+You can also specify a minimum distance between the two observations (e.g. pairs of matches that fall less than X years apart will be dropped).
+
+```{python}
+# Merge matches for different reference points
+def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+
+  # Make sure the pair of reference values are in ascending order
+  reference_vals.sort()
+
+  # Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref
+
+  # Find matches for lower reference value
+  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs[0], reference_col, group_by_col, value_col, 'below')
+
+  # Find matches for higher reference value
+  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs[1], reference_col, group_by_col, value_col, 'above')
+
+  # Merge the two sets of matches
+  merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))
+
+  # Drop obs that do not have data for both ref values
+  merged_df = merged_df.dropna()
+
+  # Drop obs where the matched data does not meet the min distance requirement
+  if not pd.isna(min_dist_between):
+
+    # Store the names of the reference column returned from the two matches
+    ref_var_high = f'{reference_col}{reference_vals[1]}'
+    ref_var_low = f'{reference_col}{reference_vals[0]}'
+
+    # Keep only rows >= to the min distance
+    merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]
+
+  return merged_df
+
+```
+
+I test this function here:
+```{python}
+test = merge_two_ref_matches(
+  df = df_wid, 
+  reference_vals = [2000, 2010], 
+  max_dist_from_refs = [5, 4],
+  min_dist_between = 9,
+  reference_col = 'Year',
+  group_by_col = 'Entity',
+  value_col = 'Gini'
+  )
+
+test.head()
+
+```
+
+#### Allowing for the different welfare concepts in the PIP data
+
+As noted in .... (add backlink) ... the PIP data contains both income and consumption observations.
+This function produces matched pairs of observations from that data in which only one pair is selected for each value of the grouping variable (i.e. each country).
+Priority is given to pairs for which both observations relate to income. Second priority is given to pairs where both observations relate to consumption.
+Only if neither pair is available then it will return a mixed pair of observations if that is available.
+```{python}
+# For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
+def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+
+  # Specify the name of the column in which the income/consumption welfare definition is stored
+  welfare_colname = 'welfare_type'
+
+  # Creat dataframes for thee scenarios:
+  # Scenario 1: only allow income data
+  df_inc_filter = df.loc[df[welfare_colname] == "income", :]
+  df_inc_filter.name = "Income"
+
+  # Scenario 2: only allow consumption data
+  df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
+  df_cons_filter.name = "Consumption"
+  # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
+  df_mixed = df.copy()
+
+  df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')
+
+  df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]
+
+  df_mixed.name = "Mixed"
+  #  Store the scneario dataframes in a list
+  df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]
+
+  # Run the matching function on each scenario
+  scenario_matches = [merge_two_ref_matches(
+    df_scenario, 
+    reference_vals, 
+    max_dist_from_refs, 
+    min_dist_between, 
+    reference_col, 
+    group_by_col, 
+    value_col) for df_scenario in df_scenarios]
+
+  # Combine the first two scenarios.
+  df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])
+
+  # Tidy up indexes
+  df_combined_matches = df_combined_matches.reset_index()
+
+  df_combined_matches = df_combined_matches.drop('level_1', axis=1)
+
+  df_combined_matches = df_combined_matches\
+    .rename(columns={"level_0": "pip_welfare"})
+
+  # Add in third scenario.
+  df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])
+
+  # add scenario name to te pip_welfare column
+  df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)
+
+  # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
+    # First count the matches
+  df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
+    # Then drop any matches from the lowest priority where there are multiple matches
+  df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
+    #  Repeat at the next level of priority
+  df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
+  df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]
+
+  # Drop the match count column
+  df_combined_matches = df_combined_matches.drop('match_count', axis=1)
+
+  return df_combined_matches
+
+```
+
+I test this function here:
+```{python}
+
+test = pip_welfare_routine(
+
+  df = df_pip,
+  reference_vals = [1986, 2016], 
+  max_dist_from_refs = [5, 5], 
+  min_dist_between = 30, 
+  reference_col = 'Year',
+  group_by_col = 'Entity',
+  value_col = 'Gini'
+
+  )
+
+test.head()
+```
+
+### Merging reference year aligned data from WID and PIP datasets
+
+```{python}
+
+def prep_wid_pip_data(reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+  pip_matches = pip_welfare_routine(
+    df = df_pip,
+    reference_vals = reference_vals,
+    max_dist_from_refs = max_dist_from_refs,
+    min_dist_between = min_dist_between,
+    reference_col = reference_col,
+    group_by_col = group_by_col,
+    value_col = value_col
+  )
+
+  wid_matches = merge_two_ref_matches(
+    df = df_wid,
+    reference_vals = reference_vals,
+    max_dist_from_refs = max_dist_from_refs,
+    min_dist_between = min_dist_between,
+    reference_col = reference_col,
+    group_by_col = group_by_col,
+    value_col = value_col
+  )
+
+  ref_pairs = pd.concat([pip_matches, wid_matches,], keys=['pip', 'wid'])
+
+  # Tidy up indexes
+  ref_pairs = ref_pairs.reset_index()
+
+  ref_pairs = ref_pairs.drop('level_1', axis=1)
+
+  ref_pairs = ref_pairs\
+    .rename(columns={"level_0": "source"})
+
+  return ref_pairs
+```
+
+## Prepare plot and tables
+
+```{python}
+#| echo: false
+
+def prep_plot_and_tables(
+  reference_vals, 
+  max_dist_from_refs,
+  min_dist_between,
+  reference_col,
+  group_by_col,
+  value_col,
+  tolerance,
+  outlier_cut_off_upper,
+  source_count_requirement):
+
+  # Other spec:
+  region_col = 'region_alt'
+
+  # ---- Prep data ---------
+
+  plot_data = prep_wid_pip_data(
+    reference_vals = reference_vals, 
+    max_dist_from_refs = max_dist_from_refs,
+    min_dist_between = min_dist_between,
+    reference_col = reference_col,
+    group_by_col = group_by_col,
+    value_col = value_col,
+  )
+
+  # Store the names of the columns to be used onthe X and Y axis
+  x_axis = f'{value_col}{reference_vals[0]}'
+  y_axis = f'{value_col}{reference_vals[1]}'
+
+  # Apply outlier cut off, if specified
+  if not pd.isna(outlier_cut_off_upper):
+    plot_data = plot_data.loc[plot_data[x_axis] <= outlier_cut_off_upper]
+    plot_data = plot_data.loc[plot_data[y_axis] <= outlier_cut_off_upper]
+
+  # Add a count by country – showing whether data is available from both sources or not
+  plot_data['source_count'] = plot_data.groupby(group_by_col)['source'].transform('count')
+
+  # Apply source requirement (whether to include only observations with data from both sources)
+  plot_data = plot_data.loc[plot_data['source_count'] >= source_count_requirement]
+
+  # Drop source_count column
+  plot_data = plot_data.drop('source_count', axis=1)
+
+  # Add in region classification
+  plot_data = pd.merge(plot_data, df_regions, how = 'left')
+
+  # Make nicely formatted table
+  plot_data_formatted = plot_data.copy()
+
+  plot_data_formatted[x_axis] = plot_data_formatted[x_axis].map('{:.2f}'.format)
+  plot_data_formatted[y_axis] = plot_data_formatted[y_axis].map('{:.2f}'.format)
+
+  plot_data_table = go.Figure(data=[go.Table(
+    header=dict(values=list(plot_data_formatted.columns)),
+    cells=dict(values=plot_data_formatted.transpose().values.tolist() ))
+    ])
+
+  # ----- Prep plot ------
+
+  # I grab what I am guessing to be the max and min tick marks on the x axist, in order to define the coordinates of a 'tolerance' ribbon.
+  x_min = plot_data[x_axis].min()
+  x_min_floor = np.floor(x_min * 10) / 10
+
+  x_max = plot_data[x_axis].max()
+  x_max_ceiling = np.ceil(x_max * 10) / 10
+
+  # Set the coordinates of the shaded ribbon
+  shaded_coord = pd.DataFrame({'x': [x_min_floor, x_max_ceiling, x_max_ceiling, x_min_floor], 
+    'y': [x_min_floor-tolerance, x_max_ceiling-tolerance, x_max_ceiling+tolerance , x_min_floor+tolerance]})
+
+  plot = (ggplot(plot_data
+  , aes(x_axis, y_axis))
+  + geom_point(aes(color=region_col))
+  + facet_wrap('~ source')
+  + geom_polygon(aes(x='x', y='y'), data=shaded_coord, fill="#FF000066"))
+
+  # ---- Prep summary table ----
+
+  # Count by rise/stable/fall bins
+
+  # Calculate the change between the two ref periods
+  plot_data['change'] = (plot_data[y_axis] - plot_data[x_axis])
+
+  # Define the thresholds using the tolerance specified
+  bin_thresholds = [-float("inf"), -tolerance, tolerance, float("inf")]
+  bin_labels = ['fall', 'stable', 'rise']
+
+  bins = pd.cut(plot_data['change'], bins=bin_thresholds, labels=bin_labels)
+
+  # By region:
+
+  # Count observations in bins defined by the thresholds
+  threshold_counts = plot_data\
+    .groupby(['source', region_col, bins])\
+    .size()\
+    .unstack()\
+    .unstack(level=region_col)\
+    .T\
+    .reset_index()\
+    .rename(columns={"change": "summary"})
+
+  # Calculate average change and n
+  mean_and_count = pd.DataFrame(plot_data.groupby(['source', region_col])['change']\
+    .agg(['mean', 'count']))\
+    .unstack()\
+    .T\
+    .reset_index()\
+    .rename(columns={"level_0": "summary"})
+
+  # rename the labels coming out of the aggregation
+  mean_and_count['summary'] = mean_and_count['summary'].replace(['mean', 'count'], ['avg change', 'n'])
+
+  # Concat together and present in a table
+  df_summary_region = pd.concat([threshold_counts, mean_and_count])
+
+  df_summary_region['summary'] = pd.Categorical(df_summary_region['summary'], categories=['fall', 'stable', 'rise','n', 'avg change'], ordered=True)
+
+  df_summary_region = df_summary_region.sort_values([region_col,'summary'])
+
+  # Set the index
+  df_summary_region = df_summary_region.set_index([region_col,'summary'])
+
+  # format to show appropriate d.p. I run a loop over each row for want of figuring out a better way!
+  repeat_length = 5
+  for i in range(0, len(df_summary_region.index.levels[0].unique())):
+    df_summary_region.iloc[i*repeat_length,:] = df_summary_region.iloc[i*repeat_length,:].map('{:.0f}'.format)
+    df_summary_region.iloc[1 + i*repeat_length,:] = df_summary_region.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
+    df_summary_region.iloc[2 + i*repeat_length,:] = df_summary_region.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
+    df_summary_region.iloc[3 + i*repeat_length,:] = df_summary_region.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
+    df_summary_region.iloc[4 + i*repeat_length,:] = df_summary_region.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)
+
+  # World summary
+
+  # Count observations in bins defined by the thresholds
+  threshold_counts = plot_data\
+    .groupby(['source', bins])\
+    .size()\
+    .unstack(level='source')\
+    .reset_index()\
+    .rename(columns={"change": "summary"})
+
+  # Calculate average change and n
+  mean_and_count = pd.DataFrame(plot_data.groupby(['source'])['change']\
+    .agg(['mean', 'count']))\
+    .T\
+    .reset_index()\
+    .rename(columns={"index": "summary"})
+
+  # rename the labels coming out of the aggregation
+  mean_and_count['summary'] = mean_and_count['summary'].replace(['mean', 'count'], ['avg change', 'n'])
+
+  # Concat together and present in a table
+  df_summary_world = pd.concat([threshold_counts, mean_and_count])
+
+  df_summary_world['summary'] = pd.Categorical(df_summary_world['summary'], categories=['fall', 'stable', 'rise','n', 'avg change'], ordered=True)
+
+  df_summary_world = df_summary_world.sort_values(['summary']).set_index('summary')
+
+  # format to show appropriate d.p. I do it row by row for want of figuring out a better way!
+  df_summary_world.iloc[0,:] = df_summary_world.iloc[0,:].map('{:.0f}'.format)
+  df_summary_world.iloc[1,:] = df_summary_world.iloc[1,:].map('{:.0f}'.format)
+  df_summary_world.iloc[2,:] = df_summary_world.iloc[2,:].map('{:.0f}'.format)
+  df_summary_world.iloc[3,:] = df_summary_world.iloc[3,:].map('{:.0f}'.format)
+  df_summary_world.iloc[4,:] = df_summary_world.iloc[4,:].map('{:.2f}'.format)
+
+  # ----- Return outputs ------
+  plot_and_tables = {
+    "table_data": plot_data,
+    "table": plot_data_table,
+    "plot": plot,
+    "summary_world": df_summary_world,
+    "summary_region": df_summary_region
+  }
+
+  return plot_and_tables
+
+```
+
+```{python}
+
+# Specifications
+reference_vals = [1990, 2018]
+max_dist_from_refs = [4, 4]
+min_dist_between = 20
+value_col = 'Bottom_50_share'
+# 'Gini'
+# 'Top_10_share'
+# 'Bottom_50_share'
+# 'MLD'
+# 'P90_P10_ratio' – not that helpful due to accuracy of low incomes
+
+# Data from both sources is needed?
+  # If this = 1 then a country will be incldued even if it is only availalbe from one source.
+  # If = 2, then only countries with data for both sources will be included
+source_count_requirement = 2
+
+# Set the value which constitutes a significant change in the value being measured
+  # E.g. for Gini tolerance = 0.02 (2 points)
+
+if value_col == 'Gini':
+  tolerance = 0.02
+  outlier_cut_off_upper = None
+
+if value_col == 'Top_10_share':
+  tolerance = 1
+  outlier_cut_off_upper = None
+
+if value_col == 'Bottom_50_share':
+  tolerance = 1
+  outlier_cut_off_upper = None
+
+if value_col == 'P90_P10_ratio':
+  tolerance = 0.5
+  outlier_cut_off_upper = 40
+
+# These parameters are unlikely to change:
+reference_col = 'Year'
+group_by_col = 'Entity'
+
+plot_and_tables = prep_plot_and_tables(
+  reference_vals = reference_vals, 
+  max_dist_from_refs = max_dist_from_refs,
+  min_dist_between = min_dist_between,
+  reference_col = reference_col,
+  group_by_col = group_by_col,
+  value_col = value_col,
+  tolerance = tolerance,
+  outlier_cut_off_upper = outlier_cut_off_upper,
+  source_count_requirement = source_count_requirement
+  )
+
+```
+
+```{python}
+plot_and_tables['plot']
+```
+
+```{python}
+plot_and_tables['summary']
+```
+
+:::{.column-page}
+```{python}
+
+plot_and_tables['table'].show()
+
+```
+
+```{python}
+
+plot_data_results = plot_and_tables['table_data']
+
+```
+:::
 

### 2023-01-18 PhD_pages/data_appendices/3_C_calc_factor_shares.qmd
restructuring and data work on paper 1

@@
 editor: visual
+execute:
+  freeze: true  # never re-render during project render
 ---
+
+# To-do:
+
+- I moved the data folder, but haven't yet relected this in the code below.
 

### 2023-01-20 PhD_pages/data_appendices/1_A_prepare_PIP_and_WID.qmd
Adding to Paper 1 data analysis, shifting to python

@@
 warning: false
+
 ---
@@
 #  A local version, in case internet isn't available 
-df_pip<- read.csv("PhD_pages/data_appendices/data/original/pip_dataset.csv")
+df_pip<- read.csv("data/original/pip_dataset.csv")
 
@@
 write.csv(df_pip, 
-      "PhD_pages/data_appendices/data/clean/pip.csv", row.names=FALSE)
+      "data/clean/pip.csv", row.names=FALSE)
 
@@
 # A local version, in case internet isn't available 
-df_wid<- read.csv("PhD_pages/data_appendices/data/original/wid.csv")
+df_wid<- read.csv("data/original/wid.csv")
 
@@
 write.csv(df_wid, 
-      "PhD_pages/data_appendices/data/clean/wid.csv", row.names=FALSE)
+      "data/clean/wid.csv", row.names=FALSE)
 
@@
 
-df_regions<- read.csv("PhD_pages/data_appendices/data/original/WID_regions_country_standardized.csv")
+# df_regions<- read.csv("data/original/WID_regions_country_standardized.csv")
+
+```
+
+```{python}
+
+df_regions = pd.read_csv("data/original/WID_regions_country_standardized.csv")
 
@@
 
-df_regions<- df_regions %>%
-  select(Our.World.In.Data.Name, `region..standard.`, sub.region) %>%
-  rename(Entity = Our.World.In.Data.Name,
-        WID_region = `region..standard.`) %>%
-  mutate(region_alt = if_else(sub.region == 'Western Europe',
-  'Western Europe',
-  WID_region)) %>%
-  mutate(region_alt = if_else((sub.region == 'Eastern Europe' | WID_region == "Russia & Central Asia"),
-  'Eastern Europe & Central Asia',
-  WID_region)) %>%
-  select(Entity, WID_region, region_alt)
+# df_regions<- df_regions %>%
+#   select(Our.World.In.Data.Name, `region..standard.`, sub.region) %>%
+#   rename(Entity = Our.World.In.Data.Name,
+#         WID_region = `region..standard.`) %>%
+#   mutate(region_alt = if_else(sub.region == 'Western Europe',
+#   'Western Europe',
+#   WID_region)) %>%
+#   mutate(region_alt = if_else((sub.region == 'Eastern Europe' | WID_region == "Russia & Central Asia"),
+#   'Eastern Europe & Central Asia',
+#   WID_region)) %>%
+#   select(Entity, WID_region, region_alt)
 
-kable(
-  df_regions
-)
+# kable(
+#   df_regions
+# )
+```
+
+```{python}
+
+df_regions = df_regions[['Our World In Data Name', 'region (standard)', 'sub-region']].rename(columns={'Our World In Data Name': 'Entity', 'region (standard)': 'WID_region'})
+
+# Define an alt region classification
+df_regions['region_alt'] = df_regions['WID_region']
+
+df_regions.loc[df_regions['sub-region'] == 'Western Europe','region_alt'] = 'Western Europe'
+
+df_regions.loc[(df_regions['sub-region'] == 'Eastern Europe') | (df_regions['WID_region'] == 'Russia & Central Asia'),'region_alt'] = 'Eastern Europe & Central Asia'
+
 ```
@@
 ```{r}
-# Merge in the regions dataset
-df_pip<- left_join(df_pip, df_regions)
-df_wid<- left_join(df_wid, df_regions)
+# # Merge in the regions dataset
+# df_pip<- left_join(df_pip, df_regions)
+# df_wid<- left_join(df_wid, df_regions)
 
-# PIP check
-kable(
-  df_pip %>% 
-    select(Entity, WID_region, region_alt) %>%
-    unique() %>%
-    filter(is.na(WID_region))
-)
+# # PIP check
+# kable(
+#   df_pip %>% 
+#     select(Entity, WID_region, region_alt) %>%
+#     unique() %>%
+#     filter(is.na(WID_region))
+# )
 
-# WID check
-kable(
-  df_wid %>% 
-    select(Entity, WID_region, region_alt) %>%
-    unique() %>%
-    filter(is.na(WID_region))
-)
+# # WID check
+# kable(
+#   df_wid %>% 
+#     select(Entity, WID_region, region_alt) %>%
+#     unique() %>%
+#     filter(is.na(WID_region))
+# )
 
@@
 ```{r}
-write.csv(df_regions, 
-      "PhD_pages/data_appendices/data/clean/WID_region_mapping.csv", row.names=FALSE)
+# write.csv(df_regions, 
+#       "data/clean/WID_region_mapping.csv", row.names=FALSE)
 
@@
 
+```{python}
+
+df_regions.to_csv('data/clean/WID_region_mapping.csv')
+
+```
 :::
@@
 
+Check we have population for each country
+```{python}
+# df_pop = pd.read_csv('data/clean/population.csv')
+
+# df_wid = pd.read_csv("data/clean/wid.csv")
+# df_pip = pd.read_csv("data/clean/pip.csv")
+
+# check_pip = pd.merge(df_pip, df_pop, how = 'left')
+# check_pip[check_pip['population'].isna()]
+
+# check_wid = pd.merge(df_wid, df_pop, how = 'left')
+# check_wid[check_wid['population'].isna()]
+
+# # There's no population for Zanzibar in the OWID data
+# # For now I make up a population series based on two data points from Wikipedia. At some point I can come back and replaces this with the WID population data
+
+# zanz_pop_1978 = 476111
+# zanz_pop_2022 = 1889773
+
+# # Assume a constant growth rate
+# zanz_g = pow(zanz_pop_2022/zanz_pop_1978, 1/(2022-1978))-1
+
+# year = 2022
+# pop = zanz_pop_1978 * pow(1+zanz_g, year - 1978)
+
+# df_zanz_pop = df_wid[df_wid['Entity'] == "Zanzibar"].copy()
+# df_zanz_pop['population'] = zanz_pop_1978 * pow(1+zanz_g, df_zanz_pop['Year'] - 1978)
+
+# df_zanz_pop = df_zanz_pop[['Entity', 'Year', 'population']]
+
+# df_pop = pd.concat([df_pop, df_zanz_pop])
+
+```
+
 Save the population data
 ```{python}
-# df_pop.to_csv('PhD_pages/data_appendices/data/clean/population.csv')
+# df_pop.to_csv('data/clean/population.csv')
 
 ```
+
 :::

### 2023-01-20 PhD_pages/data_appendices/1_B_within_country_trends.qmd
Adding to Paper 1 data analysis, shifting to python

@@
   - shinylive
+
 ---
@@
 df_regions = pd.read_csv("data/clean/WID_region_mapping.csv")
+df_pop = pd.read_csv("data/clean/population.csv")
 
@@
 # df_regions = pd.read_csv("PhD_pages/data_appendices/data/clean/WID_region_mapping.csv")
+# df_pop = pd.read_csv("PhD_pages/data_appendices/data/clean/population.csv")
 
@@
 
+  x_ref_val = f'{reference_col}{reference_vals[0]}'
+  y_ref_val = f'{reference_col}{reference_vals[1]}'
+
   # Apply outlier cut off, if specified
@@
   plot_data = pd.merge(plot_data, df_regions, how = 'left')
+
+  # Add in population data
+    # For first ref
+  df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[0], ['Entity', 'population'] ]
+
+  plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')
+
+  plot_data = plot_data.rename(columns={'population':'population_ref1'})
+
+    # For second ref
+  df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[1], ['Entity', 'population'] ]
+
+  plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')
+
+  plot_data = plot_data.rename(columns={'population':'population_ref2'})
+
+    # Calculate average population two reference periods
+  plot_data['avg_pop'] = (plot_data['population_ref1'] + plot_data['population_ref2'])/2
 
@@
   plot = (ggplot(plot_data
-  , aes(x_axis, y_axis))
-  + geom_point(aes(color=region_col))
+  , aes(x_axis, y_axis, alpha=0.5))
+  + geom_point(aes(color=region_col, size='avg_pop', alpha = 0.6))
   + facet_wrap('~ source')
-  + geom_polygon(aes(x='x', y='y'), data=shaded_coord, fill="#FF000066"))
+  + geom_polygon(aes(x='x', y='y', alpha = 0.2), data=shaded_coord)
+  + theme_light())
 
   # ---- Prep summary table ----
-
-  # Count by rise/stable/fall bins
 
@@
 
-  # Define the thresholds using the tolerance specified
-  bin_thresholds = [-float("inf"), -tolerance, tolerance, float("inf")]
-  bin_labels = ['fall', 'stable', 'rise']
+  group_yes_vars = ['source', region_col]
+  group_no_vars = ['source']
 
-  bins = pd.cut(plot_data['change'], bins=bin_thresholds, labels=bin_labels)
+  group_scenarios = [group_no_vars, group_yes_vars]
 
-  # By region:
+  aggs = []
 
-  # Count observations in bins defined by the thresholds
-  threshold_counts = plot_data\
-    .groupby(['source', region_col, bins])\
-    .size()\
-    .unstack()\
-    .unstack(level=region_col)\
-    .T\
-    .reset_index()\
-    .rename(columns={"change": "summary"})
+  for group_vars in group_scenarios:
 
-  # Calculate average change and n
-  mean_and_count = pd.DataFrame(plot_data.groupby(['source', region_col])['change']\
-    .agg(['mean', 'count']))\
-    .unstack()\
-    .T\
-    .reset_index()\
-    .rename(columns={"level_0": "summary"})
+  # Prepare pop-weighted average change
+    agg_level = plot_data.copy()
 
-  # rename the labels coming out of the aggregation
-  mean_and_count['summary'] = mean_and_count['summary'].replace(['mean', 'count'], ['avg change', 'n'])
+    # Calulate pop weights by source and region
+    agg_level['regional_pop_weights'] = agg_level.groupby(group_vars)['avg_pop'].transform(lambda x: x/x.sum())
 
-  # Concat together and present in a table
-  df_summary_region = pd.concat([threshold_counts, mean_and_count])
+    agg_level['global_pop_weights'] = agg_level.groupby('source')['avg_pop'].transform(lambda x: x/x.sum())
 
-  df_summary_region['summary'] = pd.Categorical(df_summary_region['summary'], categories=['fall', 'stable', 'rise','n', 'avg change'], ordered=True)
+    # Multiply the change by the pop weights 
+    agg_level['region_weighted_change'] = agg_level['change'] * agg_level['regional_pop_weights']
 
-  df_summary_region = df_summary_region.sort_values([region_col,'summary'])
+  # Prepare fall/stable/rise categories
+    agg_level['fall'] = agg_level['change'] < -tolerance
 
-  # Set the index
-  df_summary_region = df_summary_region.set_index([region_col,'summary'])
+    agg_level['stable'] = (agg_level['change'] <= tolerance) & (agg_level['change'] >= -tolerance)
+
+    agg_level['rise'] = agg_level['change'] > tolerance
+
+    agg_level[['fall', 'stable', 'rise']] = agg_level[['fall', 'stable', 'rise']].astype(int)
+
+    agg_level = agg_level.groupby(group_vars)[['change', 'region_weighted_change', 'global_pop_weights', 'fall', 'stable', 'rise']]\
+      .agg(['sum', 'mean', 'count'])
+
+  # Aggregate by source and region
+    if group_vars == group_yes_vars:
+
+      agg_level = agg_level\
+        .unstack().T.reset_index()
+
+    elif group_vars == group_no_vars:
+
+      agg_level = agg_level\
+        .T.reset_index()
+
+      agg_level[region_col] = "World"
+
+    aggs.append(agg_level)
+
+  df_summary = pd.concat(aggs)
+
+    # filter for the aggregations we need and label in the 'summary' column
+  df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'mean'), 'summary'] = 'avg change'
+
+  df_summary.loc[(df_summary['level_0'] == 'region_weighted_change') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Regional pop-weighted avg change'
+
+  df_summary.loc[(df_summary['level_0'] == 'global_pop_weights') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Global pop weights'
+
+  df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'count'), 'summary'] = 'n'
+
+  df_summary.loc[(df_summary['level_0'] == 'fall') & (df_summary['level_1'] == 'sum'), 'summary'] = 'fall'
+
+  df_summary.loc[(df_summary['level_0'] == 'stable') & (df_summary['level_1'] == 'sum'), 'summary'] = 'stable'
+
+  df_summary.loc[(df_summary['level_0'] == 'rise') & (df_summary['level_1'] == 'sum'), 'summary'] = 'rise'
+
+  df_summary = df_summary[df_summary['summary'].notnull()]
+
+  df_summary = df_summary.drop(['level_0', 'level_1'], axis = 1)
+
+  df_summary['summary'] = pd.Categorical(df_summary['summary'], categories=['fall', 'stable', 'rise','n', 'avg change', 'Regional pop-weighted avg change', 'Global pop weights'], ordered=True)
+
+  df_summary = df_summary.sort_values([region_col,'summary']).set_index([region_col,'summary'])
 
   # format to show appropriate d.p. I run a loop over each row for want of figuring out a better way!
-  repeat_length = 5
-  for i in range(0, len(df_summary_region.index.levels[0].unique())):
-    df_summary_region.iloc[i*repeat_length,:] = df_summary_region.iloc[i*repeat_length,:].map('{:.0f}'.format)
-    df_summary_region.iloc[1 + i*repeat_length,:] = df_summary_region.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
-    df_summary_region.iloc[2 + i*repeat_length,:] = df_summary_region.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
-    df_summary_region.iloc[3 + i*repeat_length,:] = df_summary_region.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
-    df_summary_region.iloc[4 + i*repeat_length,:] = df_summary_region.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)
-
-  # World summary
-
-  # Count observations in bins defined by the thresholds
-  threshold_counts = plot_data\
-    .groupby(['source', bins])\
-    .size()\
-    .unstack(level='source')\
-    .reset_index()\
-    .rename(columns={"change": "summary"})
-
-  # Calculate average change and n
-  mean_and_count = pd.DataFrame(plot_data.groupby(['source'])['change']\
-    .agg(['mean', 'count']))\
-    .T\
-    .reset_index()\
-    .rename(columns={"index": "summary"})
-
-  # rename the labels coming out of the aggregation
-  mean_and_count['summary'] = mean_and_count['summary'].replace(['mean', 'count'], ['avg change', 'n'])
-
-  # Concat together and present in a table
-  df_summary_world = pd.concat([threshold_counts, mean_and_count])
-
-  df_summary_world['summary'] = pd.Categorical(df_summary_world['summary'], categories=['fall', 'stable', 'rise','n', 'avg change'], ordered=True)
-
-  df_summary_world = df_summary_world.sort_values(['summary']).set_index('summary')
-
-  # format to show appropriate d.p. I do it row by row for want of figuring out a better way!
-  df_summary_world.iloc[0,:] = df_summary_world.iloc[0,:].map('{:.0f}'.format)
-  df_summary_world.iloc[1,:] = df_summary_world.iloc[1,:].map('{:.0f}'.format)
-  df_summary_world.iloc[2,:] = df_summary_world.iloc[2,:].map('{:.0f}'.format)
-  df_summary_world.iloc[3,:] = df_summary_world.iloc[3,:].map('{:.0f}'.format)
-  df_summary_world.iloc[4,:] = df_summary_world.iloc[4,:].map('{:.2f}'.format)
+  repeat_length = 7
+  for i in range(0, len(df_summary.index.levels[0].unique())):
+    df_summary.iloc[i*repeat_length,:] = df_summary.iloc[i*repeat_length,:].map('{:.0f}'.format)
+    df_summary.iloc[1 + i*repeat_length,:] = df_summary.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
+    df_summary.iloc[2 + i*repeat_length,:] = df_summary.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
+    df_summary.iloc[3 + i*repeat_length,:] = df_summary.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
+    df_summary.iloc[4 + i*repeat_length,:] = df_summary.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)
+    df_summary.iloc[5 + i*repeat_length,:] = df_summary.iloc[5 + i*repeat_length,:].map('{:.2f}'.format)
+    df_summary.iloc[6 + i*repeat_length,:] = df_summary.iloc[6 + i*repeat_length,:].map('{:.2f}'.format)
 
@@
     "plot": plot,
-    "summary_world": df_summary_world,
-    "summary_region": df_summary_region
+    "summary": df_summary
   }
@@
 ```{python}
+pd.set_option('display.max_rows', None)
 
+```
+
+## 1990-2017
+```{python}
 # Specifications
-reference_vals = [1990, 2018]
+reference_vals = [1990, 2017]
 max_dist_from_refs = [4, 4]
-min_dist_between = 20
-value_col = 'Bottom_50_share'
+min_dist_between = 23
+
+# These parameters are unlikely to change:
+reference_col = 'Year'
+group_by_col = 'Entity'
+
+```
+
+::: {.panel-tabset}
+## Gini
+
+```{python}
+# | echo: false
+# Specifications
+
+value_col = 'Gini'
 # 'Gini'
@@
 
+# Set the value which constitutes a significant change in the value being measured
+  # E.g. for Gini tolerance = 0.02 (2 points)
+
+tolerance = 0.02
+outlier_cut_off_upper = None
+
+```
+::: {.panel-tabset}
+## All countries
+```{python}
+# | echo: false
+
 # Data from both sources is needed?
@@
   # If = 2, then only countries with data for both sources will be included
-source_count_requirement = 2
-
-# Set the value which constitutes a significant change in the value being measured
-  # E.g. for Gini tolerance = 0.02 (2 points)
-
-if value_col == 'Gini':
-  tolerance = 0.02
-  outlier_cut_off_upper = None
-
-if value_col == 'Top_10_share':
-  tolerance = 1
-  outlier_cut_off_upper = None
-
-if value_col == 'Bottom_50_share':
-  tolerance = 1
-  outlier_cut_off_upper = None
-
-if value_col == 'P90_P10_ratio':
-  tolerance = 0.5
-  outlier_cut_off_upper = 40
-
-# These parameters are unlikely to change:
-reference_col = 'Year'
-group_by_col = 'Entity'
+source_count_requirement = 1
 
@@
 ```
-
+::: {.panel-tabset}
+## Plot
 ```{python}
@@
 
+## Summary
 ```{python}
@@
 
-:::{.column-page}
-```{python}
-
-plot_and_tables['table'].show()
-
-```
-
+## Data
 ```{python}
@@
 plot_data_results = plot_and_tables['table_data']
+plot_data_results
+```
+:::
+
+## Countries with data for both sources
+
+```{python}
+# | echo: false
+
+# Data from both sources is needed?
+  # If this = 1 then a country will be incldued even if it is only availalbe from one source.
+  # If = 2, then only countries with data for both sources will be included
+source_count_requirement = 2
+
+plot_and_tables = prep_plot_and_tables(
+  reference_vals = reference_vals, 
+  max_dist_from_refs = max_dist_from_refs,
+  min_dist_between = min_dist_between,
+  reference_col = reference_col,
+  group_by_col = group_by_col,
+  value_col = value_col,
+  tolerance = tolerance,
+  outlier_cut_off_upper = outlier_cut_off_upper,
+  source_count_requirement = source_count_requirement
+  )
 
 ```
+::: {.panel-tabset}
+## Plot
+```{python}
+plot_and_tables['plot']
+```
+
+## Summary
+```{python}
+plot_and_tables['summary']
+```
+
+## Data
+```{python}
+
+plot_data_results = plot_and_tables['table_data']
+plot_data_results
+```
+:::
+:::
+
+## Top 10% share
+
+```{python}
+# | echo: false
+# Specifications
+
+value_col = 'Top_10_share'
+
+# Set the value which constitutes a significant change in the value being measured
+  # E.g. for Gini tolerance = 0.02 (2 points)
+
+tolerance = 1
+outlier_cut_off_upper = None
+
+```
+::: {.panel-tabset}
+## All countries
+```{python}
+# | echo: false
+
+# Data from both sources is needed?
+  # If this = 1 then a country will be incldued even if it is only availalbe from one source.
+  # If = 2, then only countries with data for both sources will be included
+source_count_requirement = 1
+
+plot_and_tables = prep_plot_and_tables(
+  reference_vals = reference_vals, 
+  max_dist_from_refs = max_dist_from_refs,
+  min_dist_between = min_dist_between,
+  reference_col = reference_col,
+  group_by_col = group_by_col,
+  value_col = value_col,
+  tolerance = tolerance,
+  outlier_cut_off_upper = outlier_cut_off_upper,
+  source_count_requirement = source_count_requirement
+  )
+
+```
+::: {.panel-tabset}
+## Plot
+```{python}
+plot_and_tables['plot']
+```
+
+## Summary
+```{python}
+plot_and_tables['summary']
+```
+
+## Data
+```{python}
+
+plot_data_results = plot_and_tables['table_data']
+plot_data_results
+```
+:::
+
+## Countries with data for both sources
+
+```{python}
+# | echo: false
+
+# Data from both sources is needed?
+  # If this = 1 then a country will be incldued even if it is only availalbe from one source.
+  # If = 2, then only countries with data for both sources will be included
+source_count_requirement = 2
+
+plot_and_tables = prep_plot_and_tables(
+  reference_vals = reference_vals, 
+  max_dist_from_refs = max_dist_from_refs,
+  min_dist_between = min_dist_between,
+  reference_col = reference_col,
+  group_by_col = group_by_col,
+  value_col = value_col,
+  tolerance = tolerance,
+  outlier_cut_off_upper = outlier_cut_off_upper,
+  source_count_requirement = source_count_requirement
+  )
+
+```
+::: {.panel-tabset}
+## Plot
+```{python}
+plot_and_tables['plot']
+```
+
+## Summary
+```{python}
+plot_and_tables['summary']
+```
+
+## Data
+```{python}
+
+plot_data_results = plot_and_tables['table_data']
+plot_data_results
+```
+:::
+
+:::
+
+Next measure
+
 :::

### 2023-01-20 PhD_pages/paper_pages/paper1.qmd
Adding to Paper 1 data analysis, shifting to python

--- 2022-10-11 PhD_pages/paper_pages/paper1.qmd
@@
 title: "Paper 1: Triangulating trends in inequality around the world from secondary databases"
+format: html
 ---
 
-```{r, echo=FALSE, message = FALSE, warning=FALSE}
+::: {.callout-note collapse="true" appearance="minimal"}
+### Load packages
 
+```{r}
 library(tidyverse)
-library(trackdown)
+library(knitr)
+library(data.table)
 
 ```
+:::
 
@@
 
-### Other sources
-
-Should we use Solt? As another check on results?
-
-What about GCIP? For trying to flip between income and consumption...
-
 ### Coverage
 
-s
+Give an overview of coverage in each dataset.
 
-## Methods: Comparing databases
+## Comparing databases
 
-### Lining up years
+### Comparing metrics
 
-Method 1: Reference years
+Does it make a difference whether we look at Gini or top 10%?
 
-Method 2: Fully extrapolating/interpolating
+People worry about whether Gini masks what's happening at the top.
 
-The equivalent of the reference year approach where there is no max threshold gap.
+Empirically: 
+    - Make a linear model of top 10% on Gini, for each source separately
+    - Make predicted top 10%
+    - Take a look at the R2
+    - Investigate individual countries with very poor predictions
 
-Method 3: For the time trend, use a regression with country fixed effects to allow for entry and exit from the sample.
+Sub-Saharan Africa: example where the Gini fell in more cases than the top 10%.
 
-### Comparing welfare concepts
+### Since 1990
 
-Getting a *global* view of inequalities within countries faces a Catch-22 situation: If you want broad coverage you get very heterogeneous data; if you want comparable data, or to be able to consider different welfare concepts consistently, the coverage you get is much smaller.
+Similarities:
 
-Here I'll take two avenues.
+Convergence – Smaller rises in coutries that started out with high inequality. Driven regionally. (We could do a test for this.)
 
-Firstly, what we're going for with the 'triangulation' approach is dominance - if different welfare metrics agree
+Indeed for some regions very similar picture:
 
-Second we will cross reference to attempts to 'standardise' across these differences: using Solt and GCIP.
-How valid that is I don't know.
-But taking the two approaches together we get as close as we can to robust observations.
+    - MENA - Roughly 2 point fall, both Gini and top 10%. Only one country with a rise: Not total agreeement on coverage (one country either side)
 
-### A range of inequality metrics
+    - South and South East Asia – Very similar not weighting by population. But wildly different when weighting by population, due to the much much stronger rise
 
-### Lorenz curves
+    -  Sub-Saharan Afric – Very similar across sources.
+Top 10: Roughly even mix of rises and falling countries, but with a 3 point fall on average.
+Gini: More falls, 5 point on average.
+(Mention data quality issues: e.g.
+Kenya big fall.
+Lesoto big fall.
+But how comparable are the surveys??
+Same surveys underlie WID...)
 
-Dominance implies any (well-behaved) summary measure of inequality will agree on the ranking.
+    - Western Europe – Rises in a majority of countries, but relatively moderate of 1-2 points (Whether Gini of top 10).
 
-Operationalize this by taking 99 points on the Lorenze curve: the percentiles.
+Differences:
 
-This leaves out the *very* top.
-But this would be quite a warped conception of inequality.
-I.e.
-if you have two distributions, A and B, and the share of income received by every whole percentile threshold in B is higher than A, but the Lorenze curves crossed within the top percentile - to claim that inequality is higher in B you'd need a very extreme conception of inequality.
+Verry different bottom line at the global level:
 
-# Results
+- PIP: roughly equal numbers falling and rising.
+0 average change – though a modest rise when pop weighted (2 points).
+Same whether Gini or Top 10.
+That's without LAtin America (Which WID is missing).
+Includung LatAm tips nudges it (with about a 1 point fall unweighter and 1 point rise weighted)
 
-### Outliers
-
+- WID: Far more countries with rising inequality.
+On average a 3 point rise (whether top 10 or Gini) – [Better to express these relative to the level of ineq.
+i.e.
+% increase rather than % point increase?
+A point rise means 'less' in WID terms because inequality levels are much higher.] But the most stark difference comes when we weight by population.
+Some key populous countries: India, China, US?, Indonesia?
 Russia
 
+That difference is driven by particular regions:
+-
+
+More countries with rises. LArger rises. Especially larger populations. Average view very different:
+
+Countries where the sources paint a radically different picture that affect the aggregate results:
+
+- India: Insane rise in WID 34->56
+- China:
+- Russia: Massive drop in PIP. Insane rise in WID.
+- USA: Much larger rise in WID.
+
+What explains the different trends in these countries? (We can maybe look in more detail, at different welfare concepts etc.)
+
+It sometimes goes the other way: Indonesia – Rise in PIP, fall in WID.
+
+:::
+
+#### Method 2: Use a regression to obtain the time trend
+
+### Results for a range of inequality metrics
+
+::: {.callout-note collapse="true" appearance="minimal"} 
+## Calculations for comparisons
+
+## Explaining differences between PIP and WID data
+
+### For a subset of countries we can bridge between the PIP and WID data
+
+Besides the underlying data sources, there are many differences between what is being measured in the PIP and WID data - notably the welfare concept and unit of analysis/equivalization.
+
+In order to get a better understanding of where differences between the datasets arise, for a subset of countries we can bridge between the two somewhat. For example:
+
+-   WID has post-tax income series for some countries
+-   We can use LIS to look at how changing the welfare concept, unit of analysis, equivalisation affects things.
+(For equivalisation, we could also possibly use the newly available PIP harmonized microdata).
+
+### More in-depth discussion of particular outliers
+
+-   Russia is a very notable case where the two datasets tell very different stories.
+
 # Conclusion

### 2023-01-20 presentations/DSPI_CSP_masters_Jan2023.qmd
Adding to Paper 1 data analysis, shifting to python

---
title: "From $1.90 to $2.15: the updated International Poverty Line"
author: "Joe Hasell"
format: revealjs
---

## The International Poverty Line {.smaller}
:::: {.columns}

::: {.column width="30%"}
Left column

- some point
- another point
:::

::: {.column width="70%"}
```{=html}
<iframe src="https://ourworldindata.org/grapher/share-of-population-in-extreme-poverty?country=BGD~BOL~MDG~IND~CHN~ETH" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>
```

:::

::::

## The IPL has been updated from $1.90 to $2.15

## Basic story {.smaller}

:::: {.columns}

::: {.column width="30%"}

- An adjustment for inflation
- $1 in 2017 prices < $1 in 2011 prices
- The value of the IPL has, broadly speaking, been maintined
- Global extreme poverty very similar measured in either terms.
:::

::: {.column width="70%"}

```{=html}
<iframe src="https://ourworldindata.org/explorers/poverty-explorer-2011-vs-2017-ppp?time=2013..latest&hideControls=true&Metric=Share+in+poverty&International-%24=Compare+2017+and+2011+prices&Poverty+line=International+Poverty+Line&Household+survey+data+type=Show+data+from+both+income+and+expenditure+surveys&country=~OWID_WRL" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>
```

:::

::::

### 2023-01-22 presentations/DSPI_CSP_masters_Jan2023/DSPI_CSP_masters_Jan2023.qmd
Adding DSPI presentation revealjs slides

---
title: "From $1.90 to $2.15: the updated International Poverty Line"
author: "Joe Hasell"
format:
  revealjs:
    scrollable: true
---

## The International Poverty Line {.smaller}
:::: {.columns}

::: {.column width="30%"}

::: {.incremental}
- Set by the World Bank
- Reflects definitions of poverty adopted in the **world's poorest countries**.
- Adopted by the UN as the first SDG – *Indicator 1.1.1*

- **Recently updated from $1.90 to $2.15**
:::

:::

::: {.column width="70%"}
```{=html}
<iframe src="https://ourworldindata.org/grapher/share-of-population-in-extreme-poverty?country=BGD~BOL~MDG~IND~CHN~ETH" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>
```

:::

::::

## Outline

1. Why has the IPL been updated?
2. How does this affect our understanding of global poverty?

::: {.incremental}
- **Basic picture:** 
    Just a change of units, Limited impact on global estimates
- **Major caveat I:** Updated PPPs, Substantial changes for individual countries
- **Major caveat II:** 
    Interaction with how WB sets the IPL. The real value of higher WB poverty lines *did* change.

:::

# Basic story
A change of units

## A change of units {.smaller}

:::: {.columns}

::: {.column width="30%"}

::: {.incremental}

- International-$ at **2011** -> **2017** prices
- Change reflects inflation: $1 in 2017 prices < $1 in 2011 prices
- The *value* of the IPL (quantity of goods and services it buys) has, broadly speaking, been maintined

:::

:::

::: {.column width="70%"}

:::

::::

## A change of units {.smaller}

:::: {.columns}

::: {.column width="30%"}

- International-$ at **2011** -> **2017** prices
- Change reflects inflation: $1 in 2017 prices < $1 in 2011 prices
- The *value* of the IPL (quantity of goods and services it buys) has, broadly speaking, been maintined

- Global extreme poverty very similar measured in either terms.

:::

::: {.column width="70%"}

```{=html}
<iframe src="https://ourworldindata.org/explorers/poverty-explorer-2011-vs-2017-ppp?time=2013..latest&hideControls=true&Metric=Share+in+poverty&International-%24=Compare+2017+and+2011+prices&Poverty+line=International+Poverty+Line&Household+survey+data+type=Show+data+from+both+income+and+expenditure+surveys&country=~OWID_WRL" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>
```

:::

::::

<!-- xxxxxxxxxxxxxx -->
# Major caveat I
Country-level data

## Why *not* update prices? {.smaller}

:::: {.columns}

::: {.column width="30%"}

Flip the question: Why did the World Bank *not* update it's prices sooner?

:::

::: {.column width="70%"}

:::

::::

## Why *not* update prices? {.smaller}

:::: {.columns}

::: {.column width="30%"}

Flip the question: Why did the World Bank *not* update it's prices sooner?

Short answer: They were asked not to by the *Atkinson Commission (2017)*.

:::

::: {.column width="70%"}

:::

::::

## Why *not* update prices? {.smaller}

:::: {.columns}

::: {.column width="30%"}

Flip the question: Why did the World Bank *not* update it's prices sooner?

Short answer: They were asked not to by the *Atkinson Commission (2017)*.

This time round the update in prices had a very small effect on global poverty estimates.

But in previous updates the changes were much more substantial.

:::

::: {.column width="70%"}

![Global poverty fell by half on a Tuesday in 2014...](CGDEV_screenshot.png)

:::

::::

## Understanding international-${.smaller}

- Converting local currencies to international-$ means making **two** price adjustments:
  - inflation within individual countries
  - differences in the cost of living between countries

## Understanding international-$ {.smaller}

- Converting local currencies to international-$ means making **two** price adjustments:
  - inflation within individual countries
  - differences in the cost of living between countries

- To make the second adjustment the WB uses Purchasing Power Parity rates
  - A snapshot of price difference across coutries in a particular year
  - Data gathered by the ICP in 'rounds': 2017, 2011, 2005...
  - US is the 'numeraire': 
    - 1 int-$ in 2017 prices is the *quantity of goods and services that 1 USD could buy in the US in 2017.*
    - The real value of 2017 vs 2011 int-$ fell by the same amount that prices increased in the US

## Understanding international-$ {.smaller}

- Converting local currencies to international-$ means making **two** price adjustments:
  - inflation within individual countries
  - differences in the cost of living between countries

- To make the second adjustment the WB uses Purchasing Power Parity rates
  - A snapshot of price difference across coutries in a particular year
  - Data gathered by the ICP in 'rounds': 2017, 2011, 2005...
  - US is the 'numeraire': 
    - 1 int-$ in 2017 prices is the *quantity of goods and services that 1 USD could buy in the US in 2017.*
    - The real value of 2017 vs 2011 int-$ fell by the same amount that prices increased in the US

- Note that *'in 2017 prices'* thus has a **dual** meaning:
  - the base year to which the within-country inflation adjustment is made
  - the ICP round – i.e. the year cross-country price differences were observed.

## PPP 'shocks' {.smaller}

When PPPs from a new ICP round are released, our view of relative prices between coutries can change *a lot*.

I think of this as like a 'shock' to our understanding about the world.

With the new PPPs we update our view of how prices compare between countries.

## PPP 'shocks' {.smaller}

:::: {.columns}

::: {.column width="30%"}

Crucially – as our view of relative **prices** change, so does our view of relative **incomes**.

:::

::: {.column width="70%"}

```{=html}
<iframe src="https://ourworldindata.org/explorers/poverty-explorer-2011-vs-2017-ppp?yScale=log&hideControls=true&Metric=Mean+income+or+consumption&International-%24=Compare+2017+and+2011+prices&Poverty+line=International+Poverty+Line&Household+survey+data+type=Show+data+from+both+income+and+consumption+surveys&country=AGO~NGA~VNM~USA~BRA~GHA" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>
```

:::

::::

## Why do these PPP 'shocks' arise? {.smaller}

::: {.incremental}

- Changes in ICP methodology, quality/availability of price data
- Genuine differences in how prices have changed in different countries (i.e. different inflation rates)
  - BUT! These should roughly be accouted for by the inflation adjustment bit of converting local currencies to international-$!
  - BUT^2^!! It's part of the arithmetic construction of PPPs that a change in the relative PPP rate between two countries needn't exactly map on to their relative rates of inflation (this is needed to provide 'multilateral' PPP rates – i.e. transitivity)
:::

## PPP 'shocks': impact on poverty estimates {.smaller}

:::: {.columns}

::: {.column width="30%"}

As our view of relative **prices** and **incomes** is updated with a new set of PPPs – so is our view of the extent of **poverty** in each country

In some extreme cases, poverty rates look very different when measured in 2017 and 2011 international-$.

:::

::: {.column width="70%"}

```{=html}
<iframe src="https://ourworldindata.org/explorers/poverty-explorer-2011-vs-2017-ppp?hideControls=true&Metric=Share+in+poverty&International-%24=Compare+2017+and+2011+prices&Poverty+line=International+Poverty+Line&Household+survey+data+type=Show+data+from+both+income+and+consumption+surveys&country=AGO~NGA~VNM~USA~BRA~GHA" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>
```

:::

::::

## PPP 'shocks': impact on poverty estimates {.smaller}

The shift from 2011 to 2017 PPPs makes such a little difference to global poverty estimates because:
  - the country-level shocks cancel out to some extent 
  - the shocks were less major than in previous rounds

<!-- xxxxxxxxxxxxxx -->

# Major caveat II
Setting the World Bank poverty lines

## Setting the IPL {.smaller}

:::: {.columns}

::: {.column width="30%"}

The World Bank sets the IPL to reflect the poverty lines adopted in the world's poorest countries.

More specifically, it is set at the **median** poverty line among 'low-income' countries (as per World Bank's classification).

In 2017 prices, that gives the new IPL value of $2.15 per day.

:::

::: {.column width="70%"}

```{=html}
<iframe src="https://ourworldindata.org/grapher/national-poverty-line-vs-gdp-per-capita" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>
```

:::

::::

## Setting the IPL {.smaller}

With all the chopping and changing of relative price levels, the real value of the IPL *could have* changed with the new PPPs – if the median value of low-income countries' poverty lines had changed.

It's sort of by coincidence that didn't happen!

The median poverty line increased by roughly the same amount (13%) that the value of the international $ fell (9% – the amount of inflation in the US between 2011 and 2017).

It's only this conincidence that meant the real value of the IPL was maintained.

## Higher World Bank poverty lines {.smaller}

:::: {.columns}

::: {.column width="30%"}

The World Bank also sets two higher poverty lines of **$3.65** and **$6.85** – the median poverty line among *lower-* and *upper-middle income* countries respectively.

In shifting to 2017 PPPs, these were updated from $3.20 and $5.50 when calculated with 2011 PPPs (back in 2016).
:::

::: {.column width="70%"}

```{=html}
<iframe src="https://ourworldindata.org/grapher/national-poverty-line-vs-gdp-per-capita" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>
```

:::

::::

## Higher World Bank poverty lines {.smaller}

:::: {.columns}

::: {.column width="30%"}

For the upper-middle income poverty line – which rose from $5.50 to $6.85 (by 25%) – the 'happy coincidence' didn't happen.

Global poverty estimates for this poverty line changed considerably across the change in units because – unlike the IPL – the *real value* of this poverty line changed too.

:::

::: {.column width="70%"}

```{=html}
<iframe src="https://ourworldindata.org/explorers/poverty-explorer-2011-vs-2017-ppp?hideControls=true&Metric=Share+in+poverty&International-%24=Compare+2017+and+2011+prices&Poverty+line=Upper-middle+income+poverty+line&Household+survey+data+type=Show+data+from+both+income+and+expenditure+surveys&country=~OWID_WRL" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>
```

:::

::::

## Summary {.smaller}

- The real value of the IPL didnt change (though that's not true for higher line)
- Unsurprisingly, global poverty rates look very similar.
- But this is partly due to some 'cancelling out' across countries
- The new PPPs give us a revised understanding of how incomes – and hence poverty levels – compare in different countries.
- The shift from 2011 to 2017 PPPs is, however, was much less of a 'shock' than earlier ICP rounds.

### 2023-01-22 PhD_pages/data_appendices/1_B_within_country_trends.qmd
Moving code to shinylive app

@@
 import pandas as pd
+import numpy as np
+
+#  ---------------------------------
+#  ---------- Data prep functions -------------
+#  ----------------------------------
+
+#  Function get matching for ref years
+def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):
+
+  df = df.loc[:, [reference_col, group_by_col, value_col]]
+
+  # Drop NAs
+  df = df.dropna()
+
+  # Calculate absolute distance from reference value
+  df['ref_diff'] = abs(df[reference_col] - reference_val)
+
+  # Drop any rows with a distance beyond threshold
+  if not pd.isna(max_dist_from_ref):
+    df = df.loc[df['ref_diff'] <= max_dist_from_ref]
+
+  # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
+  df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)
+
+  # Settle tie-breaks
+  if tie_break == 'below':
+    df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)
+
+  elif tie_break == 'above':
+    df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)
+
+  df = df.drop('ref_diff', axis=1)
+
+  return df
+
+# Merge matches for different reference points
+def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+
+# Make sure the pair of reference values are in ascending order
+  reference_vals.sort()
+
+# Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref
+
+# Find matches for lower reference value
+  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs, reference_col, group_by_col, value_col, 'below')
+
+# Find matches for higher reference value
+  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs, reference_col, group_by_col, value_col, 'above')
+
+# Merge the two sets of matches
+  merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))
+
+# Drop obs that do not have data for both ref values
+  merged_df = merged_df.dropna()
+
+# Drop obs where the matched data does not meet the min distance requirement
+  if not pd.isna(min_dist_between):
+
+  # Store the names of the reference column returned from the two matches
+    ref_var_high = f'{reference_col}{reference_vals[1]}'
+    ref_var_low = f'{reference_col}{reference_vals[0]}'
+
+  # Keep only rows >= to the min distance
+    merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]
+
+  return merged_df
+
+  # For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
+def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+
+    # Specify the name of the column in which the income/consumption welfare definition is stored
+    welfare_colname = 'welfare_type'
+
+    # Creat dataframes for thee scenarios:
+    # Scenario 1: only allow income data
+    df_inc_filter = df.loc[df[welfare_colname] == "income", :]
+    df_inc_filter.name = "Income"
+
+    # Scenario 2: only allow consumption data
+    df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
+    df_cons_filter.name = "Consumption"
+    # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
+    df_mixed = df.copy()
+
+    df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')
+
+    df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]
+
+    df_mixed.name = "Mixed"
+    #  Store the scneario dataframes in a list
+    df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]
+
+    # Run the matching function on each scenario
+    scenario_matches = [merge_two_ref_matches(
+      df_scenario, 
+      reference_vals, 
+      max_dist_from_refs, 
+      min_dist_between, 
+      reference_col, 
+      group_by_col, 
+      value_col) for df_scenario in df_scenarios]
+
+    # Combine the first two scenarios.
+    df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])
+
+    # Tidy up indexes
+    df_combined_matches = df_combined_matches.reset_index()
+
+    df_combined_matches = df_combined_matches.drop('level_1', axis=1)
+
+    df_combined_matches = df_combined_matches\
+      .rename(columns={"level_0": "pip_welfare"})
+
+    # Add in third scenario.
+    df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])
+
+    # add scenario name to te pip_welfare column
+    df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)
+
+    # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
+      # First count the matches
+    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
+      # Then drop any matches from the lowest priority where there are multiple matches
+    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
+      #  Repeat at the next level of priority
+    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
+    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]
+
+    # Drop the match count column
+    df_combined_matches = df_combined_matches.drop('match_count', axis=1)
+
+    return df_combined_matches
+
+#  ---------------------------------
+#  ---------- Specifications -------------
+#  ----------------------------------
+
+# These parameters are not controlled in the app.
+reference_col = 'Year'
+group_by_col = 'Entity'
+
+tolerance = 1
+outlier_cut_off_upper = None
+
+source_count_requirement = 2
+
+#  ---------------------------------
+#  ---------- App code -------------
+#  ----------------------------------
+
+def panel_box(*args, **kwargs):
+    return ui.div(
+        ui.div(*args, class_="card-body"),
+        **kwargs,
+        class_="card mb-3",
+    )
 
 app_ui = ui.page_fluid(
-    ui.output_plot("example_plot"),
+    {"class": "p-4"},
+    ui.row(
+        ui.column(
+            4,
+            panel_box(
+                ui.input_slider("reference_vals", "Reference years", 1975, 2020, value=[1990,2018]),
+                ui.input_numeric(
+                            "max_dist_from_refs",
+                            "=/- (years)",
+                            4,
+                        ),
+                ui.input_numeric(
+                            "min_dist_between",
+                            "Min distance between observations (years)",
+                            23,
+                        ),
+                ui.input_action_button(
+                    "download", "Download data and image", class_="btn-primary w-100"
+                ),
+            ),
+            ui.navset_tab_card(
+                ui.nav(
+                    "Compare sources",
+                    ui.input_select(
+                "metric",
+                "Metric",
+                {
+                    "Gini": "Gini",
+                    "Top_10_share": "Top 10p.c. share",
+                },
+            ),
+            ),
+                ui.nav(
+                  "Compare metrics"),
+            ),
+        ),
+        ui.column(
+            8,
+            ui.output_plot("example_plot")
+            ),
+    ),
 )
@@
 
+      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'
+
+      df_pip = pd.read_csv(open_url(url))
+
+      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'
+
+      df_wid = pd.read_csv(open_url(url))
+
+      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/WID_region_mapping.csv'
+
+      df_regions = pd.read_csv(open_url(url))
+
+      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'
+
+      df_pop = pd.read_csv(open_url(url))
+
+      def prep_wid_pip_data(reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+
+        pip_matches = pip_welfare_routine(
+          df = df_pip,
+          reference_vals = reference_vals,
+          max_dist_from_refs = max_dist_from_refs,
+          min_dist_between = min_dist_between,
+          reference_col = reference_col,
+          group_by_col = group_by_col,
+          value_col = value_col
+        )
+
+        wid_matches = merge_two_ref_matches(
+          df = df_wid,
+          reference_vals = reference_vals,
+          max_dist_from_refs = max_dist_from_refs,
+          min_dist_between = min_dist_between,
+          reference_col = reference_col,
+          group_by_col = group_by_col,
+          value_col = value_col
+        )
+
+        ref_pairs = pd.concat([pip_matches, wid_matches,], keys=['pip', 'wid'])
+
+        # Tidy up indexes
+        ref_pairs = ref_pairs.reset_index()
+
+        ref_pairs = ref_pairs.drop('level_1', axis=1)
+
+        ref_pairs = ref_pairs\
+          .rename(columns={"level_0": "source"})
+
+        return ref_pairs
+
+      def prep_plot_and_tables(
+        reference_vals, 
+        max_dist_from_refs,
+        min_dist_between,
+        reference_col,
+        group_by_col,
+        value_col,
+        tolerance,
+        outlier_cut_off_upper,
+        source_count_requirement):
+
+        # Other spec:
+        region_col = 'region_alt'
+
+        # ---- Prep data ---------
+
+        plot_data = prep_wid_pip_data(
+          reference_vals = reference_vals, 
+          max_dist_from_refs = max_dist_from_refs,
+          min_dist_between = min_dist_between,
+          reference_col = reference_col,
+          group_by_col = group_by_col,
+          value_col = value_col,
+        )
+
+        # Store the names of the columns to be used onthe X and Y axis
+        x_axis = f'{value_col}{reference_vals[0]}'
+        y_axis = f'{value_col}{reference_vals[1]}'
+
+        x_ref_val = f'{reference_col}{reference_vals[0]}'
+        y_ref_val = f'{reference_col}{reference_vals[1]}'
+
+        # Apply outlier cut off, if specified
+        if not pd.isna(outlier_cut_off_upper):
+          plot_data = plot_data.loc[plot_data[x_axis] <= outlier_cut_off_upper]
+          plot_data = plot_data.loc[plot_data[y_axis] <= outlier_cut_off_upper]
+
+        # Add a count by country – showing whether data is available from both sources or not
+        plot_data['source_count'] = plot_data.groupby(group_by_col)['source'].transform('count')
+
+        # Apply source requirement (whether to include only observations with data from both sources)
+        plot_data = plot_data.loc[plot_data['source_count'] >= source_count_requirement]
+
+        # Drop source_count column
+        plot_data = plot_data.drop('source_count', axis=1)
+
+        # Add in region classification
+        plot_data = pd.merge(plot_data, df_regions, how = 'left')
+
+        # Add in population data
+          # For first ref
+        df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[0], ['Entity', 'population'] ]
+
+        plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')
+
+        plot_data = plot_data.rename(columns={'population':'population_ref1'})
+
+          # For second ref
+        df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[1], ['Entity', 'population'] ]
+
+        plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')
+
+        plot_data = plot_data.rename(columns={'population':'population_ref2'})
+
+          # Calculate average population two reference periods
+        plot_data['avg_pop'] = (plot_data['population_ref1'] + plot_data['population_ref2'])/2
+
+        # ----- Prep plot ------
+
+        # I grab what I am guessing to be the max and min tick marks on the x axist, in order to define the coordinates of a 'tolerance' ribbon.
+        x_min = plot_data[x_axis].min()
+        x_min_floor = np.floor(x_min * 10) / 10
+
+        x_max = plot_data[x_axis].max()
+        x_max_ceiling = np.ceil(x_max * 10) / 10
+
+        # Set the coordinates of the shaded ribbon
+        shaded_coord = pd.DataFrame({'x': [x_min_floor, x_max_ceiling, x_max_ceiling, x_min_floor], 
+          'y': [x_min_floor-tolerance, x_max_ceiling-tolerance, x_max_ceiling+tolerance , x_min_floor+tolerance]})
+
+        plot = (ggplot(plot_data
+        , aes(x_axis, y_axis, alpha=0.5))
+        + geom_point(aes(color=region_col, size='avg_pop', alpha = 0.6))
+        + facet_wrap('~ source')
+        + geom_polygon(aes(x='x', y='y', alpha = 0.2), data=shaded_coord)
+        + theme_light()
+        + theme(legend_position='none')
+        )
+
+        # ---- Prep summary table ----
+
+        # Calculate the change between the two ref periods
+        plot_data['change'] = (plot_data[y_axis] - plot_data[x_axis])
+
+        group_yes_vars = ['source', region_col]
+        group_no_vars = ['source']
+
+        group_scenarios = [group_no_vars, group_yes_vars]
+
+        aggs = []
+
+        for group_vars in group_scenarios:
+
+        # Prepare pop-weighted average change
+          agg_level = plot_data.copy()
+
+          # Calulate pop weights by source and region
+          agg_level['regional_pop_weights'] = agg_level.groupby(group_vars)['avg_pop'].transform(lambda x: x/x.sum())
+
+          agg_level['global_pop_weights'] = agg_level.groupby('source')['avg_pop'].transform(lambda x: x/x.sum())
+
+          # Multiply the change by the pop weights 
+          agg_level['region_weighted_change'] = agg_level['change'] * agg_level['regional_pop_weights']
+
+        # Prepare fall/stable/rise categories
+          agg_level['fall'] = agg_level['change'] < -tolerance
+
+          agg_level['stable'] = (agg_level['change'] <= tolerance) & (agg_level['change'] >= -tolerance)
+
+          agg_level['rise'] = agg_level['change'] > tolerance
+
+          agg_level[['fall', 'stable', 'rise']] = agg_level[['fall', 'stable', 'rise']].astype(int)
+
+          agg_level = agg_level.groupby(group_vars)[['change', 'region_weighted_change', 'global_pop_weights', 'fall', 'stable', 'rise']]\
+            .agg(['sum', 'mean', 'count'])
+
+        # Aggregate by source and region
+          if group_vars == group_yes_vars:
+
+            agg_level = agg_level\
+              .unstack().T.reset_index()
+
+          elif group_vars == group_no_vars:
+
+            agg_level = agg_level\
+              .T.reset_index()
+
+            agg_level[region_col] = "World"
+
+          aggs.append(agg_level)
+
+        df_summary = pd.concat(aggs)
+
+          # filter for the aggregations we need and label in the 'summary' column
+        df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'mean'), 'summary'] = 'avg change'
+
+        df_summary.loc[(df_summary['level_0'] == 'region_weighted_change') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Regional pop-weighted avg change'
+
+        df_summary.loc[(df_summary['level_0'] == 'global_pop_weights') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Global pop weights'
+
+        df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'count'), 'summary'] = 'n'
+
+        df_summary.loc[(df_summary['level_0'] == 'fall') & (df_summary['level_1'] == 'sum'), 'summary'] = 'fall'
+
+        df_summary.loc[(df_summary['level_0'] == 'stable') & (df_summary['level_1'] == 'sum'), 'summary'] = 'stable'
+
+        df_summary.loc[(df_summary['level_0'] == 'rise') & (df_summary['level_1'] == 'sum'), 'summary'] = 'rise'
+
+        df_summary = df_summary[df_summary['summary'].notnull()]
+
+        df_summary = df_summary.drop(['level_0', 'level_1'], axis = 1)
+
+        df_summary['summary'] = pd.Categorical(df_summary['summary'], categories=['fall', 'stable', 'rise','n', 'avg change', 'Regional pop-weighted avg change', 'Global pop weights'], ordered=True)
+
+        df_summary = df_summary.sort_values([region_col,'summary']).set_index([region_col,'summary'])
+
+        # format to show appropriate d.p. I run a loop over each row for want of figuring out a better way!
+        repeat_length = 7
+        for i in range(0, len(df_summary.index.levels[0].unique())):
+          df_summary.iloc[i*repeat_length,:] = df_summary.iloc[i*repeat_length,:].map('{:.0f}'.format)
+          df_summary.iloc[1 + i*repeat_length,:] = df_summary.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
+          df_summary.iloc[2 + i*repeat_length,:] = df_summary.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
+          df_summary.iloc[3 + i*repeat_length,:] = df_summary.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
+          df_summary.iloc[4 + i*repeat_length,:] = df_summary.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)
+          df_summary.iloc[5 + i*repeat_length,:] = df_summary.iloc[5 + i*repeat_length,:].map('{:.2f}'.format)
+          df_summary.iloc[6 + i*repeat_length,:] = df_summary.iloc[6 + i*repeat_length,:].map('{:.2f}'.format)
+
+        # ----- Return outputs ------
+        plot_and_tables = {
+          "table_data": plot_data,
+          "plot": plot,
+          "summary": df_summary
+        }
+
+        return plot_and_tables
+
+      plot_and_tables = prep_plot_and_tables(
+          reference_vals = list(input.reference_vals()), 
+          max_dist_from_refs = input.max_dist_from_refs(),
+          min_dist_between = input.min_dist_between(),
+          reference_col = reference_col,
+          group_by_col = group_by_col,
+          value_col = input.metric(),
+          tolerance = tolerance,
+          outlier_cut_off_upper = outlier_cut_off_upper,
+          source_count_requirement = source_count_requirement
+          )
+
+      plot = plot_and_tables['plot']
+
+      return plot
+
+app = App(app_ui, server)
+
+```
+
+```{shinylive-python}
+#| standalone: true
+#| viewerHeight: 420
+
+from shiny import *
+from plotnine import *
+from pyodide.http import open_url
+import pandas as pd
+
+app_ui = ui.page_fluid(
+    ui.output_plot("example_plot"),
+)
+
+def server(input, output, session):
+    @output
+    @render.plot
+    def example_plot():
+
         url = 'https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/JoeHasell/PIP/data/ppp_2017/final/PIP_data_public_download/full_dataset/cons_only/poverty_cons_only.csv'
@@
 ```
+
+```{python}
+
+      # #  -----------------------------------
+      # #  ------- Data prep functions -------
+      # #  -----------------------------------
+
+      # #  Function get matching for ref years
+      # def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):
+
+      #     df = df.loc[:, [reference_col, group_by_col, value_col]]
+
+      #     # Drop NAs
+      #     df = df.dropna()
+
+      #     # Calculate absolute distance from reference value
+      #     df['ref_diff'] = abs(df[reference_col] - reference_val)
+
+      #     # Drop any rows with a distance beyond threshold
+      #     if not pd.isna(max_dist_from_ref):
+      #       df = df.loc[df['ref_diff'] <= max_dist_from_ref]
+
+      #     # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
+      #     df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)
+
+      #     # Settle tie-breaks
+      #     if tie_break == 'below':
+      #       df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)
+
+      #     elif tie_break == 'above':
+      #       df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)
+
+      #     df = df.drop('ref_diff', axis=1)
+
+      #     return df
+
+      #   def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+
+      #     # Make sure the pair of reference values are in ascending order
+      #     reference_vals.sort()
+
+      #     # Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref
+
+      #     # Find matches for lower reference value
+      #     lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs[0], reference_col, group_by_col, value_col, 'below')
+
+      #     # Find matches for higher reference value
+      #     higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs[1], reference_col, group_by_col, value_col, 'above')
+
+      #     # Merge the two sets of matches
+      #     merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))
+
+      #     # Drop obs that do not have data for both ref values
+      #     merged_df = merged_df.dropna()
+
+      #     # Drop obs where the matched data does not meet the min distance requirement
+      #     if not pd.isna(min_dist_between):
+
+      #       # Store the names of the reference column returned from the two matches
+      #       ref_var_high = f'{reference_col}{reference_vals[1]}'
+      #       ref_var_low = f'{reference_col}{reference_vals[0]}'
+
+      #      # Keep only rows >= to the min distance
+      #       merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]
+
+      #     return merged_df
+
+      #   def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+
+      #     # Specify the name of the column in which the income/consumption welfare definition is stored
+      #     welfare_colname = 'welfare_type'
+
+      #     # Creat dataframes for thee scenarios:
+      #     # Scenario 1: only allow income data
+      #     df_inc_filter = df.loc[df[welfare_colname] == "income", :]
+      #     df_inc_filter.name = "Income"
+
+      #     # Scenario 2: only allow consumption data
+      #     df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
+      #     df_cons_filter.name = "Consumption"
+      #     # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
+      #     df_mixed = df.copy()
+
+      #     df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')
+
+      #     df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]
+
+      #     df_mixed.name = "Mixed"
+      #     #  Store the scneario dataframes in a list
+      #     df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]
+
+      #     # Run the matching function on each scenario
+      #     scenario_matches = [merge_two_ref_matches(
+      #       df_scenario, 
+      #       reference_vals, 
+      #       max_dist_from_refs, 
+      #       min_dist_between, 
+      #       reference_col, 
+      #       group_by_col, 
+      #       value_col) for df_scenario in df_scenarios]
+
+      #     # Combine the first two scenarios.
+      #     df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])
+
+      #     # Tidy up indexes
+      #     df_combined_matches = df_combined_matches.reset_index()
+
+      #     df_combined_matches = df_combined_matches.drop('level_1', axis=1)
+
+      #     df_combined_matches = df_combined_matches\
+      #         .rename(columns={"level_0": "pip_welfare"})
+
+      #     # Add in third scenario.
+      #     df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])
+
+      #     # add scenario name to te pip_welfare column
+      #     df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)
+
+      #     # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
+      #     # First count the matches
+      #     df_combined_matches['match_count'] =        df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
+      #     # Then drop any matches from the lowest priority where there are multiple matches
+      #     df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
+      #     #  Repeat at the next level of priority
+      #     df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
+      #     df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]
+
+      #     # Drop the match count column
+      #     df_combined_matches = df_combined_matches.drop('match_count', axis=1)
+
+      #     return df_combined_matches
+
+```

### 2023-01-27 PhD_pages/data_appendices/1_A_prepare_PIP_and_WID.qmd
Further work on comparison app (allow any combo of variables)

@@
   mutate(Bottom_50_share = decile1_share + decile2_share + decile3_share + decile4_share + decile5_share,
-          P90_P10_ratio = decile9_thr/decile1_thr)
+          P90_P10_ratio = decile9_thr/decile1_thr,
+          gini = gini * 100)
 
@@
 df_wid<- df_wid %>%
-    mutate(P90_P10_ratio = `P90...income.threshold.of.the.top.10.`/P10...income.threshold)
+    mutate(P90_P10_ratio = `P90...income.threshold.of.the.top.10.`/P10...income.threshold,
+    Gini = Gini.coefficient * 100)
 
@@
 df_wid<- df_wid %>%
-    rename("Gini" = "Gini.coefficient",
-         "Top_10_share" = "P90.P100...share.of.the.top.10.",
+    rename("Top_10_share" = "P90.P100...share.of.the.top.10.",
          "Bottom_50_share"= "P0.P50...share.of.the.bottom.50.",
@@
 :::
+
+## LIS data
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Load WID data
+We load the data from a file prepared from the WID Stata package by Our World in Data.
+
+```{r}
+
+from owid import catalog
+
+# results = catalog.find('population')
+# owid_pop = results[results['path'] == 'garden/owid/latest/key_indicators/population']
+
+# df_pop = owid_pop.iloc[0].load().reset_index().rename(columns={'year':'Year', 'country': 'Entity'})
+
+# df_pop.head()
+
+# df_wid<- read.csv("https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/PabloArriagada/WID/data/final/wid_pretax.csv")
+
+# A local version, in case internet isn't available 
+# df_wid<- read.csv("data/original/wid.csv")
+
+```
 
@@
 ::: {.callout-note collapse="true" appearance="minimal"}
-### Region classfications
+### Population
 

### 2023-01-27 PhD_pages/data_appendices/1_B_within_country_trends.qmd
Further work on comparison app (allow any combo of variables)

--- 2023-01-20 PhD_pages/data_appendices/1_B_within_country_trends.qmd
@@
 
+from plotnine import *
 import pandas as pd
 import numpy as np
-from plotnine import *
-import plotly.graph_objects as go
-```
-
-Load in the data
-```{python}
-
-df_wid = pd.read_csv("data/clean/wid.csv")
-df_pip = pd.read_csv("data/clean/pip.csv")
-df_regions = pd.read_csv("data/clean/WID_region_mapping.csv")
-df_pop = pd.read_csv("data/clean/population.csv")
 
@@
 
+This first block of code.....
+
 ```{python}
 
-# df_wid = pd.read_csv("PhD_pages/data_appendices/data/clean/wid.csv")
-# df_pip = pd.read_csv("PhD_pages/data_appendices/data/clean/pip.csv")
-# df_regions = pd.read_csv("PhD_pages/data_appendices/data/clean/WID_region_mapping.csv")
-# df_pop = pd.read_csv("PhD_pages/data_appendices/data/clean/population.csv")
-
-```
-
-A key difficulty of comparing trends is that the data is incomplete coverage.
-The paper uses two methods to calculate and compare aggregate within-country trends across incomplete data: using reference years and using year fixed effects in a regression.
-
-## Method 1: Reference years
-
-I need to allow for the income/consumption issue (currently the data includes both – i.e. multi observations per country year).
-
-### Mapping data to a reference year
-This function grabs the observation, by a grouping variable (e.g. by country), for which a reference variable (e.g. year) is closest to a reference value (e.g. 2002).
-You can specify a maximum distance from the reference value, beyond which no match will be returned.
-Because there may be tie-breaks – matches equally distant above or below the reference value – there is an argmuent to specify how these tie-breaks should resolved.
-```{python}
+#  ---------------------------------
+#  ---------- SECTION 1: Data prep functions -------------
+#  ----------------------------------
 
@@
 
+  df = df\
+    .rename(columns={value_col: "value"})
+
   return df
 
-```
-
-I test this function here:
-```{python}
-
-test = closest_to_reference(
-  df = df_wid, 
-  reference_val = 2002,
-  max_dist_from_ref = 5,
-  reference_col = 'Year',
-  group_by_col = 'Entity',
-  value_col = 'Gini',
-  tie_break = 'below'
-  )
-
-test.head()
-
-```
-
-### Obtaining a pair of observations
-
-This function runs the `closest_to_reference` function twice, over a pair of reference values.
-Tie-breaks are settled so as to maximise the gap between the two reference values.
-You can also specify a minimum distance between the two observations (e.g. pairs of matches that fall less than X years apart will be dropped).
-
-```{python}
 # Merge matches for different reference points
@@
 # Find matches for lower reference value
-  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs[0], reference_col, group_by_col, value_col, 'below')
+  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs, reference_col, group_by_col, value_col, 'below')
 
 # Find matches for higher reference value
-  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs[1], reference_col, group_by_col, value_col, 'above')
+  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs, reference_col, group_by_col, value_col, 'above')
 
@@
 
-```
-
-I test this function here:
-```{python}
-test = merge_two_ref_matches(
-  df = df_wid, 
-  reference_vals = [2000, 2010], 
-  max_dist_from_refs = [5, 4],
-  min_dist_between = 9,
-  reference_col = 'Year',
-  group_by_col = 'Entity',
-  value_col = 'Gini'
-  )
-
-test.head()
-
-```
-
-#### Allowing for the different welfare concepts in the PIP data
-
-As noted in .... (add backlink) ... the PIP data contains both income and consumption observations.
-This function produces matched pairs of observations from that data in which only one pair is selected for each value of the grouping variable (i.e. each country).
-Priority is given to pairs for which both observations relate to income. Second priority is given to pairs where both observations relate to consumption.
-Only if neither pair is available then it will return a mixed pair of observations if that is available.
-```{python}
   # For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
@@
 
-I test this function here:
 ```{python}
+#  ---------------------------------
+#  ---------- SECTION 2: Specifications -------------
+#  ----------------------------------
 
-test = pip_welfare_routine(
+# These parameters are not controlled in the app.
+reference_col = 'Year'
+group_by_col = 'Entity'
 
-  df = df_pip,
-  reference_vals = [1986, 2016], 
-  max_dist_from_refs = [5, 5], 
-  min_dist_between = 30, 
-  reference_col = 'Year',
-  group_by_col = 'Entity',
-  value_col = 'Gini'
+tolerance = 1
+outlier_cut_off_upper = None
 
-  )
+source_count_requirement = "all"
+# source_count_requirement = "any"
 
-test.head()
+reference_vals = [1990,2018]
+max_dist_from_refs = 4
+min_dist_between = 5
+
+data = {"PIP": df_pip,
+        "WID": df_wid}
+
+data_to_match = {
+  "sources": ["PIP", "WID", "PIP"],
+  "var_names": ["Gini", "Top_10_share", "Top_10_share"]
+  }
+
 ```
 
-### Merging reference year aligned data from WID and PIP datasets
+This second block of code we read in the data (this needs to be done in a different way in )
 
@@
 
+#  ---------------------------------
+#  ---------- SECTION 3: Read in data -------------
+#  ----------------------------------
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'
+
+df_pip = pd.read_csv(url)
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'
+
+df_wid = pd.read_csv(url)
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/WID_region_mapping.csv'
+
+df_regions = pd.read_csv(url)
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'
+
+df_pop = pd.read_csv(url)
+
+```
+
+```{python}
+
+#  ---------------------------------
+#  ---------- SECTION 4: Run data functions and produce plot -------------
+#  ----------------------------------
+
+# Combine the source and varname strings element-wise to produce keys that will be used for the concat later
+source_var_keys = [i + j for i, j in zip(data_to_match["sources"], data_to_match["var_names"])]
+
+def prep_and_merge_multiple_vars(data_to_match, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col):
+
+  matches = []
+
+  for i in range(len(data_to_match['sources'])):
+
+    source = data_to_match["sources"][i]
+    var = data_to_match["var_names"][i]
+    print(i)
+    print(source)
+    print(var)
+    df = data[source]
+
+    if source == "PIP":
+
+      matched = pip_welfare_routine(
+          df = df,
+          reference_vals = reference_vals,
+          max_dist_from_refs = max_dist_from_refs,
+          min_dist_between = min_dist_between,
+          reference_col = reference_col,
+          group_by_col = group_by_col,
+          value_col = var
+        )
+
+    else:
+
+      matched = merge_two_ref_matches(
+          df = df,
+          reference_vals = reference_vals,
+          max_dist_from_refs = max_dist_from_refs,
+          min_dist_between = min_dist_between,
+          reference_col = reference_col,
+          group_by_col = group_by_col,
+          value_col = var
+        )
+
+    matches.append(matched)
+
+  ref_matches = pd.concat(matches, keys=source_var_keys)
+
+  # Tidy up indexes
+  ref_matches = ref_matches.reset_index()
+
+  ref_matches = ref_matches.drop('level_1', axis=1)
+
+  ref_matches = ref_matches\
+    .rename(columns={"level_0": "source_var"})
+
+  return ref_matches
+
+def prep_plot_and_tables(
+  data_to_match,
+  reference_vals, 
+  max_dist_from_refs,
+  min_dist_between,
+  reference_col,
+  group_by_col,
+  tolerance,
+  outlier_cut_off_upper,
+  source_count_requirement):
+
+  # Other spec:
+  region_col = 'region_alt'
+
+  # ---- Prep data ---------
+
+  plot_data = prep_and_merge_multiple_vars(
+    data_to_match = data_to_match,
+    reference_vals = reference_vals, 
+    max_dist_from_refs = max_dist_from_refs,
+    min_dist_between = min_dist_between,
+    reference_col = reference_col,
+    group_by_col = group_by_col
+  )
+
+  # Store the names of the columns to be used onthe X and Y axis
+  x_axis = f'value{reference_vals[0]}'
+  y_axis = f'value{reference_vals[1]}'
+
+  x_ref_val = f'{reference_col}{reference_vals[0]}'
+  y_ref_val = f'{reference_col}{reference_vals[1]}'
+
+  # Apply outlier cut off, if specified
+  if not pd.isna(outlier_cut_off_upper):
+    plot_data = plot_data.loc[plot_data[x_axis] <= outlier_cut_off_upper]
+    plot_data = plot_data.loc[plot_data[y_axis] <= outlier_cut_off_upper]
+
+  # Add a count by country – counting whether data is available from each source or not
+  plot_data['source_count'] = plot_data.groupby(group_by_col)['source_var'].transform('count')
+
+  # Apply source requirement (whether to include only observations with data from all sources)
+  if source_count_requirement == "all":
+    plot_data = plot_data.loc[plot_data['source_count'] == plot_data['source_count'].max()]
+
+  # Drop source_count column
+  plot_data = plot_data.drop('source_count', axis=1)
+
+  # Add in region classification
+  plot_data = pd.merge(plot_data, df_regions, how = 'left')
+
+  # Add in population data
+    # For first ref
+  df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[0], ['Entity', 'population'] ]
+
+  plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')
+
+  plot_data = plot_data.rename(columns={'population':'population_ref1'})
+
+    # For second ref
+  df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[1], ['Entity', 'population'] ]
+
+  plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')
+
+  plot_data = plot_data.rename(columns={'population':'population_ref2'})
+
+    # Calculate average population two reference periods
+  plot_data['avg_pop'] = (plot_data['population_ref1'] + plot_data['population_ref2'])/2
+
+  # ----- Prep plot ------
+
+  # I grab what I am guessing to be the max and min tick marks on the x axist, in order to define the coordinates of a 'tolerance' ribbon.
+  x_min = plot_data[x_axis].min()
+  x_min_floor = np.floor(x_min * 10) / 10
+
+  x_max = plot_data[x_axis].max()
+  x_max_ceiling = np.ceil(x_max * 10) / 10
+
+  # Set the coordinates of the shaded ribbon
+  shaded_coord = pd.DataFrame({'x': [x_min_floor, x_max_ceiling, x_max_ceiling, x_min_floor], 
+    'y': [x_min_floor-tolerance, x_max_ceiling-tolerance, x_max_ceiling+tolerance , x_min_floor+tolerance]})
+
+  plot = (ggplot(plot_data
+  , aes(x_axis, y_axis, alpha=0.5))
+  + geom_point(aes(color=region_col, size='avg_pop', alpha = 0.6))
+  + facet_wrap('~ source_var')
+  + geom_polygon(aes(x='x', y='y', alpha = 0.2), data=shaded_coord)
+  + theme_light()
+  + theme(legend_position='none')
+  )
+
+  # ---- Prep summary table ----
+
+  # Calculate the change between the two ref periods
+  plot_data['change'] = (plot_data[y_axis] - plot_data[x_axis])
+
+  group_yes_vars = ['source_var', region_col]
+  group_no_vars = ['source_var']
+
+  group_scenarios = [group_no_vars, group_yes_vars]
+
+  aggs = []
+
+  for group_vars in group_scenarios:
+
+  # Prepare pop-weighted average change
+    agg_level = plot_data.copy()
+
+    # Calulate pop weights by source and region
+    agg_level['regional_pop_weights'] = agg_level.groupby(group_vars)['avg_pop'].transform(lambda x: x/x.sum())
+
+    agg_level['global_pop_weights'] = agg_level.groupby('source_var')['avg_pop'].transform(lambda x: x/x.sum())
+
+    # Multiply the change by the pop weights 
+    agg_level['region_weighted_change'] = agg_level['change'] * agg_level['regional_pop_weights']
+
+  # Prepare fall/stable/rise categories
+    agg_level['fall'] = agg_level['change'] < -tolerance
+
+    agg_level['stable'] = (agg_level['change'] <= tolerance) & (agg_level['change'] >= -tolerance)
+
+    agg_level['rise'] = agg_level['change'] > tolerance
+
+    agg_level[['fall', 'stable', 'rise']] = agg_level[['fall', 'stable', 'rise']].astype(int)
+
+    agg_level = agg_level.groupby(group_vars)[['change', 'region_weighted_change', 'global_pop_weights', 'fall', 'stable', 'rise']]\
+      .agg(['sum', 'mean', 'count'])
+
+  # Aggregate by source and region
+    if group_vars == group_yes_vars:
+
+      agg_level = agg_level\
+        .unstack().T.reset_index()
+
+    elif group_vars == group_no_vars:
+
+      agg_level = agg_level\
+        .T.reset_index()
+
+      agg_level[region_col] = "World"
+
+    aggs.append(agg_level)
+
+  df_summary = pd.concat(aggs)
+
+    # filter for the aggregations we need and label in the 'summary' column
+  df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'mean'), 'summary'] = 'avg change'
+
+  df_summary.loc[(df_summary['level_0'] == 'region_weighted_change') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Regional pop-weighted avg change'
+
+  df_summary.loc[(df_summary['level_0'] == 'global_pop_weights') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Global pop weights'
+
+  df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'count'), 'summary'] = 'n'
+
+  df_summary.loc[(df_summary['level_0'] == 'fall') & (df_summary['level_1'] == 'sum'), 'summary'] = 'fall'
+
+  df_summary.loc[(df_summary['level_0'] == 'stable') & (df_summary['level_1'] == 'sum'), 'summary'] = 'stable'
+
+  df_summary.loc[(df_summary['level_0'] == 'rise') & (df_summary['level_1'] == 'sum'), 'summary'] = 'rise'
+
+  df_summary = df_summary[df_summary['summary'].notnull()]
+
+  df_summary = df_summary.drop(['level_0', 'level_1'], axis = 1)
+
+  df_summary['summary'] = pd.Categorical(df_summary['summary'], categories=['fall', 'stable', 'rise','n', 'avg change', 'Regional pop-weighted avg change', 'Global pop weights'], ordered=True)
+
+  df_summary = df_summary.sort_values([region_col,'summary']).set_index([region_col,'summary'])
+
+  # format to show appropriate d.p. I run a loop over each row for want of figuring out a better way!
+  repeat_length = 7
+  for i in range(0, len(df_summary.index.levels[0].unique())):
+    df_summary.iloc[i*repeat_length,:] = df_summary.iloc[i*repeat_length,:].map('{:.0f}'.format)
+    df_summary.iloc[1 + i*repeat_length,:] = df_summary.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
+    df_summary.iloc[2 + i*repeat_length,:] = df_summary.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
+    df_summary.iloc[3 + i*repeat_length,:] = df_summary.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
+    df_summary.iloc[4 + i*repeat_length,:] = df_summary.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)
+    df_summary.iloc[5 + i*repeat_length,:] = df_summary.iloc[5 + i*repeat_length,:].map('{:.2f}'.format)
+    df_summary.iloc[6 + i*repeat_length,:] = df_summary.iloc[6 + i*repeat_length,:].map('{:.2f}'.format)
+
+  # ----- Return outputs ------
+  plot_and_tables = {
+    "table_data": plot_data,
+    "plot": plot,
+    "summary": df_summary
+  }
+
+  return plot_and_tables
+
+```
+
+```{python}
+plot_and_tables = prep_plot_and_tables(
+    data_to_match = data_to_match,
+    reference_vals = reference_vals, 
+    max_dist_from_refs = max_dist_from_refs,
+    min_dist_between = min_dist_between,
+    reference_col = reference_col,
+    group_by_col = group_by_col,
+    tolerance = tolerance,
+    outlier_cut_off_upper = outlier_cut_off_upper,
+    source_count_requirement = source_count_requirement
+    )
+
+plot = plot_and_tables['plot']
+summary = plot_and_tables['summary']
+
+```
+
+```{python}
+plot
+
+```
+
+```{python}
+summary
+
+```
+
+## Explore the data
+
+I plan to build a Shiny app to help compare trends across datasets (using Shinylive – built on Shiny for Python - ).
+
+Here is a test app just so I can test the wiring of how such an app works.
+
+```{shinylive-python}
+#| standalone: true
+#| viewerHeight: 420
+
+from shiny import *
+from plotnine import *
+from pyodide.http import open_url
+import pandas as pd
+import numpy as np
+
+#  ---------------------------------
+#  ---------- Data prep functions -------------
+#  ----------------------------------
+
+#  Function get matching for ref years
+def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):
+
+  df = df.loc[:, [reference_col, group_by_col, value_col]]
+
+  # Drop NAs
+  df = df.dropna()
+
+  # Calculate absolute distance from reference value
+  df['ref_diff'] = abs(df[reference_col] - reference_val)
+
+  # Drop any rows with a distance beyond threshold
+  if not pd.isna(max_dist_from_ref):
+    df = df.loc[df['ref_diff'] <= max_dist_from_ref]
+
+  # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
+  df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)
+
+  # Settle tie-breaks
+  if tie_break == 'below':
+    df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)
+
+  elif tie_break == 'above':
+    df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)
+
+  df = df.drop('ref_diff', axis=1)
+
+  return df
+
+# Merge matches for different reference points
+def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+
+# Make sure the pair of reference values are in ascending order
+  reference_vals.sort()
+
+# Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref
+
+# Find matches for lower reference value
+  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs, reference_col, group_by_col, value_col, 'below')
+
+# Find matches for higher reference value
+  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs, reference_col, group_by_col, value_col, 'above')
+
+# Merge the two sets of matches
+  merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))
+
+# Drop obs that do not have data for both ref values
+  merged_df = merged_df.dropna()
+
+# Drop obs where the matched data does not meet the min distance requirement
+  if not pd.isna(min_dist_between):
+
+  # Store the names of the reference column returned from the two matches
+    ref_var_high = f'{reference_col}{reference_vals[1]}'
+    ref_var_low = f'{reference_col}{reference_vals[0]}'
+
+  # Keep only rows >= to the min distance
+    merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]
+
+  return merged_df
+
+  # For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
+def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+
+    # Specify the name of the column in which the income/consumption welfare definition is stored
+    welfare_colname = 'welfare_type'
+
+    # Creat dataframes for thee scenarios:
+    # Scenario 1: only allow income data
+    df_inc_filter = df.loc[df[welfare_colname] == "income", :]
+    df_inc_filter.name = "Income"
+
+    # Scenario 2: only allow consumption data
+    df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
+    df_cons_filter.name = "Consumption"
+    # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
+    df_mixed = df.copy()
+
+    df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')
+
+    df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]
+
+    df_mixed.name = "Mixed"
+    #  Store the scneario dataframes in a list
+    df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]
+
+    # Run the matching function on each scenario
+    scenario_matches = [merge_two_ref_matches(
+      df_scenario, 
+      reference_vals, 
+      max_dist_from_refs, 
+      min_dist_between, 
+      reference_col, 
+      group_by_col, 
+      value_col) for df_scenario in df_scenarios]
+
+    # Combine the first two scenarios.
+    df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])
+
+    # Tidy up indexes
+    df_combined_matches = df_combined_matches.reset_index()
+
+    df_combined_matches = df_combined_matches.drop('level_1', axis=1)
+
+    df_combined_matches = df_combined_matches\
+      .rename(columns={"level_0": "pip_welfare"})
+
+    # Add in third scenario.
+    df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])
+
+    # add scenario name to te pip_welfare column
+    df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)
+
+    # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
+      # First count the matches
+    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
+      # Then drop any matches from the lowest priority where there are multiple matches
+    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
+      #  Repeat at the next level of priority
+    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
+    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]
+
+    # Drop the match count column
+    df_combined_matches = df_combined_matches.drop('match_count', axis=1)
+
+    return df_combined_matches
+
+#  ---------------------------------
+#  ---------- Specifications -------------
+#  ----------------------------------
+
+# These parameters are not controlled in the app.
+reference_col = 'Year'
+group_by_col = 'Entity'
+
+tolerance = 1
+outlier_cut_off_upper = None
+
+source_count_requirement = 2
+
+#  ---------------------------------
+#  ---------- App code -------------
+#  ----------------------------------
+
+def panel_box(*args, **kwargs):
+    return ui.div(
+        ui.div(*args, class_="card-body"),
+        **kwargs,
+        class_="card mb-3",
+    )
+
+app_ui = ui.page_fluid(
+    {"class": "p-4"},
+    ui.row(
+        ui.column(
+            4,
+            panel_box(
+                ui.input_slider("reference_vals", "Reference years", 1975, 2020, value=[1990,2018]),
+                ui.input_numeric(
+                            "max_dist_from_refs",
+                            "=/- (years)",
+                            4,
+                        ),
+                ui.input_numeric(
+                            "min_dist_between",
+                            "Min distance between observations (years)",
+                            23,
+                        ),
+                ui.input_action_button(
+                    "download", "Download data and image", class_="btn-primary w-100"
+                ),
+            ),
+            ui.navset_tab_card(
+                ui.nav(
+                    "Compare sources",
+                    ui.input_select(
+                "metric",
+                "Metric",
+                {
+                    "Gini": "Gini",
+                    "Top_10_share": "Top 10p.c. share",
+                },
+            ),
+            ),
+                ui.nav(
+                  "Compare metrics"),
+            ),
+        ),
+        ui.column(
+            8,
+            ui.output_plot("example_plot")
+            ),
+    ),
+)
+
+def server(input, output, session):
+    @output
+    @render.plot
+    def example_plot():
+
+      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'
+
+      df_pip = pd.read_csv(open_url(url))
+
+      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'
+
+      df_wid = pd.read_csv(open_url(url))
+
+      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/WID_region_mapping.csv'
+
+      df_regions = pd.read_csv(open_url(url))
+
+      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'
+
+      df_pop = pd.read_csv(open_url(url))
+
       def prep_wid_pip_data(reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+
         pip_matches = pip_welfare_routine(
@@
         return ref_pairs
-```
-
-## Prepare plot and tables
-
-```{python}
-#| echo: false
 
@@
 
-  # Make nicely formatted table
-  plot_data_formatted = plot_data.copy()
-
-  plot_data_formatted[x_axis] = plot_data_formatted[x_axis].map('{:.2f}'.format)
-  plot_data_formatted[y_axis] = plot_data_formatted[y_axis].map('{:.2f}'.format)
-
-  plot_data_table = go.Figure(data=[go.Table(
-    header=dict(values=list(plot_data_formatted.columns)),
-    cells=dict(values=plot_data_formatted.transpose().values.tolist() ))
-    ])
-
         # ----- Prep plot ------
@@
         + geom_polygon(aes(x='x', y='y', alpha = 0.2), data=shaded_coord)
-  + theme_light())
+        + theme_light()
+        + theme(legend_position='none')
+        )
 
@@
           "table_data": plot_data,
-    "table": plot_data_table,
           "plot": plot,
@@
 
-```
-
-```{python}
-pd.set_option('display.max_rows', None)
-
-```
-
-## 1990-2017
-```{python}
-# Specifications
-reference_vals = [1990, 2017]
-max_dist_from_refs = [4, 4]
-min_dist_between = 23
-
-# These parameters are unlikely to change:
-reference_col = 'Year'
-group_by_col = 'Entity'
-
-```
-
-::: {.panel-tabset}
-## Gini
-
-```{python}
-# | echo: false
-# Specifications
-
-value_col = 'Gini'
-# 'Gini'
-# 'Top_10_share'
-# 'Bottom_50_share'
-# 'MLD'
-# 'P90_P10_ratio' – not that helpful due to accuracy of low incomes
-
-# Set the value which constitutes a significant change in the value being measured
-  # E.g. for Gini tolerance = 0.02 (2 points)
-
-tolerance = 0.02
-outlier_cut_off_upper = None
-
-```
-::: {.panel-tabset}
-## All countries
-```{python}
-# | echo: false
-
-# Data from both sources is needed?
-  # If this = 1 then a country will be incldued even if it is only availalbe from one source.
-  # If = 2, then only countries with data for both sources will be included
-source_count_requirement = 1
-
       plot_and_tables = prep_plot_and_tables(
-  reference_vals = reference_vals, 
-  max_dist_from_refs = max_dist_from_refs,
-  min_dist_between = min_dist_between,
+          reference_vals = list(input.reference_vals()), 
+          max_dist_from_refs = input.max_dist_from_refs(),
+          min_dist_between = input.min_dist_between(),
           reference_col = reference_col,
           group_by_col = group_by_col,
-  value_col = value_col,
+          value_col = input.metric(),
           tolerance = tolerance,
@@
 
-```
-::: {.panel-tabset}
-## Plot
-```{python}
-plot_and_tables['plot']
-```
-
-## Summary
-```{python}
-plot_and_tables['summary']
-```
-
-## Data
-```{python}
-
-plot_data_results = plot_and_tables['table_data']
-plot_data_results
-```
-:::
-
-## Countries with data for both sources
-
-```{python}
-# | echo: false
-
-# Data from both sources is needed?
-  # If this = 1 then a country will be incldued even if it is only availalbe from one source.
-  # If = 2, then only countries with data for both sources will be included
-source_count_requirement = 2
-
-plot_and_tables = prep_plot_and_tables(
-  reference_vals = reference_vals, 
-  max_dist_from_refs = max_dist_from_refs,
-  min_dist_between = min_dist_between,
-  reference_col = reference_col,
-  group_by_col = group_by_col,
-  value_col = value_col,
-  tolerance = tolerance,
-  outlier_cut_off_upper = outlier_cut_off_upper,
-  source_count_requirement = source_count_requirement
-  )
-
-```
-::: {.panel-tabset}
-## Plot
-```{python}
-plot_and_tables['plot']
-```
-
-## Summary
-```{python}
-plot_and_tables['summary']
-```
-
-## Data
-```{python}
-
-plot_data_results = plot_and_tables['table_data']
-plot_data_results
-```
-:::
-:::
-
-## Top 10% share
-
-```{python}
-# | echo: false
-# Specifications
-
-value_col = 'Top_10_share'
-
-# Set the value which constitutes a significant change in the value being measured
-  # E.g. for Gini tolerance = 0.02 (2 points)
-
-tolerance = 1
-outlier_cut_off_upper = None
-
-```
-::: {.panel-tabset}
-## All countries
-```{python}
-# | echo: false
-
-# Data from both sources is needed?
-  # If this = 1 then a country will be incldued even if it is only availalbe from one source.
-  # If = 2, then only countries with data for both sources will be included
-source_count_requirement = 1
-
-plot_and_tables = prep_plot_and_tables(
-  reference_vals = reference_vals, 
-  max_dist_from_refs = max_dist_from_refs,
-  min_dist_between = min_dist_between,
-  reference_col = reference_col,
-  group_by_col = group_by_col,
-  value_col = value_col,
-  tolerance = tolerance,
-  outlier_cut_off_upper = outlier_cut_off_upper,
-  source_count_requirement = source_count_requirement
-  )
-
-```
-::: {.panel-tabset}
-## Plot
-```{python}
-plot_and_tables['plot']
-```
-
-## Summary
-```{python}
-plot_and_tables['summary']
-```
-
-## Data
-```{python}
-
-plot_data_results = plot_and_tables['table_data']
-plot_data_results
-```
-:::
-
-## Countries with data for both sources
-
-```{python}
-# | echo: false
-
-# Data from both sources is needed?
-  # If this = 1 then a country will be incldued even if it is only availalbe from one source.
-  # If = 2, then only countries with data for both sources will be included
-source_count_requirement = 2
-
-plot_and_tables = prep_plot_and_tables(
-  reference_vals = reference_vals, 
-  max_dist_from_refs = max_dist_from_refs,
-  min_dist_between = min_dist_between,
-  reference_col = reference_col,
-  group_by_col = group_by_col,
-  value_col = value_col,
-  tolerance = tolerance,
-  outlier_cut_off_upper = outlier_cut_off_upper,
-  source_count_requirement = source_count_requirement
-  )
-
-```
-::: {.panel-tabset}
-## Plot
-```{python}
-plot_and_tables['plot']
-```
-
-## Summary
-```{python}
-plot_and_tables['summary']
-```
-
-## Data
-```{python}
-
-plot_data_results = plot_and_tables['table_data']
-plot_data_results
-```
-:::
-
-:::
-
-Next measure
-
-:::
-
-## Method 2: regression analysis
-
-## Explore the date
-
-I plan to build a Shiny app to help compare trends across datasets (using Shinylive – built on Shiny for Python - ).
-
-Here is a test app just so I can test the wiring of how such an app works.
-
-```{shinylive-python}
-#| standalone: true
-#| viewerHeight: 420
-
-from shiny import *
-from plotnine import *
-from pyodide.http import open_url
-import pandas as pd
-
-app_ui = ui.page_fluid(
-    ui.output_plot("example_plot"),
-)
-
-def server(input, output, session):
-    @output
-    @render.plot
-    def example_plot():
-
-        url = 'https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/JoeHasell/PIP/data/ppp_2017/final/PIP_data_public_download/full_dataset/cons_only/poverty_cons_only.csv'
-
-        df = pd.read_csv(open_url(url))
-
-        plot = (ggplot(df, aes('Year', 'headcount_ratio_365', color = 'Entity'))
-        + geom_line())
-
-        # d = {'col1': [1, 2], 'col2': [3, 4]}
-        # df = pd.DataFrame(data=d)
-        # plot = (ggplot(df, aes('col1', 'col2'))
-        # + geom_point())
+      plot = plot_and_tables['plot']
 

### 2023-01-29 PhD_pages/data_appendices/1_B_prepare_WID.qmd
Paper 1, more conversion to python

---
title: "Appendix 1.B: Preparing WID inequality data"
format: html
warning: false

---

::: {.callout-note collapse="true" appearance="minimal"}
### Packages and settings

```{python}

import pandas as pd

pd.set_option('display.max_rows', None)
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Load WID data
We load the data from a file prepared from the WID Stata package by Our World in Data.

```{python}
# df_wid<- read.csv("https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/PabloArriagada/WID/data/final/wid_pretax.csv")

df_wid = pd.read_csv("data/original/wid.csv")

```

The dataframe looks like this:
```{python}
df_wid.head(50)
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Drop subnational, regional and historical geographical entities

Here we create a list of entities in the WID data that do not correspond to countries in existence today. These we drop from the data.

```{python}

drop_entities =  [
  "China - rural",
  "China - urban",
  "East Germany",
  "Other Russia and Central Asia",
  "Other East Asia",
  "Other Western Europe",
  "Other Latin America",
  "Other MENA",
  "Other South & South-East Asia",
  "Other Sub-Saharan Africa",
  "Africa",
  "Asia",
  "Europe",
  "Oceania",
  "Central Asia",
  "East Africa",
  "East Asia",
  "South Asia",
  "Eastern Europe",
  "Middle Africa",
  "North Africa",
  "North America",
  "South-East Asia",
  "South Africa region",
  "West Africa",
  "West Asia",
  "Western Europe",
  "European Union",
  "World",
  "Asia (excluding Middle East)",
  "North America and Oceania",
  "Sub-Saharan Africa",
  "Latin America",
  "Middle East",
  "MENA",
  "Russia and Central Asia",
  "South & South-East Asia"
]

df_wid = df_wid.loc[~df_wid["Entity"].isin(drop_entities)]

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Calculate some additional inequality metrics

```{python}
# Gini as index
df_wid["Gini"] = df_wid['Gini.coefficient'] * 100

```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Rename and select needed variables

```{python}

df_wid = df_wid.rename(columns={
    'P90.P100...share.of.the.top.10.':'Top_10_share',
    'P0.P50...share.of.the.bottom.50.':'Bottom_50_share',
    'P90...income.threshold.of.the.top.10.':'P90',
    'P10...income.threshold':'P10'
  })

keep_cols = [
  "Year",
  "Entity",
  "Gini",
  "Top_10_share",
  "Bottom_50_share"
]

df_wid = df_wid[keep_cols]

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### See the prepared WID data

The first 100 rows of the dataframe looks like this:
```{python}
df_wid.head(100)
```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Save the data
I save the prepared data locally:
```{python}
df_wid.to_csv("data/clean/wid.csv", index=False)
```
:::

### 2023-01-29 PhD_pages/data_appendices/1_C_prepare_LIS.qmd
Paper 1, more conversion to python

--- 2023-01-18 PhD_pages/data_appendices/1_A_prepare_PIP_and_WID.qmd
@@
 ---
-title: "Appendix 1.A: Preparing PIP and WID inequality data"
+title: "Appendix 1.C: Preparing LIS inequality data"
 format: html
 warning: false
+
 ---
-
-# Processing steps
-
-The titles in each of the boxes below summarise the steps taken to prepare the PIP and WID data. In each case you can expand the box to see the code used.
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Load packages
-
-```{r}
-library(tidyverse)
-library(knitr)
-
-```
 
@@
 import pandas as pd
-```
-:::
 
-## PIP data
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Load PIP data
-We load the data from a file prepared from the PIP API by Our World in Data.
-```{r}
-# df_pip<- read.csv("https://raw.githubusercontent.com/owid/poverty-data/main/datasets/pip_dataset.csv")
-
-#  A local version, in case internet isn't available 
-df_pip<- read.csv("PhD_pages/data_appendices/data/original/pip_dataset.csv")
-
+pd.set_option('display.max_rows', None)
 ```
 
-The dataframe looks like this:
-```{r}
-kable(
-
-    head(df_pip, 50)
-
-)
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Drop sub-national data
-
-The PIP data includes observations for national, urban and rural populations (`reporting_level`):
-
-```{r}
-kable(
-    df_pip %>%
-        count(reporting_level)
-)
-```
-
-Here we add a count of the number of reporting levels per country-year.
-```{r}
-
-count_reporting_levels<- df_pip %>%
-    group_by(country, year) %>%
-    count(reporting_level) %>%
-    group_by(country, year) %>%
-    summarise(n_reporting_levels = n())
-
-df_pip<- left_join(df_pip, count_reporting_levels) %>%
-    relocate(n_reporting_levels, .after=reporting_level)
-
-kable(
-
-    head(df_pip, 20)
-
-)
-```
-
-We only want one reporting level for each country-year.
-
-That is already true for most country-years. But some include all three reporting levels.
-```{r}
-kable(
-
-    df_pip %>%
-        count(n_reporting_levels)
-
-)
-
-```
-
-Here are the countries which include all three reporting levels:
-```{r}
-multi_reporting_level_countries<- 
-    df_pip %>%
-        filter(n_reporting_levels == 3) %>%
-        select(country) %>%
-        unique() %>%
-        pull(country)
-
-print(multi_reporting_level_countries)
-
-```
-
-For the country-years with 3 reporting levels, we keep only the national estimates. 
-```{r}
-df_pip<- df_pip %>%
-    filter(n_reporting_levels != 3 | reporting_level == 'national' )
-
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Drop regional data
-
-The dataset includes aggregated estiates for world regions.
-
-Here we define a list of these aggregate entities and drop them from the data.
-```{r}
-drop_aggs<- c(
-  "East Asia and Pacific",
-  "South Asia",
-  "Europe and Central Asia",
-  "High income countries",
-  "Latin America and the Caribbean",
-  "Middle East and North Africa",
-  "Sub-Saharan Africa",
-  "World"
-)
-
-df_pip<- df_pip %>%
-  filter(!country %in% drop_aggs)
-
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Select 2017 PPPs
-
-The data is provided by the World Bank in both 2011 and 2017 PPPs.
-
-We will be using 2017 PPPs for Paper 1.
+We load the data from a file prepared from the WID Stata package by Our World in Data.
 
@@
 
-df_pip<- df_pip %>%
-  filter(ppp_version == 2017)
+from owid import catalog
 
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Drop extremely low P10 values
-
-There are some very low values of P10. These lead to exaggerated values of the P90/P10 ratio.
-
-```{r}
-
-kable(
-    df_pip %>% filter(decile1_thr<0.02)
-
-  )
-
-```
-
-We set these values of P10 to NA, and thereby precluding a value for P90/P10 ratio when it is calculated below.
-
-```{r}
-
-df_pip<- df_pip %>%
-  mutate(decile1_thr = if_else(decile1_thr<0.02,
-                              as.numeric(NA),
-                              decile1_thr))
-
-kable(
-    df_pip %>% filter(is.na(decile1_thr))
-
-  )
-
-```
-
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Calculate additional variables
-
-```{r}
-
-df_pip<- df_pip %>%
-  mutate(Bottom_50_share = decile1_share + decile2_share + decile3_share + decile4_share + decile5_share,
-          P90_P10_ratio = decile9_thr/decile1_thr)
-
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Rename and select needed columns
-```{r}
-
-df_pip<- df_pip %>%
-  rename("Gini" = "gini",
-         "Top_10_share" = "decile10_share",
-         "MLD" = "mld",
-         "Entity" = "country",
-         "Year" = "year")
-
-df_pip<- df_pip %>%
-    select(
-        Year,
-        Entity,
-        Gini,
-        Top_10_share,
-        Bottom_50_share,
-        MLD,
-        P90_P10_ratio,
-        welfare_type)
-
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### See the prepared PIP data
-
-The first 100 rows of the dataframe looks like this:
-```{r}
-kable(
-
-    head(df_pip, 100)
-
-)
-```
-
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Save the data
-I save the prepared data locally:
-```{r}
-
-write.csv(df_pip, 
-      "PhD_pages/data_appendices/data/clean/pip.csv", row.names=FALSE)
-
-```
-:::
-
-## WID data
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Load WID data
-We load the data from a file prepared from the WID Stata package by Our World in Data.
-
-```{r}
-# df_wid<- read.csv("https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/PabloArriagada/WID/data/final/wid_pretax.csv")
-
-# A local version, in case internet isn't available 
-df_wid<- read.csv("PhD_pages/data_appendices/data/original/wid.csv")
-
-```
-
-The dataframe looks like this:
-```{r}
-kable(
-
-    head(df_wid, 50)
-
-)
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Drop subnational, regional and historical geographical entities
-
-Here we create a list of entities in the WID data that do not correspond to countries in existence today. These we drop from the data.
-```{r}
-drop_entities<- c(
-"China - rural",
-"China - urban",
-"East Germany",
-"Other Russia and Central Asia",
-"Other East Asia",
-"Other Western Europe",
-"Other Latin America",
-"Other MENA",
-"Other South & South-East Asia",
-"Other Sub-Saharan Africa",
-"Africa",
-"Asia",
-"Europe",
-"Oceania",
-"Central Asia",
-"East Africa",
-"East Asia",
-"South Asia",
-"Eastern Europe",
-"Middle Africa",
-"North Africa",
-"North America",
-"South-East Asia",
-"South Africa region",
-"West Africa",
-"West Asia",
-"Western Europe",
-"European Union",
-"World",
-"Asia (excluding Middle East)",
-"North America and Oceania",
-"Sub-Saharan Africa",
-"Latin America",
-"Middle East",
-"MENA",
-"Russia and Central Asia",
-"South & South-East Asia"
-)
-
-df_wid<- df_wid %>%
-  filter(!Entity %in% drop_entities)
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Drop extremely low P10 values
-
-There are some very low values of P10. These lead to exaggerated values of the P90/P10 ratio. Here are all values less than $1 a day (using pretax national income)
-
-```{r}
-
-kable(
-    df_wid %>% filter(P10...income.threshold/365<1)
-
-  )
-
-```
-
-We set these values of P10 to NA, and thereby precluding a value for P90/P10 ratio when it is calculated below.
-
-```{r}
-
-df_wid<- df_wid %>%
-  mutate(P10...income.threshold = if_else(P10...income.threshold/365<1,
-                              as.numeric(NA),
-                              P10...income.threshold))
-
-                              kable(
-    df_wid %>% filter(is.na(P10...income.threshold))
-
-  )
-
-```
-
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Calculate some additional inequality metrics
-
-```{r}
-
-#  Calculate P90_P10_ratio
-df_wid<- df_wid %>%
-    mutate(P90_P10_ratio = `P90...income.threshold.of.the.top.10.`/P10...income.threshold)
-
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Rename and select needed variables
-
-```{r}
-df_wid<- df_wid %>%
-    rename("Gini" = "Gini.coefficient",
-         "Top_10_share" = "P90.P100...share.of.the.top.10.",
-         "Bottom_50_share"= "P0.P50...share.of.the.bottom.50.",
-         "P90" = "P90...income.threshold.of.the.top.10.",
-         "P10" = "P10...income.threshold") %>%
-    select(
-        Year,
-        Entity,
-        Gini,
-        Top_10_share,
-        # MLD,
-        P90_P10_ratio,
-        Bottom_50_share,
-        P90,
-        P10)
-
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### See the prepared WID data
-
-The first 100 rows of the dataframe looks like this:
-```{r}
-kable(
-
-    head(df_wid, 100)
-
-)
-```
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Save the data
-I save the prepared data locally:
-```{r}
-
-write.csv(df_wid, 
-      "PhD_pages/data_appendices/data/clean/wid.csv", row.names=FALSE)
-
-```
-:::
-:::
-
-## Additional variables
-::: {.callout-note collapse="true" appearance="minimal"}
-### Region classfications
-For regional classifications I take WID definitions, as specified on their website [here](https://wid.world/codes-dictionary/#country-code).
-
-I read in a csv prepared from the table in that webpage, which has been mapped to the Our World in Data standarized country names used throughout this project.
-```{r}
-
-df_regions<- read.csv("PhD_pages/data_appendices/data/original/WID_regions_country_standardized.csv")
-
-```
-
-I create a clean region mapping table using two regional definitions.
-One is the 'standard' definitions used by WID.
-In an alternative set I split Western and Eastern Europe, ccombining the latter with Russia & Central Asia to make a joint Eastern Europe and Central Asia region.
-```{r}
-
-df_regions<- df_regions %>%
-  select(Our.World.In.Data.Name, `region..standard.`, sub.region) %>%
-  rename(Entity = Our.World.In.Data.Name,
-        WID_region = `region..standard.`) %>%
-  mutate(region_alt = if_else(sub.region == 'Western Europe',
-  'Western Europe',
-  WID_region)) %>%
-  mutate(region_alt = if_else((sub.region == 'Eastern Europe' | WID_region == "Russia & Central Asia"),
-  'Eastern Europe & Central Asia',
-  WID_region)) %>%
-  select(Entity, WID_region, region_alt)
-
-kable(
-  df_regions
-)
-```
-
-Check that no countries are not matched for either dataset.
-```{r}
-# Merge in the regions dataset
-df_pip<- left_join(df_pip, df_regions)
-df_wid<- left_join(df_wid, df_regions)
-
-# PIP check
-kable(
-  df_pip %>% 
-    select(Entity, WID_region, region_alt) %>%
-    unique() %>%
-    filter(is.na(WID_region))
-)
-
-# WID check
-kable(
-  df_wid %>% 
-    select(Entity, WID_region, region_alt) %>%
-    unique() %>%
-    filter(is.na(WID_region))
-)
-
-```
-
-Save the region mapping
-```{r}
-write.csv(df_regions, 
-      "PhD_pages/data_appendices/data/clean/WID_region_mapping.csv", row.names=FALSE)
-
-```
-
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Region classfications
-
-```{python}
-# For some reason I can't get the owid package working when rendering... only when running in the interactive window in vs code
 # results = catalog.find('population')
@@
 
-```
+# df_wid<- read.csv("https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/PabloArriagada/WID/data/final/wid_pretax.csv")
 
-Save the population data
-```{python}
-# df_pop.to_csv('PhD_pages/data_appendices/data/clean/population.csv')
+# A local version, in case internet isn't available 
+# df_wid<- read.csv("data/original/wid.csv")
 
 ```
-:::
-
-## Important notes
-
-### Income and consumption data in PIP
-The PIP data includes a mix of income and consumption observations – including both for some country-years.
-
-### Coverage
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Compare countries available across datasets
-```{r}
-
-pip_countries<- df_pip %>%
-  select(Entity) %>%
-  unique() %>%
-  mutate(Entity_pip = Entity)
-
-wid_countries<- df_wid %>%
-  select(Entity) %>%
-  unique() %>%
-  mutate(Entity_wid = Entity)
-
-countries<- full_join(pip_countries, wid_countries, keep = TRUE) %>%
-  select(Entity_pip, Entity_wid)
-
-```
-
-Countries that are in the PIP data but not the WID data:
-```{r}
-kable(
-  countries %>%
-       filter(is.na(Entity_wid))
-)
-```
-Countries that are in the WID data but not the PIP data:
-```{r}
-kable(
-  countries %>%
-       filter(is.na(Entity_pip))
-)
-```
-:::
-
-### Outliers
-::: {.callout-note collapse="true" appearance="minimal"}
-### P90/P10 ratio
-
-For the P90/P10 ratio, a few observations have extreme values. In the PIP data
-```{r}
-kable(
-
-  df_pip %>% arrange(-P90_P10_ratio) %>% head()
-)
-
-```
-
-And the WID data:
-
-```{r}
-kable(
-
-  df_wid %>% arrange(-P90_P10_ratio) %>% head(100)
-)
-
-```
-
-:::

### 2023-01-29 PhD_pages/data_appendices/1_D_prepare_population.qmd
Paper 1, more conversion to python

---
title: "Appendix 1.D: Preparing population data"
format: html
warning: false

---

::: {.callout-note collapse="true" appearance="minimal"}
### Packages and settings
```{python}
from owid import catalog
import pandas as pd

pd.set_option('display.max_rows', None)
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Load population data from Our World in Data

I grab the main population series prepared by Our World in Data.
```{python}

results = catalog.find('population')
owid_pop = results[results['path'] == 'garden/owid/latest/key_indicators/population']

df_pop = owid_pop.iloc[0].load().reset_index().rename(columns={'year':'Year', 'country': 'Entity'})

df_pop.head()

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Check there is population data for each country in WID, PIP, and LIS.

All countries in the PIP data are matched:
```{python}

df_pip = pd.read_csv("data/clean/pip.csv")

check_pip = pd.merge(df_pip, df_pop, how = 'left')
check_pip[check_pip['population'].isna()]

```

There is no population data for Zanzibar, present in the WID data.

```{python}
df_wid = pd.read_csv("data/clean/wid.csv")

check_wid = pd.merge(df_wid, df_pop, how = 'left')
check_wid[check_wid['population'].isna()]

```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Add in Zanzibar population

WID release their own population data, but this currently insn't gathered in our data pipeline. For now I make up a population series for Zanzibar based on two data points from Wikipedia. At some point I will come back and replaces this with the WID population data.

```{python}
zanz_pop_1978 = 476111
zanz_pop_2022 = 1889773

# Assume a constant growth rate
zanz_g = pow(zanz_pop_2022/zanz_pop_1978, 1/(2022-1978))-1

year = 2022
pop = zanz_pop_1978 * pow(1+zanz_g, year - 1978)

df_zanz_pop = df_wid[df_wid['Entity'] == "Zanzibar"].copy()
df_zanz_pop['population'] = zanz_pop_1978 * pow(1+zanz_g, df_zanz_pop['Year'] - 1978)

df_zanz_pop = df_zanz_pop[['Entity', 'Year', 'population']]

df_pop = pd.concat([df_pop, df_zanz_pop])

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### See the prepared data

The first 100 rows of the dataframe looks like this:
```{python}
df_pop.head(100)
```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Save
```{python}
df_pop.to_csv('data/clean/population.csv')

```

:::

### 2023-01-29 PhD_pages/data_appendices/1_E_prepare_region_classifications.qmd
Paper 1, more conversion to python

--- 2023-01-18 PhD_pages/data_appendices/1_A_prepare_PIP_and_WID.qmd
@@
 ---
-title: "Appendix 1.A: Preparing PIP and WID inequality data"
+title: "Appendix 1.E: Preparing country region classifications"
 format: html
 warning: false
+
 ---
 
-# Processing steps
+::: {.callout-note collapse="true" appearance="minimal"}
+### Packages and settings
+```{python}
+from owid import catalog
+import pandas as pd
 
-The titles in each of the boxes below summarise the steps taken to prepare the PIP and WID data. In each case you can expand the box to see the code used.
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Load packages
-
-```{r}
-library(tidyverse)
-library(knitr)
-
-```
-
-```{python}
-# For some reason I can't get the owid package working when rendering... only when running in the interactive window in vs code
-# from owid import catalog
-import pandas as pd
-```
-:::
-
-## PIP data
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Load PIP data
-We load the data from a file prepared from the PIP API by Our World in Data.
-```{r}
-# df_pip<- read.csv("https://raw.githubusercontent.com/owid/poverty-data/main/datasets/pip_dataset.csv")
-
-#  A local version, in case internet isn't available 
-df_pip<- read.csv("PhD_pages/data_appendices/data/original/pip_dataset.csv")
-
-```
-
-The dataframe looks like this:
-```{r}
-kable(
-
-    head(df_pip, 50)
-
-)
+pd.set_option('display.max_rows', None)
 ```
@@
 ::: {.callout-note collapse="true" appearance="minimal"}
-### Drop sub-national data
+### Load WID region classifications
 
-The PIP data includes observations for national, urban and rural populations (`reporting_level`):
+For regional classifications I take WID definitions, as specified on their website [here](https://wid.world/codes-dictionary/#country-code).
 
-```{r}
-kable(
-    df_pip %>%
-        count(reporting_level)
-)
-```
+I read in a csv prepared from the table in that webpage, which has been mapped to the Our World in Data standarized country names used throughout this project.
 
-Here we add a count of the number of reporting levels per country-year.
-```{r}
+```{python}
 
-count_reporting_levels<- df_pip %>%
-    group_by(country, year) %>%
-    count(reporting_level) %>%
-    group_by(country, year) %>%
-    summarise(n_reporting_levels = n())
-
-df_pip<- left_join(df_pip, count_reporting_levels) %>%
-    relocate(n_reporting_levels, .after=reporting_level)
-
-kable(
-
-    head(df_pip, 20)
-
-)
-```
-
-We only want one reporting level for each country-year.
-
-That is already true for most country-years. But some include all three reporting levels.
-```{r}
-kable(
-
-    df_pip %>%
-        count(n_reporting_levels)
-
-)
-
-```
-
-Here are the countries which include all three reporting levels:
-```{r}
-multi_reporting_level_countries<- 
-    df_pip %>%
-        filter(n_reporting_levels == 3) %>%
-        select(country) %>%
-        unique() %>%
-        pull(country)
-
-print(multi_reporting_level_countries)
-
-```
-
-For the country-years with 3 reporting levels, we keep only the national estimates. 
-```{r}
-df_pip<- df_pip %>%
-    filter(n_reporting_levels != 3 | reporting_level == 'national' )
-
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Drop regional data
-
-The dataset includes aggregated estiates for world regions.
-
-Here we define a list of these aggregate entities and drop them from the data.
-```{r}
-drop_aggs<- c(
-  "East Asia and Pacific",
-  "South Asia",
-  "Europe and Central Asia",
-  "High income countries",
-  "Latin America and the Caribbean",
-  "Middle East and North Africa",
-  "Sub-Saharan Africa",
-  "World"
-)
-
-df_pip<- df_pip %>%
-  filter(!country %in% drop_aggs)
-
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Select 2017 PPPs
-
-The data is provided by the World Bank in both 2011 and 2017 PPPs.
-
-We will be using 2017 PPPs for Paper 1.
-
-```{r}
-
-df_pip<- df_pip %>%
-  filter(ppp_version == 2017)
-
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Drop extremely low P10 values
-
-There are some very low values of P10. These lead to exaggerated values of the P90/P10 ratio.
-
-```{r}
-
-kable(
-    df_pip %>% filter(decile1_thr<0.02)
-
-  )
-
-```
-
-We set these values of P10 to NA, and thereby precluding a value for P90/P10 ratio when it is calculated below.
-
-```{r}
-
-df_pip<- df_pip %>%
-  mutate(decile1_thr = if_else(decile1_thr<0.02,
-                              as.numeric(NA),
-                              decile1_thr))
-
-kable(
-    df_pip %>% filter(is.na(decile1_thr))
-
-  )
+df_regions = pd.read_csv("data/original/WID_regions_country_standardized.csv")
 
@@
 ::: {.callout-note collapse="true" appearance="minimal"}
-### Calculate additional variables
+### Define two region classifications
 
-```{r}
+I create a clean region mapping table using two regional definitions.
 
-df_pip<- df_pip %>%
-  mutate(Bottom_50_share = decile1_share + decile2_share + decile3_share + decile4_share + decile5_share,
-          P90_P10_ratio = decile9_thr/decile1_thr)
+One is the 'standard' definitions used by WID.
+
+In an alternative set I split Western and Eastern Europe, ccombining the latter with Russia & Central Asia to make a joint Eastern Europe and Central Asia region.
+
+```{python}
+
+df_regions = df_regions[['Our World In Data Name', 'region (standard)', 'sub-region']].rename(columns={'Our World In Data Name': 'Entity', 'region (standard)': 'WID_region'})
+
+# Define an alt region classification
+df_regions['region_alt'] = df_regions['WID_region']
+
+df_regions.loc[df_regions['sub-region'] == 'Western Europe','region_alt'] = 'Western Europe'
+
+df_regions.loc[(df_regions['sub-region'] == 'Eastern Europe') | (df_regions['WID_region'] == 'Russia & Central Asia'),'region_alt'] = 'Eastern Europe & Central Asia'
+
+df_regions = df_regions[["Entity", "WID_region", "region_alt"]]
 
@@
 ::: {.callout-note collapse="true" appearance="minimal"}
-### Rename and select needed columns
-```{r}
+### See the prepared data
 
-df_pip<- df_pip %>%
-  rename("Gini" = "gini",
-         "Top_10_share" = "decile10_share",
-         "MLD" = "mld",
-         "Entity" = "country",
-         "Year" = "year")
-
-df_pip<- df_pip %>%
-    select(
-        Year,
-        Entity,
-        Gini,
-        Top_10_share,
-        Bottom_50_share,
-        MLD,
-        P90_P10_ratio,
-        welfare_type)
-
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### See the prepared PIP data
-
-The first 100 rows of the dataframe looks like this:
-```{r}
-kable(
-
-    head(df_pip, 100)
-
-)
+```{python}
+df_regions
 ```
@@
 ::: {.callout-note collapse="true" appearance="minimal"}
-### Save the data
-I save the prepared data locally:
-```{r}
-
-write.csv(df_pip, 
-      "PhD_pages/data_appendices/data/clean/pip.csv", row.names=FALSE)
-
-```
-:::
-
-## WID data
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Load WID data
-We load the data from a file prepared from the WID Stata package by Our World in Data.
-
-```{r}
-# df_wid<- read.csv("https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/PabloArriagada/WID/data/final/wid_pretax.csv")
-
-# A local version, in case internet isn't available 
-df_wid<- read.csv("PhD_pages/data_appendices/data/original/wid.csv")
-
-```
-
-The dataframe looks like this:
-```{r}
-kable(
-
-    head(df_wid, 50)
-
-)
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Drop subnational, regional and historical geographical entities
-
-Here we create a list of entities in the WID data that do not correspond to countries in existence today. These we drop from the data.
-```{r}
-drop_entities<- c(
-"China - rural",
-"China - urban",
-"East Germany",
-"Other Russia and Central Asia",
-"Other East Asia",
-"Other Western Europe",
-"Other Latin America",
-"Other MENA",
-"Other South & South-East Asia",
-"Other Sub-Saharan Africa",
-"Africa",
-"Asia",
-"Europe",
-"Oceania",
-"Central Asia",
-"East Africa",
-"East Asia",
-"South Asia",
-"Eastern Europe",
-"Middle Africa",
-"North Africa",
-"North America",
-"South-East Asia",
-"South Africa region",
-"West Africa",
-"West Asia",
-"Western Europe",
-"European Union",
-"World",
-"Asia (excluding Middle East)",
-"North America and Oceania",
-"Sub-Saharan Africa",
-"Latin America",
-"Middle East",
-"MENA",
-"Russia and Central Asia",
-"South & South-East Asia"
-)
-
-df_wid<- df_wid %>%
-  filter(!Entity %in% drop_entities)
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Drop extremely low P10 values
-
-There are some very low values of P10. These lead to exaggerated values of the P90/P10 ratio. Here are all values less than $1 a day (using pretax national income)
-
-```{r}
-
-kable(
-    df_wid %>% filter(P10...income.threshold/365<1)
-
-  )
-
-```
-
-We set these values of P10 to NA, and thereby precluding a value for P90/P10 ratio when it is calculated below.
-
-```{r}
-
-df_wid<- df_wid %>%
-  mutate(P10...income.threshold = if_else(P10...income.threshold/365<1,
-                              as.numeric(NA),
-                              P10...income.threshold))
-
-                              kable(
-    df_wid %>% filter(is.na(P10...income.threshold))
-
-  )
+### Save
+```{python}
+df_regions.to_csv('data/clean/region_mapping.csv')
 
@@
 :::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Calculate some additional inequality metrics
-
-```{r}
-
-#  Calculate P90_P10_ratio
-df_wid<- df_wid %>%
-    mutate(P90_P10_ratio = `P90...income.threshold.of.the.top.10.`/P10...income.threshold)
-
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Rename and select needed variables
-
-```{r}
-df_wid<- df_wid %>%
-    rename("Gini" = "Gini.coefficient",
-         "Top_10_share" = "P90.P100...share.of.the.top.10.",
-         "Bottom_50_share"= "P0.P50...share.of.the.bottom.50.",
-         "P90" = "P90...income.threshold.of.the.top.10.",
-         "P10" = "P10...income.threshold") %>%
-    select(
-        Year,
-        Entity,
-        Gini,
-        Top_10_share,
-        # MLD,
-        P90_P10_ratio,
-        Bottom_50_share,
-        P90,
-        P10)
-
-```
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### See the prepared WID data
-
-The first 100 rows of the dataframe looks like this:
-```{r}
-kable(
-
-    head(df_wid, 100)
-
-)
-```
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Save the data
-I save the prepared data locally:
-```{r}
-
-write.csv(df_wid, 
-      "PhD_pages/data_appendices/data/clean/wid.csv", row.names=FALSE)
-
-```
-:::
-:::
-
-## Additional variables
-::: {.callout-note collapse="true" appearance="minimal"}
-### Region classfications
-For regional classifications I take WID definitions, as specified on their website [here](https://wid.world/codes-dictionary/#country-code).
-
-I read in a csv prepared from the table in that webpage, which has been mapped to the Our World in Data standarized country names used throughout this project.
-```{r}
-
-df_regions<- read.csv("PhD_pages/data_appendices/data/original/WID_regions_country_standardized.csv")
-
-```
-
-I create a clean region mapping table using two regional definitions.
-One is the 'standard' definitions used by WID.
-In an alternative set I split Western and Eastern Europe, ccombining the latter with Russia & Central Asia to make a joint Eastern Europe and Central Asia region.
-```{r}
-
-df_regions<- df_regions %>%
-  select(Our.World.In.Data.Name, `region..standard.`, sub.region) %>%
-  rename(Entity = Our.World.In.Data.Name,
-        WID_region = `region..standard.`) %>%
-  mutate(region_alt = if_else(sub.region == 'Western Europe',
-  'Western Europe',
-  WID_region)) %>%
-  mutate(region_alt = if_else((sub.region == 'Eastern Europe' | WID_region == "Russia & Central Asia"),
-  'Eastern Europe & Central Asia',
-  WID_region)) %>%
-  select(Entity, WID_region, region_alt)
-
-kable(
-  df_regions
-)
-```
-
-Check that no countries are not matched for either dataset.
-```{r}
-# Merge in the regions dataset
-df_pip<- left_join(df_pip, df_regions)
-df_wid<- left_join(df_wid, df_regions)
-
-# PIP check
-kable(
-  df_pip %>% 
-    select(Entity, WID_region, region_alt) %>%
-    unique() %>%
-    filter(is.na(WID_region))
-)
-
-# WID check
-kable(
-  df_wid %>% 
-    select(Entity, WID_region, region_alt) %>%
-    unique() %>%
-    filter(is.na(WID_region))
-)
-
-```
-
-Save the region mapping
-```{r}
-write.csv(df_regions, 
-      "PhD_pages/data_appendices/data/clean/WID_region_mapping.csv", row.names=FALSE)
-
-```
-
-:::
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Region classfications
-
-```{python}
-# For some reason I can't get the owid package working when rendering... only when running in the interactive window in vs code
-# results = catalog.find('population')
-# owid_pop = results[results['path'] == 'garden/owid/latest/key_indicators/population']
-
-# df_pop = owid_pop.iloc[0].load().reset_index().rename(columns={'year':'Year', 'country': 'Entity'})
-
-# df_pop.head()
-
-```
-
-Save the population data
-```{python}
-# df_pop.to_csv('PhD_pages/data_appendices/data/clean/population.csv')
-
-```
-:::
-
-## Important notes
-
-### Income and consumption data in PIP
-The PIP data includes a mix of income and consumption observations – including both for some country-years.
-
-### Coverage
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Compare countries available across datasets
-```{r}
-
-pip_countries<- df_pip %>%
-  select(Entity) %>%
-  unique() %>%
-  mutate(Entity_pip = Entity)
-
-wid_countries<- df_wid %>%
-  select(Entity) %>%
-  unique() %>%
-  mutate(Entity_wid = Entity)
-
-countries<- full_join(pip_countries, wid_countries, keep = TRUE) %>%
-  select(Entity_pip, Entity_wid)
-
-```
-
-Countries that are in the PIP data but not the WID data:
-```{r}
-kable(
-  countries %>%
-       filter(is.na(Entity_wid))
-)
-```
-Countries that are in the WID data but not the PIP data:
-```{r}
-kable(
-  countries %>%
-       filter(is.na(Entity_pip))
-)
-```
-:::
-
-### Outliers
-::: {.callout-note collapse="true" appearance="minimal"}
-### P90/P10 ratio
-
-For the P90/P10 ratio, a few observations have extreme values. In the PIP data
-```{r}
-kable(
-
-  df_pip %>% arrange(-P90_P10_ratio) %>% head()
-)
-
-```
-
-And the WID data:
-
-```{r}
-kable(
-
-  df_wid %>% arrange(-P90_P10_ratio) %>% head(100)
-)
-
-```
-
-:::

### 2023-01-29 PhD_pages/data_appendices/1_F_within_country_trends.qmd
Paper 1, more conversion to python

--- 2023-01-27 PhD_pages/data_appendices/1_B_within_country_trends.qmd
@@
 ---
-title: "Appendix 1.B: Analysing global trends in within-country inequality"
+title: "Appendix 1.F: Analysing global trends in within-country inequality"
 format: html
@@
 
+This second block of code we read in the data (this needs to be done in a different way in the shiny app...)
+
+```{python}
+
+#  ---------------------------------
+#  ---------- SECTION 2: Read in data -------------
+#  ----------------------------------
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'
+
+df_pip = pd.read_csv(url)
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'
+
+df_wid = pd.read_csv(url)
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/region_mapping.csv'
+
+df_regions = pd.read_csv(url)
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'
+
+df_pop = pd.read_csv(url)
+
+```
+
+In a third block, we set specifications for the data matching. This is different in the app because some parameters are controlled by the app.
+
 ```{python}
 #  ---------------------------------
-#  ---------- SECTION 2: Specifications -------------
+#  ---------- SECTION 3: Specifications -------------
 #  ----------------------------------
@@
 
-This second block of code we read in the data (this needs to be done in a different way in )
-
-```{python}
-
-#  ---------------------------------
-#  ---------- SECTION 3: Read in data -------------
-#  ----------------------------------
-
-url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'
-
-df_pip = pd.read_csv(url)
-
-url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'
-
-df_wid = pd.read_csv(url)
-
-url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/WID_region_mapping.csv'
-
-df_regions = pd.read_csv(url)
-
-url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'
-
-df_pop = pd.read_csv(url)
-
-```
+In this code block we run the code and prepare final plots and tables.
 
@@
 ```{shinylive-python}
-#| standalone: true
-#| viewerHeight: 420
+# #| standalone: true
+# #| viewerHeight: 420
 
-from shiny import *
-from plotnine import *
-from pyodide.http import open_url
-import pandas as pd
-import numpy as np
+# from shiny import *
+# from plotnine import *
+# from pyodide.http import open_url
+# import pandas as pd
+# import numpy as np
 
-#  ---------------------------------
-#  ---------- Data prep functions -------------
-#  ----------------------------------
+# #  ---------------------------------
+# #  ---------- Data prep functions -------------
+# #  ----------------------------------
 
-#  Function get matching for ref years
-def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):
+# #  Function get matching for ref years
+# def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):
 
-  df = df.loc[:, [reference_col, group_by_col, value_col]]
+#   df = df.loc[:, [reference_col, group_by_col, value_col]]
 
-  # Drop NAs
-  df = df.dropna()
+#   # Drop NAs
+#   df = df.dropna()
 
-  # Calculate absolute distance from reference value
-  df['ref_diff'] = abs(df[reference_col] - reference_val)
+#   # Calculate absolute distance from reference value
+#   df['ref_diff'] = abs(df[reference_col] - reference_val)
 
-  # Drop any rows with a distance beyond threshold
-  if not pd.isna(max_dist_from_ref):
-    df = df.loc[df['ref_diff'] <= max_dist_from_ref]
+#   # Drop any rows with a distance beyond threshold
+#   if not pd.isna(max_dist_from_ref):
+#     df = df.loc[df['ref_diff'] <= max_dist_from_ref]
 
-  # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
-  df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)
+#   # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
+#   df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)
 
-  # Settle tie-breaks
-  if tie_break == 'below':
-    df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)
+#   # Settle tie-breaks
+#   if tie_break == 'below':
+#     df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)
 
-  elif tie_break == 'above':
-    df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)
+#   elif tie_break == 'above':
+#     df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)
 
-  df = df.drop('ref_diff', axis=1)
+#   df = df.drop('ref_diff', axis=1)
 
-  return df
+#   return df
 
-# Merge matches for different reference points
-def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+# # Merge matches for different reference points
+# def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
 
-# Make sure the pair of reference values are in ascending order
-  reference_vals.sort()
+# # Make sure the pair of reference values are in ascending order
+#   reference_vals.sort()
 
-# Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref
+# # Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref
 
-# Find matches for lower reference value
-  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs, reference_col, group_by_col, value_col, 'below')
+# # Find matches for lower reference value
+#   lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs, reference_col, group_by_col, value_col, 'below')
 
-# Find matches for higher reference value
-  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs, reference_col, group_by_col, value_col, 'above')
+# # Find matches for higher reference value
+#   higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs, reference_col, group_by_col, value_col, 'above')
 
-# Merge the two sets of matches
-  merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))
+# # Merge the two sets of matches
+#   merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))
 
-# Drop obs that do not have data for both ref values
-  merged_df = merged_df.dropna()
+# # Drop obs that do not have data for both ref values
+#   merged_df = merged_df.dropna()
 
-# Drop obs where the matched data does not meet the min distance requirement
-  if not pd.isna(min_dist_between):
+# # Drop obs where the matched data does not meet the min distance requirement
+#   if not pd.isna(min_dist_between):
 
-  # Store the names of the reference column returned from the two matches
-    ref_var_high = f'{reference_col}{reference_vals[1]}'
-    ref_var_low = f'{reference_col}{reference_vals[0]}'
+#   # Store the names of the reference column returned from the two matches
+#     ref_var_high = f'{reference_col}{reference_vals[1]}'
+#     ref_var_low = f'{reference_col}{reference_vals[0]}'
 
-  # Keep only rows >= to the min distance
-    merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]
+#   # Keep only rows >= to the min distance
+#     merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]
 
-  return merged_df
+#   return merged_df
 
-  # For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
-def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+#   # For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
+# def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
 
-    # Specify the name of the column in which the income/consumption welfare definition is stored
-    welfare_colname = 'welfare_type'
+#     # Specify the name of the column in which the income/consumption welfare definition is stored
+#     welfare_colname = 'welfare_type'
 
-    # Creat dataframes for thee scenarios:
-    # Scenario 1: only allow income data
-    df_inc_filter = df.loc[df[welfare_colname] == "income", :]
-    df_inc_filter.name = "Income"
+#     # Creat dataframes for thee scenarios:
+#     # Scenario 1: only allow income data
+#     df_inc_filter = df.loc[df[welfare_colname] == "income", :]
+#     df_inc_filter.name = "Income"
 
-    # Scenario 2: only allow consumption data
-    df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
-    df_cons_filter.name = "Consumption"
-    # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
-    df_mixed = df.copy()
+#     # Scenario 2: only allow consumption data
+#     df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
+#     df_cons_filter.name = "Consumption"
+#     # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
+#     df_mixed = df.copy()
 
-    df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')
+#     df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')
 
-    df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]
+#     df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]
 
-    df_mixed.name = "Mixed"
-    #  Store the scneario dataframes in a list
-    df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]
+#     df_mixed.name = "Mixed"
+#     #  Store the scneario dataframes in a list
+#     df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]
 
-    # Run the matching function on each scenario
-    scenario_matches = [merge_two_ref_matches(
-      df_scenario, 
-      reference_vals, 
-      max_dist_from_refs, 
-      min_dist_between, 
-      reference_col, 
-      group_by_col, 
-      value_col) for df_scenario in df_scenarios]
+#     # Run the matching function on each scenario
+#     scenario_matches = [merge_two_ref_matches(
+#       df_scenario, 
+#       reference_vals, 
+#       max_dist_from_refs, 
+#       min_dist_between, 
+#       reference_col, 
+#       group_by_col, 
+#       value_col) for df_scenario in df_scenarios]
 
-    # Combine the first two scenarios.
-    df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])
+#     # Combine the first two scenarios.
+#     df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])
 
-    # Tidy up indexes
-    df_combined_matches = df_combined_matches.reset_index()
+#     # Tidy up indexes
+#     df_combined_matches = df_combined_matches.reset_index()
 
-    df_combined_matches = df_combined_matches.drop('level_1', axis=1)
+#     df_combined_matches = df_combined_matches.drop('level_1', axis=1)
 
-    df_combined_matches = df_combined_matches\
-      .rename(columns={"level_0": "pip_welfare"})
+#     df_combined_matches = df_combined_matches\
+#       .rename(columns={"level_0": "pip_welfare"})
 
-    # Add in third scenario.
-    df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])
+#     # Add in third scenario.
+#     df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])
 
-    # add scenario name to te pip_welfare column
-    df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)
+#     # add scenario name to te pip_welfare column
+#     df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)
 
-    # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
-      # First count the matches
-    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
-      # Then drop any matches from the lowest priority where there are multiple matches
-    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
-      #  Repeat at the next level of priority
-    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
-    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]
+#     # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
+#       # First count the matches
+#     df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
+#       # Then drop any matches from the lowest priority where there are multiple matches
+#     df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
+#       #  Repeat at the next level of priority
+#     df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
+#     df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]
 
-    # Drop the match count column
-    df_combined_matches = df_combined_matches.drop('match_count', axis=1)
+#     # Drop the match count column
+#     df_combined_matches = df_combined_matches.drop('match_count', axis=1)
 
-    return df_combined_matches
+#     return df_combined_matches
 
-#  ---------------------------------
-#  ---------- Specifications -------------
-#  ----------------------------------
+# #  ---------------------------------
+# #  ---------- Specifications -------------
+# #  ----------------------------------
 
-# These parameters are not controlled in the app.
-reference_col = 'Year'
-group_by_col = 'Entity'
+# # These parameters are not controlled in the app.
+# reference_col = 'Year'
+# group_by_col = 'Entity'
 
-tolerance = 1
-outlier_cut_off_upper = None
+# tolerance = 1
+# outlier_cut_off_upper = None
 
-source_count_requirement = 2
+# source_count_requirement = 2
 
-#  ---------------------------------
-#  ---------- App code -------------
-#  ----------------------------------
+# #  ---------------------------------
+# #  ---------- App code -------------
+# #  ----------------------------------
 
-def panel_box(*args, **kwargs):
-    return ui.div(
-        ui.div(*args, class_="card-body"),
-        **kwargs,
-        class_="card mb-3",
-    )
+# def panel_box(*args, **kwargs):
+#     return ui.div(
+#         ui.div(*args, class_="card-body"),
+#         **kwargs,
+#         class_="card mb-3",
+#     )
 
-app_ui = ui.page_fluid(
-    {"class": "p-4"},
-    ui.row(
-        ui.column(
-            4,
-            panel_box(
-                ui.input_slider("reference_vals", "Reference years", 1975, 2020, value=[1990,2018]),
-                ui.input_numeric(
-                            "max_dist_from_refs",
-                            "=/- (years)",
-                            4,
-                        ),
-                ui.input_numeric(
-                            "min_dist_between",
-                            "Min distance between observations (years)",
-                            23,
-                        ),
-                ui.input_action_button(
-                    "download", "Download data and image", class_="btn-primary w-100"
-                ),
-            ),
-            ui.navset_tab_card(
-                ui.nav(
-                    "Compare sources",
-                    ui.input_select(
-                "metric",
-                "Metric",
-                {
-                    "Gini": "Gini",
-                    "Top_10_share": "Top 10p.c. share",
-                },
-            ),
-            ),
-                ui.nav(
-                  "Compare metrics"),
-            ),
-        ),
-        ui.column(
-            8,
-            ui.output_plot("example_plot")
-            ),
-    ),
-)
+# app_ui = ui.page_fluid(
+#     {"class": "p-4"},
+#     ui.row(
+#         ui.column(
+#             4,
+#             panel_box(
+#                 ui.input_slider("reference_vals", "Reference years", 1975, 2020, value=[1990,2018]),
+#                 ui.input_numeric(
+#                             "max_dist_from_refs",
+#                             "=/- (years)",
+#                             4,
+#                         ),
+#                 ui.input_numeric(
+#                             "min_dist_between",
+#                             "Min distance between observations (years)",
+#                             23,
+#                         ),
+#                 ui.input_action_button(
+#                     "download", "Download data and image", class_="btn-primary w-100"
+#                 ),
+#             ),
+#             ui.navset_tab_card(
+#                 ui.nav(
+#                     "Compare sources",
+#                     ui.input_select(
+#                 "metric",
+#                 "Metric",
+#                 {
+#                     "Gini": "Gini",
+#                     "Top_10_share": "Top 10p.c. share",
+#                 },
+#             ),
+#             ),
+#                 ui.nav(
+#                   "Compare metrics"),
+#             ),
+#         ),
+#         ui.column(
+#             8,
+#             ui.output_plot("example_plot")
+#             ),
+#     ),
+# )
 
-def server(input, output, session):
-    @output
-    @render.plot
-    def example_plot():
+# def server(input, output, session):
+#     @output
+#     @render.plot
+#     def example_plot():
 
-      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'
+#       url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'
 
-      df_pip = pd.read_csv(open_url(url))
+#       df_pip = pd.read_csv(open_url(url))
 
-      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'
+#       url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'
 
-      df_wid = pd.read_csv(open_url(url))
+#       df_wid = pd.read_csv(open_url(url))
 
-      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/WID_region_mapping.csv'
+#       url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/region_mapping.csv'
 
-      df_regions = pd.read_csv(open_url(url))
+#       df_regions = pd.read_csv(open_url(url))
 
-      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'
+#       url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'
 
-      df_pop = pd.read_csv(open_url(url))
+#       df_pop = pd.read_csv(open_url(url))
 
-      def prep_wid_pip_data(reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+#       def prep_wid_pip_data(reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
 
-        pip_matches = pip_welfare_routine(
-          df = df_pip,
-          reference_vals = reference_vals,
-          max_dist_from_refs = max_dist_from_refs,
-          min_dist_between = min_dist_between,
-          reference_col = reference_col,
-          group_by_col = group_by_col,
-          value_col = value_col
-        )
+#         pip_matches = pip_welfare_routine(
+#           df = df_pip,
+#           reference_vals = reference_vals,
+#           max_dist_from_refs = max_dist_from_refs,
+#           min_dist_between = min_dist_between,
+#           reference_col = reference_col,
+#           group_by_col = group_by_col,
+#           value_col = value_col
+#         )
 
-        wid_matches = merge_two_ref_matches(
-          df = df_wid,
-          reference_vals = reference_vals,
-          max_dist_from_refs = max_dist_from_refs,
-          min_dist_between = min_dist_between,
-          reference_col = reference_col,
-          group_by_col = group_by_col,
-          value_col = value_col
-        )
+#         wid_matches = merge_two_ref_matches(
+#           df = df_wid,
+#           reference_vals = reference_vals,
+#           max_dist_from_refs = max_dist_from_refs,
+#           min_dist_between = min_dist_between,
+#           reference_col = reference_col,
+#           group_by_col = group_by_col,
+#           value_col = value_col
+#         )
 
-        ref_pairs = pd.concat([pip_matches, wid_matches,], keys=['pip', 'wid'])
+#         ref_pairs = pd.concat([pip_matches, wid_matches,], keys=['pip', 'wid'])
 
-        # Tidy up indexes
-        ref_pairs = ref_pairs.reset_index()
+#         # Tidy up indexes
+#         ref_pairs = ref_pairs.reset_index()
 
-        ref_pairs = ref_pairs.drop('level_1', axis=1)
+#         ref_pairs = ref_pairs.drop('level_1', axis=1)
 
-        ref_pairs = ref_pairs\
-          .rename(columns={"level_0": "source"})
+#         ref_pairs = ref_pairs\
+#           .rename(columns={"level_0": "source"})
 
-        return ref_pairs
+#         return ref_pairs
 
-      def prep_plot_and_tables(
-        reference_vals, 
-        max_dist_from_refs,
-        min_dist_between,
-        reference_col,
-        group_by_col,
-        value_col,
-        tolerance,
-        outlier_cut_off_upper,
-        source_count_requirement):
+#       def prep_plot_and_tables(
+#         reference_vals, 
+#         max_dist_from_refs,
+#         min_dist_between,
+#         reference_col,
+#         group_by_col,
+#         value_col,
+#         tolerance,
+#         outlier_cut_off_upper,
+#         source_count_requirement):
 
-        # Other spec:
-        region_col = 'region_alt'
+#         # Other spec:
+#         region_col = 'region_alt'
 
-        # ---- Prep data ---------
+#         # ---- Prep data ---------
 
-        plot_data = prep_wid_pip_data(
-          reference_vals = reference_vals, 
-          max_dist_from_refs = max_dist_from_refs,
-          min_dist_between = min_dist_between,
-          reference_col = reference_col,
-          group_by_col = group_by_col,
-          value_col = value_col,
-        )
+#         plot_data = prep_wid_pip_data(
+#           reference_vals = reference_vals, 
+#           max_dist_from_refs = max_dist_from_refs,
+#           min_dist_between = min_dist_between,
+#           reference_col = reference_col,
+#           group_by_col = group_by_col,
+#           value_col = value_col,
+#         )
 
-        # Store the names of the columns to be used onthe X and Y axis
-        x_axis = f'{value_col}{reference_vals[0]}'
-        y_axis = f'{value_col}{reference_vals[1]}'
+#         # Store the names of the columns to be used onthe X and Y axis
+#         x_axis = f'{value_col}{reference_vals[0]}'
+#         y_axis = f'{value_col}{reference_vals[1]}'
 
-        x_ref_val = f'{reference_col}{reference_vals[0]}'
-        y_ref_val = f'{reference_col}{reference_vals[1]}'
+#         x_ref_val = f'{reference_col}{reference_vals[0]}'
+#         y_ref_val = f'{reference_col}{reference_vals[1]}'
 
-        # Apply outlier cut off, if specified
-        if not pd.isna(outlier_cut_off_upper):
-          plot_data = plot_data.loc[plot_data[x_axis] <= outlier_cut_off_upper]
-          plot_data = plot_data.loc[plot_data[y_axis] <= outlier_cut_off_upper]
+#         # Apply outlier cut off, if specified
+#         if not pd.isna(outlier_cut_off_upper):
+#           plot_data = plot_data.loc[plot_data[x_axis] <= outlier_cut_off_upper]
+#           plot_data = plot_data.loc[plot_data[y_axis] <= outlier_cut_off_upper]
 
-        # Add a count by country – showing whether data is available from both sources or not
-        plot_data['source_count'] = plot_data.groupby(group_by_col)['source'].transform('count')
+#         # Add a count by country – showing whether data is available from both sources or not
+#         plot_data['source_count'] = plot_data.groupby(group_by_col)['source'].transform('count')
 
-        # Apply source requirement (whether to include only observations with data from both sources)
-        plot_data = plot_data.loc[plot_data['source_count'] >= source_count_requirement]
+#         # Apply source requirement (whether to include only observations with data from both sources)
+#         plot_data = plot_data.loc[plot_data['source_count'] >= source_count_requirement]
 
-        # Drop source_count column
-        plot_data = plot_data.drop('source_count', axis=1)
+#         # Drop source_count column
+#         plot_data = plot_data.drop('source_count', axis=1)
 
-        # Add in region classification
-        plot_data = pd.merge(plot_data, df_regions, how = 'left')
+#         # Add in region classification
+#         plot_data = pd.merge(plot_data, df_regions, how = 'left')
 
-        # Add in population data
-          # For first ref
-        df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[0], ['Entity', 'population'] ]
+#         # Add in population data
+#           # For first ref
+#         df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[0], ['Entity', 'population'] ]
 
-        plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')
+#         plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')
 
-        plot_data = plot_data.rename(columns={'population':'population_ref1'})
+#         plot_data = plot_data.rename(columns={'population':'population_ref1'})
 
-          # For second ref
-        df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[1], ['Entity', 'population'] ]
+#           # For second ref
+#         df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[1], ['Entity', 'population'] ]
 
-        plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')
+#         plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')
 
-        plot_data = plot_data.rename(columns={'population':'population_ref2'})
+#         plot_data = plot_data.rename(columns={'population':'population_ref2'})
 
-          # Calculate average population two reference periods
-        plot_data['avg_pop'] = (plot_data['population_ref1'] + plot_data['population_ref2'])/2
+#           # Calculate average population two reference periods
+#         plot_data['avg_pop'] = (plot_data['population_ref1'] + plot_data['population_ref2'])/2
 
-        # ----- Prep plot ------
+#         # ----- Prep plot ------
 
-        # I grab what I am guessing to be the max and min tick marks on the x axist, in order to define the coordinates of a 'tolerance' ribbon.
-        x_min = plot_data[x_axis].min()
-        x_min_floor = np.floor(x_min * 10) / 10
+#         # I grab what I am guessing to be the max and min tick marks on the x axist, in order to define the coordinates of a 'tolerance' ribbon.
+#         x_min = plot_data[x_axis].min()
+#         x_min_floor = np.floor(x_min * 10) / 10
 
-        x_max = plot_data[x_axis].max()
-        x_max_ceiling = np.ceil(x_max * 10) / 10
+#         x_max = plot_data[x_axis].max()
+#         x_max_ceiling = np.ceil(x_max * 10) / 10
 
-        # Set the coordinates of the shaded ribbon
-        shaded_coord = pd.DataFrame({'x': [x_min_floor, x_max_ceiling, x_max_ceiling, x_min_floor], 
-          'y': [x_min_floor-tolerance, x_max_ceiling-tolerance, x_max_ceiling+tolerance , x_min_floor+tolerance]})
+#         # Set the coordinates of the shaded ribbon
+#         shaded_coord = pd.DataFrame({'x': [x_min_floor, x_max_ceiling, x_max_ceiling, x_min_floor], 
+#           'y': [x_min_floor-tolerance, x_max_ceiling-tolerance, x_max_ceiling+tolerance , x_min_floor+tolerance]})
 
-        plot = (ggplot(plot_data
-        , aes(x_axis, y_axis, alpha=0.5))
-        + geom_point(aes(color=region_col, size='avg_pop', alpha = 0.6))
-        + facet_wrap('~ source')
-        + geom_polygon(aes(x='x', y='y', alpha = 0.2), data=shaded_coord)
-        + theme_light()
-        + theme(legend_position='none')
-        )
+#         plot = (ggplot(plot_data
+#         , aes(x_axis, y_axis, alpha=0.5))
+#         + geom_point(aes(color=region_col, size='avg_pop', alpha = 0.6))
+#         + facet_wrap('~ source')
+#         + geom_polygon(aes(x='x', y='y', alpha = 0.2), data=shaded_coord)
+#         + theme_light()
+#         + theme(legend_position='none')
+#         )
 
-        # ---- Prep summary table ----
+#         # ---- Prep summary table ----
 
-        # Calculate the change between the two ref periods
-        plot_data['change'] = (plot_data[y_axis] - plot_data[x_axis])
+#         # Calculate the change between the two ref periods
+#         plot_data['change'] = (plot_data[y_axis] - plot_data[x_axis])
 
-        group_yes_vars = ['source', region_col]
-        group_no_vars = ['source']
+#         group_yes_vars = ['source', region_col]
+#         group_no_vars = ['source']
 
-        group_scenarios = [group_no_vars, group_yes_vars]
+#         group_scenarios = [group_no_vars, group_yes_vars]
 
-        aggs = []
+#         aggs = []
 
-        for group_vars in group_scenarios:
+#         for group_vars in group_scenarios:
 
-        # Prepare pop-weighted average change
-          agg_level = plot_data.copy()
+#         # Prepare pop-weighted average change
+#           agg_level = plot_data.copy()
 
-          # Calulate pop weights by source and region
-          agg_level['regional_pop_weights'] = agg_level.groupby(group_vars)['avg_pop'].transform(lambda x: x/x.sum())
+#           # Calulate pop weights by source and region
+#           agg_level['regional_pop_weights'] = agg_level.groupby(group_vars)['avg_pop'].transform(lambda x: x/x.sum())
 
-          agg_level['global_pop_weights'] = agg_level.groupby('source')['avg_pop'].transform(lambda x: x/x.sum())
+#           agg_level['global_pop_weights'] = agg_level.groupby('source')['avg_pop'].transform(lambda x: x/x.sum())
 
-          # Multiply the change by the pop weights 
-          agg_level['region_weighted_change'] = agg_level['change'] * agg_level['regional_pop_weights']
+#           # Multiply the change by the pop weights 
+#           agg_level['region_weighted_change'] = agg_level['change'] * agg_level['regional_pop_weights']
 
-        # Prepare fall/stable/rise categories
-          agg_level['fall'] = agg_level['change'] < -tolerance
+#         # Prepare fall/stable/rise categories
+#           agg_level['fall'] = agg_level['change'] < -tolerance
 
-          agg_level['stable'] = (agg_level['change'] <= tolerance) & (agg_level['change'] >= -tolerance)
+#           agg_level['stable'] = (agg_level['change'] <= tolerance) & (agg_level['change'] >= -tolerance)
 
-          agg_level['rise'] = agg_level['change'] > tolerance
+#           agg_level['rise'] = agg_level['change'] > tolerance
 
-          agg_level[['fall', 'stable', 'rise']] = agg_level[['fall', 'stable', 'rise']].astype(int)
+#           agg_level[['fall', 'stable', 'rise']] = agg_level[['fall', 'stable', 'rise']].astype(int)
 
-          agg_level = agg_level.groupby(group_vars)[['change', 'region_weighted_change', 'global_pop_weights', 'fall', 'stable', 'rise']]\
-            .agg(['sum', 'mean', 'count'])
+#           agg_level = agg_level.groupby(group_vars)[['change', 'region_weighted_change', 'global_pop_weights', 'fall', 'stable', 'rise']]\
+#             .agg(['sum', 'mean', 'count'])
 
-        # Aggregate by source and region
-          if group_vars == group_yes_vars:
+#         # Aggregate by source and region
+#           if group_vars == group_yes_vars:
 
-            agg_level = agg_level\
-              .unstack().T.reset_index()
+#             agg_level = agg_level\
+#               .unstack().T.reset_index()
 
-          elif group_vars == group_no_vars:
+#           elif group_vars == group_no_vars:
 
-            agg_level = agg_level\
-              .T.reset_index()
+#             agg_level = agg_level\
+#               .T.reset_index()
 
-            agg_level[region_col] = "World"
+#             agg_level[region_col] = "World"
 
-          aggs.append(agg_level)
+#           aggs.append(agg_level)
 
-        df_summary = pd.concat(aggs)
+#         df_summary = pd.concat(aggs)
 
-          # filter for the aggregations we need and label in the 'summary' column
-        df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'mean'), 'summary'] = 'avg change'
+#           # filter for the aggregations we need and label in the 'summary' column
+#         df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'mean'), 'summary'] = 'avg change'
 
-        df_summary.loc[(df_summary['level_0'] == 'region_weighted_change') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Regional pop-weighted avg change'
+#         df_summary.loc[(df_summary['level_0'] == 'region_weighted_change') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Regional pop-weighted avg change'
 
-        df_summary.loc[(df_summary['level_0'] == 'global_pop_weights') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Global pop weights'
+#         df_summary.loc[(df_summary['level_0'] == 'global_pop_weights') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Global pop weights'
 
-        df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'count'), 'summary'] = 'n'
+#         df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'count'), 'summary'] = 'n'
 
-        df_summary.loc[(df_summary['level_0'] == 'fall') & (df_summary['level_1'] == 'sum'), 'summary'] = 'fall'
+#         df_summary.loc[(df_summary['level_0'] == 'fall') & (df_summary['level_1'] == 'sum'), 'summary'] = 'fall'
 
-        df_summary.loc[(df_summary['level_0'] == 'stable') & (df_summary['level_1'] == 'sum'), 'summary'] = 'stable'
+#         df_summary.loc[(df_summary['level_0'] == 'stable') & (df_summary['level_1'] == 'sum'), 'summary'] = 'stable'
 
-        df_summary.loc[(df_summary['level_0'] == 'rise') & (df_summary['level_1'] == 'sum'), 'summary'] = 'rise'
+#         df_summary.loc[(df_summary['level_0'] == 'rise') & (df_summary['level_1'] == 'sum'), 'summary'] = 'rise'
 
-        df_summary = df_summary[df_summary['summary'].notnull()]
+#         df_summary = df_summary[df_summary['summary'].notnull()]
 
-        df_summary = df_summary.drop(['level_0', 'level_1'], axis = 1)
+#         df_summary = df_summary.drop(['level_0', 'level_1'], axis = 1)
 
-        df_summary['summary'] = pd.Categorical(df_summary['summary'], categories=['fall', 'stable', 'rise','n', 'avg change', 'Regional pop-weighted avg change', 'Global pop weights'], ordered=True)
+#         df_summary['summary'] = pd.Categorical(df_summary['summary'], categories=['fall', 'stable', 'rise','n', 'avg change', 'Regional pop-weighted avg change', 'Global pop weights'], ordered=True)
 
-        df_summary = df_summary.sort_values([region_col,'summary']).set_index([region_col,'summary'])
+#         df_summary = df_summary.sort_values([region_col,'summary']).set_index([region_col,'summary'])
 
-        # format to show appropriate d.p. I run a loop over each row for want of figuring out a better way!
-        repeat_length = 7
-        for i in range(0, len(df_summary.index.levels[0].unique())):
-          df_summary.iloc[i*repeat_length,:] = df_summary.iloc[i*repeat_length,:].map('{:.0f}'.format)
-          df_summary.iloc[1 + i*repeat_length,:] = df_summary.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
-          df_summary.iloc[2 + i*repeat_length,:] = df_summary.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
-          df_summary.iloc[3 + i*repeat_length,:] = df_summary.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
-          df_summary.iloc[4 + i*repeat_length,:] = df_summary.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)
-          df_summary.iloc[5 + i*repeat_length,:] = df_summary.iloc[5 + i*repeat_length,:].map('{:.2f}'.format)
-          df_summary.iloc[6 + i*repeat_length,:] = df_summary.iloc[6 + i*repeat_length,:].map('{:.2f}'.format)
+#         # format to show appropriate d.p. I run a loop over each row for want of figuring out a better way!
+#         repeat_length = 7
+#         for i in range(0, len(df_summary.index.levels[0].unique())):
+#           df_summary.iloc[i*repeat_length,:] = df_summary.iloc[i*repeat_length,:].map('{:.0f}'.format)
+#           df_summary.iloc[1 + i*repeat_length,:] = df_summary.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
+#           df_summary.iloc[2 + i*repeat_length,:] = df_summary.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
+#           df_summary.iloc[3 + i*repeat_length,:] = df_summary.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
+#           df_summary.iloc[4 + i*repeat_length,:] = df_summary.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)
+#           df_summary.iloc[5 + i*repeat_length,:] = df_summary.iloc[5 + i*repeat_length,:].map('{:.2f}'.format)
+#           df_summary.iloc[6 + i*repeat_length,:] = df_summary.iloc[6 + i*repeat_length,:].map('{:.2f}'.format)
 
-        # ----- Return outputs ------
-        plot_and_tables = {
-          "table_data": plot_data,
-          "plot": plot,
-          "summary": df_summary
-        }
+#         # ----- Return outputs ------
+#         plot_and_tables = {
+#           "table_data": plot_data,
+#           "plot": plot,
+#           "summary": df_summary
+#         }
 
-        return plot_and_tables
+#         return plot_and_tables
 
-      plot_and_tables = prep_plot_and_tables(
-          reference_vals = list(input.reference_vals()), 
-          max_dist_from_refs = input.max_dist_from_refs(),
-          min_dist_between = input.min_dist_between(),
-          reference_col = reference_col,
-          group_by_col = group_by_col,
-          value_col = input.metric(),
-          tolerance = tolerance,
-          outlier_cut_off_upper = outlier_cut_off_upper,
-          source_count_requirement = source_count_requirement
-          )
+#       plot_and_tables = prep_plot_and_tables(
+#           reference_vals = list(input.reference_vals()), 
+#           max_dist_from_refs = input.max_dist_from_refs(),
+#           min_dist_between = input.min_dist_between(),
+#           reference_col = reference_col,
+#           group_by_col = group_by_col,
+#           value_col = input.metric(),
+#           tolerance = tolerance,
+#           outlier_cut_off_upper = outlier_cut_off_upper,
+#           source_count_requirement = source_count_requirement
+#           )
 
-      plot = plot_and_tables['plot']
+#       plot = plot_and_tables['plot']
 
-      return plot
+#       return plot
 
-app = App(app_ui, server)
+# app = App(app_ui, server)
 

### 2023-01-29 PhD_pages/data_appendices/1_E_prepare_region_classifications.qmd
Paper 1: Adding LIS data (and restructuring data prep files)

@@
 ```{python}
-from owid import catalog
 import pandas as pd
-
 pd.set_option('display.max_rows', None)

### 2023-01-29 PhD_pages/data_appendices/1_F_within_country_trends.qmd
Paper 1: Adding LIS data (and restructuring data prep files)

--- 2023-01-27 PhD_pages/data_appendices/1_B_within_country_trends.qmd
@@
 
-      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/WID_region_mapping.csv'
+      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/region_mapping.csv'
 

### 2023-01-29 PhD_pages/data_appendices/1_F_within_country_trends.qmd
Paper 1: Adding LIS data into the mix

@@
 
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/lis.csv'
+
+df_lis = pd.read_csv(url)
+
 url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/region_mapping.csv'
@@
 data = {"PIP": df_pip,
-        "WID": df_wid}
+        "WID": df_wid,
+        "LIS": df_lis}
 
 data_to_match = {
-  "sources": ["PIP", "WID", "PIP"],
-  "var_names": ["Gini", "Top_10_share", "Top_10_share"]
+  "sources": ["PIP", "WID", "LIS"],
+  "var_names": ["Top_10_share", "Top_10_share_pretax", "Top_10_share_equiv_disposable"]
   }

### 2023-02-01 PhD_pages/data_appendices/1_F_within_country_trends.qmd
Paper 1: shifting adapted code back into shiny app

@@
 #  ---------------------------------
-#  ---------- Data prep functions -------------
+#  ---------- SECTION 1: Data prep functions -------------
 #  ----------------------------------
@@
   df = df.drop('ref_diff', axis=1)
+
+  df = df\
+    .rename(columns={value_col: "value"})
 
@@
 #  ---------------------------------
-#  ---------- Specifications -------------
+#  ---------- SECTION 2: Read in data -------------
+#  ----------------------------------
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'
+
+df_pip = pd.read_csv(open_url(url))
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'
+
+df_wid = pd.read_csv(open_url(url))
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/lis.csv'
+
+df_lis = pd.read_csv(open_url(url))
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/region_mapping.csv'
+
+df_regions = pd.read_csv(open_url(url))
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'
+
+df_pop = pd.read_csv(open_url(url))
+
+#  ---------------------------------
+#  ---------- SECTION 3: Specifications -------------
 #  ----------------------------------
@@
 
-source_count_requirement = 2
+data = {"PIP": df_pip,
+        "WID": df_wid,
+        "LIS": df_lis}
+
+# make an empty list to add all the varnames
+all_vars = list()
+
+# A string to use to separate the source name from the var name
+sep = ": "
+
+# for each df in 'data'
+for k,v in data.items():
+
+  # make a list of varnames
+  vars = list(v.columns)
+
+  # remove the id cols
+  vars.remove(reference_col)
+  vars.remove(group_by_col)
+
+  # append the name of the source (the key of 'data') at the front
+  vars = [f'{k}{sep}' + var for var in vars]
+
+  # append to the running list
+  all_vars = all_vars + vars
 
@@
                         ),
+                ui.input_selectize(
+                  "metrics", "Metrics", all_vars, multiple=True
+                  ),
+                ui.input_checkbox("all_metrics_requirement", "Show only countries with data for all metrics", True),
+                ui.input_action_button(
+                    "download", "Download data and image", class_="btn-primary w-100"
+                ),
+            ),
+        ),
+        ui.column(
+            8,
+            ui.output_plot("example_plot")
+            ),
+    ),
+)
+
+def server(input, output, session):
+    @output
+    @render.plot
+    def example_plot():
+
+      #  ---------------------------------
+      #  ---------- SECTION 4: Run data functions and produce plot -------------
+      #  ----------------------------------
+
+      def prep_and_merge_multiple_vars(data_to_match, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col):
+
+        # Combine the source and varname strings element-wise to produce keys that will be used for the concat later
+        source_var_keys = [i + j for i, j in zip(data_to_match["sources"], data_to_match["var_names"])]
+
+        matches = []
+
+        for i in range(len(data_to_match['sources'])):
+
+          source = data_to_match["sources"][i]
+          var = data_to_match["var_names"][i]
+
+          df = data[source]
+
+          if source == "PIP":
+
+            matched = pip_welfare_routine(
+                df = df,
+                reference_vals = reference_vals,
+                max_dist_from_refs = max_dist_from_refs,
+                min_dist_between = min_dist_between,
+                reference_col = reference_col,
+                group_by_col = group_by_col,
+                value_col = var
+              )
+
+          else:
+
+            matched = merge_two_ref_matches(
+                df = df,
+                reference_vals = reference_vals,
+                max_dist_from_refs = max_dist_from_refs,
+                min_dist_between = min_dist_between,
+                reference_col = reference_col,
+                group_by_col = group_by_col,
+                value_col = var
+              )
+
+          matches.append(matched)
+
+        ref_matches = pd.concat(matches, keys=source_var_keys)
+
+        # Tidy up indexes
+        ref_matches = ref_matches.reset_index()
+
+        ref_matches = ref_matches.drop('level_1', axis=1)
+
+        ref_matches = ref_matches\
+          .rename(columns={"level_0": "source_var"})
+
+        return ref_matches
+
+      def prep_plot_and_tables(
+        data_to_match,
+        reference_vals, 
+        max_dist_from_refs,
+        min_dist_between,
+        reference_col,
+        group_by_col,
+        tolerance,
+        outlier_cut_off_upper,
+        all_metrics_requirement):
+
+        # Other spec:
+        region_col = 'region_alt'
+
+        # ---- Prep data ---------
+
+        plot_data = prep_and_merge_multiple_vars(
+          data_to_match = data_to_match,
+          reference_vals = reference_vals, 
+          max_dist_from_refs = max_dist_from_refs,
+          min_dist_between = min_dist_between,
+          reference_col = reference_col,
+          group_by_col = group_by_col
+        )
+
+        # Store the names of the columns to be used onthe X and Y axis
+        x_axis = f'value{reference_vals[0]}'
+        y_axis = f'value{reference_vals[1]}'
+
+        x_ref_val = f'{reference_col}{reference_vals[0]}'
+        y_ref_val = f'{reference_col}{reference_vals[1]}'
+
+        # Apply outlier cut off, if specified
+        if not pd.isna(outlier_cut_off_upper):
+          plot_data = plot_data.loc[plot_data[x_axis] <= outlier_cut_off_upper]
+          plot_data = plot_data.loc[plot_data[y_axis] <= outlier_cut_off_upper]
+
+        # Add a count by country – counting whether data is available from each source or not
+        plot_data['source_count'] = plot_data.groupby(group_by_col)['source_var'].transform('count')
+
+        # Apply metrics requirement (whether to include only observations with data from all metrics)
+        if all_metrics_requirement:
+          plot_data = plot_data.loc[plot_data['source_count'] == plot_data['source_count'].max()]
+
+        # Drop source_count column
+        plot_data = plot_data.drop('source_count', axis=1)
+
+        # Add in region classification
+        plot_data = pd.merge(plot_data, df_regions, how = 'left')
+
+        # Add in population data
+          # For first ref
+        df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[0], ['Entity', 'population'] ]
+
+        plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')
+
+        plot_data = plot_data.rename(columns={'population':'population_ref1'})
+
+          # For second ref
+        df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[1], ['Entity', 'population'] ]
+
+        plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')
+
+        plot_data = plot_data.rename(columns={'population':'population_ref2'})
+
+          # Calculate average population two reference periods
+        plot_data['avg_pop'] = (plot_data['population_ref1'] + plot_data['population_ref2'])/2
+
+        # ----- Prep plot ------
+
+        # I grab what I am guessing to be the max and min tick marks on the x axist, in order to define the coordinates of a 'tolerance' ribbon.
+        x_min = plot_data[x_axis].min()
+        x_min_floor = np.floor(x_min * 10) / 10
+
+        x_max = plot_data[x_axis].max()
+        x_max_ceiling = np.ceil(x_max * 10) / 10
+
+        # Set the coordinates of the shaded ribbon
+        shaded_coord = pd.DataFrame({'x': [x_min_floor, x_max_ceiling, x_max_ceiling, x_min_floor], 
+          'y': [x_min_floor-tolerance, x_max_ceiling-tolerance, x_max_ceiling+tolerance , x_min_floor+tolerance]})
+
+        plot = (ggplot(plot_data
+        , aes(x_axis, y_axis, alpha=0.5))
+        + geom_point(aes(color=region_col, size='avg_pop', alpha = 0.6))
+        + facet_wrap('~ source_var')
+        + geom_polygon(aes(x='x', y='y', alpha = 0.2), data=shaded_coord)
+        + theme_light()
+        + theme(legend_position='none')
+        )
+
+        # ---- Prep summary table ----
+
+        # Calculate the change between the two ref periods
+        plot_data['change'] = (plot_data[y_axis] - plot_data[x_axis])
+
+        group_yes_vars = ['source_var', region_col]
+        group_no_vars = ['source_var']
+
+        group_scenarios = [group_no_vars, group_yes_vars]
+
+        aggs = []
+
+        for group_vars in group_scenarios:
+
+        # Prepare pop-weighted average change
+          agg_level = plot_data.copy()
+
+          # Calulate pop weights by source and region
+          agg_level['regional_pop_weights'] = agg_level.groupby(group_vars)['avg_pop'].transform(lambda x: x/x.sum())
+
+          agg_level['global_pop_weights'] = agg_level.groupby('source_var')['avg_pop'].transform(lambda x: x/x.sum())
+
+          # Multiply the change by the pop weights 
+          agg_level['region_weighted_change'] = agg_level['change'] * agg_level['regional_pop_weights']
+
+        # Prepare fall/stable/rise categories
+          agg_level['fall'] = agg_level['change'] < -tolerance
+
+          agg_level['stable'] = (agg_level['change'] <= tolerance) & (agg_level['change'] >= -tolerance)
+
+          agg_level['rise'] = agg_level['change'] > tolerance
+
+          agg_level[['fall', 'stable', 'rise']] = agg_level[['fall', 'stable', 'rise']].astype(int)
+
+          agg_level = agg_level.groupby(group_vars)[['change', 'region_weighted_change', 'global_pop_weights', 'fall', 'stable', 'rise']]\
+            .agg(['sum', 'mean', 'count'])
+
+        # Aggregate by source and region
+          if group_vars == group_yes_vars:
+
+            agg_level = agg_level\
+              .unstack().T.reset_index()
+
+          elif group_vars == group_no_vars:
+
+            agg_level = agg_level\
+              .T.reset_index()
+
+            agg_level[region_col] = "World"
+
+          aggs.append(agg_level)
+
+        df_summary = pd.concat(aggs)
+
+          # filter for the aggregations we need and label in the 'summary' column
+        df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'mean'), 'summary'] = 'avg change'
+
+        df_summary.loc[(df_summary['level_0'] == 'region_weighted_change') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Regional pop-weighted avg change'
+
+        df_summary.loc[(df_summary['level_0'] == 'global_pop_weights') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Global pop weights'
+
+        df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'count'), 'summary'] = 'n'
+
+        df_summary.loc[(df_summary['level_0'] == 'fall') & (df_summary['level_1'] == 'sum'), 'summary'] = 'fall'
+
+        df_summary.loc[(df_summary['level_0'] == 'stable') & (df_summary['level_1'] == 'sum'), 'summary'] = 'stable'
+
+        df_summary.loc[(df_summary['level_0'] == 'rise') & (df_summary['level_1'] == 'sum'), 'summary'] = 'rise'
+
+        df_summary = df_summary[df_summary['summary'].notnull()]
+
+        df_summary = df_summary.drop(['level_0', 'level_1'], axis = 1)
+
+        df_summary['summary'] = pd.Categorical(df_summary['summary'], categories=['fall', 'stable', 'rise','n', 'avg change', 'Regional pop-weighted avg change', 'Global pop weights'], ordered=True)
+
+        df_summary = df_summary.sort_values([region_col,'summary']).set_index([region_col,'summary'])
+
+        # format to show appropriate d.p. I run a loop over each row for want of figuring out a better way!
+        repeat_length = 7
+        for i in range(0, len(df_summary.index.levels[0].unique())):
+          df_summary.iloc[i*repeat_length,:] = df_summary.iloc[i*repeat_length,:].map('{:.0f}'.format)
+          df_summary.iloc[1 + i*repeat_length,:] = df_summary.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
+          df_summary.iloc[2 + i*repeat_length,:] = df_summary.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
+          df_summary.iloc[3 + i*repeat_length,:] = df_summary.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
+          df_summary.iloc[4 + i*repeat_length,:] = df_summary.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)
+          df_summary.iloc[5 + i*repeat_length,:] = df_summary.iloc[5 + i*repeat_length,:].map('{:.2f}'.format)
+          df_summary.iloc[6 + i*repeat_length,:] = df_summary.iloc[6 + i*repeat_length,:].map('{:.2f}'.format)
+
+        # ----- Return outputs ------
+        plot_and_tables = {
+          "table_data": plot_data,
+          "plot": plot,
+          "summary": df_summary
+        }
+
+        return plot_and_tables
+
+      def prep_selected_vars(selected_vars):
+
+        # selected_vars = [
+        #   'LIS: Top_10_share_equiv_market',
+        #   'WID: Bottom_50_share_posttax_dis',
+        #   'PIP: Top_10_share']
+
+        selected_sources = [var.split(sep, 1)[0] for var in selected_vars]
+        selected_var_names = [var.split(sep, 1)[1] for var in selected_vars]
+
+        match_data = {
+          "sources": selected_sources,
+          "var_names": selected_var_names
+          }
+
+        return match_data
+
+      data_to_match = prep_selected_vars(input.metrics())
+
+      plot_and_tables = prep_plot_and_tables(
+        data_to_match = data_to_match,
+        reference_vals = list(input.reference_vals()), 
+        max_dist_from_refs = input.max_dist_from_refs(),
+        min_dist_between = input.min_dist_between(),
+        reference_col = reference_col,
+        group_by_col = group_by_col,
+        tolerance = tolerance,
+        outlier_cut_off_upper = outlier_cut_off_upper,
+        all_metrics_requirement = input.all_metrics_requirement()
+        )
+
+      plot = plot_and_tables['plot']
+
+      return plot
+
+app = App(app_ui, server)
+
+```
+
+```{shinylive-python}
+#| standalone: true
+#| viewerHeight: 420
+
+from shiny import *
+from plotnine import *
+from pyodide.http import open_url
+import pandas as pd
+import numpy as np
+
+#  ---------------------------------
+#  ---------- Data prep functions -------------
+#  ----------------------------------
+
+#  Function get matching for ref years
+def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):
+
+  df = df.loc[:, [reference_col, group_by_col, value_col]]
+
+  # Drop NAs
+  df = df.dropna()
+
+  # Calculate absolute distance from reference value
+  df['ref_diff'] = abs(df[reference_col] - reference_val)
+
+  # Drop any rows with a distance beyond threshold
+  if not pd.isna(max_dist_from_ref):
+    df = df.loc[df['ref_diff'] <= max_dist_from_ref]
+
+  # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
+  df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)
+
+  # Settle tie-breaks
+  if tie_break == 'below':
+    df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)
+
+  elif tie_break == 'above':
+    df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)
+
+  df = df.drop('ref_diff', axis=1)
+
+  return df
+
+# Merge matches for different reference points
+def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+
+# Make sure the pair of reference values are in ascending order
+  reference_vals.sort()
+
+# Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref
+
+# Find matches for lower reference value
+  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs, reference_col, group_by_col, value_col, 'below')
+
+# Find matches for higher reference value
+  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs, reference_col, group_by_col, value_col, 'above')
+
+# Merge the two sets of matches
+  merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))
+
+# Drop obs that do not have data for both ref values
+  merged_df = merged_df.dropna()
+
+# Drop obs where the matched data does not meet the min distance requirement
+  if not pd.isna(min_dist_between):
+
+  # Store the names of the reference column returned from the two matches
+    ref_var_high = f'{reference_col}{reference_vals[1]}'
+    ref_var_low = f'{reference_col}{reference_vals[0]}'
+
+  # Keep only rows >= to the min distance
+    merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]
+
+  return merged_df
+
+  # For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
+def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+
+    # Specify the name of the column in which the income/consumption welfare definition is stored
+    welfare_colname = 'welfare_type'
+
+    # Creat dataframes for thee scenarios:
+    # Scenario 1: only allow income data
+    df_inc_filter = df.loc[df[welfare_colname] == "income", :]
+    df_inc_filter.name = "Income"
+
+    # Scenario 2: only allow consumption data
+    df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
+    df_cons_filter.name = "Consumption"
+    # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
+    df_mixed = df.copy()
+
+    df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')
+
+    df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]
+
+    df_mixed.name = "Mixed"
+    #  Store the scneario dataframes in a list
+    df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]
+
+    # Run the matching function on each scenario
+    scenario_matches = [merge_two_ref_matches(
+      df_scenario, 
+      reference_vals, 
+      max_dist_from_refs, 
+      min_dist_between, 
+      reference_col, 
+      group_by_col, 
+      value_col) for df_scenario in df_scenarios]
+
+    # Combine the first two scenarios.
+    df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])
+
+    # Tidy up indexes
+    df_combined_matches = df_combined_matches.reset_index()
+
+    df_combined_matches = df_combined_matches.drop('level_1', axis=1)
+
+    df_combined_matches = df_combined_matches\
+      .rename(columns={"level_0": "pip_welfare"})
+
+    # Add in third scenario.
+    df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])
+
+    # add scenario name to te pip_welfare column
+    df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)
+
+    # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
+      # First count the matches
+    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
+      # Then drop any matches from the lowest priority where there are multiple matches
+    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
+      #  Repeat at the next level of priority
+    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
+    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]
+
+    # Drop the match count column
+    df_combined_matches = df_combined_matches.drop('match_count', axis=1)
+
+    return df_combined_matches
+
+#  ---------------------------------
+#  ---------- Specifications -------------
+#  ----------------------------------
+
+# These parameters are not controlled in the app.
+reference_col = 'Year'
+group_by_col = 'Entity'
+
+tolerance = 1
+outlier_cut_off_upper = None
+
+source_count_requirement = 2
+
+#  ---------------------------------
+#  ---------- App code -------------
+#  ----------------------------------
+
+def panel_box(*args, **kwargs):
+    return ui.div(
+        ui.div(*args, class_="card-body"),
+        **kwargs,
+        class_="card mb-3",
+    )
+
+app_ui = ui.page_fluid(
+    {"class": "p-4"},
+    ui.row(
+        ui.column(
+            4,
+            panel_box(
+                ui.input_slider("reference_vals", "Reference years", 1975, 2020, value=[1990,2018]),
+                ui.input_numeric(
+                            "max_dist_from_refs",
+                            "=/- (years)",
+                            4,
+                        ),
+                ui.input_numeric(
+                            "min_dist_between",
+                            "Min distance between observations (years)",
+                            23,
+                        ),
                 ui.input_action_button(

### 2023-02-01 PhD_pages/data_appendices/1_F_within_country_trends.qmd
Tidying names and layout

@@
 #| standalone: true
-#| viewerHeight: 420
+#| viewerHeight: 1000
 
@@
 ```
-
-```{shinylive-python}
-#| standalone: true
-#| viewerHeight: 420
-
-from shiny import *
-from plotnine import *
-from pyodide.http import open_url
-import pandas as pd
-import numpy as np
-
-#  ---------------------------------
-#  ---------- Data prep functions -------------
-#  ----------------------------------
-
-#  Function get matching for ref years
-def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):
-
-  df = df.loc[:, [reference_col, group_by_col, value_col]]
-
-  # Drop NAs
-  df = df.dropna()
-
-  # Calculate absolute distance from reference value
-  df['ref_diff'] = abs(df[reference_col] - reference_val)
-
-  # Drop any rows with a distance beyond threshold
-  if not pd.isna(max_dist_from_ref):
-    df = df.loc[df['ref_diff'] <= max_dist_from_ref]
-
-  # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
-  df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)
-
-  # Settle tie-breaks
-  if tie_break == 'below':
-    df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)
-
-  elif tie_break == 'above':
-    df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)
-
-  df = df.drop('ref_diff', axis=1)
-
-  return df
-
-# Merge matches for different reference points
-def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
-
-# Make sure the pair of reference values are in ascending order
-  reference_vals.sort()
-
-# Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref
-
-# Find matches for lower reference value
-  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs, reference_col, group_by_col, value_col, 'below')
-
-# Find matches for higher reference value
-  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs, reference_col, group_by_col, value_col, 'above')
-
-# Merge the two sets of matches
-  merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))
-
-# Drop obs that do not have data for both ref values
-  merged_df = merged_df.dropna()
-
-# Drop obs where the matched data does not meet the min distance requirement
-  if not pd.isna(min_dist_between):
-
-  # Store the names of the reference column returned from the two matches
-    ref_var_high = f'{reference_col}{reference_vals[1]}'
-    ref_var_low = f'{reference_col}{reference_vals[0]}'
-
-  # Keep only rows >= to the min distance
-    merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]
-
-  return merged_df
-
-  # For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
-def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
-
-    # Specify the name of the column in which the income/consumption welfare definition is stored
-    welfare_colname = 'welfare_type'
-
-    # Creat dataframes for thee scenarios:
-    # Scenario 1: only allow income data
-    df_inc_filter = df.loc[df[welfare_colname] == "income", :]
-    df_inc_filter.name = "Income"
-
-    # Scenario 2: only allow consumption data
-    df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
-    df_cons_filter.name = "Consumption"
-    # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
-    df_mixed = df.copy()
-
-    df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')
-
-    df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]
-
-    df_mixed.name = "Mixed"
-    #  Store the scneario dataframes in a list
-    df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]
-
-    # Run the matching function on each scenario
-    scenario_matches = [merge_two_ref_matches(
-      df_scenario, 
-      reference_vals, 
-      max_dist_from_refs, 
-      min_dist_between, 
-      reference_col, 
-      group_by_col, 
-      value_col) for df_scenario in df_scenarios]
-
-    # Combine the first two scenarios.
-    df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])
-
-    # Tidy up indexes
-    df_combined_matches = df_combined_matches.reset_index()
-
-    df_combined_matches = df_combined_matches.drop('level_1', axis=1)
-
-    df_combined_matches = df_combined_matches\
-      .rename(columns={"level_0": "pip_welfare"})
-
-    # Add in third scenario.
-    df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])
-
-    # add scenario name to te pip_welfare column
-    df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)
-
-    # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
-      # First count the matches
-    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
-      # Then drop any matches from the lowest priority where there are multiple matches
-    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
-      #  Repeat at the next level of priority
-    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
-    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]
-
-    # Drop the match count column
-    df_combined_matches = df_combined_matches.drop('match_count', axis=1)
-
-    return df_combined_matches
-
-#  ---------------------------------
-#  ---------- Specifications -------------
-#  ----------------------------------
-
-# These parameters are not controlled in the app.
-reference_col = 'Year'
-group_by_col = 'Entity'
-
-tolerance = 1
-outlier_cut_off_upper = None
-
-source_count_requirement = 2
-
-#  ---------------------------------
-#  ---------- App code -------------
-#  ----------------------------------
-
-def panel_box(*args, **kwargs):
-    return ui.div(
-        ui.div(*args, class_="card-body"),
-        **kwargs,
-        class_="card mb-3",
-    )
-
-app_ui = ui.page_fluid(
-    {"class": "p-4"},
-    ui.row(
-        ui.column(
-            4,
-            panel_box(
-                ui.input_slider("reference_vals", "Reference years", 1975, 2020, value=[1990,2018]),
-                ui.input_numeric(
-                            "max_dist_from_refs",
-                            "=/- (years)",
-                            4,
-                        ),
-                ui.input_numeric(
-                            "min_dist_between",
-                            "Min distance between observations (years)",
-                            23,
-                        ),
-                ui.input_action_button(
-                    "download", "Download data and image", class_="btn-primary w-100"
-                ),
-            ),
-            ui.navset_tab_card(
-                ui.nav(
-                    "Compare sources",
-                    ui.input_select(
-                "metric",
-                "Metric",
-                {
-                    "Gini": "Gini",
-                    "Top_10_share": "Top 10p.c. share",
-                },
-            ),
-            ),
-                ui.nav(
-                  "Compare metrics"),
-            ),
-        ),
-        ui.column(
-            8,
-            ui.output_plot("example_plot")
-            ),
-    ),
-)
-
-def server(input, output, session):
-    @output
-    @render.plot
-    def example_plot():
-
-      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'
-
-      df_pip = pd.read_csv(open_url(url))
-
-      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'
-
-      df_wid = pd.read_csv(open_url(url))
-
-      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/region_mapping.csv'
-
-      df_regions = pd.read_csv(open_url(url))
-
-      url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'
-
-      df_pop = pd.read_csv(open_url(url))
-
-      def prep_wid_pip_data(reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
-
-        pip_matches = pip_welfare_routine(
-          df = df_pip,
-          reference_vals = reference_vals,
-          max_dist_from_refs = max_dist_from_refs,
-          min_dist_between = min_dist_between,
-          reference_col = reference_col,
-          group_by_col = group_by_col,
-          value_col = value_col
-        )
-
-        wid_matches = merge_two_ref_matches(
-          df = df_wid,
-          reference_vals = reference_vals,
-          max_dist_from_refs = max_dist_from_refs,
-          min_dist_between = min_dist_between,
-          reference_col = reference_col,
-          group_by_col = group_by_col,
-          value_col = value_col
-        )
-
-        ref_pairs = pd.concat([pip_matches, wid_matches,], keys=['pip', 'wid'])
-
-        # Tidy up indexes
-        ref_pairs = ref_pairs.reset_index()
-
-        ref_pairs = ref_pairs.drop('level_1', axis=1)
-
-        ref_pairs = ref_pairs\
-          .rename(columns={"level_0": "source"})
-
-        return ref_pairs
-
-      def prep_plot_and_tables(
-        reference_vals, 
-        max_dist_from_refs,
-        min_dist_between,
-        reference_col,
-        group_by_col,
-        value_col,
-        tolerance,
-        outlier_cut_off_upper,
-        source_count_requirement):
-
-        # Other spec:
-        region_col = 'region_alt'
-
-        # ---- Prep data ---------
-
-        plot_data = prep_wid_pip_data(
-          reference_vals = reference_vals, 
-          max_dist_from_refs = max_dist_from_refs,
-          min_dist_between = min_dist_between,
-          reference_col = reference_col,
-          group_by_col = group_by_col,
-          value_col = value_col,
-        )
-
-        # Store the names of the columns to be used onthe X and Y axis
-        x_axis = f'{value_col}{reference_vals[0]}'
-        y_axis = f'{value_col}{reference_vals[1]}'
-
-        x_ref_val = f'{reference_col}{reference_vals[0]}'
-        y_ref_val = f'{reference_col}{reference_vals[1]}'
-
-        # Apply outlier cut off, if specified
-        if not pd.isna(outlier_cut_off_upper):
-          plot_data = plot_data.loc[plot_data[x_axis] <= outlier_cut_off_upper]
-          plot_data = plot_data.loc[plot_data[y_axis] <= outlier_cut_off_upper]
-
-        # Add a count by country – showing whether data is available from both sources or not
-        plot_data['source_count'] = plot_data.groupby(group_by_col)['source'].transform('count')
-
-        # Apply source requirement (whether to include only observations with data from both sources)
-        plot_data = plot_data.loc[plot_data['source_count'] >= source_count_requirement]
-
-        # Drop source_count column
-        plot_data = plot_data.drop('source_count', axis=1)
-
-        # Add in region classification
-        plot_data = pd.merge(plot_data, df_regions, how = 'left')
-
-        # Add in population data
-          # For first ref
-        df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[0], ['Entity', 'population'] ]
-
-        plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')
-
-        plot_data = plot_data.rename(columns={'population':'population_ref1'})
-
-          # For second ref
-        df_pop_ref = df_pop.loc[df_pop['Year'] == reference_vals[1], ['Entity', 'population'] ]
-
-        plot_data = pd.merge(plot_data, df_pop_ref, how = 'left')
-
-        plot_data = plot_data.rename(columns={'population':'population_ref2'})
-
-          # Calculate average population two reference periods
-        plot_data['avg_pop'] = (plot_data['population_ref1'] + plot_data['population_ref2'])/2
-
-        # ----- Prep plot ------
-
-        # I grab what I am guessing to be the max and min tick marks on the x axist, in order to define the coordinates of a 'tolerance' ribbon.
-        x_min = plot_data[x_axis].min()
-        x_min_floor = np.floor(x_min * 10) / 10
-
-        x_max = plot_data[x_axis].max()
-        x_max_ceiling = np.ceil(x_max * 10) / 10
-
-        # Set the coordinates of the shaded ribbon
-        shaded_coord = pd.DataFrame({'x': [x_min_floor, x_max_ceiling, x_max_ceiling, x_min_floor], 
-          'y': [x_min_floor-tolerance, x_max_ceiling-tolerance, x_max_ceiling+tolerance , x_min_floor+tolerance]})
-
-        plot = (ggplot(plot_data
-        , aes(x_axis, y_axis, alpha=0.5))
-        + geom_point(aes(color=region_col, size='avg_pop', alpha = 0.6))
-        + facet_wrap('~ source')
-        + geom_polygon(aes(x='x', y='y', alpha = 0.2), data=shaded_coord)
-        + theme_light()
-        + theme(legend_position='none')
-        )
-
-        # ---- Prep summary table ----
-
-        # Calculate the change between the two ref periods
-        plot_data['change'] = (plot_data[y_axis] - plot_data[x_axis])
-
-        group_yes_vars = ['source', region_col]
-        group_no_vars = ['source']
-
-        group_scenarios = [group_no_vars, group_yes_vars]
-
-        aggs = []
-
-        for group_vars in group_scenarios:
-
-        # Prepare pop-weighted average change
-          agg_level = plot_data.copy()
-
-          # Calulate pop weights by source and region
-          agg_level['regional_pop_weights'] = agg_level.groupby(group_vars)['avg_pop'].transform(lambda x: x/x.sum())
-
-          agg_level['global_pop_weights'] = agg_level.groupby('source')['avg_pop'].transform(lambda x: x/x.sum())
-
-          # Multiply the change by the pop weights 
-          agg_level['region_weighted_change'] = agg_level['change'] * agg_level['regional_pop_weights']
-
-        # Prepare fall/stable/rise categories
-          agg_level['fall'] = agg_level['change'] < -tolerance
-
-          agg_level['stable'] = (agg_level['change'] <= tolerance) & (agg_level['change'] >= -tolerance)
-
-          agg_level['rise'] = agg_level['change'] > tolerance
-
-          agg_level[['fall', 'stable', 'rise']] = agg_level[['fall', 'stable', 'rise']].astype(int)
-
-          agg_level = agg_level.groupby(group_vars)[['change', 'region_weighted_change', 'global_pop_weights', 'fall', 'stable', 'rise']]\
-            .agg(['sum', 'mean', 'count'])
-
-        # Aggregate by source and region
-          if group_vars == group_yes_vars:
-
-            agg_level = agg_level\
-              .unstack().T.reset_index()
-
-          elif group_vars == group_no_vars:
-
-            agg_level = agg_level\
-              .T.reset_index()
-
-            agg_level[region_col] = "World"
-
-          aggs.append(agg_level)
-
-        df_summary = pd.concat(aggs)
-
-          # filter for the aggregations we need and label in the 'summary' column
-        df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'mean'), 'summary'] = 'avg change'
-
-        df_summary.loc[(df_summary['level_0'] == 'region_weighted_change') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Regional pop-weighted avg change'
-
-        df_summary.loc[(df_summary['level_0'] == 'global_pop_weights') & (df_summary['level_1'] == 'sum'), 'summary'] = 'Global pop weights'
-
-        df_summary.loc[(df_summary['level_0'] == 'change') & (df_summary['level_1'] == 'count'), 'summary'] = 'n'
-
-        df_summary.loc[(df_summary['level_0'] == 'fall') & (df_summary['level_1'] == 'sum'), 'summary'] = 'fall'
-
-        df_summary.loc[(df_summary['level_0'] == 'stable') & (df_summary['level_1'] == 'sum'), 'summary'] = 'stable'
-
-        df_summary.loc[(df_summary['level_0'] == 'rise') & (df_summary['level_1'] == 'sum'), 'summary'] = 'rise'
-
-        df_summary = df_summary[df_summary['summary'].notnull()]
-
-        df_summary = df_summary.drop(['level_0', 'level_1'], axis = 1)
-
-        df_summary['summary'] = pd.Categorical(df_summary['summary'], categories=['fall', 'stable', 'rise','n', 'avg change', 'Regional pop-weighted avg change', 'Global pop weights'], ordered=True)
-
-        df_summary = df_summary.sort_values([region_col,'summary']).set_index([region_col,'summary'])
-
-        # format to show appropriate d.p. I run a loop over each row for want of figuring out a better way!
-        repeat_length = 7
-        for i in range(0, len(df_summary.index.levels[0].unique())):
-          df_summary.iloc[i*repeat_length,:] = df_summary.iloc[i*repeat_length,:].map('{:.0f}'.format)
-          df_summary.iloc[1 + i*repeat_length,:] = df_summary.iloc[1 + i*repeat_length,:].map('{:.0f}'.format)
-          df_summary.iloc[2 + i*repeat_length,:] = df_summary.iloc[2 + i*repeat_length,:].map('{:.0f}'.format)
-          df_summary.iloc[3 + i*repeat_length,:] = df_summary.iloc[3 + i*repeat_length,:].map('{:.0f}'.format)
-          df_summary.iloc[4 + i*repeat_length,:] = df_summary.iloc[4 + i*repeat_length,:].map('{:.2f}'.format)
-          df_summary.iloc[5 + i*repeat_length,:] = df_summary.iloc[5 + i*repeat_length,:].map('{:.2f}'.format)
-          df_summary.iloc[6 + i*repeat_length,:] = df_summary.iloc[6 + i*repeat_length,:].map('{:.2f}'.format)
-
-        # ----- Return outputs ------
-        plot_and_tables = {
-          "table_data": plot_data,
-          "plot": plot,
-          "summary": df_summary
-        }
-
-        return plot_and_tables
-
-      plot_and_tables = prep_plot_and_tables(
-          reference_vals = list(input.reference_vals()), 
-          max_dist_from_refs = input.max_dist_from_refs(),
-          min_dist_between = input.min_dist_between(),
-          reference_col = reference_col,
-          group_by_col = group_by_col,
-          value_col = input.metric(),
-          tolerance = tolerance,
-          outlier_cut_off_upper = outlier_cut_off_upper,
-          source_count_requirement = source_count_requirement
-          )
-
-      plot = plot_and_tables['plot']
-
-      return plot
-
-app = App(app_ui, server)
-
-```

### 2023-02-09 PhD_pages/data_appendices/1_F_within_country_trends.qmd
Paper 1: reorganise app and initial drafting of paper

@@
   - shinylive
+sidebar: false
 
 ---
-```{python}
-
+
+:::{.column-page}
+
+```{shinylive-python}
+#| standalone: true
+#| viewerHeight: 1000
+
+## file: app.py
+
+import jinja2
+from shiny import *
+import pandas as pd
+from pyodide.http import open_url
+import numpy as np
 from plotnine import *
-import pandas as pd
-import numpy as np
-
-```
-
-This first block of code.....
-
-```{python}
 
@@
 
-```
-
-This second block of code we read in the data (this needs to be done in a different way in the shiny app...)
-
-```{python}
-
-#  ---------------------------------
-#  ---------- SECTION 2: Read in data -------------
-#  ----------------------------------
-
-url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'
-
-df_pip = pd.read_csv(url)
-
-url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'
-
-df_wid = pd.read_csv(url)
-
-url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/lis.csv'
-
-df_lis = pd.read_csv(url)
-
-url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/region_mapping.csv'
-
-df_regions = pd.read_csv(url)
-
-url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'
-
-df_pop = pd.read_csv(url)
-
-```
-
-In a third block, we set specifications for the data matching. This is different in the app because some parameters are controlled by the app.
-
-```{python}
-#  ---------------------------------
-#  ---------- SECTION 3: Specifications -------------
-#  ----------------------------------
-
-# These parameters are not controlled in the app.
-reference_col = 'Year'
-group_by_col = 'Entity'
-
-tolerance = 1
-outlier_cut_off_upper = None
-
-source_count_requirement = "all"
-# source_count_requirement = "any"
-
-reference_vals = [1990,2018]
-max_dist_from_refs = 4
-min_dist_between = 5
-
-data = {"PIP": df_pip,
-        "WID": df_wid,
-        "LIS": df_lis}
-
-data_to_match = {
-  "sources": ["PIP", "WID", "LIS"],
-  "var_names": ["Top_10_share", "Top_10_share_pretax", "Top_10_share_equiv_disposable"]
+def prep_selected_vars(selected_vars):
+
+  # selected_vars = [
+  #   'LIS: Top_10_share_equiv_market',
+  #   'WID: Bottom_50_share_posttax_dis',
+  #   'PIP: Top_10_share']
+
+  selected_sources = [var.split(sep, 1)[0] for var in selected_vars]
+  selected_var_names = [var.split(sep, 1)[1] for var in selected_vars]
+
+  match_data = {
+    "sources": selected_sources,
+    "var_names": selected_var_names
           }
 
-```
-
-In this code block we run the code and prepare final plots and tables.
-
-```{python}
-
-#  ---------------------------------
-#  ---------- SECTION 4: Run data functions and produce plot -------------
-#  ----------------------------------
+  return match_data
+
+def prep_data(
+      data_to_match,
+      reference_vals,
+      max_dist_from_refs,
+      min_dist_between,
+      all_metrics_requirement,
+      reference_col,
+      group_by_col,
+      tolerance,
+      outlier_cut_off_upper):
 
@@
 
-def prep_and_merge_multiple_vars(data_to_match, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col):
-
   matches = []
@@
     var = data_to_match["var_names"][i]
-    print(i)
-    print(source)
-    print(var)
+
     df = data[source]
@@
 
-  ref_matches = pd.concat(matches, keys=source_var_keys)
+  plot_data = pd.concat(matches, keys=source_var_keys)
 
   # Tidy up indexes
-  ref_matches = ref_matches.reset_index()
-
-  ref_matches = ref_matches.drop('level_1', axis=1)
-
-  ref_matches = ref_matches\
+  plot_data = plot_data.reset_index()
+
+  plot_data = plot_data.drop('level_1', axis=1)
+
+  plot_data = plot_data\
     .rename(columns={"level_0": "source_var"})
-
-  return ref_matches
-
-def prep_plot_and_tables(
-  data_to_match,
-  reference_vals, 
-  max_dist_from_refs,
-  min_dist_between,
-  reference_col,
-  group_by_col,
-  tolerance,
-  outlier_cut_off_upper,
-  source_count_requirement):
-
-  # Other spec:
-  region_col = 'region_alt'
-
-  # ---- Prep data ---------
-
-  plot_data = prep_and_merge_multiple_vars(
-    data_to_match = data_to_match,
-    reference_vals = reference_vals, 
-    max_dist_from_refs = max_dist_from_refs,
-    min_dist_between = min_dist_between,
-    reference_col = reference_col,
-    group_by_col = group_by_col
-  )
 
@@
 
-  # Apply source requirement (whether to include only observations with data from all sources)
-  if source_count_requirement == "all":
+  # Apply metrics requirement (whether to include only observations with data from all metrics)
+  if all_metrics_requirement:
     plot_data = plot_data.loc[plot_data['source_count'] == plot_data['source_count'].max()]
@@
 
-  # ----- Prep plot ------
-
-  # I grab what I am guessing to be the max and min tick marks on the x axist, in order to define the coordinates of a 'tolerance' ribbon.
-  x_min = plot_data[x_axis].min()
-  x_min_floor = np.floor(x_min * 10) / 10
-
-  x_max = plot_data[x_axis].max()
-  x_max_ceiling = np.ceil(x_max * 10) / 10
-
-  # Set the coordinates of the shaded ribbon
-  shaded_coord = pd.DataFrame({'x': [x_min_floor, x_max_ceiling, x_max_ceiling, x_min_floor], 
-    'y': [x_min_floor-tolerance, x_max_ceiling-tolerance, x_max_ceiling+tolerance , x_min_floor+tolerance]})
-
-  plot = (ggplot(plot_data
-  , aes(x_axis, y_axis, alpha=0.5))
-  + geom_point(aes(color=region_col, size='avg_pop', alpha = 0.6))
-  + facet_wrap('~ source_var')
-  + geom_polygon(aes(x='x', y='y', alpha = 0.2), data=shaded_coord)
-  + theme_light()
-  + theme(legend_position='none')
-  )
-
-  # ---- Prep summary table ----
+  #  PREP SUMMARY TABLE
 
@@
 
-  # ----- Return outputs ------
-  plot_and_tables = {
-    "table_data": plot_data,
-    "plot": plot,
-    "summary": df_summary
-  }
-
-  return plot_and_tables
-
-```
-
-```{python}
-plot_and_tables = prep_plot_and_tables(
-    data_to_match = data_to_match,
-    reference_vals = reference_vals, 
-    max_dist_from_refs = max_dist_from_refs,
-    min_dist_between = min_dist_between,
-    reference_col = reference_col,
-    group_by_col = group_by_col,
-    tolerance = tolerance,
-    outlier_cut_off_upper = outlier_cut_off_upper,
-    source_count_requirement = source_count_requirement
-    )
-
-plot = plot_and_tables['plot']
-summary = plot_and_tables['summary']
-
-```
-
-```{python}
-plot
-
-```
-
-```{python}
-summary
-
-```
-
-## Explore the data
-
-I plan to build a Shiny app to help compare trends across datasets (using Shinylive – built on Shiny for Python - ).
-
-Here is a test app just so I can test the wiring of how such an app works.
-
-```{shinylive-python}
-#| standalone: true
-#| viewerHeight: 1000
-
-from shiny import *
-from plotnine import *
-from pyodide.http import open_url
-import pandas as pd
-import numpy as np
+  return plot_data, df_summary
 
 #  ---------------------------------
-#  ---------- SECTION 1: Data prep functions -------------
+#  ---------- SECTION 2: Specifications -------------
 #  ----------------------------------
-
-#  Function get matching for ref years
-def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):
-
-  df = df.loc[:, [reference_col, group_by_col, value_col]]
-
-  # Drop NAs
-  df = df.dropna()
-
-  # Calculate absolute distance from reference value
-  df['ref_diff'] = abs(df[reference_col] - reference_val)
-
-  # Drop any rows with a distance beyond threshold
-  if not pd.isna(max_dist_from_ref):
-    df = df.loc[df['ref_diff'] <= max_dist_from_ref]
-
-  # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
-  df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)
-
-  # Settle tie-breaks
-  if tie_break == 'below':
-    df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)
-
-  elif tie_break == 'above':
-    df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)
-
-  df = df.drop('ref_diff', axis=1)
-
-  df = df\
-    .rename(columns={value_col: "value"})
-
-  return df
-
-# Merge matches for different reference points
-def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
-
-# Make sure the pair of reference values are in ascending order
-  reference_vals.sort()
-
-# Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref
-
-# Find matches for lower reference value
-  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs, reference_col, group_by_col, value_col, 'below')
-
-# Find matches for higher reference value
-  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs, reference_col, group_by_col, value_col, 'above')
-
-# Merge the two sets of matches
-  merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))
-
-# Drop obs that do not have data for both ref values
-  merged_df = merged_df.dropna()
-
-# Drop obs where the matched data does not meet the min distance requirement
-  if not pd.isna(min_dist_between):
-
-  # Store the names of the reference column returned from the two matches
-    ref_var_high = f'{reference_col}{reference_vals[1]}'
-    ref_var_low = f'{reference_col}{reference_vals[0]}'
-
-  # Keep only rows >= to the min distance
-    merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]
-
-  return merged_df
-
-  # For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
-def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
-
-    # Specify the name of the column in which the income/consumption welfare definition is stored
-    welfare_colname = 'welfare_type'
-
-    # Creat dataframes for thee scenarios:
-    # Scenario 1: only allow income data
-    df_inc_filter = df.loc[df[welfare_colname] == "income", :]
-    df_inc_filter.name = "Income"
-
-    # Scenario 2: only allow consumption data
-    df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
-    df_cons_filter.name = "Consumption"
-    # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
-    df_mixed = df.copy()
-
-    df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')
-
-    df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]
-
-    df_mixed.name = "Mixed"
-    #  Store the scneario dataframes in a list
-    df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]
-
-    # Run the matching function on each scenario
-    scenario_matches = [merge_two_ref_matches(
-      df_scenario, 
-      reference_vals, 
-      max_dist_from_refs, 
-      min_dist_between, 
-      reference_col, 
-      group_by_col, 
-      value_col) for df_scenario in df_scenarios]
-
-    # Combine the first two scenarios.
-    df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])
-
-    # Tidy up indexes
-    df_combined_matches = df_combined_matches.reset_index()
-
-    df_combined_matches = df_combined_matches.drop('level_1', axis=1)
-
-    df_combined_matches = df_combined_matches\
-      .rename(columns={"level_0": "pip_welfare"})
-
-    # Add in third scenario.
-    df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])
-
-    # add scenario name to te pip_welfare column
-    df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)
-
-    # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
-      # First count the matches
-    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
-      # Then drop any matches from the lowest priority where there are multiple matches
-    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
-      #  Repeat at the next level of priority
-    df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
-    df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]
-
-    # Drop the match count column
-    df_combined_matches = df_combined_matches.drop('match_count', axis=1)
-
-    return df_combined_matches
-
-#  ---------------------------------
-#  ---------- SECTION 2: Read in data -------------
-#  ----------------------------------
-
-url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'
-
-df_pip = pd.read_csv(open_url(url))
-
-url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'
-
-df_wid = pd.read_csv(open_url(url))
-
-url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/lis.csv'
-
-df_lis = pd.read_csv(open_url(url))
-
-url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/region_mapping.csv'
-
-df_regions = pd.read_csv(open_url(url))
-
-url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'
-
-df_pop = pd.read_csv(open_url(url))
-
-#  ---------------------------------
-#  ---------- SECTION 3: Specifications -------------
-#  ----------------------------------
-
 # These parameters are not controlled in the app.
+
 reference_col = 'Year'
@@
 outlier_cut_off_upper = None
+region_col = 'region_alt'
+
+#  ---------------------------------
+#  ---------- SECTION 3: Read in data -------------
+#  ----------------------------------
+# Note when testing outside of shinylive, this needs to be adjusted to a normal pd_read, without the pyodide step.
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'
+
+df_pip = pd.read_csv(open_url(url))
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'
+
+df_wid = pd.read_csv(open_url(url))
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/lis.csv'
+
+df_lis = pd.read_csv(open_url(url))
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/region_mapping.csv'
+
+df_regions = pd.read_csv(open_url(url))
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'
+
+df_pop = pd.read_csv(open_url(url))
+
+#  ---------------------------------
+#  ---------- SECTION 4: Gather data into dictionary and collete varnames -------------
+#  ----------------------------------
 
@@
   all_vars = all_vars + vars
+
+#  ---------------------------------
+#  ---------------------------------
 
@@
             8,
-            ui.output_plot("example_plot")
+            ui.output_plot("plot"),
+            ui.output_table("table")
             ),
@@
 def server(input, output, session):
+
+    @reactive.Calc
+    def reactive_data():
+
+      data_to_match = prep_selected_vars(input.metrics())
+
+      reference_vals = list(input.reference_vals())
+
+      max_dist_from_refs = input.max_dist_from_refs()
+      min_dist_between = input.min_dist_between()
+      all_metrics_requirement = input.all_metrics_requirement()
+
+      prepped_data = prep_data(
+        data_to_match = data_to_match,
+        reference_vals = reference_vals,
+        max_dist_from_refs = max_dist_from_refs,
+        min_dist_between = min_dist_between,
+        all_metrics_requirement = all_metrics_requirement,
+        reference_col = reference_col,
+        group_by_col = group_by_col,
+        tolerance = tolerance,
+        outlier_cut_off_upper = outlier_cut_off_upper)
+
+      return prepped_data
+
     @output
     @render.plot
-    def example_plot():
+    def plot():
+
+        reference_vals = list(input.reference_vals())
+
+        prepped_data = reactive_data()
+
+        plot_data = prepped_data[0]
+
+        # Store the names of the columns to be used onthe X and Y axis
+        x_axis = f'value{reference_vals[0]}'
+        y_axis = f'value{reference_vals[1]}'
+
+        # I grab what I am guessing to be the max and min tick marks on the x axist, in order to define the coordinates of a 'tolerance' ribbon.
+        x_min = plot_data[x_axis].min()
+        x_min_floor = np.floor(x_min * 10) / 10
+
+        x_max = plot_data[x_axis].max()
+        x_max_ceiling = np.ceil(x_max * 10) / 10
+
+        plot = (ggplot(plot_data
+        , aes(x_axis, y_axis, alpha=0.5))
+        + geom_point(aes(color=region_col, size='avg_pop', alpha = 0.6))
+        + geom_abline(intercept=0,slope=1) 
+        + facet_wrap('~ source_var')
+        + theme_light()
+        + theme(legend_position='none')
+        )
+
+        return plot
+
+    @output
+    @render.table
+    def table():
+
+        prepped_data = reactive_data()
+
+        summary = prepped_data[1].reset_index()
+
+        return summary
+
+app = App(app_ui, server)
+
+## file: requirements.txt
+Jinja2
+
+```
+
+:::
+
+TESTING!!!
+These are the same sections as above, to be used for testing.
+
+```{python}
+
+from plotnine import *
+import plotly.express as px
+import pandas as pd
+import numpy as np
+
+```
+
+```{python}
 
 #  ---------------------------------
-      #  ---------- SECTION 4: Run data functions and produce plot -------------
+#  ---------- SECTION 1: Data prep functions -------------
 #  ----------------------------------
 
-      def prep_and_merge_multiple_vars(data_to_match, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col):
+#  Function get matching for ref years
+def closest_to_reference(df, reference_val, max_dist_from_ref, reference_col, group_by_col, value_col, tie_break):
+
+  df = df.loc[:, [reference_col, group_by_col, value_col]]
+
+  # Drop NAs
+  df = df.dropna()
+
+  # Calculate absolute distance from reference value
+  df['ref_diff'] = abs(df[reference_col] - reference_val)
+
+  # Drop any rows with a distance beyond threshold
+  if not pd.isna(max_dist_from_ref):
+    df = df.loc[df['ref_diff'] <= max_dist_from_ref]
+
+  # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
+  df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)
+
+  # Settle tie-breaks
+  if tie_break == 'below':
+    df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)
+
+  elif tie_break == 'above':
+    df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)
+
+  df = df.drop('ref_diff', axis=1)
+
+  df = df\
+    .rename(columns={value_col: "value"})
+
+  return df
+
+# Merge matches for different reference points
+def merge_two_ref_matches(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+
+# Make sure the pair of reference values are in ascending order
+  reference_vals.sort()
+
+# Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref
+
+# Find matches for lower reference value
+  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs, reference_col, group_by_col, value_col, 'below')
+
+# Find matches for higher reference value
+  higher_ref_matches = closest_to_reference(df, reference_vals[1], max_dist_from_refs, reference_col, group_by_col, value_col, 'above')
+
+# Merge the two sets of matches
+  merged_df = pd.merge(lower_ref_matches, higher_ref_matches, on=group_by_col, suffixes=(reference_vals[0], reference_vals[1]))
+
+# Drop obs that do not have data for both ref values
+  merged_df = merged_df.dropna()
+
+# Drop obs where the matched data does not meet the min distance requirement
+  if not pd.isna(min_dist_between):
+
+  # Store the names of the reference column returned from the two matches
+    ref_var_high = f'{reference_col}{reference_vals[1]}'
+    ref_var_low = f'{reference_col}{reference_vals[0]}'
+
+  # Keep only rows >= to the min distance
+    merged_df = merged_df.loc[(merged_df[ref_var_high] - merged_df[ref_var_low]) >= min_dist_between, :]
+
+  return merged_df
+
+  # For PIP run this three times – first filtering data for just consumpion only, then with income only, then with a dataset that prefers income over consumption
+def pip_welfare_routine(df, reference_vals, max_dist_from_refs, min_dist_between, reference_col, group_by_col, value_col):
+
+  # Specify the name of the column in which the income/consumption welfare definition is stored
+  welfare_colname = 'welfare_type'
+
+  # Creat dataframes for thee scenarios:
+  # Scenario 1: only allow income data
+  df_inc_filter = df.loc[df[welfare_colname] == "income", :]
+  df_inc_filter.name = "Income"
+
+  # Scenario 2: only allow consumption data
+  df_cons_filter = df.loc[df[welfare_colname] == "consumption", :]
+  df_cons_filter.name = "Consumption"
+  # Scenario 3: allow a mix – dropping consumption data where income data is available in the same year
+  df_mixed = df.copy()
+
+  df_mixed['welfare_count'] = df_mixed.groupby([reference_col, group_by_col])[welfare_colname].transform('count')
+
+  df_mixed = df_mixed.loc[(df_mixed['welfare_count'] == 1) | (df_mixed[welfare_colname] == "income")]
+
+  df_mixed.name = "Mixed"
+  #  Store the scneario dataframes in a list
+  df_scenarios = [df_inc_filter, df_cons_filter, df_mixed]
+
+  # Run the matching function on each scenario
+  scenario_matches = [merge_two_ref_matches(
+    df_scenario, 
+    reference_vals, 
+    max_dist_from_refs, 
+    min_dist_between, 
+    reference_col, 
+    group_by_col, 
+    value_col) for df_scenario in df_scenarios]
+
+  # Combine the first two scenarios.
+  df_combined_matches = pd.concat([scenario_matches[0], scenario_matches[1]], keys=[df_scenarios[0].name, df_scenarios[1].name])
+
+  # Tidy up indexes
+  df_combined_matches = df_combined_matches.reset_index()
+
+  df_combined_matches = df_combined_matches.drop('level_1', axis=1)
+
+  df_combined_matches = df_combined_matches\
+    .rename(columns={"level_0": "pip_welfare"})
+
+  # Add in third scenario.
+  df_combined_matches = pd.concat([df_combined_matches, scenario_matches[2]])
+
+  # add scenario name to te pip_welfare column
+  df_combined_matches['pip_welfare'] = df_combined_matches['pip_welfare'].fillna(df_scenarios[2].name)
+
+  # Keep only one match per group (e.g. per Country) - in the priority laid out in the df_scenarios list above (income only -> consumption only -> mixed)
+    # First count the matches
+  df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
+    # Then drop any matches from the lowest priority where there are multiple matches
+  df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[2].name)]
+    #  Repeat at the next level of priority
+  df_combined_matches['match_count'] = df_combined_matches.groupby(group_by_col)['pip_welfare'].transform('count')
+  df_combined_matches = df_combined_matches.loc[(df_combined_matches['match_count']==1) | ~(df_combined_matches['pip_welfare']==df_scenarios[1].name)]
+
+  # Drop the match count column
+  df_combined_matches = df_combined_matches.drop('match_count', axis=1)
+
+  return df_combined_matches
+
+def prep_selected_vars(selected_vars):
+
+  # selected_vars = [
+  #   'LIS: Top_10_share_equiv_market',
+  #   'WID: Bottom_50_share_posttax_dis',
+  #   'PIP: Top_10_share']
+
+  selected_sources = [var.split(sep, 1)[0] for var in selected_vars]
+  selected_var_names = [var.split(sep, 1)[1] for var in selected_vars]
+
+  match_data = {
+    "sources": selected_sources,
+    "var_names": selected_var_names
+          }
+
+  return match_data
+
+# Specify the tolerances for each kind of metric
+def return_tolerance_for_var(var):
+
+  if 'Gini' in var:
+    tolerance = 1
+
+  elif 'Top_10_share' in var:
+    tolerance = 1
+
+  elif 'Bottom_50_share' in var:
+    tolerance = 1
+
+  else:
+    tolerance = None
+
+  return tolerance
+
+def prep_data(
+      data_to_match,
+      reference_vals,
+      max_dist_from_refs,
+      min_dist_between,
+      all_metrics_requirement,
+      reference_col,
+      group_by_col,
+      outlier_cut_off_upper):
 
@@
     var = data_to_match["var_names"][i]
+
+    tolerance = return_tolerance_for_var(
+      data_to_match["var_names"][i]
+    )
 
@@
 
-        ref_matches = pd.concat(matches, keys=source_var_keys)
+  plot_data = pd.concat(matches, keys=source_var_keys)
 
   # Tidy up indexes
-        ref_matches = ref_matches.reset_index()
-
-        ref_matches = ref_matches.drop('level_1', axis=1)
-
-        ref_matches = ref_matches\
+  plot_data = plot_data.reset_index()
+
+  plot_data = plot_data.drop('level_1', axis=1)
+
+  plot_data = plot_data\
     .rename(columns={"level_0": "source_var"})
-
-        return ref_matches
-
-      def prep_plot_and_tables(
-        data_to_match,
-        reference_vals, 
-        max_dist_from_refs,
-        min_dist_between,
-        reference_col,
-        group_by_col,
-        tolerance,
-        outlier_cut_off_upper,
-        all_metrics_requirement):
-
-        # Other spec:
-        region_col = 'region_alt'
-
-        # ---- Prep data ---------
-
-        plot_data = prep_and_merge_multiple_vars(
-          data_to_match = data_to_match,
-          reference_vals = reference_vals, 
-          max_dist_from_refs = max_dist_from_refs,
-          min_dist_between = min_dist_between,
-          reference_col = reference_col,
-          group_by_col = group_by_col
-        )
 
@@
 
-        # ----- Prep plot ------
-
-        # I grab what I am guessing to be the max and min tick marks on the x axist, in order to define the coordinates of a 'tolerance' ribbon.
-        x_min = plot_data[x_axis].min()
-        x_min_floor = np.floor(x_min * 10) / 10
-
-        x_max = plot_data[x_axis].max()
-        x_max_ceiling = np.ceil(x_max * 10) / 10
-
-        # Set the coordinates of the shaded ribbon
-        shaded_coord = pd.DataFrame({'x': [x_min_floor, x_max_ceiling, x_max_ceiling, x_min_floor], 
-          'y': [x_min_floor-tolerance, x_max_ceiling-tolerance, x_max_ceiling+tolerance , x_min_floor+tolerance]})
-
-        plot = (ggplot(plot_data
-        , aes(x_axis, y_axis, alpha=0.5))
-        + geom_point(aes(color=region_col, size='avg_pop', alpha = 0.6))
-        + facet_wrap('~ source_var')
-        + geom_polygon(aes(x='x', y='y', alpha = 0.2), data=shaded_coord)
-        + theme_light()
-        + theme(legend_position='none')
-        )
-
-        # ---- Prep summary table ----
+  #  PREP SUMMARY TABLE
 
@@
 
-        # ----- Return outputs ------
-        plot_and_tables = {
-          "table_data": plot_data,
-          "plot": plot,
-          "summary": df_summary
-        }
-
-        return plot_and_tables
-
-      def prep_selected_vars(selected_vars):
-
-        # selected_vars = [
-        #   'LIS: Top_10_share_equiv_market',
-        #   'WID: Bottom_50_share_posttax_dis',
-        #   'PIP: Top_10_share']
-
-        selected_sources = [var.split(sep, 1)[0] for var in selected_vars]
-        selected_var_names = [var.split(sep, 1)[1] for var in selected_vars]
-
-        match_data = {
-          "sources": selected_sources,
-          "var_names": selected_var_names
-          }
-
-        return match_data
-
-      data_to_match = prep_selected_vars(input.metrics())
-
-      plot_and_tables = prep_plot_and_tables(
+  return plot_data, df_summary
+
+```
+
+```{python}
+
+#  ---------------------------------
+#  ---------- SECTION 2: Specifications -------------
+#  ----------------------------------
+# These parameters are not controlled in the app.
+
+reference_col = 'Year'
+group_by_col = 'Entity'
+
+outlier_cut_off_upper = None
+region_col = 'region_alt'
+
+```
+
+```{python}
+
+#  ---------------------------------
+#  ---------- SECTION 3: Read in data -------------
+#  ----------------------------------
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/pip.csv'
+
+df_pip = pd.read_csv(url)
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/wid.csv'
+
+df_wid = pd.read_csv(url)
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/lis.csv'
+
+df_lis = pd.read_csv(url)
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/region_mapping.csv'
+
+df_regions = pd.read_csv(url)
+
+url = 'https://raw.githubusercontent.com/JoeHasell/personal_site/main/PhD_pages/data_appendices/data/clean/population.csv'
+
+df_pop = pd.read_csv(url)
+
+```
+
+```{python}
+
+#  ---------------------------------
+#  ---------- SECTION 4: Gather data into dictionary and collete varnames -------------
+#  ----------------------------------
+
+data = {"PIP": df_pip,
+        "WID": df_wid,
+        "LIS": df_lis}
+
+# make an empty list to add all the varnames
+all_vars = list()
+
+# A string to use to separate the source name from the var name
+sep = ": "
+
+# for each df in 'data'
+for k,v in data.items():
+
+  # make a list of varnames
+  vars = list(v.columns)
+
+  # remove the id cols
+  vars.remove(reference_col)
+  vars.remove(group_by_col)
+
+  # append the name of the source (the key of 'data') at the front
+  vars = [f'{k}{sep}' + var for var in vars]
+
+  # append to the running list
+  all_vars = all_vars + vars
+
+```
+
+We don't run this reactive step in the testing. We just run the prep data function in the plot and table functions below.
+
+```{python}
+# def reactive_data():
+
+#   # For testing we define manually, rather than picked up from the app inputs
+#   data_to_match = prep_selected_vars(input.metrics())
+
+#   reference_vals = list(input.reference_vals())
+
+#   max_dist_from_refs = input.max_dist_from_refs()
+#   min_dist_between = input.min_dist_between()
+#   all_metrics_requirement = input.all_metrics_requirement()
+
+#   prepped_data = prep_data(
+#       data_to_match = data_to_match,
+#       reference_vals = reference_vals,
+#       max_dist_from_refs = max_dist_from_refs,
+#       min_dist_between = min_dist_between,
+#       all_metrics_requirement = all_metrics_requirement,
+#       reference_col = reference_col,
+#       group_by_col = group_by_col,
+#       tolerance = tolerance,
+#       outlier_cut_off_upper = outlier_cut_off_upper)
+
+#   return prepped_data
+
+```
+
+```{python}
+
+def interactive_plot(
+  selected_vars,
+  reference_vals,
+  max_dist_from_refs,
+  min_dist_between,
+  all_metrics_requirement):
+
+  data_to_match = prep_selected_vars(selected_vars)
+
+  prepped_data = prep_data(
       data_to_match = data_to_match,
-        reference_vals = list(input.reference_vals()), 
-        max_dist_from_refs = input.max_dist_from_refs(),
-        min_dist_between = input.min_dist_between(),
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = min_dist_between,
+      all_metrics_requirement = all_metrics_requirement,
       reference_col = reference_col,
       group_by_col = group_by_col,
-        tolerance = tolerance,
-        outlier_cut_off_upper = outlier_cut_off_upper,
-        all_metrics_requirement = input.all_metrics_requirement()
+      outlier_cut_off_upper = outlier_cut_off_upper)
+
+  plot_data = prepped_data[0]
+
+  fig = px.scatter(
+    plot_data, 
+    x=f'value{reference_vals[0]}', 
+    y=f'value{reference_vals[1]}',
+    color=region_col, 
+    facet_col='source_var', 
+    facet_col_wrap=1, 
+    size='avg_pop',
+    hover_name=group_by_col,
+    hover_data={
+      f'value{reference_vals[0]}':':.2f',
+      f'{reference_col}{reference_vals[0]}':True,
+      f'value{reference_vals[1]}':':.2f',
+      f'{reference_col}{reference_vals[1]}':True,
+      region_col: True,
+      'avg_pop': False,
+      'source_var': False},
+      height=1000
     )
 
-      plot = plot_and_tables['plot']
-
-      return plot
-
-app = App(app_ui, server)
+  fig.update_layout(legend=dict(
+      orientation="h",
+      yanchor="bottom",
+      y=1.02,
+      xanchor="right",
+      x=1
+  ))
+
+  fig.for_each_annotation(lambda a: a.update(text=a.text.split("=",1)[1]))
+
+  fig.update_layout(
+      legend_title=""
+  )
+
+  return fig
 
 ```
+
+```{python}
+def table_test(selected_vars,
+  reference_vals,
+  max_dist_from_refs,
+  min_dist_between,
+  all_metrics_requirement):
+
+  data_to_match = prep_selected_vars(selected_vars)
+
+  prepped_data = prep_data(
+    data_to_match = data_to_match,
+    reference_vals = reference_vals,
+    max_dist_from_refs = max_dist_from_refs,
+    min_dist_between = min_dist_between,
+    all_metrics_requirement = all_metrics_requirement,
+    reference_col = reference_col,
+    group_by_col = group_by_col,
+    outlier_cut_off_upper = outlier_cut_off_upper)
+
+  summary = prepped_data[1].reset_index()
+
+  return summary
+
+```
+
+```{python}
+
+fig = interactive_plot(
+  selected_vars = [
+      'WID: Top_10_share_pretax',
+      'PIP: Top_10_share'],
+  reference_vals = [1990,2015],
+  max_dist_from_refs = 5,
+  min_dist_between = 10,
+  all_metrics_requirement = True
+)
+
+fig.show()
+
+```
+
+```{python}
+
+summary_table = table_test(
+  selected_vars = [
+      'WID: Top_10_share_pretax',
+      'PIP: Top_10_share'],
+  reference_vals = [1990,2015],
+  max_dist_from_refs = 5,
+  min_dist_between = 10,
+  all_metrics_requirement = True
+)
+
+summary_table
+
+```

### 2023-02-09 PhD_pages/paper_pages/paper1.qmd
Paper 1: reorganise app and initial drafting of paper

@@
 
-What has happened to within-country inequality globally?
+What has happened to inequality within countries over the past generation?
+
+# Introduction
+
+- A range of motives for taking a global view: SDGs; globalization – trade; ....
+
+- We don't have a clear global view.    
+  [UN summary of inequality trends](https://www.un.org/en/un75/inequality-bridging-divide)
+
+- Why is it that we don't have a clear view?
+
+    1) Most academic research and commentary has focussed on high income countries (refs), especially the US.
+There are many reasons why we should not expect trends in inequality in these countries to reflect broader global trends.
+
+    2) Hampered by the available data
+        - **Coverage** – for instance no surveys.
+        - **Comparability of data** – Some secondary databases with global coverage mixes very heterogenous datapoints (WIID) (Find Atkinson and Brandolini ref from Handbook...).
+Those with a greater degree of harmonization tend to lack global coverage (LIS, EU-SILC, SEDLAC, until recently PIP)
+        - **Some trade off between the two is inevitable** – There is no global survey of incomes.
+Countries run surveys based on their own needs and capacities.
+As such they differ.
+Saying something global means accepting a degree of heterogenity in the data.
+Similarly, definitions of income for tax purposes vary, impacting the comparability of tax data.
+
+    3) Contradictory global views are presented by different global databases.
+        - The data situation has improved drastically: thanks to harmonized databases with global coverage. In particular:
+            - PIP – recently expanded to include high income countries
+            - WID – Transition from WTID, and the DINA methodology, global coverage.
+            - LIS: some data from middle and low income countries.
+        - But the 'take home' messages from the institutions producing these databases offer very contradictory views.
+            - "In recent decades, income inequality 
+has increased in nearly all countries, 
+but at different speeds, suggesting 
+that institutions and policies matter in shaping inequality." (WIR 2018)
+            - "[Across a range of approaches] aver-
+age  inequality  within  countries  increased   in  the  1990s,  but  has  been  falling  in  recent years according to most specifications." (World Bank, 2016)
+
+        - You can see this in the latest data from these two sources, using the metric typically favoured by each: Top 1% from WID vs Gini from PIP.
+
+        - These different messages are based on a mix of many different methodological differences:
+            - different metrics, sensitive to different parts of the distribution; 
+            - Different welfare concepts; 
+            - Different underlying data sources and adjusmtents to the top (or lack thereof);
+            - Different coverage or time periods being analysed.
+
+# Literature review – Global studies of inequality
+
+- Atkinson and Brandolini (2010) argue that global inequality analyses in particular need to consider both absol ute  and relative differences.
+
+# Data review Why is our understanding of inequality around the world currently limited
+
+- **A global view is hampered by the available data**.
+
+- **But databases have improved substantially recently**
+Thanks to improvements and expansion of key global harmonized databases, we are now in a much stronger position to take a global view on inequality within countries.
+
+There are range of databases and a full review is out of the scope of this paper. Focus on the three databases of harmonized distributional data providing global or near-global coverage.
+
+- **PIP** 
+    - Survey data
+    - Mix of disposable income and consumption data (definitions vary)
+    - Per capita
+    – Coverage: now includes high income countries (Add a box/visualization with coverage). Global coverage from 1990 onwards.
+    - A range of inequality indices provided 'as standard', but the whole distribution may be queried via API.
+Either microdata or, in a minority of cases, a parametric distribution fitted to 'grouped data' (e.g.
+income deciles)
+
+- **WID** 
+    – Survey data, tax data, National accounts
+    - Coverage much greater for pre-tax national income. But post-tax concepts also available for a subset of observations.
+    - Greatest coverage for individuals – equal split where underlying data was available for couples.
+    - Coverage 
+    - Estimates of the whole distribution Generalized Pareto Interpolation. Though longer-run coverage is higher for pretax top 1% share (reliance on tax data)
+
+- **LIS**
+    - Offers less coverage, but is notable for providing access to microdata.
 

### 2023-02-09 PhD_pages/data_appendices/app_compare_ref_year_change.py

@@ -1,0 +1,36 @@
+# Run this app with `python app.py` and
+# visit http://127.0.0.1:8050/ in your web browser.
+
+from dash import Dash, html, dcc
+import plotly.express as px
+import pandas as pd
+
+app = Dash(__name__)
+
+# assume you have a "long-form" data frame
+# see https://plotly.com/python/px-arguments/ for more options
+df = pd.DataFrame({
+    "Fruit": ["Apples", "Oranges", "Bananas", "Apples", "Oranges", "Bananas"],
+    "Amount": [4, 1, 2, 2, 4, 5],
+    "City": ["SF", "SF", "SF", "Montreal", "Montreal", "Montreal"]
+})
+
+fig = px.bar(df, x="Fruit", y="Amount", color="City", barmode="group")
+
+app.layout = html.Div(children=[
+    html.H1(children='Test Dash app'),
+
+    html.Div(children='''
+        Dash: A web application framework for your data.
+    '''),
+
+    dcc.Graph(
+        id='example-graph',
+        figure=fig
+    )
+])
+
+if __name__ == '__main__':
+    app.run_server(debug=True)
+
+

### 2023-02-10 PhD_pages/Paper_1_within_country_ineq/data_appendix/1_A_prepare_PIP.qmd
Restructuring folders so that everything is contained in one complete subfolder per paper

---
title: "Appendix 1.A: Preparing PIP inequality data"
format: html
warning: false

---

# Processing steps

The titles in each of the boxes below summarise the steps taken to prepare the PIP data. In each case you can expand the box to see the code used.

::: {.callout-note collapse="true" appearance="minimal"}
### Packages and settings

```{python}

import pandas as pd

pd.set_option('display.max_rows', None)
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Load PIP data
We load the data from a file prepared from the PIP API by Our World in Data.

```{python}
# df_pip<- read.csv("https://raw.githubusercontent.com/owid/poverty-data/main/datasets/pip_dataset.csv")

#  A local version, in case internet isn't available 
df_pip = pd.read_csv("data/original/pip_dataset.csv")

```

The dataframe looks like this:
```{python}
df_pip.head()
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Select PPP version
We will use the 2017 PPPs.
```{python}
df_pip = df_pip.loc[df_pip["ppp_version"] == 2017]
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Drop sub-national data

The PIP data includes observations for national, urban and rural populations (`reporting_level`):

```{python}
df_pip.groupby('reporting_level').count()
```

Here we add a count of the number of reporting levels per country-year.
```{python}
df_pip['n_reporting_level'] = df_pip\
  .groupby(["country", "year"])['reporting_level']\
  .transform('count')

df_pip.head(20)  
```

We only want one reporting level for each country-year.

That is already true for most country-years. But some include all three reporting levels.
```{python}
df_pip.groupby('n_reporting_level').count()

```

Here are the countries which include all three reporting levels:
```{python}

multi_reporting_level_countries = df_pip.loc[df_pip['n_reporting_level'] == 3, "country"].unique()

multi_reporting_level_countries

```

For the country-years with 3 reporting levels, we keep only the national estimates. 
```{python}
df_pip = df_pip.loc[(df_pip['n_reporting_level'] != 3) | (df_pip['reporting_level'] == 'national') ]
```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Drop regional data

The dataset includes aggregated estiates for world regions.

Here we define a list of these aggregate entities and drop them from the data.

```{python}
drop_aggs = [
  "East Asia and Pacific",
  "South Asia",
  "Europe and Central Asia",
  "High income countries",
  "Latin America and the Caribbean",
  "Middle East and North Africa",
  "Sub-Saharan Africa",
  "World"
]

df_pip = df_pip.loc[~df_pip.country.isin(drop_aggs)]

```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Calculate additional variables

```{python}

# Bottom 50% share – sum shares of lower 5 deciles
sum_cols = [
  "decile1_share",
  "decile2_share",
  "decile3_share",
  "decile4_share",
  "decile5_share"]

df_pip["Bottom_50_share"] = df_pip[sum_cols].sum(axis=1)

df_pip["P90_P50_ratio"] = df_pip["decile9_thr"]/df_pip["median"]

df_pip["Gini"] = df_pip['gini'] * 100

```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Rename and select needed columns

```{python}

df_pip = df_pip.rename(columns={
    'decile10_share':'Top_10_share',
    'mld': 'MLD',
    "country": "Entity",
    "year": "Year"
  })

keep_cols = [
  "Year",
  "Entity",
  "Gini",
  "Top_10_share",
  "Bottom_50_share",
  "MLD",
  "welfare_type"
]

df_pip = df_pip[keep_cols]

```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### See the prepared PIP data

The first 100 rows of the dataframe looks like this:
```{python}
df_pip.head(100)
```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Save the data
I save the prepared data locally:
```{python}
df_pip.to_csv("data/clean/pip.csv", index=False)
```
:::

## Important notes

### Income and consumption data in PIP
The PIP data includes a mix of income and consumption observations – including both for some country-years.

### 2023-02-10 PhD_pages/Paper_1_within_country_ineq/data_appendix/1_B_prepare_WID.qmd
Restructuring folders so that everything is contained in one complete subfolder per paper

--- 2023-01-29 PhD_pages/data_appendices/1_B_prepare_WID.qmd
@@
 
+The data was prepared in three separate files according to three income concepts: post_tax disposable income, post_tax national income and pretax national income.
+
+These we append into a single file.
+
 ```{python}
@@
 
-df_wid = pd.read_csv("data/original/wid.csv")
+inc_concepts = [
+  'posttax_dis',
+  'posttax_nat',
+  'pretax'
+]
+
+# filepaths = [f'data/original/wid_{inc_concept}.csv' for inc_concept in inc_concepts]
+
+filepaths = [f'https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/PabloArriagada/WID/data/final/wid_{inc_concept}.csv' for inc_concept in inc_concepts]
+
+wid_dfs = [pd.read_csv(filepath) for filepath in filepaths]
+
+df_wid = pd.concat(wid_dfs, keys=inc_concepts)
+
+# Tidy up indexes
+df_wid = df_wid.reset_index()
+
+df_wid = df_wid.drop('level_1', axis=1)
+
+df_wid = df_wid\
+  .rename(columns={"level_0": "inc_concept"})
 
@@
 
-The dataframe looks like this:
+The resulting dataframe looks like this:
 ```{python}
@@
 # Gini as index
-df_wid["Gini"] = df_wid['Gini.coefficient'] * 100
+df_wid["Gini"] = df_wid['Gini coefficient'] * 100
 
@@
 df_wid = df_wid.rename(columns={
-    'P90.P100...share.of.the.top.10.':'Top_10_share',
-    'P0.P50...share.of.the.bottom.50.':'Bottom_50_share',
-    'P90...income.threshold.of.the.top.10.':'P90',
-    'P10...income.threshold':'P10'
+    'P90-P100 - share of the top 10%':'Top_10_share',
+    'P0-P50 - share of the bottom 50%':'Bottom_50_share',
+    'P99-P100 - share of the top 1%':'Top_1_share'
   })
@@
   "Top_10_share",
-  "Bottom_50_share"
+  "Bottom_50_share",
+  "inc_concept"
 ]
@@
 df_wid = df_wid[keep_cols]
+
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Pivot the data wider by income concept
+
+```{python}
+
+df_wid = df_wid.set_index(['Year','Entity','inc_concept']).unstack()
+
+df_wid.columns = df_wid.columns.map(lambda x: '{}_{}'.format(x[0], x[1]))
+
+df_wid = df_wid.reset_index()
+
+df_wid = df_wid.sort_values(by=['Entity', 'Year'])
 

### 2023-02-10 PhD_pages/Paper_1_within_country_ineq/data_appendix/1_C_prepare_LIS.qmd
Restructuring folders so that everything is contained in one complete subfolder per paper

---
title: "Appendix 1.C: Preparing LIS inequality data"
format: html
warning: false

---

::: {.callout-note collapse="true" appearance="minimal"}
### Packages and settings

```{python}
import pandas as pd
pd.set_option('display.max_rows', None)
```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Load LIS data

We load a local file downloaded from Our World in Data

```{python}
df_lis = pd.read_csv("data/original/Luxembourg Income Study (LIS, 2022).csv")
```

The dataframe looks like this:
```{python}
df_lis.head(50)
```
:::

::: {.callout-note collapse="true" appearance="minimal"}
### Select and rename needed variables, calculating some additional inequality metrics

```{python}

inc_concepts = {

    "short_name": [
        "equiv_disposable",
        "equiv_market",
        "percap_disposable",
        "percap_market"],
    "long_name": [
        "Disposable household income, equivalized",
        "Market income, equivalized",
        "Disposable household income, per capita",
        "Disposable household income, per capita"
    ]
}

keep_vars = []

for i in range(len(inc_concepts['short_name'])):

    short_suffix = inc_concepts['short_name'][i]
    long_suffix = inc_concepts['long_name'][i]

    # Gini
    gini_new_var = f'Gini_{short_suffix}'
    gini_old_var = f'Gini coefficient ({long_suffix})'

    df_lis[gini_new_var] = df_lis[gini_old_var] * 100

    keep_vars.append(gini_new_var)

    # Top 10% share
    top10_new_var = f'Top_10_share_{short_suffix}'
    top10_old_var = f'Richest decile - Share ({long_suffix})'

    df_lis[top10_new_var] = df_lis[top10_old_var]

    keep_vars.append(top10_new_var)

    # Bottom 50% share
    sum_cols = [
    'Poorest decile - Share',
    '2nd decile - Share',
    '3rd decile - Share',
    '4th decile - Share',
    '5th decile - Share',
    ]

    sum_cols = [sub + f' ({long_suffix})' for sub in sum_cols]

    bottom50_new_var = f'Bottom_50_share_{short_suffix}'

    df_lis[bottom50_new_var] = df_lis[sum_cols].sum(axis=1)

    keep_vars.append(bottom50_new_var)

other_vars_to_keep = ["Year", "Entity"]

keep_vars = other_vars_to_keep + keep_vars

df_lis = df_lis[keep_vars]

```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### See the prepared data

The first 100 rows of the dataframe looks like this:
```{python}
df_lis.head(100)
```

:::

::: {.callout-note collapse="true" appearance="minimal"}
### Save the data
I save the prepared data locally:
```{python}
df_lis.to_csv("data/clean/lis.csv", index=False)
```
:::

### 2023-02-10 PhD_pages/Paper_1_within_country_ineq/paper1.qmd
Restructuring folders so that everything is contained in one complete subfolder per paper

--- 2023-02-09 PhD_pages/paper_pages/paper1.qmd
@@
 
+# Outline
+
+- We need a global view on inequality
+
+- Though there is much data, we don't have a clear global view.    
+  [UN summary of inequality trends](https://www.un.org/en/un75/inequality-bridging-divide)
+
+- A global view is hampered by many factors:
+    - In research and commentary, there is a disportionate focus high-income countries, in particular the US
+
+    - The extent and quality of available data
+        - Coverage
+        - Quality and comarabilty of data
+        - Some trade off between the two is inevitable
+
+    - **The major global inequality databases occupy different 'worlds' of measurement**
+
+    It is the latter that I will be addressing. We are in a relatively rich world in terms of inequality data – thanks in particlar to efforts of harmonized databases.
+
+    But we lose much of the value – different how disconnected
+
+     – that the rich doesn't translate to as  understanding
+
+- Two worlds of inequality measurement
+
 # Introduction
@@
         - **Coverage** – for instance no surveys.
-        - **Comparability of data** – Some secondary databases with global coverage mixes very heterogenous datapoints (WIID) (Find Atkinson and Brandolini ref from Handbook...).
+        - **Quality and comarabilty of data** – Some secondary databases with global coverage mixes very heterogenous datapoints (WIID) (Find Atkinson and Brandolini ref from Handbook...).
 Those with a greater degree of harmonization tend to lack global coverage (LIS, EU-SILC, SEDLAC, until recently PIP)
@@
 that institutions and policies matter in shaping inequality." (WIR 2018)
-            - "[Across a range of approaches] aver-
-age  inequality  within  countries  increased   in  the  1990s,  but  has  been  falling  in  recent years according to most specifications." (World Bank, 2016)
+            - "[Across a range of approaches] average  inequality  within  countries  increased   in  the  1990s,  but  has  been  falling  in  recent years according to most specifications." (World Bank, 2016)
 

### 2023-02-10 PhD_pages/older_structure/data_appendices/3_A_prepare_KLEMS.qmd
Restructuring folders so that everything is contained in one complete subfolder per paper

--- 2022-11-30 PhD_pages/data_appendices/3_A_prepare_KLEMS.qmd
@@
 format: html
+execute:
+  eval: false  # never run
 ---
+
+# To-do:
+
+- I moved the data folder, but haven't yet relected this in the code below.
 
@@
 
-There are many (`r df %>% filter(VA == CAP & is.na(LAB)) %>% count() %>% as.numeric()`) country-year-industry observations where where CAP is 100% of VA, but LAB is missing.
+There are many (r df %>% filter(VA == CAP & is.na(LAB)) %>% count() %>% as.numeric()) country-year-industry observations where where CAP is 100% of VA, but LAB is missing.
 
-Most of these (`r df %>% filter(VA == CAP & is.na(LAB) & VA == 0) %>% count() %>% as.numeric()`) are where both capital income and value added equal zero.
+Most of these (r df %>% filter(VA == CAP & is.na(LAB) & VA == 0) %>% count() %>% as.numeric()) are where both capital income and value added equal zero.
 Such observations appear only for the following industries:
@@
 
-There is a similar set of country-year-industry observations (n = `r df %>% filter(VA == LAB & is.na(CAP)) %>% count() %>% as.numeric()`) where VA = LAB but CAP is NA. These we list here.
+There is a similar set of country-year-industry observations (n = r df %>% filter(VA == LAB & is.na(CAP)) %>% count() %>% as.numeric()) where VA = LAB but CAP is NA. These we list here.
 
@@
 
-There are some observations (n=`r main_industry_accounts %>% filter(is.na(VA) & is.na(CAP) & is.na(LAB)) %>% count() %>% as.numeric()`) without .... . We group these by country and industry.
+There are some observations (n=r main_industry_accounts %>% filter(is.na(VA) & is.na(CAP) & is.na(LAB)) %>% count() %>% as.numeric()`) without .... . We group these by country and industry.
 
@@
 In general the sum across industries is very close to to the reported economy-wide aggregate.
-In the table below we show all observations where the ratio deviates from 100% by more than `r tolerance*100`%.
+In the table below we show all observations where the ratio deviates from 100% by more than r tolerance*100`%.
 

### 2023-02-10 PhD_pages/older_structure/data_appendices/3_B_prepare_OECD.qmd
Restructuring folders so that everything is contained in one complete subfolder per paper

--- 2022-11-30 PhD_pages/data_appendices/3_B_prepare_OECD.qmd
@@
 
-For instance, out of `r length(main_industry_codes)` aggregate industry codes, these country years have data for fewer than 18:
+For instance, out of r length(main_industry_codes)` aggregate industry codes, these country years have data for fewer than 18:
 
@@
 
-In particular, the dataset contains only `r main_industry_accounts %>% filter(ind_code == "VU")%>%count()` country-year observations for `VU` (as compared to `r main_industry_accounts %>% filter(ind_code == "VA0")%>%count()` for `VA0`) and these are in general only NA or 0 values.
+In particular, the dataset contains only r main_industry_accounts %>% filter(ind_code == "VU")%>%count()` country-year observations for `VU` (as compared to r main_industry_accounts %>% filter(ind_code == "VA0")%>%count()` for `VA0`) and these are in general only NA or 0 values.
 

### 2023-02-10 PhD_pages/Paper_1/Presentation_INET_16_feb_inequality_around_the_world.qmd
Paper 1: Drafting presentation

---
title: "Triangulating a global view on inequality"
author: "Joe Hasell"
format: 
    revealjs:
        theme: [default, revealjs_joe_theme.scss] 
bibliography: references.bib
---

## Concepts

- Three concepts of global inequality (Milanovic):

    1. Average within-country inequality

    2. Average within-country inequality, weighting by population

    3. Inequality across the world population taken as a whole (future work)

- Income(ish) inequality, not wealth

- Relative inequality, not absolute

## Why do we need a global view?

- Cosmopolitan values (e.g. UN's SDGs)
- Opportunities and limitations for national policy to address inequality
- Priorities for international policy (taxation, migration, aid...)
- Priorities for philanthropy 
- Winners and losers of globalization ('Degrowth')

## A clear global view is lacking {.smaller}

- **Limitations in the available data**
    - Coverage
    - Quality and comarabilty
    - Much improved recently!

- **Different conceptual focus in richer and poorer countries**
    - Inequality/relative poverty in high income countries
    - Absolute poverty in poorer countries
    - Much improved recently! (Atkinson commission... Atkinson & Brandolini 2010... Ravallion 2019... 'Societal poverty' – World Bank 2018, 2020, 2022)

- **Different 'worlds' of measurement occupied by global harmonized databases**
    - Different metrics (Gini, top income shares etc.); Different welfare concepts; underlying data sources; methods (e.g. treatment of 'missing top'; coverage or time periods analysed
    - Differences somewhat bundled together
    - Yield very different 'take home' messages

## Different 'worlds' of inequality measurement {.smaller}
**World Inequality Database**

> In recent decades, **income inequality has increased in nearly all countries**, but at different speeds...
[@Alvaredo2018-qq]

> Also find a quote about the within-coutry component of global inequality

## Different 'worlds' of inequality measurement {.smaller}
**World Bank, Poverty and Inequality Platform (formerly 'Povcal')**

> [Since c.1993] the number of countries with rising inequality was **slightly higher** than the number of countries with falling inequality. Yet, the average Gini among these 91 countries actually **declined**.
[@WBpoverty2016]

> Also find a quote about the within-coutry component of global inequality

## The view from the World Inequality Database {.smaller}
Top 1% share; pre-tax 'national' income; combining survey, tax and national accounts data; generalized pareto interpolation; individuals (equal split couples)

(CHART of top 1% WID)

## The view from the World Inequality Database {.smaller}
Top 1% share; pre-tax 'national' income; combining survey, tax and national accounts data; generalized pareto interpolation; individuals (equal split couples)

(TABLE of top 1% WID)

## The view from the World Bank Poverty and Inequality Platform {.smaller}
Gini; mix of consumption and disposable income; survey data (mostly micro, some 'grouped' data); lognormal or XXX interpolation (grouped data only); per capita household

(CHART of Gini PIP)

## The view from the World Bank Poverty and Inequality Platform {.smaller}
Gini; mix of consumption and disposable income; survey data (mostly micro, some 'grouped' data); lognormal or XXX interpolation (grouped data only); per capita household

(Table of Gini PIP)

## Connecting up these 'worlds' is a low hanging fruit {preview-links=true}

- The respective institutions – WIL and World Bank – have built hugely valuable and powerful platforms.
- We are in a better position than ever to get a global view on inequality
- But the two 'worlds' of measurement makes the current data terrain is *very* confusing for policimakers, researchers in adjacent fields, and the public.

Example: a summary of inequality trends [from the UN](https://www.un.org/en/un75/inequality-bridging-divide).

## Connecting up these 'worlds' is a low hanging fruit

- Existing studies tend to be based on only one data source, or only partially bridge between the two.

    - Flagship reports of the two institutions
    - Lakner Milanovic
    - Groeningen...
    - Ravi Kanbur...

## Connecting up these 'worlds'

1. Line up what we can:
    - Coverage and time period
    - Metrics

    - (With lower coverage) Income concept
        - WID pre-tax
        - LIS?
        - Income vs consumption: GCIP

    - Methods: Hacky top income adjustment to PIP data?

2. Triangulate
    - Flag what we *don't* know – Isolate what is causing the differences between 'take home' messages
        - Particular methodological choices?
        - Particular countries?
    - Gather together what we *do* know, irrespecive of may remaining differences

## Bibliography

### 2023-02-10 presentations.qmd
Paper 1: Drafting presentation

---
title: "Presentations"
format: html
---

### From $1.90 to $2.15: the updated International Poverty Line
20/01/2023

A presentation given to Masters students in Comparative Social Policy at the Department for Social Policy, University of Oxford.

[Slides](other_presentations/DSPI_CSP_masters_Jan2023/DSPI_CSP_masters_Jan2023.qmd)

### 2023-02-22 PhD_pages/Paper_1/Presentation_INET_16_feb_inequality_around_the_world.qmd
Paper 1: preparing presentation

@@
 title: "Triangulating a global view on inequality"
-author: "Joe Hasell"
+author: "<br> <br> Joe Hasell – DPhil candidate at DSPI University of Oxford; Our World in Data<br> <br> Pablo Arriagada – Our World in Data"
 format: 
@@
 
+```{python}
+
+from plotnine import *
+import plotly.express as px
+import pandas as pd
+import numpy as np
+from IPython.display import display, HTML, IFrame, Markdown
+import dataframe_image as dfi
+import imgkit
+
+from wrangling_and_plotting_functions import *
+
+pd.set_option('display.max_rows', None)
+
+```
+
+```{python}
+
+fp = 'data_appendix/data/clean/pip.csv'
+
+df_pip = pd.read_csv(fp)
+
+fp = 'data_appendix/data/clean/wid.csv'
+
+df_wid = pd.read_csv(fp)
+
+fp = 'data_appendix/data/clean/lis.csv'
+
+df_lis = pd.read_csv(fp)
+
+fp = 'data_appendix/data/clean/region_mapping.csv'
+
+df_regions = pd.read_csv(fp)
+
+fp = 'data_appendix/data/clean/population.csv'
+
+df_pop = pd.read_csv(fp)
+
+fp = 'data_appendix/data/clean/region_population.csv'
+
+df_pop_regions = pd.read_csv(fp)
+
+# Gather data into dictionary
+source_dfs = {
+    "PIP": df_pip,
+    "WID": df_wid,
+    "LIS": df_lis
+    }
+
+tolerance_lookup =  {
+                'var_search_term': [
+                    'Gini',
+                    'Top 1pc share', 
+                    'Top 10pc share',
+                    'Bottom 50pc share'],
+                'var_tolerance': [
+                    1,
+                    1,
+                    1,
+                    1
+                ]
+            }
+
+```
+
+```{python}
+def make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement,selected_vars):
+    msg = f'{reference_vals[0]} vs {reference_vals[1]} (+/- {max_dist_from_refs} years)'
+
+    if len(selected_vars)>1:
+        if all_metrics_requirement:
+            msg = f'{msg} – Matched countries only'
+        else:
+            msg = f'{msg} – All available countries'
+
+    return(msg)
+
+```
+
 ## Concepts
+::: {.r-fit-text}
+- Three concepts of global inequality [not quite @Milanovic2007; or @AnandSegal2015-nh]
 
-- Three concepts of global inequality (Milanovic):
+    1*. Average within-country inequality
 
-    1. Average within-country inequality
+    2*. Average within-country inequality, weighting by population
 
-    2. Average within-country inequality, weighting by population
+    3. Inequality across the world population taken as a whole
+        - Future work
 
-    3. Inequality across the world population taken as a whole (future work)
+- We will be looking at:
+    - Income(ish) inequality, not wealth
+    - Relative inequality, not absolute
+:::
 
-- Income(ish) inequality, not wealth
+## {.smaller}
+### Why do we need a global view?
+::: {.r-fit-text}
+::: {.incremental}
+- Understanding the winners and losers of globalization and economic growth
+    - *"We’re fighting for Main Street, not Wall Street. We have rejected globalism and embraced patriotism."* - Donald Trump, [Aug 2020](https://www.youtube.com/watch?v=l83sbXD0nuY&t=11m15s)
+- To inform national policy to address inequality – both opportunities and limits
+- To inform international policy to address inequality (financial reporting transparency, tax havens, global taxation, migration, aid...)
+- A reflection of cosmopolitan values and outlook (e.g. UN's SDGs)
+- As a guide to 'Effective Altruism', including charitable giving
+:::
+:::
 
-- Relative inequality, not absolute
+## {.smaller}
+### A clear global view is lacking
+::: {.r-fit-text}
+::: {.incremental}
 
-## Why do we need a global view?
+1. **Limitations in the available data**
+    - Coverage
+    - Quality and comarabilty within secondary databases [@AtkinsonBrandolini2001-qt]
+    - Much improved recently!
+        - Expanded coverage of harmonized secondary databases: World Bank's *Poverty and Inequality Platform* (PIP), *World Inequality Database* (WID), *Luxembourg Income Study* (LIS)
 
-- Cosmopolitan values (e.g. UN's SDGs)
-- Opportunities and limitations for national policy to address inequality
-- Priorities for international policy (taxation, migration, aid...)
-- Priorities for philanthropy 
-- Winners and losers of globalization ('Degrowth')
+2. **Different conceptual focus in richer and poorer countries**
+    - Income inequality/relative poverty in high income countries
+    - Absolute poverty in poorer countries
+    - Much improved recently!
+        - Atkinson commission [-@Atkinson2016-xp]; 'Societal poverty' – [@World_Bank2018-eo]; UN SDGs; expansion of WID coverage
 
-## A clear global view is lacking {.smaller}
+3. **Different 'worlds' of measurement**
+    - Many methodological differences between different global databases, *often bundled together*
+        - Yield very different take-home messages about inequality and how it has evolved
 
-- **Limitations in the available data**
-    - Coverage
-    - Quality and comarabilty
-    - Much improved recently!
+:::
+:::
 
-- **Different conceptual focus in richer and poorer countries**
-    - Inequality/relative poverty in high income countries
-    - Absolute poverty in poorer countries
-    - Much improved recently! (Atkinson commission... Atkinson & Brandolini 2010... Ravallion 2019... 'Societal poverty' – World Bank 2018, 2020, 2022)
-
-- **Different 'worlds' of measurement occupied by global harmonized databases**
-    - Different metrics (Gini, top income shares etc.); Different welfare concepts; underlying data sources; methods (e.g. treatment of 'missing top'; coverage or time periods analysed
-    - Differences somewhat bundled together
-    - Yield very different 'take home' messages
-
-## Different 'worlds' of inequality measurement {.smaller}
-**World Inequality Database**
+## {.smaller}
+### Different 'worlds': *World Inequality Database*
+::: {.incremental}
 
@@
 [@Alvaredo2018-qq]
+:::
 
-> Also find a quote about the within-coutry component of global inequality
+## {.smaller}
+### Different 'worlds': *World Inequality Database*
 
-## Different 'worlds' of inequality measurement {.smaller}
-**World Bank, Poverty and Inequality Platform (formerly 'Povcal')**
+The vast majority of global inequality (concept 3) is now due to differences **within** countries.
+
+![](images/WID_global_ineq_decomp.png)
+
+## {.smaller}
+### Different 'worlds': *World Bank, Poverty and Inequality Platform*
+::: {.incremental}
 
@@
 [@WBpoverty2016]
+:::
 
-> Also find a quote about the within-coutry component of global inequality
+## {.smaller}
+### Different 'worlds': *World Bank, Poverty and Inequality Platform*
 
-## The view from the World Inequality Database {.smaller}
-Top 1% share; pre-tax 'national' income; combining survey, tax and national accounts data; generalized pareto interpolation; individuals (equal split couples)
+The vast majority of global inequality (concept 3) remains due to differences **between** countries.
 
-(CHART of top 1% WID)
+![](images/pip_global_ineq_decomp.png){fig-align="center"}
 
-## The view from the World Inequality Database {.smaller}
-Top 1% share; pre-tax 'national' income; combining survey, tax and national accounts data; generalized pareto interpolation; individuals (equal split couples)
+##
+### Different 'worlds'
+::: {.r-fit-text}
+::: {.incremental}
+In terms of data availability, quality and comparability, we are in a better position than ever.
 
-(TABLE of top 1% WID)
+Yet some of the *most basic* questions are still open:
 
-## The view from the World Bank Poverty and Inequality Platform {.smaller}
-Gini; mix of consumption and disposable income; survey data (mostly micro, some 'grouped' data); lognormal or XXX interpolation (grouped data only); per capita household
+- Is rising inequality within countries a near-ubiquitous, global phenomenon?
+    - **WID:** Yes
+    - **PIP:** No
 
-(CHART of Gini PIP)
+- Where do global inequalities lie: in differences **between** richer and poorer countries, or in differences between richer and poorer people **within** countries?
+    - **WID:** Within countries
+    - **PIP:** Between countries
+:::
+:::
 
-## The view from the World Bank Poverty and Inequality Platform {.smaller}
-Gini; mix of consumption and disposable income; survey data (mostly micro, some 'grouped' data); lognormal or XXX interpolation (grouped data only); per capita household
+## 
+### Different methodologies
 
-(Table of Gini PIP)
+::: {.r-fit-text}
 
-## Connecting up these 'worlds' is a low hanging fruit {preview-links=true}
+::: {.incremental}
 
-- The respective institutions – WIL and World Bank – have built hugely valuable and powerful platforms.
-- We are in a better position than ever to get a global view on inequality
-- But the two 'worlds' of measurement makes the current data terrain is *very* confusing for policimakers, researchers in adjacent fields, and the public.
+The different 'worlds' are separated by many methodological differences, often bundled together:
 
-Example: a summary of inequality trends [from the UN](https://www.un.org/en/un75/inequality-bridging-divide).
+- Inequality indices (and focal points in the distribution)
+    - **WID:** Focus on the top
+    - **PIP:** Focus on overall inequality (e.g. Gini)
 
-## Connecting up these 'worlds' is a low hanging fruit
+- Welfare concepts
+    - **WID:** Focus on (and greatest coverage for) pre-tax national income
+    - **PIP:** mix of disposable income and consumption
 
-- Existing studies tend to be based on only one data source, or only partially bridge between the two.
+:::
 
-    - Flagship reports of the two institutions
-    - Lakner Milanovic
-    - Groeningen...
-    - Ravi Kanbur...
+:::
 
-## Connecting up these 'worlds'
+## 
+### Different methodologies
 
+::: {.r-fit-text}
+
+::: {.incremental}
+
+- Underlying data sources
+    - **WID:** Tax + NA + survey 
+    - **PIP:** survey only
+
+- Methodologies
+    - **WID:** Adjustments for misisng top in survey data, Scaling to national accounts aggregates
+    - **PIP:** Some ex-post harmonization of welfare concepts
+
+- Coverage
+    - Differences across the two datasets, impacting the time period that can be analysed
+
+:::
+
+:::
+
+## {.smaller}
+
+**Connecting up these 'worlds' is a low-hanging fruit...**
+
+::: {.r-fit-text}
+::: {.incremental}
+
+**For policy and public debate**
+
+- The data is better than ever
+- But the two 'worlds' of measurement makes the current data terrain *very* confusing
+    - Example: a summary of inequality trends [from the UN](https://www.un.org/en/un75/inequality-bridging-divide).
+
+:::
+:::
+
+## {.smaller .scrollable}
+
+**Connecting up these 'worlds' is a low-hanging fruit...**
+
+::: {.r-fit-text}
+::: {.incremental}
+
+**For academic research**
+
+- Existing studies tend to be based on only one 'world'
+    - Flagship reports of the two institutions [@chancel2021; @World_Bank2022-vr]
+    - Kanbur et al. [-@Kanbur2022-lk]
+    - Niño-Zarazúa et al. [-@Nino-Zarazua2017-tv]
+    - van Zanden et al. [-@Van_Zanden2014-zo]
+    - Ravallion[-@Ravallion2014-mk]
+
+- Two notable counter-examples:
+    - Anand and Segal [-@AnandSegal2015-nh]
+        - Adjusts survey data to match WTID top income shares (with fairly small effect)
+        - Estimtates up to 2005
+    - Lakner and Milanovic [-@laknerMilanovic2016]
+        - Adjusts survey data to match WTID top income shares (with fairly small effect)
+        - Only considers impact on overall global inequality
+        - Estimates up to 2008
+    - Neither focus on average within-country trends (Concepts 1* and 2* above)
+    - Coverage in the WID data has expanded subsequently
+:::
+:::
+
+# Connecting up the 'worlds'
+
+## {.smaller}
+### Connecting up the 'worlds'
 1. Line up what we can:
     - Coverage and time period
-    - Metrics
+        - 1995 vs 2015 (+/-3) – good coverage 
+        - 1980 vs 2018 (+/-5) – much worse coverage, more relevant period
 
-    - (With lower coverage) Income concept
-        - WID pre-tax
-        - LIS?
-        - Income vs consumption: GCIP
+    - Inequality indices
 
-    - Methods: Hacky top income adjustment to PIP data?
+    - Income concept 
+        - Very cursory glance [unlike @MorelliYonzan2022-zn; @CarranzaNolan2022-ko]
+
+## {.smaller}
+### Connecting up the 'worlds'
 
 2. Triangulate
-    - Flag what we *don't* know – Isolate what is causing the differences between 'take home' messages
-        - Particular methodological choices?
+    - Pinpoint what we *don't* know; what is behind the differences:
         - Particular countries?
-    - Gather together what we *do* know, irrespecive of may remaining differences
+        - Particular methodological differences?
+    - Gather together what we *do* know, irrespecive of the many remaining differences
+
+# Coverage
+
+## {.smaller}
+Coverage gets very good in both datasets from ~1990.
+
+```{python}
+
+df_coverage = pd.DataFrame()
+
+data_to_check = {
+    'Source': ['PIP', 'WID'],
+    'Metric': ['Gini', 'Top 1pc share – pretax'],
+    'df': [df_pip, df_wid]
+    }
+
+max_dist_from_ref = 3
+
+for yr in range(1960, 2023):
+
+    for i in [0,1]:
+
+        matched_data = closest_to_reference(
+            df=data_to_check['df'][i], 
+            reference_val = yr, 
+            max_dist_from_ref = max_dist_from_ref, 
+            reference_col = 'Year', 
+            group_by_col = 'Entity', 
+            value_col = data_to_check['Metric'][i], 
+            tie_break = 'above')
+
+        df_pop_ref_year = df_pop[df_pop['Year']==yr].drop(columns=['Year', 'population'])
+
+        matched_data = matched_data.merge(df_pop_ref_year, how='left')
+
+        coverage_yr = pd.DataFrame({
+            'Year': [yr],
+            'No. of countries': [matched_data['world_pop_share'].count()],
+            'Share of World population': [matched_data['world_pop_share'].sum()]
+        })
+
+        source = data_to_check['Source'][i]
+        metric = data_to_check['Metric'][i]
+        coverage_yr['Data'] = f'{source} – {metric}'
+
+        df_coverage = pd.concat([df_coverage, coverage_yr])
+
+```
+
+```{python}
+
+fig = px.line(df_coverage, x="Year", y="Share of World population", color='Data')
+
+fig.update_layout(
+    title = f'Population coverage – within {max_dist_from_ref} years',
+    legend_title="",
+    legend=dict(
+    orientation="h",
+    yanchor="bottom",
+    y=1.02,
+    xanchor="right",
+    x=1
+))
+
+fig.show()
+
+```
+
+## {.smaller}
+Coverage gets very good in both datasets from ~1990.
+
+```{python}
+
+fig = px.line(df_coverage, x="Year", y="No. of countries", color='Data')
+
+fig.update_layout(
+    title = f'No. of countries – within {max_dist_from_ref} years',
+    legend_title="",
+    legend=dict(
+    orientation="h",
+    yanchor="bottom",
+    y=1.02,
+    xanchor="right",
+    x=1
+))
+
+fig.show()
+
+```
+
+## {.smaller}
+We can increase 'coverage' in 1980 by widening the net.
+
+```{python}
+
+df_coverage = pd.DataFrame()
+
+data_to_check = {
+    'Source': ['PIP', 'WID'],
+    'Metric': ['Gini', 'Top 1pc share – pretax'],
+    'df': [df_pip, df_wid]
+    }
+
+max_dist_from_ref = 5
+
+for yr in range(1960, 2023):
+
+    for i in [0,1]:
+
+        matched_data = closest_to_reference(
+            df=data_to_check['df'][i], 
+            reference_val = yr, 
+            max_dist_from_ref = max_dist_from_ref, 
+            reference_col = 'Year', 
+            group_by_col = 'Entity', 
+            value_col = data_to_check['Metric'][i], 
+            tie_break = 'above')
+
+        df_pop_ref_year = df_pop[df_pop['Year']==yr].drop(columns=['Year', 'population'])
+
+        matched_data = matched_data.merge(df_pop_ref_year, how='left')
+
+        coverage_yr = pd.DataFrame({
+            'Year': [yr],
+            'No. of countries': [matched_data['world_pop_share'].count()],
+            'Share of World population': [matched_data['world_pop_share'].sum()]
+        })
+
+        source = data_to_check['Source'][i]
+        metric = data_to_check['Metric'][i]
+        coverage_yr['Data'] = f'{source} – {metric}'
+
+        df_coverage = pd.concat([df_coverage, coverage_yr])
+
+```
+
+```{python}
+
+fig = px.line(df_coverage, x="Year", y="Share of World population", color='Data')
+
+fig.update_layout(
+    title = f'Population coverage – within {max_dist_from_ref} years',
+    legend_title="",
+    legend=dict(
+    orientation="h",
+    yanchor="bottom",
+    y=1.02,
+    xanchor="right",
+    x=1
+))
+
+fig.show()
+
+```
+
+## {.smaller}
+We can increase 'coverage' in 1980 by widening the net.
+
+```{python}
+
+fig = px.line(df_coverage, x="Year", y="No. of countries", color='Data')
+
+fig.update_layout(
+    title = f'No. of countries – within {max_dist_from_ref} years',
+    legend_title="",
+    legend=dict(
+    orientation="h",
+    yanchor="bottom",
+    y=1.02,
+    xanchor="right",
+    x=1
+))
+
+fig.show()
+
+```
+
+# 1995 vs 2015
+
+## {.smaller}
+```{python}
+# Set-up to compare WID and PIP top 10% share, all countries
+reference_vals = [1995, 2015]
+max_dist_from_refs = 3
+all_metrics_requirement = False
+selected_vars = [
+        'WID: Top 10pc share – pretax',
+        'PIP: Top 10pc share'
+        ]
+
+print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))
+
+```
+
+```{python}
+
+# Table WID vs PIP top 10% share, all countries
+
+df_summary = change_summary_table(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )
+
+df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)
+
+dfi.export(df_summary,"images/tables/pres_table4.png")
+
+```
+![](images/tables/pres_table4.png)
+
+## {.smaller}
+```{python}
+# Switch to alow only countries with matched data from both PIP and WID for top 10% share
+
+all_metrics_requirement = True
+
+print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))
+
+```
+
+```{python}
+
+# Table WID vs PIP top 10% share, only matched countries
+
+df_summary = change_summary_table(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )
+
+df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)
+
+dfi.export(df_summary,"images/tables/pres_table5.png")
+
+```
+![](images/tables/pres_table5.png)
+
+::: {.incremental}
+
+- Country coverage drops (less so population-weighted)
+- Drop in coverage not uniform across regions and source (LAC and SS Africa in PIP)
+
+:::
+
+## {.smaller}
+
+```{python}
+print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))
+
+```
+![](images/tables/pres_table5.png)
+
+::: {.incremental}
+
+Key differences:
+
+- Inequality is far higher in WID than PIP
+- A very substantial rise in the pop-weighted average in WID.
+
+:::
+
+## {.smaller}
+
+```{python}
+print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))
+
+```
+![](images/tables/pres_table5.png)
+
+::: {.incremental}
+
+Key similarities:
+
+- Rising vs falling: approx. equal in both sources
+- Unweighted avg approx. stable in both (a fall in PIP)
+- Weighted average less benign trend than unweighted average
+
+:::
+
+##
+```{python}
+
+# Scatter WID (WID vs PIP top 10% share, only matched countries)
+
+plot_change_scatter(
+      plot_var = selected_vars[0],
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )[0]
+
+```
+
+##
+```{python}
+
+# Scatter WID (WID vs PIP top 10% share, only matched countries)
+
+plot_change_scatter(
+      plot_var = selected_vars[1],
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )[0]
+
+```
+
+## {.smaller .scrollable}
+```{python}
+
+# Scatter difference (WID vs PIP top 10% share, only matched countries)
+
+plot_change_scatter(
+      plot_var = 'diff',
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )[0]
+
+```
+
+```{python}
+df_change = plot_change_scatter(
+      plot_var = 'diff',
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )[1]
+
+df_change['change_diff'] = df_change[f'change {selected_vars[0]}'] - df_change[f'change {selected_vars[1]}']
+
+df_change.to_csv('df_change.csv')
+
+```
+::: {.incremental}
+
+- More countries below than above the line
+- For the *majority* of countries the gap is not huge (28/79 > 2 pt. difference; 20/79 > 4 pt. difference)
+- *Very* large differences in a small number of coutries – which happen to be very populous
+
+:::
+
+## Top 1 vs Top 10?
+
+## {.smaller}
+
+```{python}
+# Set-up to compare WID top 1% and top 10% share, all countries
+reference_vals = [1995, 2015]
+max_dist_from_refs = 3
+all_metrics_requirement = False
+selected_vars = [
+        'WID: Top 1pc share – pretax',
+        'WID: Top 10pc share – pretax'
+        ]
+
+print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))
+
+```
+
+```{python}
+
+# Table WID top 1% and top 10% share, all countries
+
+df_summary = change_summary_table(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )
+
+df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)
+
+dfi.export(df_summary,"images/tables/pres_table_top_1_top10.png")
+
+```
+![](images/tables/pres_table_top_1_top10.png)
+
+::: {.incremental}
+
+- Differences for some individual countries – LAC
+- But for other regions, and globally, the story over time is very similar.
+
+:::
+
+# 1980 vs 2018
+
+## {.smaller}
+```{python}
+
+reference_vals = [1980, 2018]
+max_dist_from_refs = 5
+all_metrics_requirement = False
+selected_vars = [
+        'WID: Top 10pc share – pretax',
+        'PIP: Top 10pc share'
+        ]
+
+print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))
+
+```
+
+```{python}
+
+df_summary = change_summary_table(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )
+
+df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)
+
+dfi.export(df_summary,"images/tables/pres_table8.png")
+
+```
+![](images/tables/pres_table8.png)
+
+::: {.incremental}
+
+- Even before aligning coverage, the country coverage drops substantially for both datasets - so too does population coverage
+- In WID a substantial rise in the unweighted average; none in PIP
+- A huge rise in weighted average in WID; small rise in PIP
+- PIP is the binding constraint on coverage: almost no EE & CA coverage; half the number of W. European countries.
+
+:::
+
+## {.smaller}
+```{python}
+
+reference_vals = [1980, 2018]
+max_dist_from_refs = 5
+all_metrics_requirement = True
+selected_vars = [
+        'WID: Top 10pc share – pretax',
+        'PIP: Top 10pc share'
+        ]
+
+print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))
+
+```
+
+```{python}
+
+df_summary = change_summary_table(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )
+
+df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)
+
+dfi.export(df_summary,"images/tables/pres_table9.png")
+
+```
+![](images/tables/pres_table9.png)
+
+::: {.incremental}
+
+-  Very few countries available in both datasets for this period (though population coverage falls less)
+- Aligning coverage does bring the datasets somewhat closer: 
+    - smaller rise in global unweighted average in WID
+    - more balanced split between rise and fall
+- But the stark difference remains: a huge rise in weighted average in WID
+
+:::
+
+## {.smaller .scrollable}
+
+```{python}
+
+plot_change_scatter(
+      plot_var = 'diff',
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )[0]
+
+```
+
+- For those countries with data from both PIP and WID, the majority tell a surprisingly similar story about change over time.
+- Only for a small number are there *very* large differences – this includes a number of populous countries: India, China, US, Mexico
+
+## Top 1 vs Top 10?
+
+## {.smaller}
+
+```{python}
+# Set-up to compare WID top 1% and top 10% share, all countries
+all_metrics_requirement = False
+selected_vars = [
+        'WID: Top 1pc share – pretax',
+        'WID: Top 10pc share – pretax'
+        ]
+
+print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))
+
+```
+
+```{python}
+
+# Table WID top 1% and top 10% share, all countries
+
+df_summary = change_summary_table(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )
+
+df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)
+
+dfi.export(df_summary,"images/tables/pres_table_top_1_top10_1980.png")
+
+```
+![](images/tables/pres_table_top_1_top10_1980.png)
+
+::: {.incremental}
+
+- Much more divergence visible than for 1995-2015 period – especially population-weighted
+- Average top 1% doubled, or nearly, in S & SE Asia, East Asia and W. Offshoots. EE & CA quadrupled.
+
+:::
+
+## {.smaller}
+
+```{python}
+# Set-up to compare WID top 1% and top 10% share, all countries
+all_metrics_requirement = True
+selected_vars = [
+        'WID: Top 1pc share – pretax',
+        'PIP: Top 10pc share'
+        ]
+
+print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))
+
+```
+
+```{python}
+
+# Table WID top 1% and PIP top 10% share, all countries
+
+df_summary = change_summary_table(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )
+
+df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)
+
+dfi.export(df_summary,"images/tables/pres_table_WID_top_1_PIP_top10_1980.png")
+
+```
+![](images/tables/pres_table_WID_top_1_PIP_top10_1980.png)
+
+::: {.incremental}
+
+- This remains true even after aligning coverage with PIP.
+
+:::
+
+# Welfare concept & methods
+
+## LIS pre- and post-tax data
+
+## {.smaller}
+
+```{python}
+
+reference_vals = [1980, 2018]
+max_dist_from_refs = 5
+all_metrics_requirement = True
+selected_vars = [
+        'LIS: Top 10pc share – percap_disposable',
+        'LIS: Top 10pc share – percap_market',
+        ]
+
+print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))
+
+```
+
+```{python}
+
+df_summary = change_summary_table(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )
+
+df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)
+
+dfi.export(df_summary,"images/tables/LIS_on_LIS.png")
+
+```
+![](images/tables/LIS_on_LIS.png)
+
+::: {.incremental}
+
+- Very low coverage
+- For W. Europe and W. Offshoots we do see some divergence between pre- and post- tax inequality.
+
+:::
+
+## WID pre- vs post-tax data
+
+## {.smaller}
+
+```{python}
+
+reference_vals = [1980, 2018]
+max_dist_from_refs = 5
+all_metrics_requirement = True
+selected_vars = [
+        'WID: Top 10pc share – pretax',
+        'WID: Top 10pc share – posttax_nat',
+        ]
+
+print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))
+
+```
+
+```{python}
+
+df_summary = change_summary_table(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )
+
+df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)
+
+dfi.export(df_summary,"images/tables/WID_on_WID.png")
+
+```
+![](images/tables/WID_on_WID.png)
+
+::: {.incremental}
+
+- Again very low coverage
+- For both EE & CA and W. Europe, quite a similar story in pre- and post-tax trends – albeit huge rises within EE & CA.
+
+:::
+
+## Individual countries
+
+##
+```{python}
+
+def plot_one_country_comparison(selected_entity, selected_vars):
+
+    selected_source_var_dict = prep_selected_vars(selected_vars)
+
+    df_long = pd.DataFrame()
+
+    for i in range(len(selected_source_var_dict['sources'])):
+
+        source = selected_source_var_dict["sources"][i]
+        var = selected_source_var_dict["var_names"][i]
+
+        df = source_dfs[source]
+
+        if source == "PIP":
+
+            df1 = df.loc[df['welfare_type'] == "income", ['Year', 'Entity', var]]\
+                .rename(columns = {var: 'value'})
+
+            df1['data'] = f'{source}: {var} (income)'
+
+            df_long = pd.concat([df_long, df1])
+
+            df2 = df.loc[df['welfare_type'] == "consumption", ['Year', 'Entity', var, 'welfare_type']]\
+                .rename(columns = {var: 'value'})
+
+            df2['data'] = f'{source}: {var} (consumption)'
+
+            df_long = pd.concat([df_long, df2])
+
+        else:
+
+            df = df[['Year', 'Entity', var]]\
+                .rename(columns = {var: 'value'})
+
+            df['data'] = f'{source}: {var}'
+
+            df_long = pd.concat([df_long, df])
+
+    df_plot = df_long[df_long['Entity']==selected_entity]
+
+    fig = px.line(df_plot, x="Year", y="value", color='data',
+    title = f'A range of series for {selected_entity}'
+    )
+
+    fig.update_layout(showlegend=True, legend=dict(
+    title_font_family='Lato',
+    font=dict(
+        size=8
+    )
+    ))
+    fig.show()
+
+```
+
+## {.smaller}
+
+```{python}
+selected_vars = [
+        'WID: Top 10pc share – pretax',
+        'WID: Top 10pc share – posttax_nat',
+        'PIP: Top 10pc share',
+        'LIS: Top 10pc share – percap_disposable',
+        'LIS: Top 10pc share – percap_market'
+        ]
+
+plot_one_country_comparison(selected_entity = "United States", selected_vars = selected_vars)
+```
+
+::: {.incremental}
+
+- In the US, similar trends across survey and WID 1975-1995
+- Some divergence thereafter between pre- and post-tax series - also in LIS survey data
+- Very strong divergence of WID series (both pre- and post-tax) 1995 onwards.
+
+:::
+
+## {.smaller .scrollable}
+
+```{python}
+
+plot_one_country_comparison(selected_entity = "Russia", selected_vars = selected_vars)
+```
+
+::: {.incremental}
+
+- In Russia, huge uncertainty concerning the extent and duration of rising inequality over transition.
+- Quality of the data subject to a lot of discussion [@Milanovic1998-sy; @Novokmet2018-px; @Atkinson-Micklewright1992-qr]
+- More agreement from mid-2000s
+
+:::
+
+## {.smaller}
+
+```{python}
+plot_one_country_comparison(selected_entity = "Mexico", selected_vars = selected_vars)
+```
+
+::: {.incremental}
+
+- In Mexico (and other LAC countries) WID is using a linear interpolation between decadal estimates prior to ~2000.
+
+- This relates to methods (tax data, top adjustments) rather than income concept.
+
+:::
+
+## {.smaller}
+
+```{python}
+plot_one_country_comparison(selected_entity = "Brazil", selected_vars = selected_vars)
+```
+
+::: {.incremental}
+
+- Similar in Brazil, except WID goes with the survey data after 2000.
+
+:::
+
+## {.smaller .scrollable}
+
+```{python}
+plot_one_country_comparison(selected_entity = "India", selected_vars = selected_vars)
+```
+
+::: {.incremental}
+
+- In India, only two survey available in LIS.
+- Some divergence between pre- and post- tax inequality in the LIS data, but tiny compared to the divergence of WID pre-tax. 
+- Note PIP data relates to *consumption*
+- A period of particular divergence between national accounts and survey data [@Deaton2005-oy; @Sandefur2022-lx]
+
+:::
+
+## {.smaller}
+
+```{python}
+plot_one_country_comparison(selected_entity = "China", selected_vars = selected_vars)
+```
+
+::: {.incremental}
+
+- In China, some divergence of pre- and post-tax nequality apparent in LIS survey data
+- Very condensed period of divergence in WID between 2000-2005
+- Again substantial divergence between national accounts and survey means over this time [@Deaton2005-oy]
+
+:::
+
+## {.smaller}
+
+```{python}
+plot_one_country_comparison(selected_entity = "Nigeria", selected_vars = selected_vars)
+```
+
+::: {.incremental}
+
+- Nigeria – WID seems not to account for 1996 survey in which inequality apparently peaked.
+- Breaks in comparability noted by the World Bank – see this [Our World in Data chart](https://ourworldindata.org/explorers/poverty-explorer?hideControls=false&Metric=Share+in+poverty&Poverty+line=%242.15+per+day%3A+International+Poverty+Line&Household+survey+data+type=Show+data+from+both+income+and+consumption+surveys&Show+breaks+between+less+comparable+surveys=true&country=~NGA).
+
+:::
+
+# Triangulating the key facts of global inequality
+Concept 1* and 2*
+
+## {.smaller}
+### Since 1995:
+
+::: {.r-fit-text}
+::: {.incremental}
+
+**Key differences:**
+
+- Strong rise in the population-weighted average level of inequality (Concept 2*) in WID that is absent in PIP.
+
+**Key similarities:**
+
+- For a majority of countries, trends look quite similar across PIP and WID.
+- Inequality in the average country was roughly flat, with roughly as many coutries with rising inequality as with falling inequality
+- Inequality was rising particularly strongly in some very populous countries, bringing up the trend in the weighted average
+- Different indices – even a focus on the *very* top – paint very similar pictures overall.
+
+:::
+
+:::
+
+## {.smaller}
+### Since 1980
+::: {.incremental}
+::: {.r-fit-text}
+
+**Key differences:**
+
+- Significant rise in unweighted average inequality (Concept 1*) in WID that is absent in PIP.
+- Huge rise in the population-weighted average level of inequality (Concept 2*) in WID that is much more muted in PIP.
+- Stronger differences looking at the *very* top
+
+**Key similarities**
+
+- Amongst those (few) countries with data in both PIP and WID, trends look quite similar for a majority of countries.
+- Inequality was rising particularly strongly in some very populous countries, bringing up the trend in the weighted average
+
+:::
+
+:::
+
+## {.smaller .scrollable}
+### Lessons (Concept 3)
+
+::: {.r-fit-text}
+
+::: {.incremental}
+
+- The global picture – when weighting by population – is greatly affected by a small number of populous countries. Trends diverge substantialy between PIP and WID for a number of these.
+
+- These larger discrepancies in trends relate to methodological choices, adjustment procedures etc. – the income concept plays a minor role at most.
+
+- The fact that these large differences are a (substantial) minority of cases suggests they may reflect idiosynctratic data quality or methodological issues specific to these countries – and not just a general issue of the 'the missing top'.
+    - Some examples of possibly idiosyncratic issues demonstrated above
+
+- Room for the two worlds be more proactively connecting up and explaining divergences – particularly in these cases where there are large discrepencies and large populations.
+
+- Analysis of global inequality trends – when weighting by population – should factor in this potential sensitivity to localised data issues
+    - Something not currently done in the literature on global inequality
+    - Future work!
+
+:::
+
+:::
+
+##
+
+# Appendix
+
+## Countries only found in one source - 1995 vs 2015
+
+```{python}
+
+reference_vals = [1995, 2015]
+max_dist_from_refs = 3
+all_metrics_requirement = False
+selected_vars = [
+        'WID: Top 10pc share – pretax',
+        'PIP: Top 10pc share'
+        ]
+
+```
+
+```{python}
+prepped_data = prep_data(
+        source_dfs = source_dfs,
+        df_pop = df_pop,
+        df_regions = df_regions,
+        selected_vars = selected_vars,
+        reference_vals = reference_vals,
+        max_dist_from_refs = max_dist_from_refs,
+        min_dist_between = 1,
+        all_metrics_requirement = all_metrics_requirement,
+        reference_col = "Year",
+        group_by_col = "Entity",
+        outlier_cut_off_upper = None,
+        tolerance_lookup =  tolerance_lookup
+        )
+
+prepped_data['source_count'] = prepped_data.groupby('Entity')['source_var'].transform('count')
+
+prepped_data = prepped_data[prepped_data['source_count']==1]
+```
+## 
+PIP only
+```{python}
+prepped_data.loc[prepped_data['source_var'] == selected_vars[1], ['Entity', 'region', f'value{reference_vals[0]}', f'value{reference_vals[1]}', f'Year{reference_vals[0]}', f'Year{reference_vals[1]}', 'pip_welfare', f'pop{reference_vals[1]}']].sort_values(['region', 'Entity'])
+
+```
+
+## 
+WID only
+```{python}
+prepped_data.loc[prepped_data['source_var'] == selected_vars[0], ['Entity', 'region', f'value{reference_vals[0]}', f'value{reference_vals[1]}', f'Year{reference_vals[0]}', f'Year{reference_vals[1]}', f'pop{reference_vals[1]}']].sort_values(['region', 'Entity'])
+
+```
+
+## Countries only found in one source - 1980 vs 2018
+
+```{python}
+
+reference_vals = [1980, 2018]
+max_dist_from_refs = 5
+all_metrics_requirement = False
+selected_vars = [
+        'WID: Top 10pc share – pretax',
+        'PIP: Top 10pc share'
+        ]
+
+```
+
+```{python}
+prepped_data = prep_data(
+        source_dfs = source_dfs,
+        df_pop = df_pop,
+        df_regions = df_regions,
+        selected_vars = selected_vars,
+        reference_vals = reference_vals,
+        max_dist_from_refs = max_dist_from_refs,
+        min_dist_between = 1,
+        all_metrics_requirement = all_metrics_requirement,
+        reference_col = "Year",
+        group_by_col = "Entity",
+        outlier_cut_off_upper = None,
+        tolerance_lookup =  tolerance_lookup
+        )
+
+prepped_data['source_count'] = prepped_data.groupby('Entity')['source_var'].transform('count')
+
+prepped_data = prepped_data[prepped_data['source_count']==1]
+```
+
+## {.scrollable} 
+PIP only
+```{python}
+prepped_data.loc[prepped_data['source_var'] == selected_vars[1], ['Entity', 'region', f'value{reference_vals[0]}', f'value{reference_vals[1]}', f'Year{reference_vals[0]}', f'Year{reference_vals[1]}', 'pip_welfare', f'pop{reference_vals[1]}']].sort_values(['region', 'Entity'])
+
+```
+
+## {.scrollable} 
+WID only
+```{python}
+prepped_data.loc[prepped_data['source_var'] == selected_vars[0], ['Entity', 'region', f'value{reference_vals[0]}', f'value{reference_vals[1]}', f'Year{reference_vals[0]}', f'Year{reference_vals[1]}', f'pop{reference_vals[1]}']].sort_values(['region', 'Entity'])
+
+```
+
+# Ex-slide graveyard
+
+## {.smaller}
+### Different 'worlds': *WID*
+```{python}
+
+reference_vals = [1990, 2018]
+max_dist_from_refs = 5
+all_metrics_requirement = False
+selected_vars = [
+        'WID: Top 1pc share – pretax',
+        'PIP: Gini'
+        ]
+
+```
+
+```{python}
+
+plot_change_scatter(
+      plot_var = 'WID: Top 1pc share – pretax',
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )[0]
+
+```
+
+## {.smaller}
+### Different 'worlds': *World Bank PIP*
+
+```{python}
+
+plot_change_scatter(
+      plot_var = 'PIP: Gini',
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )[0]
+
+```
+
+## {.smaller}
+### Different worlds: a comparison
+
+```{python}
+
+df_summary = change_summary_table(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )
+
+df_summary
+
+df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)
+
+dfi.export(df_summary,"images/tables/pres_table1.png")
+
+```
+
+![](images/tables/pres_table1.png)
+
+# Aligning inequality indices
+
+, not so sensitive to top [Osberg...]
+
+##
+```{python}
+
+reference_vals = [1990, 2015]
+max_dist_from_refs = 3
+all_metrics_requirement = False
+selected_vars = [
+        'PIP: Gini',
+        'PIP: Top 10pc share'
+        ]
+
+print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))
+
+```
+
+```{python}
+
+df_summary = change_summary_table(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )
+
+df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)
+
+dfi.export(df_summary,"images/tables/pres_table2.png")
+
+```
+![](images/tables/pres_table2.png)
+
+##
+Observations
+- All very similar
+- MENA is interesting.
+
+##
+Find countries where the measures moved in different directions... Plot them (the whole time series.)
+
+##
+```{python}
+
+reference_vals = [1990, 2015]
+max_dist_from_refs = 3
+all_metrics_requirement = False
+selected_vars = [
+        'WID: Gini – pretax',
+        'WID: Top 1pc share – pretax'
+        ]
+
+print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))
+
+```
+
+```{python}
+
+df_summary = change_summary_table(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )
+
+df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)
+
+dfi.export(df_summary,"images/tables/pres_table3.png")
+
+```
+![](images/tables/pres_table3.png)
+
+##
+Observations
+- All very similar
+- LAC – reverses trend among small number of countries
+
+##
+Find countries where the measures moved in different directions... Plot them (the whole time series.)
+
+## 
+General relationship – regression
+
+## {.smaller}
+###
+Take home message isn't all that dependent on the measure
+- Concern that Gini misses what's happening at the extremes.
+- In individual cases there can be important differences
+- But looking globally it's less of an issue.
+
+## Alternative spec - +/-5 years
+
+##
+```{python}
+
+reference_vals = [1995, 2018]
+max_dist_from_refs = 5
+all_metrics_requirement = False
+selected_vars = [
+        'WID: Top 10pc share – pretax',
+        'PIP: Top 10pc share'
+        ]
+
+print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))
+
+```
+
+```{python}
+
+df_summary = change_summary_table(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )
+
+df_summary.drop(["Other N. America", "Other Oceania"], axis=0, inplace=True)
+
+dfi.export(df_summary,"images/tables/pres_table6.png")
+
+```
+![](images/tables/pres_table6.png)
+
+##
+```{python}
+
+all_metrics_requirement = True
+
+print(make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement, selected_vars))
+
+```
+
+```{python}
+
+df_summary = change_summary_table(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup,
+      region_col = 'region'
+      )
+
+df_summary = df_summary.drop(["Other N. America", "Other Oceania"], axis=0)
+
+dfi.export(df_summary,"images/tables/pres_table7.png")
+
+```
+![](images/tables/pres_table7.png)
 

### 2023-02-22 PhD_pages/Paper_1/data_appendix/1_A_prepare_PIP.qmd
Paper 1: preparing presentation

--- 2023-02-10 PhD_pages/Paper_1_within_country_ineq/data_appendix/1_A_prepare_PIP.qmd
@@
 
-df_pip["Bottom_50_share"] = df_pip[sum_cols].sum(axis=1)
+df_pip["Bottom 50pc share"] = df_pip[sum_cols].sum(axis=1)
 
-df_pip["P90_P50_ratio"] = df_pip["decile9_thr"]/df_pip["median"]
+df_pip["P90:P50 ratio"] = df_pip["decile9_thr"]/df_pip["median"]
 
@@
 df_pip = df_pip.rename(columns={
-    'decile10_share':'Top_10_share',
+    'decile10_share':'Top 10pc share',
     'mld': 'MLD',
@@
   "Gini",
-  "Top_10_share",
-  "Bottom_50_share",
+  "Top 10pc share",
+  "Bottom 50pc share",
   "MLD",

### 2023-02-22 PhD_pages/Paper_1/data_appendix/1_B_prepare_WID.qmd
Paper 1: preparing presentation

--- 2023-02-10 PhD_pages/Paper_1_within_country_ineq/data_appendix/1_B_prepare_WID.qmd
@@
 These we append into a single file.
+
+```{python}
+# I should shift over to this file at some point
+# fp = 'https://raw.githubusercontent.com/owid/notebooks/main/BetterDataDocs/PabloArriagada/WID/data/raw/wid_indices_992j.csv'
+
+```
 
@@
 df_wid = df_wid.rename(columns={
-    'P90-P100 - share of the top 10%':'Top_10_share',
-    'P0-P50 - share of the bottom 50%':'Bottom_50_share',
-    'P99-P100 - share of the top 1%':'Top_1_share'
+    'P90-P100 - share of the top 10%':'Top 10pc share',
+    'P0-P50 - share of the bottom 50%':'Bottom 50pc share',
+    'P99-P100 - share of the top 1%':'Top 1pc share'
   })
@@
   "Gini",
-  "Top_10_share",
-  "Bottom_50_share",
+  'Top 1pc share',
+  "Top 10pc share",
+  "Bottom 50pc share",
   "inc_concept"
@@
 
-df_wid.columns = df_wid.columns.map(lambda x: '{}_{}'.format(x[0], x[1]))
+df_wid.columns = df_wid.columns.map(lambda x: '{} – {}'.format(x[0], x[1]))
 

### 2023-02-22 PhD_pages/Paper_1/paper1.qmd
Paper 1: Drafting presentation
Paper 1: preparing presentation

--- 2023-02-10 PhD_pages/Paper_1_within_country_ineq/paper1.qmd
@@
 format: html
+bibliography: references.bib
+execute:
+  echo: false
+  warning: false
 ---
@@
 ::: {.callout-note collapse="true" appearance="minimal"}
-### Load packages
+### Load packages and data
 
-```{r}
-library(tidyverse)
-library(knitr)
-library(data.table)
+```{python}
+
+from plotnine import *
+import plotly.express as px
+import pandas as pd
+import numpy as np
+from IPython.display import display, HTML, IFrame, Markdown
+import dataframe_image as dfi
+import imgkit
+
+from wrangling_and_plotting_functions import *
+
+pd.set_option('display.max_rows', None)
 
 ```
+
+```{python}
+
+fp = 'data_appendix/data/clean/pip.csv'
+
+df_pip = pd.read_csv(fp)
+
+fp = 'data_appendix/data/clean/wid.csv'
+
+df_wid = pd.read_csv(fp)
+
+fp = 'data_appendix/data/clean/lis.csv'
+
+df_lis = pd.read_csv(fp)
+
+fp = 'data_appendix/data/clean/region_mapping.csv'
+
+df_regions = pd.read_csv(fp)
+
+fp = 'data_appendix/data/clean/population.csv'
+
+df_pop = pd.read_csv(fp)
+
+fp = 'data_appendix/data/clean/region_population.csv'
+
+df_pop_regions = pd.read_csv(fp)
+
+# Gather data into dictionary
+source_dfs = {
+    "PIP": df_pip,
+    "WID": df_wid,
+    "LIS": df_lis
+    }
+
+# This specifies the +/- for the 'stable' category in the tables below
+tolerance_lookup =  {
+                'var_search_term': [
+                    'Gini',
+                    'Top 1pc share', 
+                    'Top 10pc share',
+                    'Bottom 50pc share'],
+                'var_tolerance': [
+                    1,
+                    1,
+                    1,
+                    1
+                ]
+            }
+
+```
+
+```{python}
+def make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement,selected_vars):
+    msg = f'{reference_vals[0]} vs {reference_vals[1]} (+/- {max_dist_from_refs} years)'
+
+    if len(selected_vars)>1:
+        if all_metrics_requirement:
+            msg = f'{msg} – Matched countries only'
+        else:
+            msg = f'{msg} – All available countries'
+
+    return(msg)
+
+```
+
 :::
@@
 What has happened to inequality within countries over the past generation?
+
+Many answers for individual countries, or for particular regions. Can Our understanding of inequality trends *globally*, however, remains fragmented unclear.
+
+For many it is a given . Many papers in the social sciences, or policy notes begin 'decades of rising inequality' – without to specify where. Implying ubiquitous.
+
+Does not reflect the conflicting data.
+
+This stands in contrast with the data in PIP – an important data source. Relied on for SDGs.
+
+Confusiong terrain: e.g. where for SDGs they use PIP; but HDR they use WID.
+
+Our view remains highly fragmented. Different global datasets offer starkly differing or even contradictory views
+
+Despite the wealth of data available, our view of inequality around the world remains fragmented and unclear.
+
+These are questions that we need answers to. Winners and loosers of globalization. Policy to tackle inequality. SDGs. Philanthropy.
+
+That's what this paper is going to do. (Outline structure).
 

### 2023-02-22 presentations.qmd
Paper 1: preparing presentation

@@
 20/01/2023
+Department for Social Policy, University of Oxford
 
@@
 [Slides](other_presentations/DSPI_CSP_masters_Jan2023/DSPI_CSP_masters_Jan2023.qmd)
+
+### Triangulating a global view on inequality from secondary data sources
+16/02/2023
+Institute for New Economic Thinking, University of Oxford
+
+:::{.callout collapse="true"}
+## Abstract
+
+Many of the key questions and debates of our time necessitate a global view of economic inequality.
+Thanks to the development and continued expansion of key global databases of harmonised distributional data, we are in a better position than ever to take such a global view.
+
+And yet, on a range of basic questions concerning the nature and evolution of inequality around the world, there is surprisingly little consensus or clarity.
+Available global datasets offer starkly differing or even contradictory views.
+The variety of methods, concepts, and sources adopted in different global databases makes it difficult to interpret and synthesise the insights they offer.
+Despite the wealth of data available, our view of inequality around the world remains highly fragmented.
+
+To help address this, I line up results from three key global datasets on inequality: the World Inequality Database, the World Bank’s Poverty and Inequality Platform and the Luxembourg Income Study.
+In doing so, I try to triangulate a more cohesive and comprehensive global view of inequality – drawing together those stylised facts that are robust to choices of data source and methodology; highlighting those that are contingent on such choices; and helping to unpick the reasons for such discrepancies.
+
+:::
+
+[Interactive slides](PhD_pages/Paper_1/Presentation_INET_16_feb_inequality_around_the_world.qmd)

### 2023-02-22 PhD_pages/Paper_1/wrangling_and_plotting_functions.py

@@ -1,0 +1,891 @@
+
+#%%
+from plotnine import *
+import plotly.express as px
+import pandas as pd
+import numpy as np
+import plotly.graph_objects as go
+import math
+
+
+#  ---------------------------------
+#  ---------- SECTION 1: Data prep functions -------------
+#  ----------------------------------
+
+#%%
+#  Function to get matching observation closest to ref year
+def closest_to_reference(
+    df, 
+    reference_val, 
+    max_dist_from_ref, 
+    reference_col, 
+    group_by_col, 
+    value_col, 
+    tie_break):
+  
+  df = df.loc[:, [reference_col, group_by_col, value_col]]
+
+  # Drop NAs
+  df = df.dropna()
+
+  # Calculate absolute distance from reference value
+  df['ref_diff'] = abs(df[reference_col] - reference_val)
+
+  # Drop any rows with a distance beyond threshold
+  if not pd.isna(max_dist_from_ref):
+    df = df.loc[df['ref_diff'] <= max_dist_from_ref]
+
+  # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
+  df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)
+
+  # Settle tie-breaks
+  if tie_break == 'below':
+    df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)
+    
+  elif tie_break == 'above':
+    df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)
+
+  df = df.drop('ref_diff', axis=1)
+
+  df = df\
+    .rename(columns={value_col: "value"})
+
+
+  return df
+
+#%%
+
+# Generate matches for two reference years and merge
+def merge_two_ref_matches(
+    df, 
+    reference_vals, 
+    max_dist_from_refs, 
+    min_dist_between, 
+    reference_col, 
+    group_by_col, 
+    value_col):
+
+# Make sure the pair of reference values are in ascending order
+  reference_vals.sort()
+
+# Maximise distance between two refs by settling tie-breaks below the lowest ref and above the highest ref 
+
+# Find matches for lower reference value
+  lower_ref_matches = closest_to_reference(df, reference_vals[0], max_dist_from_refs, reference_col, group_by_col, value_col, 'below')
+
+# Find matches for higher reference value
[…]

### 2023-03-07 PhD_pages/Paper_1/Presentation_INET_16_feb_inequality_around_the_world.qmd
Drafting paper 1

@@
         theme: [default, revealjs_joe_theme.scss] 
-bibliography: references.bib
+bibliography: references_presentation.bib
 ---

### 2023-03-07 PhD_pages/Paper_1/paper1.qmd
Drafting paper 1

@@
 
-# Outline
+## Why do we need a global view on inequality?
 
-- We need a global view on inequality
+## What do I mean by global view on inequality?
 
-- Though there is much data, we don't have a clear global view.    
-  [UN summary of inequality trends](https://www.un.org/en/un75/inequality-bridging-divide)
+- Concept 1 and 2, but not 3. Companion paper.
+- Relative not absolute.
 
-- A global view is hampered by many factors:
-    - In research and commentary, there is a disportionate focus high-income countries, in particular the US
+## Global data on inequality
 
-    - The extent and quality of available data
-        - Coverage
-        - Quality and comarabilty of data
-        - Some trade off between the two is inevitable
+### Global databases
 
-    - **The major global inequality databases occupy different 'worlds' of measurement**
+A global view faces the hard limitation of the underlying data. This has two sides: firstly is there any data? Secondly the comparability of the data, across countries and over time.
 
-    It is the latter that I will be addressing. We are in a relatively rich world in terms of inequality data – thanks in particlar to efforts of harmonized databases.
+Much improved – *harmonized* secondary databases. Three of note.
 
-    But we lose much of the value – different how disconnected
+- PIP
 
-     – that the rich doesn't translate to as  understanding
+- WID
 
-- Two worlds of inequality measurement
+- LIS
 
-# Introduction
+### Two worlds of inequality measurement
 
-- A range of motives for taking a global view: SDGs; globalization – trade; ....
+#### Connecting up the worlds is valuable
 
-- We don't have a clear global view.    
-  [UN summary of inequality trends](https://www.un.org/en/un75/inequality-bridging-divide)
+- for Policy
+- for academic research
 
-- Why is it that we don't have a clear view?
+## Triangulating a global view
 
-    1) Most academic research and commentary has focussed on high income countries (refs), especially the US.
-There are many reasons why we should not expect trends in inequality in these countries to reflect broader global trends.
+### Since 1990
 
-    2) Hampered by the available data
-        - **Coverage** – for instance no surveys.
-        - **Quality and comarabilty of data** – Some secondary databases with global coverage mixes very heterogenous datapoints (WIID) (Find Atkinson and Brandolini ref from Handbook...).
-Those with a greater degree of harmonization tend to lack global coverage (LIS, EU-SILC, SEDLAC, until recently PIP)
-        - **Some trade off between the two is inevitable** – There is no global survey of incomes.
-Countries run surveys based on their own needs and capacities.
-As such they differ.
-Saying something global means accepting a degree of heterogenity in the data.
-Similarly, definitions of income for tax purposes vary, impacting the comparability of tax data.
-
-    3) Contradictory global views are presented by different global databases.
-        - The data situation has improved drastically: thanks to harmonized databases with global coverage. In particular:
-            - PIP – recently expanded to include high income countries
-            - WID – Transition from WTID, and the DINA methodology, global coverage.
-            - LIS: some data from middle and low income countries.
-        - But the 'take home' messages from the institutions producing these databases offer very contradictory views.
-            - "In recent decades, income inequality 
-has increased in nearly all countries, 
-but at different speeds, suggesting 
-that institutions and policies matter in shaping inequality." (WIR 2018)
-            - "[Across a range of approaches] average  inequality  within  countries  increased   in  the  1990s,  but  has  been  falling  in  recent years according to most specifications." (World Bank, 2016)
-
-        - You can see this in the latest data from these two sources, using the metric typically favoured by each: Top 1% from WID vs Gini from PIP.
-
-        - These different messages are based on a mix of many different methodological differences:
-            - different metrics, sensitive to different parts of the distribution; 
-            - Different welfare concepts; 
-            - Different underlying data sources and adjusmtents to the top (or lack thereof);
-            - Different coverage or time periods being analysed.
-
-# Literature review – Global studies of inequality
-
-- Atkinson and Brandolini (2010) argue that global inequality analyses in particular need to consider both absol ute  and relative differences.
-
-# Data review Why is our understanding of inequality around the world currently limited
-
-- **A global view is hampered by the available data**.
-
-- **But databases have improved substantially recently**
-Thanks to improvements and expansion of key global harmonized databases, we are now in a much stronger position to take a global view on inequality within countries.
-
-There are range of databases and a full review is out of the scope of this paper. Focus on the three databases of harmonized distributional data providing global or near-global coverage.
-
-- **PIP** 
-    - Survey data
-    - Mix of disposable income and consumption data (definitions vary)
-    - Per capita
-    – Coverage: now includes high income countries (Add a box/visualization with coverage). Global coverage from 1990 onwards.
-    - A range of inequality indices provided 'as standard', but the whole distribution may be queried via API.
-Either microdata or, in a minority of cases, a parametric distribution fitted to 'grouped data' (e.g.
-income deciles)
-
-- **WID** 
-    – Survey data, tax data, National accounts
-    - Coverage much greater for pre-tax national income. But post-tax concepts also available for a subset of observations.
-    - Greatest coverage for individuals – equal split where underlying data was available for couples.
-    - Coverage 
-    - Estimates of the whole distribution Generalized Pareto Interpolation. Though longer-run coverage is higher for pretax top 1% share (reliance on tax data)
-
-- **LIS**
-    - Offers less coverage, but is notable for providing access to microdata.
-
-Different headlines - WID suggests general rise (find a quote). PIP much more mixed... even a fall.
-
-Which world we live in matters for how we think we should tackle inequality.
-
-A secular, general rise - suggesting global forces, perhaps something intrinsic to capitalism (Piketty, 2014).
-
-A mixed picture of inequality around the world suggests reflect specific circumstances, with accordingly much more scope for national policy to influence.
-
-Both are reasonable methods and have strengths and weaknesses.
-
-I want to:
-
-a\) Triangulate the key stylized facts. Which observations are robust to across available datasets with their different methods, and which are contingent?
-
-b\) Untangle which methods, definitions affect the different conclusions arrived at.
-
-# Methods and data
-
-## Data sources
-
-### PIP
-
-### WID
-
-### LIS
-
-### Coverage
-
-Give an overview of coverage in each dataset.
-
-## Comparing databases
+### Since 1980
 

### 2023-04-01 PhD_pages/Paper_1/Presentation_INET_16_feb_inequality_around_the_world.qmd
Paper 1 – drafting and amending bib

--- 2023-02-22 PhD_pages/Paper_1/Presentation_INET_16_feb_inequality_around_the_world.qmd
@@
         theme: [default, revealjs_joe_theme.scss] 
-bibliography: references.bib
+bibliography: references_old.bib
 ---

### 2023-04-01 PhD_pages/Paper_1/paper1.qmd
Paper 1 – drafting and amending bib

@@
 format: html
-bibliography: references.bib
+bibliography: references_07_03_2023.bib
 execute:
@@
 
-What has happened to inequality within countries over the past generation?
+READING LIST:
+Bourguignon (2015)
+The share of wages in total GDP declined in a majority of countries
+(91 out of 133 with data) from 1995 to 2014 (ILO, 2016).
 
-Many answers for individual countries, or for particular regions. Can Our understanding of inequality trends *globally*, however, remains fragmented unclear.
+ Other authors
+question the pre-eminence of globalization as the main driver of global inequality, noting
+that inequality trends differ across countries at similar levels of development and that
+are equally exposed to trade (Ravallion, 2018; Corlett, 2016). Ravallion (2018) stresses
+the vital role that natio
 
-For many it is a given . Many papers in the social sciences, or policy notes begin 'decades of rising inequality' – without to specify where. Implying ubiquitous.
+Key point:  the share of income going to both the top 1 per cent and the bottom 50 per
+cent has increased in many countries in recent decades.
 
-Does not reflect the conflicting data.
+# Introduction
 
-This stands in contrast with the data in PIP – an important data source. Relied on for SDGs.
+"Inequality defines our time" – the UN Secretary-General.
+(https://www.nelsonmandela.org/news/entry/annual-lecture-2020-secretary-general-guterress-full-speech)Increasing concern for inequality an increase i, researchers and institutions have responded by .
+In particular harmonized databases of comparable , addressing concerns (Atkinson).
+Both the demand to monitor inequality globally, and our ability to do so, is greater than ever.
 
-Confusiong terrain: e.g. where for SDGs they use PIP; but HDR they use WID.
+the UN (SDGs), the World Bank and the IMF – and WID.
 
-Our view remains highly fragmented. Different global datasets offer starkly differing or even contradictory views
+Yet the outcome is a mess.
+Contradictory messages emerge concerning the most basic descriptions from the two databases in global coverage – the *World Inequality Database*, and the World Bank's *Poverty and Inequality Platform*.
 
-Despite the wealth of data available, our view of inequality around the world remains fragmented and unclear.
+For example, different publications and studies yield starkly differing conclusions as to how widespread a phenomenon is rising inequality within countries across the world.
+The flagship publications from  view from *World Inequality Lab* to be taken do we live in a world in which inequality is rising in 'nearly everywhere' [@alvaredo_world_2018; @chancel_world_2021].[^WID_quotes] PIP, much more mixed.
+This is reflected in estimates of global inequality .
 
-These are questions that we need answers to. Winners and loosers of globalization. Policy to tackle inequality. SDGs. Philanthropy.
+Decomposition – wthin country inequality now outweighs between country differences.
 
-That's what this paper is going to do. (Outline structure).
+[^WID_quotes]:
+    > Income and wealth inequalities have been on the rise nearly everywhere since the 1980s, following a series of deregulation and liberalization programs which took different forms in different countries.
+[@chancel_world_2021, pp.
+xx]
 
-## Why do we need a global view on inequality?
+    > In recent decades, income inequality has increased in nearly all countries, but at different speeds... [@alvaredo_world_2018, pp. xx]
 
-## What do I mean by global view on inequality?
+Thus despite the great data, there remains fundamental ambiguity surrounding which of two possible worlds we live in.
+In one world, rising inequality within countries has been an almost universal phenomenon, and has now come to dominate income differences between countries.
+In another, within-country inequality trends in recent decades have been mixed, and the differences in average incomes between countries remain the lion's share of the globally.
 
-- Concept 1 and 2, but not 3. Companion paper.
-- Relative not absolute.
+Which of these worlds we live in matters hugely – both in our conception, and of where to place efforts to combat inequality in the future.
+
+The first world suggests the winners from globalization and economic growth overwhelmingly and almost universally, have been a global elite.
+It suggestest global factors – such as capital accumulation (Piketty), over and above national policy  How to tackle inequality glboal level.
+
+The second world and the importance of access to both produced and human capital , national policy (Ravallion, 2018; Corlett, 2016).
+
+Many methodolodical choices concerning underlying data sources, the constraints on temporal coverage this gives rise to,  adjust, conception of inequality.
+These to some extent come bundled together when making the choice between PIP and WID.
+That is to say, the empirical question of 'world' of inequality we live in depends on which 'world' of inequality *measurement* we adopt.
+
+Whilst recommending particular aspects, for particular countries, the choice put before us requires wholesale adoption of a bundle whose relative strengths and weaknesses when applied globally is not well understood.
+
+What we can do is triangulate
+
+In this paper, I focus on trends in within-country inequality in recent decades.(In a companion paper, I look at the global distribution.) I line up what I can.
+I triangulate what we know – .
+And Identify what we don't know – the remaining differences – and try to pinpoint where the differences lie.
+
+I do so with respect to two separate time periods, chosen to tradeoffs with respect to global coverage.
+The first period – from the fall of the Berlin wall.
+Driven primarily to the steep drop off in availability before this in household survey data collected in PIP.
+
+This maximises global coverage – 100 countries covering roughly 90% of the world poplation.
+The downside is that this period misses out – transition, liberalisation and structural adjustment.
+Question begging, by avoiding specifically the period where inequality thought to rise.
+
+To address this, I analyse a second, longer period extending back to roughly 1980.
+
+(SUMMAR OF FINDINGS)
+
+## Data
+
+### PIP
+
+### WID
+
+(Coverage chart - overlayed 3/5 year)
+
+### LIS
+
+Will be used for income concept.
+i.e.
+it represents the PIP world – indeed LIS is the source for PIP for a number of high income countries.
+But the access to microdata allows us to ask more questions for a subset of countries.
+
+## Methods
+
+### Concepts of global inequality
+
+Adapted three concepts.
+
+### Aligning coverage
+
+## Comparing inequality trends
+
+The focus of this paper is on trends. But worth pointing out levels. Inequality is much higher in WID than in PIP.
+
+Again this relates to many things – pretax vs disposable income/consumption.
+Later we will try lining up income concepts more closely.
+But for global coverage we have to pull together these 'worlds' that combine many differences.
+
+What we're trying to do with triangulation is see whether they show a qualitively similar picture of whether, and how rapidly, inequality was rising or falling.
+
+But what would 'similar' be for Gini and top 1. Not a 1 point move.
+
+The main way of operationalising 'similar' is to look at Relative change: a percentage, rather than percentage point. We can put some threshold and see if there were rises or falls.
+
+No reason to expect this given the properties of the two measures.
+If we take lognormality we see...
+(MAke a chart in appendix - where we plot Gini against top 10 and top 1% assuming lognormality, and then add actual data for a particular country.) How unlognormal the top is would be another way to judge similarity.
+Rather than doing this, we bridge by aligning the indices.
+
+- We also do absolute change: a percentage point. This is more intuitive. But somewhat arbitrary – differences in levels affect assessment of differences in trends.
+
+- A third way is residuals from a regression of the two measures
+    - This we will use to see countries that are particular outliers.
 

### 2023-04-01 PhD_pages/Paper_1/paper1.qmd
Draftind Paper 1

@@
 
+In 2021, while the very richest individuals buy trips to space, roughly three billion people are unable to afford a healthy diet.
+Even  rich countries secured COVID vaccines for their entire population, low-income had barely begun.
+COVID vaccination rates  at all, the richest countries had already administered more than one dose for every citizen.
+
+Such
+
+"Inequality defines our time," states the UN Secretary-General.^[United Nations Secretary-General António Guterres made the comment in the 18th Nelson Mandela Annual Lecture in July 2020.
+See https://www.nelsonmandela.org/news/entry/annual-lecture-2020-secretary-general-guterress-full-speech.]
+
+Just as the demand for data to monitor inequality globally is greater than ever, so is its supply.
+
+In part, this increased concern is a reflection of events - of the severe and uneven impacts of the 2007-8 Global Financial Crisis, the COVID-19 pandemic and recent sharp rises in the cost of living.
+
 # Introduction
 
-"Inequality defines our time" – the UN Secretary-General.
-(https://www.nelsonmandela.org/news/entry/annual-lecture-2020-secretary-general-guterress-full-speech)Increasing concern for inequality an increase i, researchers and institutions have responded by .
-In particular harmonized databases of comparable , addressing concerns (Atkinson).
-Both the demand to monitor inequality globally, and our ability to do so, is greater than ever.
+Economic inequalities around the world are enormous. As the very richest individuals buy trips to space, roughly three billion people are unable to afford a healthy diet.
 
-the UN (SDGs), the World Bank and the IMF – and WID.
+Such stark disparities have gained more attention in recent decades, among policimakers, academics and the public alike.
+"Inequality defines our time," according to the UN Secretary-General.^[United Nations Secretary-General António Guterres made the comment in the 18th Nelson Mandela Annual Lecture in July 2020.
+See https://www.nelsonmandela.org/news/entry/annual-lecture-2020-secretary-general-guterress-full-speech.]
 
-Yet the outcome is a mess.
-Contradictory messages emerge concerning the most basic descriptions from the two databases in global coverage – the *World Inequality Database*, and the World Bank's *Poverty and Inequality Platform*.
+In part at least, this increased concern reflects an increased availability of distributional data.
+New and expanded sources of data have shed new light on extent of inequality around the world and how it is changing.
+This has enabled concerns about high or rising inequality to be expressed in new ways, formulated more precisely and subsequently monitored.
+The supply and demand for global data on inequality have been mutually reinforcing.
 
-For example, different publications and studies yield starkly differing conclusions as to how widespread a phenomenon is rising inequality within countries across the world.
-The flagship publications from  view from *World Inequality Lab* to be taken do we live in a world in which inequality is rising in 'nearly everywhere' [@alvaredo_world_2018; @chancel_world_2021].[^WID_quotes] PIP, much more mixed.
-This is reflected in estimates of global inequality .
+Two sources of data have been particularly impactful.
 
-Decomposition – wthin country inequality now outweighs between country differences.
+One is the collection of national household survey data brought together by the World Bank, published online through its *Poverty and Inequality Platform*.
+This is produced as part of the World Bank's remit to monitor global poverty and shared prosperity.
+Most extensive harmonized .
+Its impact lies in its extensive use also by the UN in setting and monitoring many of the Sustainable Development Goals.
+
+Another especially important source is the World Inequality Database.
+
+ Capital in the 21st C -> WID. "The 1 %".
+
+Both datasets have continued to improve adn expand.
+In the case of PIP, including high income countries, following recommendations of the Atkinson commission.
+In the case of WID – transition from earlier, using with more minimal requirements in terms of source data.
+
+In terms of data availability we are thus in a better position than ever to take a global viw on inequlity. Yet amidst apparent wealth of data, limited understanding of very basics.
+
+Contradictory messages emerge concerning the most basic descriptions from the two databases.
+
+The view from *World Inequality Database* and its associated global report is that inequality is rising in 'nearly everywhere' [@alvaredo_world_2018; @chancel_world_2021].[^WID_quotes] The World Bank data and its biannual report offers a much more mixed picture – in which inequality is rising and falling in roughly equal numbers of countries.
 
@@
 
+This stark difference is reflected in decompositions of global inequality into between- and within-country components. (Explain)
+
 Thus despite the great data, there remains fundamental ambiguity surrounding which of two possible worlds we live in.
@@
 The first world suggests the winners from globalization and economic growth overwhelmingly and almost universally, have been a global elite.
-It suggestest global factors – such as capital accumulation (Piketty), over and above national policy  How to tackle inequality glboal level.
+It situates inquality within countries, but emphasises global factors – including fundamental laws of capitalism with regard to capital accumulation (Piketty) – and the need for international coordination over and above national policy.
 
 The second world and the importance of access to both produced and human capital , national policy (Ravallion, 2018; Corlett, 2016).
+
+### Unpicking the two worlds
 

### 2023-04-05 PhD_pages/Paper_1/Presentation_INET_16_feb_inequality_around_the_world.qmd
Paper 1: drafting and testing out workflow for tables

--- 2023-02-22 PhD_pages/Paper_1/Presentation_INET_16_feb_inequality_around_the_world.qmd
@@
         theme: [default, revealjs_joe_theme.scss] 
-bibliography: references.bib
+bibliography: references_old.bib
+execute:
+  eval: false
 ---

### 2023-04-05 PhD_pages/Paper_1/data_appendix/1_H_figures_and_tables.qmd
Paper 1: drafting and testing out workflow for tables

---
title: "Appendix 1.H: Preparing figures and tables used in the paper"
format: html
execute:
  echo: false
  warning: false
---

## Set-up

Load packages and functions.
```{python}

from plotnine import *
import plotly.express as px
import pandas as pd
import numpy as np

import dataframe_image as dfi

from G_functions_for_figures_and_tables import *

pd.set_option('display.max_rows', None)

```

Read in data.
```{python}

fp = 'data/clean/pip.csv'

df_pip = pd.read_csv(fp)

fp = 'data/clean/wid.csv'

df_wid = pd.read_csv(fp)

fp = 'data/clean/lis.csv'

df_lis = pd.read_csv(fp)

fp = 'data/clean/region_mapping.csv'

df_regions = pd.read_csv(fp)

fp = 'data/clean/population.csv'

df_pop = pd.read_csv(fp)

fp = 'data/clean/region_population.csv'

df_pop_regions = pd.read_csv(fp)

```

Gather the inequality data into a dictionary – this is the format assumed by the functions that will later be applied.
```{python}
# Gather data into dictionary
source_dfs = {
    "PIP": df_pip,
    "WID": df_wid,
    "LIS": df_lis
    }

```

Specify a mapping between inequality variables and the 'tolerance' that will be used to catagorise whether the metric has risen or fallen substantially (e.g. +/- 1point change). The mapping is achieved by a string match - i.e. 'Gini' within 'var_search_term' will map the correspondending tolerance to all columns whose name includes the string 'Gini'.

```{python}
tolerance_lookup =  {
                'var_search_term': [
                    'Gini',
                    'Top 1pc share', 
                    'Top 10pc share',
                    'Bottom 50pc share'],
                'var_tolerance': [
                    1,
                    1,
                    1,
                    1
                ]
            }

```

This builds astring that gets added to the tables identifying what paremeters were set when the table was built.
```{python}
def make_selection_msg(reference_vals,max_dist_from_refs,all_metrics_requirement,selected_vars):
    msg = f'{reference_vals[0]} vs {reference_vals[1]} (+/- {max_dist_from_refs} years)'

    if len(selected_vars)>1:
        if all_metrics_requirement:
            msg = f'{msg} – Matched countries only'
        else:
            msg = f'{msg} – All available countries'

    return(msg)

```

## Tables

#### Table 1

This figure compares relative change in the 'typical' metrics for WID and PIP: Top 1% share (pretax) and Gini
```{python}
# Set-up to compare WID and PIP top 10% share, all countries
reference_vals = [1995, 2015]
max_dist_from_refs = 3
all_metrics_requirement = False
selected_vars = [
        'WID: Top 1pc share – pretax',
        'PIP: Gini'
        ]

all_metrics_requirement = True

```

```{python}

df_summary = change_summary_table(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_pop_regions = df_pop_regions,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )

df_summary['PIP: Gini']

dfi.export(df_summary,"table_test.png")

```
![](table_test.png)

#### Table 2

#### Table 3

## Figures

#### Figure 1

This figure compares relative change in the 'typical' metrics for WID and PIP: Top 1% share (pretax) and Gini
```{python}
# Set-up to compare WID and PIP top 10% share, all countries
reference_vals = [1995, 2015]
max_dist_from_refs = 3
all_metrics_requirement = False
selected_vars = [
        'WID: Top 1pc share – pretax',
        'PIP: Gini'
        ]

all_metrics_requirement = True

```

```{python}

# Scatter WID (WID vs PIP top 10% share, only matched countries)

plot_change_scatter(
      plot_var = 'relative_diff',
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_regions = df_regions,
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      all_metrics_requirement = all_metrics_requirement,
      reference_col = "Year",
      group_by_col = "Entity",
      outlier_cut_off_upper = None,
      tolerance_lookup =  tolerance_lookup,
      region_col = 'region'
      )[0]

```

#### Figure 2

#### Figure 3

### 2023-04-05 PhD_pages/Paper_1/paper1.qmd
Paper 1: drafting and testing out workflow for tables

@@
 format: html
-bibliography: references_07_03_2023.bib
+bibliography: references_export_2023-04-04.bib
 execute:
@@
 
-In part at least, this increased concern reflects an increased availability of distributional data.
+In part at least, this increased concern reflects an increase in the *availability* of distributional data.
 New and expanded sources of data have shed new light on extent of inequality around the world and how it is changing.
@@
 
-Two sources of data have been particularly impactful.
+Two sources of data have been particularly impactful in this regard: the World Bank's harmonized collection of household survey data, published online through its [*Poverty and Inequality Platform (PIP)*](https://pip.worldbank.org/); and the collection of inequality estimates – most notably top income shares – brought together at the [*World Inequality Database (WID)*](https://wid.world/).^[EDIT: PIP used to be Povcal.
+It draws upon the work of various regional teams.
+WID used to be WTID.]
 
-One is the collection of national household survey data brought together by the World Bank, published online through its *Poverty and Inequality Platform*.
-This is produced as part of the World Bank's remit to monitor global poverty and shared prosperity.
-Most extensive harmonized .
-Its impact lies in its extensive use also by the UN in setting and monitoring many of the Sustainable Development Goals.
+These two sources are the only regularly-updated databases providing global coverage on a relatively consistent basis.^[EDIT: Both datasets suffer important limitations with respect to comparability acrosss time and location.
+These will be discussed in the data section below.
+Both datasets, however, take significant steps to harmonize, *ex-post*, the available underlying data.
+In this respect they can be distinguished from databases such as WIID and SWIID [@atkinson_promise_2001]] Both have played substantial agenda-setting roles in academic research, policy and public debate.
 
-Another especially important source is the World Inequality Database.
+Given its extensive coverage of low- and middle-income countries, the World Bank data occupies a central place in international development policy and research.
+Most notably, the data is relied on by the UN to set and monitor progress towards many of the targets included within Sustainable Development Goals.
+It has also to date underpinned the majority of academic studies on inequality that take a global view.^[[Including @Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk] TODO: Check and more refs].
 
- Capital in the 21st C -> WID. "The 1 %".
+The *WID* database emerged as institutional focal-point to host and coordinate inequality estimates produced as part of the burgeoning literature on top incomes [@atkinson_top_2014; @piketty_capital_2014; @alvaredo_top_2013].
+This development of this data has opened new avenues for the study of inequality right across the social sciences (TODO: example refs), underpinning a significant re-evaluation of its causes and consequences [@alvaredo_top_2013].
+It has also surely had a substantial impact on the public's conception of inequality, contributing to the momentum generated by the Occupy movement and its framing of inequality in terms of 'the 1%' vs the '99%''.
 
-Both datasets have continued to improve adn expand.
-In the case of PIP, including high income countries, following recommendations of the Atkinson commission.
-In the case of WID – transition from earlier, using with more minimal requirements in terms of source data.
+#### Two worlds of inequality measurement
 
-In terms of data availability we are thus in a better position than ever to take a global viw on inequlity. Yet amidst apparent wealth of data, limited understanding of very basics.
+The two . They in some sense represent *two different worlds* of inequality measurement.
 
-Contradictory messages emerge concerning the most basic descriptions from the two databases.
+In terms of methods, source, conception of inequality and research questions.
+"The 1 %" – pretax, with many assumptions needed to get to national income.
+These come as a bundle.
+Although both institutions continue to take many steps to provide transparency and reproducibility to their estimates, they remain a black box.
 
-The view from *World Inequality Database* and its associated global report is that inequality is rising in 'nearly everywhere' [@alvaredo_world_2018; @chancel_world_2021].[^WID_quotes] The World Bank data and its biannual report offers a much more mixed picture – in which inequality is rising and falling in roughly equal numbers of countries.
+Also in the research questions the datasets aim at. Capital vs labour share. .– including fundamental laws of capitalism with regard to capital accumulation (Piketty) –
+
+Most crucially, the worlds also diverge in terms of headline messages as to inequality around the world. Contradictory messages emerge concerning the most basic descriptions from the two databases.
+
+They differ as to how ubiquitous a phenomenon rising inequality is around the world.
+The view from *World Inequality Database* and its associated global report is that inequality is rising in 'nearly everywhere' [@alvaredo_world_2018; @chancel_world_2021].[^WID_quotes] In contrast, the World Bank data and its biannual report offers a much more mixed picture – in which inequality is rising and falling in roughly equal numbers of countries.
 
@@
 
-This stark difference is reflected in decompositions of global inequality into between- and within-country components. (Explain)
+Secondly, the two worlds differ in their conception of global inequality....  decompositions of global inequality into between- and within-country components. (Explain
 
-Thus despite the great data, there remains fundamental ambiguity surrounding which of two possible worlds we live in.
-In one world, rising inequality within countries has been an almost universal phenomenon, and has now come to dominate income differences between countries.
-In another, within-country inequality trends in recent decades have been mixed, and the differences in average incomes between countries remain the lion's share of the globally.
+Incredible efforts, involving the work of hundreds of researchers.
+Both datasets have continued to improve adn expand.
+In terms of data availability we are thus in a better position than ever to take a global viw on inequlity.
+And yet, amidst this wealth of data, limited understanding of very basics concerning inequality around the world.
 
-Which of these worlds we live in matters hugely – both in our conception, and of where to place efforts to combat inequality in the future.
+There remains fundamental ambiguity surrounding which of two possible worlds we live in.
 
-The first world suggests the winners from globalization and economic growth overwhelmingly and almost universally, have been a global elite.
-It situates inquality within countries, but emphasises global factors – including fundamental laws of capitalism with regard to capital accumulation (Piketty) – and the need for international coordination over and above national policy.
+The '*WID* world' rising inequality within countries has been an almost universal phenomenon.
+suggests the winners from globalization and economic growth overwhelmingly have been a global elite.
+It sees the bulk of global inequalities as lying within nations, but emphasises global factors in terms of causes and hence the need for international coordination to combat inequality.
 
-The second world and the importance of access to both produced and human capital , national policy (Ravallion, 2018; Corlett, 2016).
+The '*PIP* world' – within-country inequality trends in recent decades have been mixed, and the differences in average incomes between countries remain the lion's share of the globally.As such the importance of national policy (Ravallion, 2018; Corlett, 2016).
+access to both produced and human capital.
 
-### Unpicking the two worlds
+Knowing which of these worlds we live in matters a great deal – for our understanding of inequality today, how it has changed, and also where efforts to combat inequality in the future are likely to be most successful.
 
@@
 These to some extent come bundled together when making the choice between PIP and WID.
-That is to say, the empirical question of 'world' of inequality we live in depends on which 'world' of inequality *measurement* we adopt.
 
-Whilst recommending particular aspects, for particular countries, the choice put before us requires wholesale adoption of a bundle whose relative strengths and weaknesses when applied globally is not well understood.
+In many particular contexts There are known strengths and weaknesses.
+However in using the datasets a global view on inequality, requires choosing between methodological 'bundles' whose relative strengths and weaknesses, applied globally, are not well understood.
+
+### Connecting the two worlds
+*(In this section I will explain the plan for the paper)*
 
@@
 In this paper, I focus on trends in within-country inequality in recent decades.(In a companion paper, I look at the global distribution.) I line up what I can.
-I triangulate what we know – .
+I triangulate what we know – What is true irrespective of which 'world' we consider.
 And Identify what we don't know – the remaining differences – and try to pinpoint where the differences lie.
@@
 
+Explain approach. (Including LIS – which is close to PIP world (indeed PIP takes data from LIS for a number of high income countries.))
+
 (SUMMAR OF FINDINGS)
 
-## Data
+## Methods and data
 
-### PIP
+### Concepts of global inequality
 
-### WID
+Adapted three concepts.
 
-(Coverage chart - overlayed 3/5 year)
+### Data
 
-### LIS
+#### PIP
+
+#### WID
+
+#### LIS
 
@@
 
-## Methods
+#### Coverage and lining-up observations
 
-### Concepts of global inequality
+(Coverage chart - overlayed 3/5 year)
 
-Adapted three concepts.
+## Results
+
+### Comparison of typical indicators
+
+- Relative change WID 1% and PIP Gini
+
+### Aligning indices
+
+- Show that measure doesn't matter for WID until top 1 (check Palma ratio)
+
+- Check out how much different metrics matter for PIP.
+
+-----
 

### 2023-04-05 _site/PhD_pages/Paper_1/data_appendix/mytable.tex
Paper 1: drafting and testing out workflow for tables

\begin{tabular}{lllllllllllllllllllllr}
\toprule
{} &           Entity &        region & change PIP: Gini & change WID: Top 1pc share – pretax & value1995 PIP: Gini & value1995 WID: Top 1pc share – pretax & value2015 PIP: Gini & value2015 WID: Top 1pc share – pretax & Year1995 PIP: Gini & Year1995 WID: Top 1pc share – pretax & Year2015 PIP: Gini & Year2015 WID: Top 1pc share – pretax & pop1995 PIP: Gini & pop1995 WID: Top 1pc share – pretax & pop2015 PIP: Gini & pop2015 WID: Top 1pc share – pretax & region PIP: Gini & region WID: Top 1pc share – pretax & pip\_welfare PIP: Gini & pip\_welfare WID: Top 1pc share – pretax &  plot\_pop\_scale \\
\midrule
0  &          Albania &       EE \& CA &        21.263669 &                          18.181818 &           27.010342 &                                  8.14 &           32.753732 &                                  9.62 &               1996 &                                 1996 &               2015 &                                 2015 &         3284370.0 &                           3284370.0 &         2882482.0 &                           2882482.0 &          EE \& CA &                            EE \& CA &           Consumption &                                     NaN &        1.066413 \\
36 &       Kazakhstan &       EE \& CA &       -24.237001 &                           3.501629 &           35.378308 &                                 12.28 &           26.803667 &                                 12.71 &               1996 &                                 1996 &               2015 &                                 2015 &        16434323.0 &                          16434323.0 &        17835908.0 &                          17835908.0 &          EE \& CA &                            EE \& CA &           Consumption &                                     NaN &        1.495755 \\
40 &           Latvia &       EE \& CA &        10.417068 &                           2.434077 &           30.977489 &                                  9.86 &           34.204435 &                                  10.1 &               1995 &                                 1995 &               2015 &                                 2015 &         2570150.0 &                           2570150.0 &         1991951.0 &                           1991951.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.040844 \\
42 &        Lithuania &       EE \& CA &        11.099766 &                          24.861265 &           33.641136 &                                  9.01 &           37.375224 &                                 11.25 &               1993 &                                 1995 &               2015 &                                 2015 &         3747269.0 &                           3747269.0 &         2963764.0 &                           2963764.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.068747 \\
29 &          Hungary &       EE \& CA &        -8.654903 &                          59.670782 &           33.289395 &                                  7.29 &            30.40823 &                                 11.64 &               1994 &                                 1995 &               2015 &                                 2015 &        10327171.0 &                          10327171.0 &         9844250.0 &                           9844250.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.266299 \\
49 &          Moldova &       EE \& CA &       -26.836525 &                          25.581395 &           36.942146 &                                  9.03 &           27.028158 &                                 11.34 &               1997 &                                 1993 &               2015 &                                 2015 &         4461238.0 &                           4461238.0 &         3277390.0 &                           3277390.0 &          EE \& CA &                            EE \& CA &           Consumption &                                     NaN &        1.077751 \\
26 &          Georgia &       EE \& CA &         -1.69401 &                          10.278373 &           37.132707 &                                 18.68 &           36.503676 &                                  20.6 &               1996 &                                 1997 &               2015 &                                 2015 &         4877682.0 &                           4877682.0 &         3771135.0 &                           3771135.0 &          EE \& CA &                            EE \& CA &           Consumption &                                     NaN &        1.091928 \\
55 &  North Macedonia &       EE \& CA &        26.542688 &                         -18.590998 &           28.128578 &                                 10.22 &           35.594659 &                                  8.32 &               1998 &                                 1996 &               2015 &                                 2015 &         1983971.0 &                           1983971.0 &         2107965.0 &                           2107965.0 &          EE \& CA &                            EE \& CA &                 Mixed &                                     NaN &        1.044175 \\
59 &           Poland &       EE \& CA &        -8.351016 &                          35.974499 &           34.644721 &                                 10.98 &           31.751535 &                                 14.93 &               1995 &                                 1995 &               2015 &                                 2015 &        38401980.0 &                          38401980.0 &        38553148.0 &                          38553148.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        2.090589 \\
60 &          Romania &       EE \& CA &        15.263691 &                          30.492572 &           31.158026 &                                 12.79 &           35.913891 &                                 16.69 &               1995 &                                 1995 &               2015 &                                 2015 &        22533138.0 &                          22533138.0 &        19906078.0 &                          19906078.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.555194 \\
18 &          Czechia &       EE \& CA &         0.265217 &                           2.520252 &            25.79901 &                                 11.11 &           25.867433 &                                 11.39 &               1996 &                                 1996 &               2015 &                                 2015 &        10311702.0 &                          10311702.0 &        10523801.0 &                          10523801.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.285810 \\
17 &          Croatia &       EE \& CA &         8.233354 &                          48.209366 &           28.748361 &                                  7.26 &           31.115315 &                                 10.76 &               1998 &                                 1998 &               2015 &                                 2015 &         4807100.0 &                           4807100.0 &         4254821.0 &                           4254821.0 &          EE \& CA &                            EE \& CA &                 Mixed &                                     NaN &        1.105815 \\
62 &         Slovakia &       EE \& CA &         0.814271 &                          69.863014 &           26.277685 &                                  6.57 &           26.491657 &                                 11.16 &               1996 &                                 1997 &               2015 &                                 2015 &         5349999.0 &                           5349999.0 &         5424449.0 &                           5424449.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.139398 \\
61 &           Russia &       EE \& CA &       -18.138815 &                          34.978229 &           46.103463 &                                 13.78 &           37.740842 &                                  18.6 &               1996 &                                 1995 &               2015 &                                 2015 &       148557456.0 &                         148557456.0 &       144668384.0 &                         144668384.0 &          EE \& CA &                            EE \& CA &           Consumption &                                     NaN &        5.137369 \\
38 &       Kyrgyzstan &       EE \& CA &       -37.466734 &                            -6.3543 &           46.371982 &                                 15.58 &           28.997915 &                                 14.59 &               1998 &                                 1998 &               2015 &                                 2015 &         4596457.0 &                           4596457.0 &         5914985.0 &                           5914985.0 &          EE \& CA &                            EE \& CA &                 Mixed &                                     NaN &        1.153482 \\
2  &          Armenia &       EE \& CA &       -27.154942 &                         -20.716921 &           44.418623 &                                 18.97 &           32.356772 &                                 15.04 &               1996 &                                 1996 &               2015 &                                 2015 &         3322788.0 &                           3322788.0 &         2878598.0 &                           2878598.0 &          EE \& CA &                            EE \& CA &                 Mixed &                                     NaN &        1.066301 \\
73 &          Ukraine &       EE \& CA &       -35.201993 &                         -41.273885 &           39.287536 &                                  15.7 &            25.45754 &                                  9.22 &               1995 &                                 1995 &               2015 &                                 2015 &        51061268.0 &                          51061268.0 &        44982568.0 &                          44982568.0 &          EE \& CA &                            EE \& CA &           Consumption &                                     NaN &        2.275190 \\
5  &          Belarus &       EE \& CA &       -19.942076 &                         -27.177419 &            32.01835 &                                  12.4 &           25.633226 &                                  9.03 &               1998 &                                 1995 &               2015 &                                 2015 &        10461694.0 &                          10461694.0 &         9700609.0 &                           9700609.0 &          EE \& CA &                            EE \& CA &           Consumption &                                     NaN &        1.262175 \\
63 &         Slovenia &       EE \& CA &       -12.939966 &                          10.232558 &           29.184993 &                                  6.45 &           25.408465 &                                  7.11 &               1993 &                                 1995 &               2015 &                                 2015 &         1991115.0 &                           1991115.0 &         2080865.0 &                           2080865.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.043397 \\
9  &         Bulgaria &       EE \& CA &        25.568016 &                           3.258508 &           30.714308 &                                 13.81 &           38.567347 &                                 14.26 &               1992 &                                 1995 &               2015 &                                 2015 &         8477584.0 &                           8477584.0 &         7309255.0 &                           7309255.0 &          EE \& CA &                            EE \& CA &                Income &                                     NaN &        1.193514 \\
50 &         Mongolia &     East Asia &        -2.559418 &                          12.735849 &            33.19534 &                                 14.84 &           32.345733 &                                 16.73 &               1995 &                                 1995 &               2016 &                                 2015 &         2330339.0 &                           2330339.0 &         2964756.0 &                           2964756.0 &        East Asia &                          East Asia &           Consumption &                                     NaN &        1.068775 \\
14 &            China &     East Asia &         9.559654 &                          49.253731 &            35.22766 &                                  9.38 &           38.595302 &                                  14.0 &               1996 &                                 1995 &               2015 &                                 2015 &      1218144384.0 &                        1218144384.0 &      1393715456.0 &                        1393715456.0 &        East Asia &                          East Asia &           Consumption &                                     NaN &       41.000000 \\
68 &           Taiwan &     East Asia &         0.506233 &                          51.237624 &           31.324143 &                                 12.12 &           31.482716 &                                 18.33 &               1995 &                                 1995 &               2016 &                                 2015 &        21455818.0 &                          21455818.0 &        23512136.0 &                          23512136.0 &        East Asia &                          East Asia &                Income &                                     NaN &        1.658731 \\
8  &           Brazil &           LAC &       -12.823454 &                          -4.170177 &           59.569725 &                                 23.74 &           51.930829 &                                 22.75 &               1995 &                                 1995 &               2015 &                                 2015 &       163515328.0 &                         163515328.0 &       205188208.0 &                         205188208.0 &              LAC &                                LAC &                Income &                                     NaN &        6.875013 \\
1  &        Argentina &           LAC &       -14.051647 &                          11.587983 &           48.904403 &                                 13.98 &           42.032529 &                                  15.6 &               1995 &                                 1995 &               2016 &                                 2015 &        34946116.0 &                          34946116.0 &        43257064.0 &                          43257064.0 &              LAC &                                LAC &                Income &                                     NaN &        2.225647 \\
15 &         Colombia &           LAC &       -10.443994 &                         -17.310587 &           56.936908 &                                 19.93 &           50.990421 &                                 16.48 &               1996 &                                 1995 &               2015 &                                 2015 &        35970104.0 &                          35970104.0 &        47119728.0 &                          47119728.0 &              LAC &                                LAC &                Income &                                     NaN &        2.336552 \\
48 &           Mexico &           LAC &         -9.71656 &                         122.610141 &           52.808923 &                                 12.03 &           47.677713 &                                 26.78 &               1994 &                                 1995 &               2016 &                                 2015 &        89969568.0 &                          89969568.0 &       120149896.0 &                         120149896.0 &              LAC &                                LAC &                Income &                                     NaN &        4.433394 \\
70 &          Tunisia &          MENA &       -21.224167 &                         -35.823529 &           41.657297 &                                  17.0 &           32.815882 &                                 10.91 &               1995 &                                 1995 &               2015 &                                 2015 &         9294106.0 &                           9294106.0 &        11557779.0 &                          11557779.0 &             MENA &                               MENA &           Consumption &                                     NaN &        1.315498 \\
20 &            Egypt &          MENA &          5.63083 &                           22.41066 &            30.12696 &                                 16.51 &           31.823358 &                                 20.21 &               1995 &                                 1995 &               2015 &                                 2015 &        64166908.0 &                          64166908.0 &        97723792.0 &                          97723792.0 &             MENA &                               MENA &           Consumption &                                     NaN &        3.789496 \\
32 &             Iran &          MENA &        -8.203486 &                         -16.334661 &            42.99837 &                                 20.08 &           39.471004 &                                  16.8 &               1994 &                                 1994 &               2015 &                                 2015 &        60794808.0 &                          60794808.0 &        81790848.0 &                          81790848.0 &             MENA &                               MENA &           Consumption &                                     NaN &        3.332029 \\
34 &           Israel &          MENA &         4.258654 &                          -8.049189 &           38.097106 &                                 17.89 &            39.71953 &                                 16.45 &               1997 &                                 1997 &               2015 &                                 2015 &         5478738.0 &                           5478738.0 &         8007777.0 &                           8007777.0 &             MENA &                               MENA &                Income &                                     NaN &        1.213570 \\
77 &            Yemen &          MENA &         4.884046 &                          11.366774 &           34.997786 &                                 21.73 &           36.707094 &                                  24.2 &               1998 &                                 1998 &               2014 &                                 2014 &        16103338.0 &                          16103338.0 &        28516552.0 &                          28516552.0 &             MENA &                               MENA &           Consumption &                                     NaN &        1.802418 \\
71 &           Turkey &          MENA &         3.725639 &                          -2.744186 &           41.315202 &                                  21.5 &           42.854458 &                                 20.91 &               1994 &                                 1995 &               2015 &                                 2015 &        59305492.0 &                          59305492.0 &        79646176.0 &                          79646176.0 &             MENA &                               MENA &           Consumption &                                     NaN &        3.270451 \\
69 &         Thailand &   S \& SE Asia &       -17.213809 &                          -15.49978 &           43.470212 &                                 22.71 &           35.987333 &                                 19.19 &               1994 &                                 1995 &               2015 &                                 2015 &        59424836.0 &                          59424836.0 &        70294408.0 &                          70294408.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        3.001944 \\
65 &        Sri Lanka &   S \& SE Asia &          11.1477 &                          26.670792 &           35.399539 &                                 16.16 &           39.345773 &                                 20.47 &               1995 &                                 1995 &               2016 &                                 2016 &        18177570.0 &                          18177570.0 &        21336692.0 &                          21336692.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        1.596270 \\
46 &         Malaysia &   S \& SE Asia &       -15.366576 &                         -24.364191 &           48.522672 &                                 19.66 &           41.066398 &                                 14.87 &               1995 &                                 1995 &               2015 &                                 2015 &        20136886.0 &                          20136886.0 &        31068834.0 &                          31068834.0 &      S \& SE Asia &                        S \& SE Asia &                Income &                                     NaN &        1.875699 \\
76 &          Vietnam &   S \& SE Asia &        -0.495084 &                          -1.432836 &           35.442207 &                                 16.75 &           35.266739 &                                 16.51 &               1997 &                                 1997 &               2016 &                                 2016 &        73759112.0 &                          73759112.0 &        92191400.0 &                          92191400.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        3.630650 \\
57 &         Pakistan &   S \& SE Asia &         9.161968 &                           3.663664 &           28.664989 &                                 16.65 &           31.291266 &                                 17.26 &               1996 &                                 1996 &               2015 &                                 2015 &       133117480.0 &                         133117480.0 &       210969296.0 &                         210969296.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        7.041000 \\
58 &      Philippines &   S \& SE Asia &        -5.915094 &                         -11.019141 &           42.888387 &                                 19.33 &           40.351499 &                                  17.2 &               1994 &                                 1994 &               2015 &                                 2015 &        69250472.0 &                          69250472.0 &       103031360.0 &                         103031360.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        3.941887 \\
39 &             Laos &   S \& SE Asia &        11.196288 &                           1.965856 &            34.89544 &                                 19.33 &           38.802434 &                                 19.71 &               1997 &                                 1997 &               2018 &                                 2018 &         4927432.0 &                           4927432.0 &         6787420.0 &                           6787420.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        1.178531 \\
4  &       Bangladesh &   S \& SE Asia &        -1.692956 &                          -3.351287 &           32.942983 &                                 16.71 &           32.385273 &                                 16.15 &               1995 &                                 1995 &               2016 &                                 2016 &       117793336.0 &                         117793336.0 &       157830000.0 &                         157830000.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        5.515265 \\
31 &        Indonesia &   S \& SE Asia &        13.395688 &                          -4.942166 &           35.625137 &                                 19.02 &           40.397369 &                                 18.08 &               1996 &                                 1995 &               2015 &                                 2015 &       198140160.0 &                         198140160.0 &       259091968.0 &                         259091968.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &        8.422698 \\
30 &            India &   S \& SE Asia &         9.497208 &                          64.794007 &           31.696184 &                                 13.35 &           34.706436 &                                  22.0 &               1993 &                                 1995 &               2015 &                                 2013 &       964279104.0 &                         964279104.0 &      1322866560.0 &                        1322866560.0 &      S \& SE Asia &                        S \& SE Asia &           Consumption &                                     NaN &       38.965787 \\
22 &         Ethiopia &     SS Africa &       -21.477605 &                         -53.452341 &           44.564513 &                                 29.69 &           34.993123 &                                 13.82 &               1995 &                                 1995 &               2015 &                                 2015 &        57476540.0 &                          57476540.0 &       102471896.0 &                         102471896.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        3.925823 \\
72 &           Uganda &     SS Africa &         9.502102 &                          11.074919 &           39.043719 &                                 15.35 &           42.753693 &                                 17.05 &               1996 &                                 1996 &               2016 &                                 2016 &        20680834.0 &                          20680834.0 &        37477356.0 &                          37477356.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        2.059701 \\
7  &         Botswana &     SS Africa &       -12.280875 &                         -36.973392 &            60.79065 &                                 36.08 &           53.325027 &                                 22.74 &               1993 &                                 1993 &               2015 &                                 2015 &         1543641.0 &                           1543641.0 &         2305177.0 &                           2305177.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.049837 \\
10 &     Burkina Faso &     SS Africa &       -26.566681 &                          -31.54678 &           48.072493 &                                 21.27 &           35.301228 &                                 14.56 &               1994 &                                 1994 &               2014 &                                 2014 &        10353268.0 &                          10353268.0 &        18718022.0 &                          18718022.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.521083 \\
11 &          Burundi &     SS Africa &        15.872792 &                           17.63767 &           33.333812 &                                 12.53 &           38.624819 &                                 14.74 &               1992 &                                 1992 &               2013 &                                 2013 &         5932789.0 &                           5932789.0 &        10727153.0 &                          10727153.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.291649 \\
12 &         Cameroon &     SS Africa &         4.936503 &                         -14.277988 &           44.446758 &                                 18.49 &           46.640874 &                                 15.85 &               1996 &                                 1996 &               2014 &                                 2014 &        13211643.0 &                          13211643.0 &        23012650.0 &                          23012650.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.644390 \\
16 &    Cote d'Ivoire &     SS Africa &         2.245793 &                           5.195599 &            40.56382 &                                 16.36 &             41.4748 &                                 17.21 &               1995 &                                 1995 &               2015 &                                 2015 &        14299727.0 &                          14299727.0 &        23596738.0 &                          23596738.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.661160 \\
21 &         Eswatini &     SS Africa &        -9.717904 &                         -38.848466 &           60.454726 &                                 31.61 &           54.579794 &                                 19.33 &               1994 &                                 1994 &               2016 &                                 2016 &          953586.0 &                            953586.0 &         1133941.0 &                           1133941.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.016209 \\
37 &            Kenya &     SS Africa &         -5.41145 &                         -19.926199 &           43.108572 &                                 18.97 &           40.775773 &                                 15.19 &               1994 &                                 1994 &               2015 &                                 2015 &        26878350.0 &                          26878350.0 &        46851496.0 &                          46851496.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        2.328851 \\
25 &           Gambia &     SS Africa &       -25.970265 &                         -26.670259 &           48.519465 &                                 18.56 &           35.918831 &                                 13.61 &               1998 &                                 1992 &               2015 &                                 2015 &         1242160.0 &                           1242160.0 &         2253137.0 &                           2253137.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.048343 \\
54 &          Nigeria &     SS Africa &       -30.897099 &                           2.930728 &           51.922246 &                                 11.26 &           35.879779 &                                 11.59 &               1996 &                                 1992 &               2015 &                                 2018 &       108187616.0 &                         108187616.0 &       183995776.0 &                         183995776.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        6.266536 \\
78 &           Zambia &     SS Africa &        18.211681 &                          21.947368 &           48.333683 &                                  19.0 &           57.136058 &                                 23.17 &               1996 &                                 1996 &               2015 &                                 2015 &         8684139.0 &                           8684139.0 &        16248231.0 &                          16248231.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.450170 \\
41 &          Lesotho &     SS Africa &        -29.03314 &                          -35.38874 &           63.240365 &                                 22.38 &           44.879701 &                                 14.46 &               1994 &                                 1994 &               2017 &                                 2017 &         1934085.0 &                           1934085.0 &         2118532.0 &                           2118532.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.044478 \\
51 &       Mozambique &     SS Africa &         0.815567 &                          29.194352 &           53.559865 &                                 24.08 &           53.996682 &                                 31.11 &               1996 &                                 1996 &               2014 &                                 2014 &        15594829.0 &                          15594829.0 &        26843252.0 &                          26843252.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.754374 \\
44 &       Madagascar &     SS Africa &        -5.761739 &                            2.29265 &           45.255698 &                                 14.83 &           42.648182 &                                 15.17 &               1993 &                                 1997 &               2012 &                                 2012 &        13869136.0 &                          13869136.0 &        24850912.0 &                          24850912.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.697170 \\
45 &           Malawi &     SS Africa &       -32.029625 &                           64.15883 &           65.755571 &                                 19.14 &           44.694308 &                                 31.42 &               1997 &                                 1997 &               2016 &                                 2016 &        10112813.0 &                          10112813.0 &        16938944.0 &                          16938944.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.470002 \\
47 &       Mauritania &     SS Africa &       -13.585089 &                         -14.205906 &           37.747363 &                                 12.53 &            32.61935 &                                 10.75 &               1995 &                                 1995 &               2014 &                                 2014 &         2380345.0 &                           2380345.0 &         3946225.0 &                           3946225.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.096955 \\
53 &            Niger &     SS Africa &       -17.457484 &                         -30.911257 &           41.525965 &                                 16.79 &           34.276576 &                                  11.6 &               1994 &                                 1994 &               2014 &                                 2014 &         9813920.0 &                           9813920.0 &        20128132.0 &                          20128132.0 &        SS Africa &                          SS Africa &           Consumption &                                     NaN &        1.561570 \\
52 &      Netherlands &     W. Europe &        -7.206846 &                           16.41286 &           30.434982 &                                  5.91 &            28.24158 &                                  6.88 &               1993 &                                 1995 &               2015 &                                 2015 &        15446670.0 &                          15446670.0 &        17041110.0 &                          17041110.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.472935 \\
74 &   United Kingdom &     W. Europe &        -6.423917 &                          21.541502 &            35.51664 &                                 10.12 &           33.235081 &                                  12.3 &               1995 &                                 1995 &               2015 &                                 2015 &        57930364.0 &                          57930364.0 &        65224368.0 &                          65224368.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        2.856373 \\
33 &          Ireland &     W. Europe &       -14.100525 &                          16.699605 &           37.045092 &                                 10.12 &            31.82154 &                                 11.81 &               1995 &                                 1995 &               2015 &                                 2015 &         3579486.0 &                           3579486.0 &         4665764.0 &                           4665764.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.117614 \\
6  &          Belgium &     W. Europe &        -2.403209 &                           5.341615 &           28.412269 &                                  8.05 &           27.729463 &                                  8.48 &               1995 &                                 1995 &               2015 &                                 2015 &        10095204.0 &                          10095204.0 &        11248306.0 &                          11248306.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.306612 \\
43 &       Luxembourg &     W. Europe &        22.185633 &                         -21.158129 &           26.944439 &                                 13.47 &           32.922233 &                                 10.62 &               1994 &                                 1994 &               2015 &                                 2015 &          407639.0 &                            407639.0 &          569415.0 &                            569415.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.000000 \\
67 &      Switzerland &     W. Europe &        -4.576537 &                          16.213494 &           33.856408 &                                  9.93 &           32.306957 &                                 11.54 &               1992 &                                 1995 &               2015 &                                 2015 &         7038302.0 &                           7038302.0 &         8281735.0 &                           8281735.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.221436 \\
66 &           Sweden &     W. Europe &        15.745782 &                          -9.957447 &           25.236506 &                                 11.75 &           29.210191 &                                 10.58 &               1995 &                                 1995 &               2015 &                                 2015 &         8758496.0 &                           8758496.0 &         9849344.0 &                           9849344.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.266445 \\
28 &           Greece &     W. Europe &        -2.637722 &                          26.701031 &           36.975572 &                                   9.7 &            36.00026 &                                 12.29 &               1995 &                                 1995 &               2015 &                                 2015 &        10727928.0 &                          10727928.0 &        10806644.0 &                          10806644.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.293931 \\
19 &          Denmark &     W. Europe &        22.746008 &                          36.029412 &           22.955896 &                                  9.52 &           28.177446 &                                 12.95 &               1995 &                                 1995 &               2015 &                                 2015 &         5235205.0 &                           5235205.0 &         5677795.0 &                           5677795.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.146672 \\
27 &          Germany &     W. Europe &         8.954429 &                          51.025057 &           28.987663 &                                  8.78 &           31.583342 &                                 13.26 &               1995 &                                 1995 &               2015 &                                 2015 &        81123264.0 &                          81123264.0 &        82073232.0 &                          82073232.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        3.340137 \\
23 &          Finland &     W. Europe &        15.528727 &                           6.696429 &           23.459137 &                                  8.96 &           27.102042 &                                  9.56 &               1995 &                                 1995 &               2015 &                                 2015 &         5107942.0 &                           5107942.0 &         5479457.0 &                           5479457.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.140977 \\
24 &           France &     W. Europe &         1.156214 &                          13.687783 &           32.330525 &                                  8.84 &           32.704335 &                                 10.05 &               1994 &                                 1995 &               2015 &                                 2015 &        57420356.0 &                          57420356.0 &        63809768.0 &                          63809768.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        2.815757 \\
56 &           Norway &     W. Europe &          5.67384 &                          16.471963 &           26.020532 &                                  8.56 &           27.496896 &                                  9.97 &               1995 &                                 1995 &               2015 &                                 2015 &         4359327.0 &                           4359327.0 &         5190357.0 &                           5190357.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        1.132676 \\
64 &            Spain &     W. Europe &        -1.289625 &                           8.311911 &            36.65412 &                                 11.67 &           36.181419 &                                 12.64 &               1995 &                                 1995 &               2015 &                                 2015 &        39814568.0 &                          39814568.0 &        46431344.0 &                          46431344.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        2.316787 \\
35 &            Italy &     W. Europe &         0.452275 &                          27.468785 &           35.238064 &                                  8.81 &           35.397436 &                                 11.23 &               1995 &                                 1995 &               2015 &                                 2014 &        56885124.0 &                          56885124.0 &        60232908.0 &                          60232908.0 &        W. Europe &                          W. Europe &                Income &                                     NaN &        2.713058 \\
13 &           Canada &  W. Offshoots &         7.968926 &                          49.239544 &           31.255736 &                                 10.52 &           33.746482 &                                  15.7 &               1994 &                                 1995 &               2015 &                                 2015 &        29289430.0 &                          29289430.0 &        35732120.0 &                          35732120.0 &     W. Offshoots &                       W. Offshoots &                Income &                                     NaN &        2.009591 \\
75 &    United States &  W. Offshoots &         3.472042 &                          30.096419 &           39.854995 &                                 14.52 &           41.238778 &                                 18.89 &               1995 &                                 1995 &               2015 &                                 2015 &       265660560.0 &                         265660560.0 &       324607776.0 &                         324607776.0 &     W. Offshoots &                       W. Offshoots &                Income &                                     NaN &       10.303787 \\
3  &        Australia &  W. Offshoots &         3.387823 &                          40.714286 &            32.58199 &                                   8.4 &           33.685811 &                                 11.82 &               1995 &                                 1995 &               2016 &                                 2015 &        18003002.0 &                          18003002.0 &        23820240.0 &                          23820240.0 &     W. Offshoots &                       W. Offshoots &                Income &                                     NaN &        1.667578 \\
\bottomrule
\end{tabular}

### 2023-04-05 PhD_pages/Paper_1/data_appendix/G_functions_for_figures_and_tables.py

@@ -1,0 +1,787 @@
+# %%
+# ---
+# title: "Functions for preparing figures and tables"
+# execute: 
+#   enabled: false
+# ---
+
+# %% [markdown]
+# The code here is duplicated in as a .py file – from which the functions are imported in data appendix H.
+
+# %%
+
+
+from plotnine import *
+import plotly.express as px
+import pandas as pd
+import numpy as np
+import plotly.graph_objects as go
+import math
+
+# %% [markdown]
+#  This function grabs the observation nearest to a reference value of a particular column (e.g. a reference year). A parameter specifies how to break ties where a value greater and lower than the reference values are the equidistant closest values.
+#  %
+#   Function to get matching observation closest to ref year
+#  Drop NAs
+#  Calculate absolute distance from reference value
+#  Drop any rows with a distance beyond threshold
+#  Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
+#  Settle tie-breaks
+
+# %%
+#  Function to get matching observation closest to ref year
+def closest_to_reference(
+    df, 
+    reference_val, 
+    max_dist_from_ref, 
+    reference_col, 
+    group_by_col, 
+    value_col, 
+    tie_break):
+  
+  df = df.loc[:, [reference_col, group_by_col, value_col]]
+
+  # Drop NAs
+  df = df.dropna()
+
+  # Calculate absolute distance from reference value
+  df['ref_diff'] = abs(df[reference_col] - reference_val)
+
+  # Drop any rows with a distance beyond threshold
+  if not pd.isna(max_dist_from_ref):
+    df = df.loc[df['ref_diff'] <= max_dist_from_ref]
+
+  # Keep closest observation to reference value – including tie-breaks (where there is a match above and below the ref value)
+  df = df[df.groupby(group_by_col)['ref_diff'].transform('min') == df['ref_diff']].reset_index(drop=True)
+
+  # Settle tie-breaks
+  if tie_break == 'below':
+    df = df[df.groupby(group_by_col)[reference_col].transform('min') == df[reference_col]].reset_index(drop=True)
+    
+  elif tie_break == 'above':
+    df = df[df.groupby(group_by_col)[reference_col].transform('max') == df[reference_col]].reset_index(drop=True)
+
+  df = df.drop('ref_diff', axis=1)
+
+  df = df\
+    .rename(columns={value_col: "value"})
+
+
+  return df
+
+# %% [markdown]
[…]

### 2023-04-19 PhD_pages/Paper_1/data_appendix/1_H_figures_and_tables.qmd
Paper 1: Further drafting and adding figures

@@
                     1
+                ], 
+                'relative_tolerance': [
+                    0.05
                 ]
@@
 
-## Tables
+## SECTION 2:
 
-#### Table 1
+### WID top 1% vs PIP Gini
+
+#### Since 1993
 
@@
 # Set-up to compare WID and PIP top 10% share, all countries
-reference_vals = [1995, 2015]
+reference_vals = [1993, 2015]
 max_dist_from_refs = 3
@@
 all_metrics_requirement = True
+
+```
+
+```{python}
+
+df_prepped = prep_data(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_regions = df_regions,
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      all_metrics_requirement = all_metrics_requirement,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup
+    )
 
@@
       tolerance_lookup =  tolerance_lookup,
+      absolute_or_relative_tolerance = "absolute",
       region_col = 'region'
@@
 
-df_summary['PIP: Gini']
+# df_summary = df_summary['WID: Top 1pc share – pretax']
+df_summary = df_summary['PIP: Gini']
 
-dfi.export(df_summary,"table_test.png")
-
+df_summary
 ```
-![](table_test.png)
 

### 2023-04-19 PhD_pages/Paper_1/paper1.qmd
Paper 1: Further drafting and adding figures

@@
 import numpy as np
-from IPython.display import display, HTML, IFrame, Markdown
-import dataframe_image as dfi
-import imgkit
+# from IPython.display import display, HTML, IFrame, Markdown
+# import dataframe_image as dfi
+# import imgkit
 
-from wrangling_and_plotting_functions import *
+from data_appendix.G_functions_for_figures_and_tables import *
 
-pd.set_option('display.max_rows', None)
+# pd.set_option('display.max_rows', None)
 
@@
 
-Two sources of data have been particularly impactful in this regard: the World Bank's harmonized collection of household survey data, published online through its [*Poverty and Inequality Platform (PIP)*](https://pip.worldbank.org/); and the collection of inequality estimates – most notably top income shares – brought together at the [*World Inequality Database (WID)*](https://wid.world/).^[EDIT: PIP used to be Povcal.
+Two sources of data have been particularly impactful in this regard.
+One is the World Bank's harmonized collection of household survey data, published online through its [*Poverty and Inequality Platform (PIP)*](https://pip.worldbank.org/).
+Another is the collection of inequality estimates, most notably top income shares, brought together at the [*World Inequality Database (WID)*](https://wid.world/).^[TODO: PIP used to be Povcal.
 It draws upon the work of various regional teams.
@@
 
-These two sources are the only regularly-updated databases providing global coverage on a relatively consistent basis.^[EDIT: Both datasets suffer important limitations with respect to comparability acrosss time and location.
+These two sources are the only regularly-updated databases that provide global coverage of distributional data on a relatively comparable basis.^[EDIT: Both datasets suffer important limitations with respect to comparability acrosss time and location.
 These will be discussed in the data section below.
 Both datasets, however, take significant steps to harmonize, *ex-post*, the available underlying data.
-In this respect they can be distinguished from databases such as WIID and SWIID [@atkinson_promise_2001]] Both have played substantial agenda-setting roles in academic research, policy and public debate.
+In this respect they can be distinguished from databases such as WIID and SWIID [@atkinson_promise_2001]]
 
-Given its extensive coverage of low- and middle-income countries, the World Bank data occupies a central place in international development policy and research.
-Most notably, the data is relied on by the UN to set and monitor progress towards many of the targets included within Sustainable Development Goals.
-It has also to date underpinned the majority of academic studies on inequality that take a global view.^[[Including @Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk] TODO: Check and more refs].
+Both have played substantial agenda-setting roles in academic research, policy and public debate.
 
-The *WID* database emerged as institutional focal-point to host and coordinate inequality estimates produced as part of the burgeoning literature on top incomes [@atkinson_top_2014; @piketty_capital_2014; @alvaredo_top_2013].
-This development of this data has opened new avenues for the study of inequality right across the social sciences (TODO: example refs), underpinning a significant re-evaluation of its causes and consequences [@alvaredo_top_2013].
-It has also surely had a substantial impact on the public's conception of inequality, contributing to the momentum generated by the Occupy movement and its framing of inequality in terms of 'the 1%' vs the '99%''.
+The World Bank data is especially prominent within international development policy, notably being used to monitor many of the UN Sustainable Development Goals.
+Academic research on global inequality has also largely relied on this data  [@Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk].
+*[TODO: Check and more refs.]*.
 
-#### Two worlds of inequality measurement
+The WID database emerged along with the burgeoning literature on top incomes, acting as a central repository and, increasingly, coordinating institution. The 1%
 
-The two . They in some sense represent *two different worlds* of inequality measurement.
+## Two worlds of inequality measurement
 
-In terms of methods, source, conception of inequality and research questions.
-"The 1 %" – pretax, with many assumptions needed to get to national income.
-These come as a bundle.
-Although both institutions continue to take many steps to provide transparency and reproducibility to their estimates, they remain a black box.
+As well as being the the two  not just two different databases. They in some sense represent *two different worlds* of inequality measurement.
 
-Also in the research questions the datasets aim at. Capital vs labour share. .– including fundamental laws of capitalism with regard to capital accumulation (Piketty) –
+Different methods. These yield different pictures of the extent of inequality and how it has changed over time. Motivated by different ideas about the salient factors shaping inequality.
 
-Most crucially, the worlds also diverge in terms of headline messages as to inequality around the world. Contradictory messages emerge concerning the most basic descriptions from the two databases.
+### PIP
+
+### WID
 
@@
 
-Secondly, the two worlds differ in their conception of global inequality....  decompositions of global inequality into between- and within-country components. (Explain
+Decompositions of global inequality into between- and within-country components.
 
-Incredible efforts, involving the work of hundreds of researchers.
-Both datasets have continued to improve adn expand.
-In terms of data availability we are thus in a better position than ever to take a global viw on inequlity.
-And yet, amidst this wealth of data, limited understanding of very basics concerning inequality around the world.
+Capital vs labour share. .– including fundamental laws of capitalism with regard to capital accumulation (Piketty) –
+
+### The disconnect between these datasets devaules them both
 
@@
 
-Many methodolodical choices concerning underlying data sources, the constraints on temporal coverage this gives rise to,  adjust, conception of inequality.
-These to some extent come bundled together when making the choice between PIP and WID.
+Trouble is that these are all bundled up. You must chose. On what basis can you make a choice?
 
-In many particular contexts There are known strengths and weaknesses.
-However in using the datasets a global view on inequality, requires choosing between methodological 'bundles' whose relative strengths and weaknesses, applied globally, are not well understood.
+Known problems on both sides.
 
-### Connecting the two worlds
+For PIP – top incomes, and lack of comparability of the survey data – notably mix of incme and consumption data.
+
+For WID – tha many assumptions.
+An emphasis placed on the incorporation of tax data, but when taking a global view confonte by the fact that such tax data is often nonexistent.
+Top income corrections based on strong assumptions.
+
+Given the presentation, the extent to which different assumptions, indices, methodological choices affects the data and – is not something that can be read off' the data.
+
+In using the datasets a global view on inequality, requires choosing wholesale between methodological 'bundles' whose relative strengths and weaknesses, applied globally, are not well understood.
+
+Also makes for a very confusing terrain for policymakers, media and the public.
+
+The two datasets are incredible efforts, involving the work of hundreds of researchers.
+Both datasets have continued to improve adn expand.
+In terms of data availability we are thus in a better position than ever to take a global viw on inequlity.
+And yet, amidst this wealth of data, limited understanding of very basics concerning inequality around the world.
+
+### Outline: Connecting the two worlds
 *(In this section I will explain the plan for the paper)*
 
-What we can do is triangulate
+In this paper I aim to bring these two worlds of inequality measurement closer. To Triangulate what they do agree on and identify where they disagree.
 
-In this paper, I focus on trends in within-country inequality in recent decades.(In a companion paper, I look at the global distribution.) I line up what I can.
-I triangulate what we know – What is true irrespective of which 'world' we consider.
-And Identify what we don't know – the remaining differences – and try to pinpoint where the differences lie.
+The first, most obvious, step is simply to compare the in a systemtatic way.
+This very basic step is curiously lacking in the literature.
+We begin by comparing trends in each world based on the inequality indices most typical in each: the Gini coefficient for the PIP data, and the top 1% share of pre-tax income for WID.
 
-I do so with respect to two separate time periods, chosen to tradeoffs with respect to global coverage.
-The first period – from the fall of the Berlin wall.
-Driven primarily to the steep drop off in availability before this in household survey data collected in PIP.
-
-This maximises global coverage – 100 countries covering roughly 90% of the world poplation.
+Given incomplete coverage in both datasets, we 'line up' estimates to reference years to see change over time.
+We use two periods, presenting two different tradeoffs with respect to data coverage.
+The first is 1993-2018??
+This maximises global coverage – 100 countries covering roughly 90% of the world poplation (Figure XX).
 The downside is that this period misses out – transition, liberalisation and structural adjustment.
@@
 
-To address this, I analyse a second, longer period extending back to roughly 1980.
+The second is 1980-2018. Here we lose coverage, especially in the PIP data (Figure XX).
 
-Explain approach. (Including LIS – which is close to PIP world (indeed PIP takes data from LIS for a number of high income countries.))
+The two measures selected – Gini coefficient of disposable income/consumption, and the top 1% share of pre-tax income – are where the PIP and WID worlds respectively are at their furthest apart.
+Yet in terms of broad global trends, there is much that even these very different measures can agree on: - x, - y, - z.
 
-(SUMMAR OF FINDINGS)
+There are big differences though:
+- Somewhat more pronounced rise among advanced economies.
+- Big divergence in trend some of the most populous countries
 
-## Methods and data
+The rest of the paper is about understanding where those differences emerge.
 
-### Concepts of global inequality
+- Inequality indices chosen. We align. Compare Top 1%. And Gini.
+    - Explain results.
 
-Adapted three concepts.
+- We look at different income concepts for the small number of coutnries for which such data exists.
+To this end we also include data from the Luxembourg Income Study.
+Similar to PIP in that it relies on survey data.
+But provides access to microdata, allowing for the construction of different welfare aggregates.
+We focus on the countries with the biggest difference in trend and biggest population.
+I.e.
+those that contribute most to the difference in the weighted an unweigted global averages.
 
-### Data
+    What I won't do is a detailed lining up.
+    Although both institutions continue to take many steps to provide transparency and reproducibility to their estimates, they remain a black box. Makes it impossible to do this lining up. 
+    In many particular contexts – where detailed lining up is possible – there relative strengths and weaknesses may be clear.
+(This is what Morllei does.
+And Nolan).
+But at the global level the makes this impractical.
 
-#### PIP
+    - Summarise: really depends on the country.
 
-#### WID
+----
 
-#### LIS
-
-Will be used for income concept.
-i.e.
-it represents the PIP world – indeed LIS is the source for PIP for a number of high income countries.
-But the access to microdata allows us to ask more questions for a subset of countries.
-
-#### Coverage and lining-up observations
-
-(Coverage chart - overlayed 3/5 year)
-
-## Results
-
-### Comparison of typical indicators
-
-- Relative change WID 1% and PIP Gini
-
-### Aligning indices
-
-- Show that measure doesn't matter for WID until top 1 (check Palma ratio)
-
-- Check out how much different metrics matter for PIP.
-
------
-
-### Aligning coverage
-
-## Comparing inequality trends
-
-The focus of this paper is on trends. But worth pointing out levels. Inequality is much higher in WID than in PIP.
-
-Again this relates to many things – pretax vs disposable income/consumption.
-Later we will try lining up income concepts more closely.
-But for global coverage we have to pull together these 'worlds' that combine many differences.
+# 2. Comparing the two worlds
 
@@
 
-No reason to expect this given the properties of the two measures.
+Possibly a footnote: No reason to expect this given the properties of the two measures.
 If we take lognormality we see...
@@
 
-- We also do absolute change: a percentage point. This is more intuitive. But somewhat arbitrary – differences in levels affect assessment of differences in trends.
+- We also do absolute change: a percentage point.
+This is more intuitive.
+But somewhat arbitrary – differences in levels affect assessment of differences in trends.
+(Or possibly empahasise this measure if there is little difference to the relative threshold.)
 
@@
     - This we will use to see countries that are particular outliers.
+
+To get an overview: I'll use the unweigted and population weighted average. The first represents change in inequality in the average country.
+
+All broken down by region.
+
+## Inequality since 1993
+
+In @fig-pip-wid-1993-2015-all-obs we plot the level of inequality in 1993 and 2015 against each other, using the inequality measures 'typical' of each dataset.
+
+To help with the interpretation, a band is shown along the 45-degree line.
+Countries falling within this band saw little change in inequality.
+Countries falling above it saw a substantive rise in inequality between the two periods.
+Countries falling below it saw a substantive fall.
+'Substantive' here is defined as a change of more than 2 percentage points in the case of the PIP Gini coefficient, and more than 1 percentage point in the case of the WID top 1% share.
+
+From the visual inspection, in conjunction with @tbl-pip-wid-1993-2015-all-obs-averages.
+In both sets of data, we see a mix of countries with rising and falling inequality.
+In the case of WID top 1% share, far more rises than falls.
+In both we see some convergence – .
+This is largely driven by regional dynamics.
+
+This is confirmed in summary tables.  and this ...
+Tolerance: different thresholds due to different range of data. In the appendix we consider relative thresholds (+/-5%), which makes little difference.
+
+@tbl-pip-wid-1993-2015-all-obs-absolute-count and @tbl-pip-wid-1993-2015-matching-obs-absolute-count do this...
+
+Considering all available observations in each dataset, we see: 
+- 
+-
+- Much bigger rise in the weigted average in WID.
+
+Whilst the number of countries available in the two datasets is very similar, they are not the same countries.
+In particular, LAC is largely absent from WID.
+(Footnote about LAC in WID??
+Are the global results based on linear interpolation?...
+An interpolation that contradicts the survey data?) Also fewer Sub-Saharan African countries.
+In PIP, fewer rich countries in Western Europe and its offshoots.
+
+Importantly, ...
+are not neutral with respect to.The regions where WID is lacking coverage – LAC and SSA – are regions in which inequality was generally falling, according to the PIP Gini coefficient.
+The reverse is also the case: Regions that PIP are lacking – high income countries in W.
+Europe and offshoots – were regions where inequality was generally rising, according to the WID top 1% shares data.
+
+As such, differences in coverage already go some way to explaining the differences between the two datasets. 
+@fig-pip-wid-1993-2015-matching-obs and @tbl-pip-wid-1993-2015-matching-obs-averages show the data, restricting the observations to those available in both datasets.
+This brings the two series closer.
+Explain.
+(Coverage falls)
+
+It is worth emphasising the similarities:
+-
+-
+-
+
+Nevertheless, there are clearly differences in what – even after aligning coverage. Are these about the metric, or other methodological differences? WE will see later on.
+
+```{python}
+# Set-up to compare WID and PIP top 10% share, all observations
+reference_vals = [1993, 2015]
+max_dist_from_refs = 3
+
+selected_vars = [
+        'PIP: Gini',
+        'WID: Top 1pc share – pretax'
+        ]
+
+output_dict = prep_data_and_change_summary(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      region_col = 'region',
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup
+      )
+
+# Store data for OWID upload – prepared further in appendix below
+df_owid_1993_2015 = output_dict['owid_format']
+
+fp = 'data_appendix/figures_and_tables/Compare WID and PIP inequality data 1993 vs 2015 (Joe temp).csv'
+
+df_owid_1993_2015.to_csv(fp, index=False)
+
+```
+
+#### Figures 1-2
+
+::: {.panel-tabset}
+## All available observations
+
+![PIP Gini and WID Top 1% shahre in 1993 vs 21015 – All available observations](images/figures/PIPGini-vs-WIDTop1-1993-vs-2015-01.png){#fig-pip-wid-1993-2015-all-obs}
+
+## Only observations available in both datasets
+![PIP Gini and WID Top 1% shahre in 1993 vs 21015 – Matching observations only](images/figures/PIPGini-vs-WIDTop1-1993-vs-2015-02.png){#fig-pip-wid-1993-2015-matching-obs}
+
+:::
+
+#### Tables 1-2
+::: {.panel-tabset}
+## All available observations
+
+```{python}
+#| label: tbl-pip-wid-1993-2015-all-obs-absolute-count
+#| tbl-cap: Change in inequality 1993 vs 2015 - All availalbe observations
+output_dict['all_sources_False']['absolute_counts']
+
+```
+
+## Only observations available in both datasets
+
+```{python}
+#| label: tbl-pip-wid-1993-2015-matching-obs-absolute-count
+#| tbl-cap: Change in inequality 1993 vs 2015 - Matching observations only
+output_dict['all_sources_True']['absolute_counts']
+
+```
+
+:::
+
+#### Tables 3-4
+
+::: {.panel-tabset}
+## All available observations
+
+```{python}
+#| label: tbl-pip-wid-1993-2015-all-obs-averages
+#| tbl-cap: Unweighted and weighted average in 1993 vs 2015 - All availalbe observations
+output_dict['all_sources_False']['averages']
+
+```
+
+## Only observations available in both datasets
+
+```{python}
+#| label: tbl-pip-wid-1993-2015-matching-obs-averages
+#| tbl-cap: Unweighted and weighted average in 1993 vs 2015 – Matching observations only
+output_dict['all_sources_True']['averages']
+
+```
+
+:::
+
+```{=html}
+<!-- Add icon library -->
+<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">
+
+<a href="data_appendix/figures_and_tables/PIP: Gini vs WID: Top 1pc share – pretax – 1993 to 2015 – allObsFalse.csv" download="Gini_vs_Top1_1993_2015.csv">
+<!-- Auto width -->
+<button class="btn"><i class="fa fa-download"></i> Download full data</button>
+</a>
+```
+
+#### Figures 1-4
+
+```{=html}
+
+<html>
+<body>
+
+<input type="radio" name="1993-vs-2015_radio" value="PIP" checked="checked" onclick="chartSwitchFunction1993_2015_img()">&nbsp;PIP: Gini coefficient
+&nbsp;
+&nbsp;
+<input type="radio" name="1993-vs-2015_radio" value="WID" onclick="chartSwitchFunction1993_2015_img()">&nbsp;WID: Top 1% share
+
+<br>
+<label for="1993-vs-2015_radio_obsCheck">Show only observations available in both datasets:</label> 
+<input type="checkbox" id="1993-vs-2015_radio_obsCheck" onclick="chartSwitchFunction1993_2015_img()">
+
+<img src="images/figures/pip-gini-1993-vs-2015-all-obs.png" id="1993-vs-2015_plot" width="500">
+<br>
+
+</body>
+<script>
+function chartSwitchFunction1993_2015_img() {
+
+  var ele = document.getElementsByName('1993-vs-2015_radio');
+
+  var checkBox = document.getElementById("1993-vs-2015_radio_obsCheck");
+
+  if (checkBox.checked == true){
+    if(ele[0].checked)
+        document.getElementById("1993-vs-2015_plot").src = "images/figures/pip-gini-1993-vs-2015-matching-obs.png";
+
+    if(ele[1].checked)
+        document.getElementById("1993-vs-2015_plot").src = "images/figures/wid-top-1-1993-vs-2015-matching-obs.png";
+
+  } else {
+
+        if(ele[0].checked)
+        document.getElementById("1993-vs-2015_plot").src = "images/figures/pip-gini-1993-vs-2015-all-obs.png";
+
+    if(ele[1].checked)
+        document.getElementById("1993-vs-2015_plot").src = "images/figures/wid-top-1-1993-vs-2015-all-obs.png";
+  }
+}
+
+</script>
+</html>
+
+```
+
+```{python}
+output_dict['all_sources_False']['owid_format']
+
+# prepped_data_dict['prepped_data']
+
+# id_vars_duplicates = df[df.duplicated(subset=['source_var', 'Entity'])]
+
+# id_vars_duplicates
+```
+
+```{python}
+output_dict['all_sources_False']['prepped_data']
+
+```
+
+- Averages + pop coverage
+
+- Country changes (also see which countries strongly disagree in terms of direction)
+
+## Inequality since 1980
+
+```{python}
+# Set-up to compare WID and PIP top 10% share, all observations
+reference_vals = [1980, 2018]
+max_dist_from_refs = 5
+
+selected_vars = [
+        'PIP: Gini',
+        'WID: Top 1pc share – pretax'
+        ]
+
+output_dict = prep_data_and_change_summary(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      region_col = 'region',
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup
+      )
+
+# Store data for OWID upload – prepared further in appendix below
+df_owid_1980_2018 = output_dict['owid_format']
+
+fp = 'data_appendix/figures_and_tables/Compare WID and PIP inequality data 1980 vs 2018 (Joe temp).csv'
+
+df_owid_1980_2018.to_csv(fp, index=False)
+
+```
+
+#### Figures X-Y
+
+```{=html}
+
+<html>
+<body>
+
+<input type="radio" name="1980_2018_radio" value="PIP" checked="checked" onclick="chartSwitchFunction1980_2018()">&nbsp;PIP: Gini coefficient
+&nbsp;
+&nbsp;
+<input type="radio" name="1980_2018_radio" value="WID" onclick="chartSwitchFunction1980_2018()">&nbsp;WID: Top 1% share
+
+<br>
+<label for="obsCheck">Show only observations available in both datasets:</label> 
+<input type="checkbox" id="obsCheck" onclick="chartSwitchFunction1980_2018()">
+
+<iframe id="1980_2018_plot" src="https://ourworldindata.org/grapher/pip-gini-1980-vs-2018-all-obs" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>;
+
+</body>
+<script>
+function chartSwitchFunction1980_2018() {
+
+  var ele = document.getElementsByName('1980_2018_radio');
+
+  var checkBox = document.getElementById("obsCheck");
+
+  if (checkBox.checked == true){
+    if(ele[0].checked)
+        document.getElementById("1980_2018_plot").src = "https://ourworldindata.org/grapher/pip-gini-1980-vs-2018-matching-obs";
+
+    if(ele[1].checked)
+        document.getElementById("1980_2018_plot").src = "https://ourworldindata.org/grapher/wid-top-1-1980-vs-2018-matching-obs";
+
+  } else {
+
+        if(ele[0].checked)
+        document.getElementById("1980_2018_plot").src = "https://ourworldindata.org/grapher/pip-gini-1980-vs-2018-all-obs";
+
+    if(ele[1].checked)
+        document.getElementById("1980_2018_plot").src = "https://ourworldindata.org/grapher/wid-top-1-1980-vs-2018-all-obs";
+  }
+}
+
+</script>
+</html>
+
+```
+
+# 3. Aligning indices
+
+- Show that measure doesn't matter for WID until top 1 (check Palma ratio)
+
+- Check out how much different metrics matter for PIP.
+
+- Do a 1% vs 1% comparison
+
+- Do a Gini vs Gini comparison
+
+All for 1993 and 1980.
+
+Does it make a difference whether we look at Gini or top 10%?
+
+People worry about whether Gini masks what's happening at the top.
+
+Empirically: 
+    - Make a linear model of top 10% on Gini, for each source separately
+    - Make predicted top 10%
+    - Take a look at the R2
+    - Investigate individual countries with very poor predictions
+
+Sub-Saharan Africa: example where the Gini fell in more cases than the top 10%.
+
+# Bridging the gap for particular countries
+
+-----
+
+### Aligning coverage
 
@@
 ### Comparing metrics
-
-Does it make a difference whether we look at Gini or top 10%?
-
-People worry about whether Gini masks what's happening at the top.
-
-Empirically: 
-    - Make a linear model of top 10% on Gini, for each source separately
-    - Make predicted top 10%
-    - Take a look at the R2
-    - Investigate individual countries with very poor predictions
-
-Sub-Saharan Africa: example where the Gini fell in more cases than the top 10%.
 
@@
 # Conclusion
+
+# Appendix
+```{python}
+
+# df_owid = pd.concat([df_owid_1993_2015,df_owid_1980_2018])
+
+```
+
+```{python}
+# Prep data to upload to OWID for charts
+
+# owid_upload_year_values = {}
+# owid_upload_change = {}
+
+# reference_vals = [1993, 2015]
+# max_dist_from_refs = 3
+
+# for all_metrics_requirement in [False, True]:
+
+#     prepped_data = prep_data(
+#           source_dfs = source_dfs,
+#           df_pop = df_pop,
+#           df_regions = df_regions,
+#           region_col = 'region',
+#           selected_vars = selected_vars,
+#           reference_vals = reference_vals,
+#           max_dist_from_refs = max_dist_from_refs,
+#           min_dist_between = 1,
+#           all_metrics_requirement = all_metrics_requirement,
+#           reference_col = 'Year',
+#           group_by_col = 'Entity',
+#           tolerance_lookup = tolerance_lookup,
+#           outlier_cut_off_upper = None
+#           )
+
+#     keep_cols1 = [
+#         'source_var', 
+#         'Entity',
+#         f'Year{reference_vals[0]}', 
+#         f'value{reference_vals[0]}', 
+#         f'Year{reference_vals[1]}', 
+#         f'value{reference_vals[1]}',
+#         'region']
+
+#     keep_cols2 = [
+#         'source_var', 
+#         'Entity',
+#         f'Year{reference_vals[0]}', 
+#         f'value{reference_vals[0]}', 
+#         f'Year{reference_vals[1]}', 
+#         f'value{reference_vals[1]}',
+#         'region']
+
+#     year_values = prepped_data[keep_cols1]
+
+#     year_values = year_values.pivot_table(
+#         index=['Entity', 'region'], 
+#         columns='source_var',
+#         values=[
+#             f'Year{reference_vals[0]}', 
+#             f'value{reference_vals[0]}', 
+#             f'Year{reference_vals[1]}', 
+#             f'value{reference_vals[1]}'
+#         ]).reset_index()
+
+#     year_values.columns = ['_'.join([str(col) for col in cols]) for cols in year_values.columns]
+
+#     year_values = year_values.rename(columns={
+#         "Entity_": "Entity", 
+#         "region_": "region"})
+
+#     year_values = pd.melt(
+#         year_values, 
+#         id_vars=['Entity', 'region'], 
+#         value_vars=[
+#             f'Year{reference_vals[0]}_{selected_vars[0]}',
+#             f'Year{reference_vals[1]}_{selected_vars[0]}',
+#             f'value{reference_vals[0]}_{selected_vars[0]}',
+#             f'value{reference_vals[1]}_{selected_vars[0]}',
+#             f'Year{reference_vals[0]}_{selected_vars[1]}',
+#             f'Year{reference_vals[1]}_{selected_vars[1]}',
+#             f'value{reference_vals[0]}_{selected_vars[1]}',
+#             f'value{reference_vals[1]}_{selected_vars[1]}'
+#         ])
+
+#     year_values[['var', 'source_var']] = year_values['variable'].str.split('_', expand=True)
+#     # Extract year and group information from variable column
+#     year_values[['var', 'reference']] = year_values['var'].str.extract(r'(\D+)(\d+)$')
+#     # owid_format['reference'] = owid_format['reference'].astype(int)
+
+#     #Reshape from long to wide
+#     year_values=pd.pivot(year_values, index=['Entity', 'region', 'reference' ], columns = ['var','source_var'] ,values = 'value').reset_index()
+
+#     owid_upload_year_values[f'compare_{reference_vals[0]}_{reference_vals[1]}_allObs{all_metrics_requirement}'] = year_values
+
+    # change_values = prepped_data[keep_cols2]
+
+    # owid_upload_change[f'compare_{reference_vals[0]}_{reference_vals[1]}_allObs{all_metrics_requirement}'] = change_values
+
+# df_owid_year_values = pd.merge(
+#     owid_upload_year_values[f'compare_{reference_vals[0]}_{reference_vals[1]}_allObsFalse'],
+#     owid_upload_year_values[f'compare_{reference_vals[0]}_{reference_vals[1]}_allObsTrue'],
+#     on=['Entity', reference_col, region_col], how='outer', suffixes=(' – all observations', ' – only matching observations'))
+
+# df_owid = owid_upload_year_values[f'compare_{reference_vals[0]}_{reference_vals[1]}_allObsFalse']
+# # df_owid = pd.merge(df_owid_1993_2015,df_owid_1980_2018, on=[group_by_col, reference_col, region_col], how='outer', suffixes=(' – all observations', ' – only matching observations'))
+
+# # df_owid = pd.concat([df_owid_1993_2015,df_owid_1980_2018])
+
+# fp = 'data_appendix/figures_and_tables/Compare WID and PIP inequality data (Joe temp).csv'
+
+# df_owid.to_csv(fp, index=False)
+
+```
+
+```{python}
+#  # Reshape the data for OWID upload
+
+#     year_values = pd.melt(prepped_data, id_vars=['source_var', 'Entity', 'region' ], value_vars=[f'Year{reference_vals[0]}', f'Year{reference_vals[1]}', f'value{reference_vals[0]}', f'value{reference_vals[1]}'])
+
+#     # Extract year and group information from variable column
+#     year_values[['var', 'reference']] = year_values['variable'].str.extract(r'(\D+)(\d+)$')
+
+#     year_values['reference'] = year_values['reference'].astype(int)
+
+#     #Reshape from long to wide
+#     year_values=pd.pivot(year_values, index=['source_var', 'Entity', 'region', 'reference' ], columns = 'var' ,values = 'value').reset_index()
+
+#     # Drop the original reference column
+#     year_values = year_values.drop('reference', axis=1)
+
+#     # Pivot the melted dataframe wider – one value column per source_var
+#     owid_format = owid_format.pivot_table(index=[group_by_col, region_col, reference_col], columns='source_var', values='value').reset_index()
+
+#     summaries_dict['owid_format'] = owid_format
+
+#     output_dict[f'all_sources_{all_metrics_requirement}'] = summaries_dict
+
+#     # Merge the all_sources- True and False data for OWID format
+#     owid_format = pd.merge(output_dict['all_sources_False']['owid_format'], output_dict['all_sources_True']['owid_format'], on=[group_by_col, reference_col, region_col], how='outer', suffixes=(' – all observations', ' – only matching observations'))
+
+#     owid_format = owid_format.rename(columns={group_by_col: "country", reference_col: "year"})
+
+#     # Reorder columns
+#     cols = list(owid_format.columns) # Get the list of column names
+#     cols.remove('country') # Remove col1 from the list
+#     cols.remove('year') # Remove col2 from the list
+#     cols = ['country', 'year'] + cols # Add col1 and col2 to the front of the list
+#     owid_format = owid_format[cols] # Reorder columns in the DataFrame
+
+#     owid_format['year'] = owid_format['year'].astype(int) # Year needs to be integer for OWID
+
+```
+
+## Data coverage
+
+## PIP Gini vs WID Top 1% share
+
+### 1993-2015

### 2023-04-19 PhD_pages/Paper_1/data_appendix/G_functions_for_figures_and_tables.py

@@ -1,12 +1,3 @@
-# %%
-# ---
-# title: "Functions for preparing figures and tables"
-# execute: 
-#   enabled: false
-# ---
-
-# %% [markdown]
-# The code here is duplicated in as a .py file – from which the functions are imported in data appendix H.
 
 # %%
 
@@ -228,6 +219,7 @@ def prep_data(
       source_dfs,
       df_pop,
       df_regions,
+      region_col,
       selected_vars,
       reference_vals,
       max_dist_from_refs,
@@ -329,6 +321,7 @@ def prep_data(
     prepped_data = prepped_data.drop('source_count', axis=1)
 
     # Add in region classification
+    df_regions = df_regions[[group_by_col, region_col]]
     prepped_data = pd.merge(prepped_data, df_regions, how = 'left')
 
     # Add in population data
@@ -346,8 +339,17 @@ def prep_data(
 
     prepped_data = prepped_data.rename(columns={'population':f'pop{reference_vals[1]}'})
 
-    # Add a tolerance column - demarking substantial rise or fall
-    # Uses a dictionary passed as an argument to this function to
+
+
+
+    # Calculate the change between the two ref periods
+    prepped_data['change'] = (prepped_data[f'value{reference_vals[1]}'] - prepped_data[f'value{reference_vals[0]}'])
+
+
+    # ABSOLUTE TOLERANCE: 
+    # Add a tolerance column – the change the demarks a substnatial rise or fall.
+    # Here this is measured in terms of absolute change.
+    # It uses a dictionary passed as an argument to this function to
     # search for vars with particualr strings ('Gini') and the sub in
     # the tolerance specified in the dictionary
     for i in range(len(tolerance_lookup['var_search_term'])):
@@ -355,9 +357,32 @@ def prep_data(
         prepped_data.loc[prepped_data['source_var']\
             .str.contains(
                 tolerance_lookup['var_search_term'][i]
-                ), 'tolerance'] = tolerance_lookup['var_tolerance'][i]
+                ), 'absolute_tolerance'] = tolerance_lookup['var_tolerance'][i]
 
 
+    # RELATIVE TOLERANCE: 
+    # Add a tolerance column – the change the demarks a substnatial rise or fall.
+    # Here this is measured in terms of relative change.
+    # It uses a single value [0-1] irrespective of the variable.
[…]

### 2023-04-26 PhD_pages/Paper_1/paper1.qmd
Paper 1 : drafting

@@
 
-READING LIST:
-Bourguignon (2015)
-The share of wages in total GDP declined in a majority of countries
-(91 out of 133 with data) from 1995 to 2014 (ILO, 2016).
+```{python}
+# READING LIST:
+# Bourguignon (2015)
+# The share of wages in total GDP declined in a majority of countries
+# (91 out of 133 with data) from 1995 to 2014 (ILO, 2016).
 
- Other authors
-question the pre-eminence of globalization as the main driver of global inequality, noting
-that inequality trends differ across countries at similar levels of development and that
-are equally exposed to trade (Ravallion, 2018; Corlett, 2016). Ravallion (2018) stresses
-the vital role that natio
+#  Other authors
+# question the pre-eminence of globalization as the main driver of global inequality, noting
+# that inequality trends differ across countries at similar levels of development and that
+# are equally exposed to trade (Ravallion, 2018; Corlett, 2016). Ravallion (2018) stresses
+# the vital role that natio
 
-Key point:  the share of income going to both the top 1 per cent and the bottom 50 per
-cent has increased in many countries in recent decades.
+# Key point:  the share of income going to both the top 1 per cent and the bottom 50 per
+# cent has increased in many countries in recent decades.
 
-In 2021, while the very richest individuals buy trips to space, roughly three billion people are unable to afford a healthy diet.
+# In 2021, while the very richest individuals buy trips to space, roughly three billion people are unable to afford a healthy diet.
 Even  rich countries secured COVID vaccines for their entire population, low-income had barely begun.
@@
 
-Such
+# Such
 
-"Inequality defines our time," states the UN Secretary-General.^[United Nations Secretary-General António Guterres made the comment in the 18th Nelson Mandela Annual Lecture in July 2020.
+# "Inequality defines our time," states the UN Secretary-General.^[United Nations Secretary-General António Guterres made the comment in the 18th Nelson Mandela Annual Lecture in July 2020.
 See https://www.nelsonmandela.org/news/entry/annual-lecture-2020-secretary-general-guterress-full-speech.]
 
-Just as the demand for data to monitor inequality globally is greater than ever, so is its supply.
+# Just as the demand for data to monitor inequality globally is greater than ever, so is its supply.
 
-In part, this increased concern is a reflection of events - of the severe and uneven impacts of the 2007-8 Global Financial Crisis, the COVID-19 pandemic and recent sharp rises in the cost of living.
+# In part, this increased concern is a reflection of events - of the severe and uneven impacts of the 2007-8 Global Financial Crisis, the COVID-19 pandemic and recent sharp rises in the cost of living.
+
+# That research, and *WID* in particular, has shaped   Policy and advocacy been used to draw attention to the high and, in many cases, growing concentration of income among the very richest in society.
+
+# Thomas Piketty's *Capital in the 21st Century* provides a model in which, through capital accumulation, rising inequality may be understood as fundamental to the nature of capitalism.
+Such a hypothesis gains substantial motivation from the upward trends apparent in long-run top income series presented in the book and subsequently incorporated within the *WID* database.
+It starkly contrasts with orthodox accounts emphasising more contingent processes such as wage premia driven by some combination of skills-biased technology or globalisation.
+
+# Extensive coverage of low- and middle-income countries which, depending on the country and year relates to either household disposable income or consumption.
+
+# A distinctive innovation of this approach has been the incorporation of tax data.
+This allows a better view of the top of the pre-tax income distribution, and in particular of income earned on capital, which is often poorly captured by survey data [@lustig_missing_2019].
+It has also enabled the estimation of inequality for much earlier periods than possible with survey data, albeit with lower geographical coverage.
+
+```
 
@@
 
-Economic inequalities around the world are enormous. As the very richest individuals buy trips to space, roughly three billion people are unable to afford a healthy diet.
+Economic inequalities around the world are vast.
+The world's richest 1% enjoy an average income that is around 150 times that of the poorer half of the world's population [@chancel_world_2021].
+While the richest may afford private flights to space, roughly three billion people are unable to afford even just the cost of a healthy diet [@herforth_cost_2020].
 
-Such stark disparities have gained more attention in recent decades, among policimakers, academics and the public alike.
+Such stark disparities have become an increasing focal point among policimakers, academics and the public alike, emphasised by the uneven impacts of the 2007-8 financial crisis, the COVID-19 pandemic, and recent jumps in inflation.
 "Inequality defines our time," according to the UN Secretary-General.^[United Nations Secretary-General António Guterres made the comment in the 18th Nelson Mandela Annual Lecture in July 2020.
@@
 
-In part at least, this increased concern reflects an increase in the *availability* of distributional data.
-New and expanded sources of data have shed new light on extent of inequality around the world and how it is changing.
-This has enabled concerns about high or rising inequality to be expressed in new ways, formulated more precisely and subsequently monitored.
-The supply and demand for global data on inequality have been mutually reinforcing.
+In addition to recent events, the increased attention paid to economic inequalities is surely owed in part to the increased *availability* of distributional data.
+New and expanded sources of data have shed new light on the extent of inequality around the world and how it is changing.
+This has enabled concerns about high or rising inequality to be voiced with more clarity and force.
+The supply and demand for global data on inequality have in this way been mutually reinforcing.
 
@@
 One is the World Bank's harmonized collection of household survey data, published online through its [*Poverty and Inequality Platform (PIP)*](https://pip.worldbank.org/).
-Another is the collection of inequality estimates, most notably top income shares, brought together at the [*World Inequality Database (WID)*](https://wid.world/).^[TODO: PIP used to be Povcal.
-It draws upon the work of various regional teams.
-WID used to be WTID.]
+Another is the [*World Inequality Database (WID)*](https://wid.world/), which emerged out of the burgeoning literature on top incomes th incorporation of tax data [@atkinson_top_2014].
 
-These two sources are the only regularly-updated databases that provide global coverage of distributional data on a relatively comparable basis.^[EDIT: Both datasets suffer important limitations with respect to comparability acrosss time and location.
+These two sources are the only regularly-updated databases providing global coverage of distributional data on a relatively comparable basis.^[TODO: Both datasets suffer important limitations with respect to comparability acrosss time and location.
 These will be discussed in the data section below.
@@
 Both have played substantial agenda-setting roles in academic research, policy and public debate.
+Their use and impact, however, has been confined to largely separate arenas, reflecting differences in the concepts of inequality they employ and the coverage they provide.
 
-The World Bank data is especially prominent within international development policy, notably being used to monitor many of the UN Sustainable Development Goals.
-Academic research on global inequality has also largely relied on this data  [@Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk].
-*[TODO: Check and more refs.]*.
+The *PIP* data has typically been the favoured choice for investigating the whole distribution of welfare across households, particularly at a global scale.
+() The *WID* data at the top, focus on role of capital incomes.
 
-The WID database emerged along with the burgeoning literature on top incomes, acting as a central repository and, increasingly, coordinating institution. The 1%
+Moreover, the many methodological differences separating the two datasets prevent users combining the insights that they each offer individually into a coherent empirical view informed by both.
+In choosing datasets, we are simultaneously choosing underlying data source, methods, welfare concept, and coverage.
+The two datasets represent distinct and, to a large extent, incommensurable 'worlds' of inequality measurement.
 
-## Two worlds of inequality measurement
+This disconnect is all the more pertinent because they disagree on basics. Firstly. Secondly.
+
+Very confusing terrain for people to understand. We don't know which of the two worlds we live in. Depends on a choice about which little can be determined.
+
+This paper is the first of two which will try to ocnnect them up.
+
+Section 2. Section 3...
+
+### Two 'worlds' of inequality measurement
+
+The two datasets differ along a range of dimensions, including the underlying sources relied on, the measures of welfare used, the summary indicators commonly favoured, and the temporal and geographic coverage available.
+
+These differences come largely as a bundle.
+Choosing between the *WID* or *PIP* data implies not only a technical choice between data sources and methodolodies, but also a more fundamental choice as to which concept of inequality to consider, for which places and time periods.
+This has a led to a degree of specialisation in their use.
+
+The *PIP* data has typically been the favoured choice for investigating the whole distribution of welfare across households, particularly at a global scale [@Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @milanovic_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk].
+This includes the World Bank's own research and monitoring activties concerning global poverty, inequality and growth across the distribution [see for example @world_bank_poverty_2016; @world_bank_poverty_2022].
+The dataset is also relied on by the United Nations to set and monitor many of the targets in this area that form part of its *Sustainable Development Goals*.
+
+The *WID* data has more commonly been used to understand top dynamics before taxes.
+Notwithstanding substantial efforts to global, high income countries due to the data required.
+in which today's rich countries are overrepresented .
+In policy, advocacy and public debate, the data used to highlight the very large and, in many cases, growing advantage enjoyed by the richest individuals in these societies.
+In academia, top incomes data long-run processes fundamental – capital accumulation.[@atkinson_top_2007; @piketty_capital_2014]  On political economy channels, social conflict between capital and labour, wage bargaining norms and the interaction with marginal tax rates.
+More generally the data have led to a fundamental revaluation of the causes of inequality that move beyond orthodox accounts given purely in terms of marginal productivities of workers with heterogeneous skills or education [@alvaredo_top_2013; @piketty_capital_2014-1].
+
+Thus whilst both the *PIP* and *WID* datasets have played substantial agenda-setting roles in academic research, policy and public debate, this has tended to occur in largely separate arenas.
+Moreoever, no way to join up the insights they offer.
+Very confusing terrain for people to understand.
+We are left with two distinct and, to a large extent, incommensurable 'worlds' of inequality measurement.
+
+Recent extensions have brought down some of these barriers. PIP now includes high-income countries. WID now global coverage, and post-tax income concepts. In
+
+At the same time, these developments in the two datasets have made this disconnect between them all the more apparent. Fundamental differences on the basics. Two in particular.
+
+1
+2
+
+Whilst for individual countries, the connections may be well understood, globally the value of the different assumptions is not really known. We have to pick blindly.
+
+### Connecting up the worlds
+
+This is the first of two papers that seeks to join up.
+
+The dataset is used for the Bank's own research and monitoring activities which, whilst traditionally focussing on absolute poverty, have more recently encompassed trends in inequality, relative poverty and growth across the distribution.
+Notably, the data is alo used to set and monitor many of the targets, on both poverty and inequality, that form part of the UN's Sustainable Development Goals.
+More generally, the source has been indispensible for research taking a global view on inequality [@Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @milanovic_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk].
+
+The *WID* database emerged out of the burgeoning literature on top incomes [@atkinson_top_2014] and has come to play a central and, to some extent, coordinating role for this research agenda.
+The key innovation of the 'top incomes' approach has been the incorporation of tax data, as either an alternative or complement to the more orthodox reliance on household survey data.
+The primary advantage is that tax data can provide a better view of the top of the income distribution, which is often poorly captured by survey data [@lustig_missing_2019].
+However, the choice of data source has typically brought with it a number of other differences that mark out the 'top incomes' approach.
+This includes a particular focus on the top of the distribution (for which tax data is more universally available), and on income measured before tax (in terms of which the underlying data is gathered).
+This contrasts with more orthodox approaches to inequality measurement in which disposable income has tended to be seen as the more relevant measure of welfare [@canberra_group_canberra_2011], and greater attention is given to measures of overall inequality, such as the Gini coefficient, or to relative poverty.
+
+Another benefit of tax data is the much longer timeframes it is sometimes available for.
+The estimates gathered in *WID* provide a long-run perspective on the evolution of inequality over the twentieth century, albeit for a more limited geography in which today's rich countries are somewhat overrepresented [@atkinson_top_2007; @piketty_capital_2014].
+
+Moreover, the top incomes methodology carried forward by *WID* developed with a distinctive set of questions concerning the *causes* of inequality in mind.
+Particular attention is given to the distribution of wealth and capital incomes as drivers of personal income inequality, in contrast to orthodox accounts focussed purely on the dispersion of wages [@alvaredo_top_2013; @piketty_capital_2014-1].
+This ...
+hidden from view given the concentration of capital incomes at the top, captured poorly by survey data.
+
+### Two 'worlds' of inequality measurement
+Tw worlds – methods, conceptions of inequality, time periods and geographic coverage and research questions come bundled.
+
+Always unsatisfying. Very difficult to unpick. Peple do it for individual countries..
+
+A wholeslae comparison was generally avoided due to the very different coverage. Used in parallel.
+
+As the databases have developed and expanded, however, this division of labour between the two datasets has become increasingly blurred.
+PIP includes high-income countries.
+WID coverage is global.
+DINA focus on comparability.
+Increasingly, the datasets are used to answer the same range of questions, speak to the same world.
+
+Where they now meet, the result is very confusing, one that, despite the great strides forward, leaves uncertainty around the most basic questions.
+
+This is most visible in two fundamental disagreements between the two 'worlds'.
+
+Firstly, they give very different views of how ubiquitous a phenomenon rising inequality is within countries around the world.
+The view from *World Inequality Database* and its associated global report is that inequality has risen 'nearly everywhere' in recent decades [@alvaredo_world_2018; @chancel_world_2021].
+In contrast, the World Bank data and its associated biannual report offers a much more mixed picture – in which inequality is rising and falling in roughly equal numbers of countries.
+
+Secondly, the datasets paint as to how such inequalities *within* countries compare to the differences in average incomes seen *between* richer and poorer countries.
+
+Most analyses of global income inequality to date have relied heavily on the *PIP* data [@Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @milanovic_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk].^[TODO: Add footnote explaining transition from Povcal] *[TODO: Check and more refs.]*.
+A consistent result from this literature is that inequalities within countries remain small in comparison to the much larger differences seen between countries.
+This is despite the convergence seen in recent decades between countries, given higher economic growth rates across low- and middle-income countries.
+According to such research, a person's income remains largely determined by the country in which they happen to have been born [@milanovic_global_2015].
+
+More recent analysis of the WID data flips this established result on its head. Chancel... estimate xx.
+
+The datasets are enormous efforts, involving the work of hundreds of researchers.
+They are hugely valuable offering an unprecidented perspective on global inequality and how it is evolving.
+And yet the disconnect between the two devalues them both.
+Amidst the wealth of data, basic facts are not known.
+And the terrain is incredibly confusing to navigate for policimakers and the concerned public.
+
+### Connecting up the worlds
+
+This paper, along with a companion paper, seeks to join up these two worlds. This first paper picks up the first question – within country. A companion paper considers global inequality.
+
+Section 2 of this paper introduces the two worlds.
+Section 3 compares within country trends between two periods.
+It begins with the inequlaity metrics for which these datasets are most associated, and in which the worlds fall furthest apart: Gini, Top 1% share.
+Already
+Section 4 aligns coverage. It is not neutral. Already this goes some way to explaining the gap between the two datasets.
+Section 5 aligns inequality indices. [Summary.]
+Section 5 considers the role of welfare metrics and specific data quality issues for a subset of countries where there is most disagreement.
+
+Summary of main conclusion!?
+
+# 2. Two worlds of inequality measurement
 
@@
 ### The disconnect between these datasets devaules them both
+
+These datasets are incredibly valuable – playing to motivate action.
+The example in the opening sentence above illustrates this well.
+It is only thanks to the WID data that we can say X.
+It is only thanks to the PIP data that we can say that three billion cannot afford a healthy diet.
+
+Often it is presented that there is a division of labour between these dataset – WID is for inequality, especially at the top.PIP is for poverty.
+
+But the point is that it is only one world that we live in. The disconnect is not something we can live with, but rather presents a challenge to both datasets.
 
@@
 The two measures selected – Gini coefficient of disposable income/consumption, and the top 1% share of pre-tax income – are where the PIP and WID worlds respectively are at their furthest apart.
-Yet in terms of broad global trends, there is much that even these very different measures can agree on: - x, - y, - z.
+Yet in terms of broad global trends, there is much that even these very different measures can agree on:
+- x
+- y
+- z.
 
@@
 
-The rest of the paper is about understanding where those differences emerge.
+Subsequent sections attempt to bridge the gaps in order to learn where those differences emerge.
+
+- Section 2: Aligning coverage. This reduces the gap already. Coverage of countries is not... with respect to inequality trends.
 
@@
 
-- We look at different income concepts for the small number of coutnries for which such data exists.
+- Welfare concept.
+We look at different income concepts for the small number of coutnries for which such data exists.
 To this end we also include data from the Luxembourg Income Study.
@@
 Although both institutions continue to take many steps to provide transparency and reproducibility to their estimates, they remain a black box. Makes it impossible to do this lining up.
+
 In many particular contexts – where detailed lining up is possible – there relative strengths and weaknesses may be clear.
@@
 
-# 2. Comparing the two worlds
+# 3. Comparing the two worlds
 
-What we're trying to do with triangulation is see whether they show a qualitively similar picture of whether, and how rapidly, inequality was rising or falling.
-
-But what would 'similar' be for Gini and top 1. Not a 1 point move.
-
-The main way of operationalising 'similar' is to look at Relative change: a percentage, rather than percentage point. We can put some threshold and see if there were rises or falls.
-
-Possibly a footnote: No reason to expect this given the properties of the two measures.
-If we take lognormality we see...
-(MAke a chart in appendix - where we plot Gini against top 10 and top 1% assuming lognormality, and then add actual data for a particular country.) How unlognormal the top is would be another way to judge similarity.
-Rather than doing this, we bridge by aligning the indices.
-
-- We also do absolute change: a percentage point.
-This is more intuitive.
-But somewhat arbitrary – differences in levels affect assessment of differences in trends.
-(Or possibly empahasise this measure if there is little difference to the relative threshold.)
-
-- A third way is residuals from a regression of the two measures
-    - This we will use to see countries that are particular outliers.
-
-To get an overview: I'll use the unweigted and population weighted average. The first represents change in inequality in the average country.
-
-All broken down by region.
+We begin by comparing the metrics where the worlds are furthest apart. PIP Gini. Wid Top 1% share.
 
@@
 
-In @fig-pip-wid-1993-2015-all-obs we plot the level of inequality in 1993 and 2015 against each other, using the inequality measures 'typical' of each dataset.
+@fig-pip-wid-1993-2015-all-obs comoares the level of inequality, as measured by the two metrics, in or around 1993 and 2015.
+A band is shown along the 45-degree line.
+Countries falling within this band saw little change in inequality.
+For countries above the band, inequality in 2015 was substantively higher than in 1993.
+Countries falling below the band instead saw a substantive fall.
 
-To help with the interpretation, a band is shown along the 45-degree line.
-Countries falling within this band saw little change in inequality.
-Countries falling above it saw a substantive rise in inequality between the two periods.
-Countries falling below it saw a substantive fall.
-'Substantive' here is defined as a change of more than 2 percentage points in the case of the PIP Gini coefficient, and more than 1 percentage point in the case of the WID top 1% share.
+@tbl-pip-wid-1993-2015-all-obs-absolute-count provides a summary of the .
+A 'substantive' change here is defined as a movement of more than 2 percentage points in the case of the PIP Gini coefficient, and more than 1 percentage point in the case of the WID top 1% share.
+[TODO: add table in appendix showing that using relative thresholds (+/-5% across both metrics) makes little difference.]
 
-From the visual inspection, in conjunction with @tbl-pip-wid-1993-2015-all-obs-averages.
-In both sets of data, we see a mix of countries with rising and falling inequality.
-In the case of WID top 1% share, far more rises than falls.
-In both we see some convergence – .
-This is largely driven by regional dynamics.
+According to PIP estimates for the Gini coefficient, inequality was rising and falling in roughly an equal number of countries over the period (broadly replicating the analysis presented in @world_bank_poverty_2016).
 
-This is confirmed in summary tables.  and this ...
-Tolerance: different thresholds due to different range of data. In the appendix we consider relative thresholds (+/-5%), which makes little difference.
+@tbl-pip-wid-1993-2015-all-obs-averages presents averages for the two reference years.
+The unweighted average Gini coefficient within the PIP data fell somewhat over the period, from 39.8 to 30.0 points.
+As we see from @fig-pip-wid-1993-2015-all-obs (top panel), contributing to this average decline are a number of countries showing very substantial falls – in some cases declines of 10 points or more.
+These are seen mostly in Sub-Saharan Africa, in addition to a number of Latin American and Eastern European and Central Asian countries.
+[TODO: Mention particular questions around survey comparability in SSA – this is a subject that will be taken up in section X.]
 
-@tbl-pip-wid-1993-2015-all-obs-absolute-count and @tbl-pip-wid-1993-2015-matching-obs-absolute-count do this...
+WID top 1% share estimates show decidedly more cases of rising inequality – in roughly two-thirds of countries.The average, again, is pulled down by some particularly substantial falls, which in the WID data are more exclusively concentrated in Sub-Saharan Africa.
+This leaves the average top 1% share remarkably stable over the period, showing a modest rise from 14.5% to 14.8%.
 
-Considering all available observations in each dataset, we see: 
-- 
--
-- Much bigger rise in the weigted average in WID.
+In both datasets, we see that inequality trends were less benign in particularly populous countries.
+This is apparent from a comparison of the unweighted and population-weighted averages (@tbl-pip-wid-1993-2015-all-obs-averages).
+The moderate fall seen in the unweigted average PIP Gini coefficient is no longer visible in the weighted average.
+In the WID data, in contrast to the unweighted average, the weighted average rose by more than a third, from 13% to 17%.
 
-Whilst the number of countries available in the two datasets is very similar, they are not the same countries.
-In particular, LAC is largely absent from WID.
-(Footnote about LAC in WID??
-Are the global results based on linear interpolation?...
-An interpolation that contradicts the survey data?) Also fewer Sub-Saharan African countries.
-In PIP, fewer rich countries in Western Europe and its offshoots.
+As @fig-pip-vs-wid-1993-2015 illustrates, it is also in particularly populous countries where we see some of the starkest differences between the two datasets.
+The figure plots the change in the two metrics, along with the line of best fit.
+[TODO: 
+- Plot also for relative change (%, not %-pt).
+- But what should we expect? Plot the relationship between Gini change and top 1% share change assuming lognormality. Possibly this makes more sense as part of section 3 on inequality indices.]
 
-Importantly, ...
-are not neutral with respect to.The regions where WID is lacking coverage – LAC and SSA – are regions in which inequality was generally falling, according to the PIP Gini coefficient.
-The reverse is also the case: Regions that PIP are lacking – high income countries in W.
-Europe and offshoots – were regions where inequality was generally rising, according to the WID top 1% shares data.
+This includes India, United States, for which far more substantial rises are seen in the WID top 1% share.
+In Brazil and Nigeria, substantial falls present in the PIP Gini are not reflected in the WID data.
+Mexico and Russia stand out in particular in seeing large movements in opposite directions within the two datasets.
 
-As such, differences in coverage already go some way to explaining the differences between the two datasets. 
-@fig-pip-wid-1993-2015-matching-obs and @tbl-pip-wid-1993-2015-matching-obs-averages show the data, restricting the observations to those available in both datasets.
-This brings the two series closer.
-Explain.
-(Coverage falls)
-
-It is worth emphasising the similarities:
--
--
--
-
-Nevertheless, there are clearly differences in what – even after aligning coverage. Are these about the metric, or other methodological differences? WE will see later on.
+Regionally [TODO: summarise regional trends].
 
@@
 
-```{=html}
-<!-- Add icon library -->
-<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">
+```{python}
+import statsmodels.formula.api as smf
+from matplotlib import pyplot as plt
 
-<a href="data_appendix/figures_and_tables/PIP: Gini vs WID: Top 1pc share – pretax – 1993 to 2015 – allObsFalse.csv" download="Gini_vs_Top1_1993_2015.csv">
-<!-- Auto width -->
-<button class="btn"><i class="fa fa-download"></i> Download full data</button>
-</a>
-```
+change_compare = output_dict['all_sources_True']['prepped_data'][['source_var', 'Entity', 'change', 'region']]
 
-#### Figures 1-4
+change_compare = change_compare.pivot_table(index=['Entity', 'region'], columns='source_var', values='change').reset_index()
 
-```{=html}
+change_compare = change_compare.rename(columns={"WID: Top 1pc share – pretax": 'change_WID', 'PIP: Gini': 'change_PIP'})
 
-<html>
-<body>
+# Save the data to upload to WID for plot
+change_compare_owid = change_compare
+change_compare_owid['year'] = 2000
+change_compare_owid = change_compare_owid.rename(columns={'Entity': 'country'})
 
-<input type="radio" name="1993-vs-2015_radio" value="PIP" checked="checked" onclick="chartSwitchFunction1993_2015_img()">&nbsp;PIP: Gini coefficient
-&nbsp;
-&nbsp;
-<input type="radio" name="1993-vs-2015_radio" value="WID" onclick="chartSwitchFunction1993_2015_img()">&nbsp;WID: Top 1% share
+fp = 'data_appendix/figures_and_tables/Compare WID and PIP CHANGE in inequality 1993 vs 2015 (Joe temp).csv'
 
-<br>
-<label for="1993-vs-2015_radio_obsCheck">Show only observations available in both datasets:</label> 
-<input type="checkbox" id="1993-vs-2015_radio_obsCheck" onclick="chartSwitchFunction1993_2015_img()">
+change_compare_owid.to_csv(fp, index=False)
 
-<img src="images/figures/pip-gini-1993-vs-2015-all-obs.png" id="1993-vs-2015_plot" width="500">
-<br>
-
-</body>
-<script>
-function chartSwitchFunction1993_2015_img() {
-
-  var ele = document.getElementsByName('1993-vs-2015_radio');
-
-  var checkBox = document.getElementById("1993-vs-2015_radio_obsCheck");
-
-  if (checkBox.checked == true){
-    if(ele[0].checked)
-        document.getElementById("1993-vs-2015_plot").src = "images/figures/pip-gini-1993-vs-2015-matching-obs.png";
-
-    if(ele[1].checked)
-        document.getElementById("1993-vs-2015_plot").src = "images/figures/wid-top-1-1993-vs-2015-matching-obs.png";
-
-  } else {
-
-        if(ele[0].checked)
-        document.getElementById("1993-vs-2015_plot").src = "images/figures/pip-gini-1993-vs-2015-all-obs.png";
-
-    if(ele[1].checked)
-        document.getElementById("1993-vs-2015_plot").src = "images/figures/wid-top-1-1993-vs-2015-all-obs.png";
-  }
-}
-
-</script>
-</html>
+# Initialise and fit linear regression model using `statsmodels`
+model = smf.ols('change_WID ~ change_PIP', data=change_compare)
+model = model.fit()
+# model.params
 
@@
 
-```{python}
-output_dict['all_sources_False']['owid_format']
-
-# prepped_data_dict['prepped_data']
-
-# id_vars_duplicates = df[df.duplicated(subset=['source_var', 'Entity'])]
-
-# id_vars_duplicates
-```
-
-```{python}
-output_dict['all_sources_False']['prepped_data']
-
-```
-
-- Averages + pop coverage
-
-- Country changes (also see which countries strongly disagree in terms of direction)
+![Change in PIP Gini vs WID Top 1% share in 1993 vs 2015](images/figures/Change-1993-2015.svg){#fig-pip-vs-wid-1993-2015}
 
@@
 
-```{=html}
+#### Tables 3-4
 
-<html>
-<body>
+::: {.panel-tabset}
+## All available observations
 
-<input type="radio" name="1980_2018_radio" value="PIP" checked="checked" onclick="chartSwitchFunction1980_2018()">&nbsp;PIP: Gini coefficient
-&nbsp;
-&nbsp;
-<input type="radio" name="1980_2018_radio" value="WID" onclick="chartSwitchFunction1980_2018()">&nbsp;WID: Top 1% share
-
-<br>
-<label for="obsCheck">Show only observations available in both datasets:</label> 
-<input type="checkbox" id="obsCheck" onclick="chartSwitchFunction1980_2018()">
-
-<iframe id="1980_2018_plot" src="https://ourworldindata.org/grapher/pip-gini-1980-vs-2018-all-obs" loading="lazy" style="width: 100%; height: 600px; border: 0px none;"></iframe>;
-
-</body>
-<script>
-function chartSwitchFunction1980_2018() {
-
-  var ele = document.getElementsByName('1980_2018_radio');
-
-  var checkBox = document.getElementById("obsCheck");
-
-  if (checkBox.checked == true){
-    if(ele[0].checked)
-        document.getElementById("1980_2018_plot").src = "https://ourworldindata.org/grapher/pip-gini-1980-vs-2018-matching-obs";
-
-    if(ele[1].checked)
-        document.getElementById("1980_2018_plot").src = "https://ourworldindata.org/grapher/wid-top-1-1980-vs-2018-matching-obs";
-
-  } else {
-
-        if(ele[0].checked)
-        document.getElementById("1980_2018_plot").src = "https://ourworldindata.org/grapher/pip-gini-1980-vs-2018-all-obs";
-
-    if(ele[1].checked)
-        document.getElementById("1980_2018_plot").src = "https://ourworldindata.org/grapher/wid-top-1-1980-vs-2018-all-obs";
-  }
-}
-
-</script>
-</html>
+```{python}
+#| label: tbl-pip-wid-1980-2018-all-obs-averages
+#| tbl-cap: Unweighted and weighted average in 1980 vs 2018 - All availalbe observations
+output_dict['all_sources_False']['averages']
 
 ```
+
+## Only observations available in both datasets
+
+```{python}
+#| label: tbl-pip-wid-1980-2018-matching-obs-averages
+#| tbl-cap: Unweighted and weighted average in 1980 vs 2018– Matching observations only
+output_dict['all_sources_True']['averages']
+
+```
+
+:::
+
+## Similarities and differences
+
+Summary, across both periods. (Reflect this back in the intro.)
+
+# 3. Aligning coverage
+
+Whilst the number of countries available in the two datasets is very similar, they are not the same countries.
+In particular, LAC is largely absent from WID.
+(Footnote about LAC in WID??
+Are the global results based on linear interpolation?...
+An interpolation that contradicts the survey data?) Also fewer Sub-Saharan African countries.
+In PIP, fewer rich countries in Western Europe and its offshoots.
+
+Importantly, ...
+are not neutral with respect to.The regions where WID is lacking coverage – LAC and SSA – are regions in which inequality was generally falling, according to the PIP Gini coefficient.
+The reverse is also the case: Regions that PIP are lacking – high income countries in W.
+Europe and offshoots – were regions where inequality was generally rising, according to the WID top 1% shares data.
+
+As such, differences in coverage already go some way to explaining the differences between the two datasets. 
+@fig-pip-wid-1993-2015-matching-obs and @tbl-pip-wid-1993-2015-matching-obs-averages show the data, restricting the observations to those available in both datasets.
+This brings the two series closer.
+Explain.
+(Coverage falls)
+
+Summarise remaining differences.
+
+Plot – Just like figure 1, but show only the data points missing from WID in the PIP pannel. And visa versa.
 
@@
 
+  - A table, using relative threshold (+/-5%) Different measures: Gini, Palma, top 10, top 1.
+
+  - Regressions – all pairs of the four measures. See the r2 and note outliers for each.
+
 - Check out how much different metrics matter for PIP.
 
-- Do a 1% vs 1% comparison
+  - Same table and regressions.
 
-- Do a Gini vs Gini comparison
+- For each of the 4 measures, do a direct comparison across sources.
 
-All for 1993 and 1980.
+  - For both 1993 and 1980: matching observations only and only global results.
 
-Does it make a difference whether we look at Gini or top 10%?
+(Sub-Saharan Africa: example where the Gini fell in more cases than the top 10%.?)
 
-People worry about whether Gini masks what's happening at the top.
+## Similarities and differences
 
-Empirically: 
-    - Make a linear model of top 10% on Gini, for each source separately
-    - Make predicted top 10%
-    - Take a look at the R2
-    - Investigate individual countries with very poor predictions
-
-Sub-Saharan Africa: example where the Gini fell in more cases than the top 10%.
+Summarise where we're at after aligning the inequality metric.
 
 # Bridging the gap for particular countries
+
+## Sub-Saharan Africa
+Notes on the WID methodology. Basically no tax data. Based on guesses about how you shift consumption to pre-tax income, and then top income adjustment.
+
+Plot: make a direct comparison for this region with the same index. NEcessarily tethered to the survey data (unlike say Mexico) The changes are amplified in WID (falls are larger, rises are larger).
+
+Also note the quality/comparability of the PIP survey data.
 
@@
 ### 1993-2015
+
+```{=html}
+<!-- Add icon library -->
+
+<!-- 
+<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/4.7.0/css/font-awesome.min.css">
+-->
+
+<!--
+<a href="data_appendix/figures_and_tables/PIP: Gini vs WID: Top 1pc share – pretax – 1993 to 2015 – allObsFalse.csv" download="Gini_vs_Top1_1993_2015.csv">
+    -->
+<!-- Auto width -->
+<!--
+<button class="btn"><i class="fa fa-download"></i> Download full data</button>
+</a> 
+-->
+```
+
+```{=html}
+<!-- 
+<html>
+<body>
+
+<input type="radio" name="1993-vs-2015_radio" value="PIP" checked="checked" onclick="chartSwitchFunction1993_2015_img()">&nbsp;PIP: Gini coefficient
+&nbsp;
+&nbsp;
+<input type="radio" name="1993-vs-2015_radio" value="WID" onclick="chartSwitchFunction1993_2015_img()">&nbsp;WID: Top 1% share
+
+<br>
+<label for="1993-vs-2015_radio_obsCheck">Show only observations available in both datasets:</label> 
+<input type="checkbox" id="1993-vs-2015_radio_obsCheck" onclick="chartSwitchFunction1993_2015_img()">
+
+<img src="images/figures/pip-gini-1993-vs-2015-all-obs.png" id="1993-vs-2015_plot" width="500">
+<br>
+
+</body>
+<script>
+function chartSwitchFunction1993_2015_img() {
+
+  var ele = document.getElementsByName('1993-vs-2015_radio');
+
+  var checkBox = document.getElementById("1993-vs-2015_radio_obsCheck");
+
+  if (checkBox.checked == true){
+    if(ele[0].checked)
+        document.getElementById("1993-vs-2015_plot").src = "images/figures/pip-gini-1993-vs-2015-matching-obs.png";
+
+    if(ele[1].checked)
+        document.getElementById("1993-vs-2015_plot").src = "images/figures/wid-top-1-1993-vs-2015-matching-obs.png";
+
+  } else {
+
+        if(ele[0].checked)
+        document.getElementById("1993-vs-2015_plot").src = "images/figures/pip-gini-1993-vs-2015-all-obs.png";
+
+    if(ele[1].checked)
+        document.getElementById("1993-vs-2015_plot").src = "images/figures/wid-top-1-1993-vs-2015-all-obs.png";
+  }
+}
+
+</script>
+</html> -->
+
+```

### 2023-04-26 PhD_pages/Paper_1/paper1.qmd
Paper 1: drafting

@@
 One is the World Bank's harmonized collection of household survey data, published online through its [*Poverty and Inequality Platform (PIP)*](https://pip.worldbank.org/).
-Another is the [*World Inequality Database (WID)*](https://wid.world/), which emerged out of the burgeoning literature on top incomes th incorporation of tax data [@atkinson_top_2014].
+Another is the [*World Inequality Database (WID)*](https://wid.world/), which emerged out of the burgeoning literature on top incomes [@atkinson_top_2014].
 
@@
 Both datasets, however, take significant steps to harmonize, *ex-post*, the available underlying data.
-In this respect they can be distinguished from databases such as WIID and SWIID [@atkinson_promise_2001]]
-
-Both have played substantial agenda-setting roles in academic research, policy and public debate.
-Their use and impact, however, has been confined to largely separate arenas, reflecting differences in the concepts of inequality they employ and the coverage they provide.
-
-The *PIP* data has typically been the favoured choice for investigating the whole distribution of welfare across households, particularly at a global scale.
-() The *WID* data at the top, focus on role of capital incomes.
-
-Moreover, the many methodological differences separating the two datasets prevent users combining the insights that they each offer individually into a coherent empirical view informed by both.
-In choosing datasets, we are simultaneously choosing underlying data source, methods, welfare concept, and coverage.
+In this respect they can be distinguished from databases such as WIID and SWIID [@atkinson_promise_2001]] Both have played substantial agenda-setting roles in academic research, policy and public debate.
+
+Their use and impact, however, has taken place within largely separate arenas.
+Many methodological differences separate the two datasets, including the  underlying data sources they rely on, the methods and assumptions applied to those sources, the welfare concept, the summary indices of inequality favoured, and coverage they provide.
+This lack of alignment between the datasets greatly limits our ability to combine the insights provided by each into a coherent empirical view informed by both.
 The two datasets represent distinct and, to a large extent, incommensurable 'worlds' of inequality measurement.
 
-This disconnect is all the more pertinent because they disagree on basics. Firstly. Secondly.
-
-Very confusing terrain for people to understand. We don't know which of the two worlds we live in. Depends on a choice about which little can be determined.
-
-This paper is the first of two which will try to ocnnect them up.
-
-Section 2. Section 3...
-
-### Two 'worlds' of inequality measurement
+Moreover, these two worlds of measurement yield different views concerning the nature and evolution of inequality.
+
+Firstly, while the *WID* data and its associated global reports suggest rising inequality was a near-ubiquitous phenomenon [@alvaredo_world_2018; @chancel_world_2021], in contrast the World Bank *PIP* data and its associated reports offer a considerably more mixed picture, in which inequality has been rising and falling in roughly equal numbers of countries [@world_bank_poverty_2016].[^WID_PIP_quotes]
+
+[^WID_PIP_quotes]:
+    > Income and wealth inequalities have been on the rise nearly everywhere since the 1980s, following a series of deregulation and liberalization programs which took different forms in different countries.
+[@chancel_world_2021, pp.
+xx]
+
+    > In recent decades, income inequality has increased in nearly all countries, but at different speeds... [@alvaredo_world_2018, pp. xx]
+
+    > [Since c.1993] the number of countries with rising inequality was **slightly higher** than the number of countries with falling inequality.
+Yet, the average Gini among these 91 countries actually **declined**.
+[@world_bank_poverty_2016, pp.
+xx]
+
+Secondly, the datasets provide contrasting accounts as to how inequalities *within* countries compare to the differences in average incomes seen *between* countries.
+Most analyses of global income inequality to date have relied heavily on the *PIP* data [@Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @milanovic_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk].
+A consistent result from these studies is that, despite recent convergence [@patel_new_2021], differences in average incomes between countries continue to make up the lion's share of global inequality.
+On such accounts, a person's income is largely determined by the country in which they happen to have been born [@milanovic_global_2015].
+More recent analysis of the *WID* data, however, flips this established result on its head.
+@chancel_global_2021 find within-country inequalities instead 'dominate' inequality between countries, contributing 78% in a decompositon of the global Theil index.
+
+Thus despite the greatly improved range of distributional data at our disposal, considerable ambiguity continues to surround some of the most basic descriptive questions concerning inequality around the world.
+
+This is the first of a pair of papers that try to address this problem by bridging between the *PIP* and *WID* datasets, and the two 'worlds' of measurement they manifest.
+The novelty of the papers lies preciesly in considering both datasets in conjunction, in contrast to the many studies that consider only one in isolation.
+In taking such a combined view, we are able to 'triangulate' a more robust picture of inequality around the world by highlighting where the datasets agree and reinforce one another, and pinpointing more precisely where and why they do not.
+
+The present paper the focuses on trends in within-country inequality.
+A companion paper considers overall global inequality, factoring in differences in average incomes between countries alongside the within-country inequalities analysed here.
+
+The rest of the paper is organized as follows.
+Section 2 provides a more detailed description of the *PIP* and *WID* datasets and their main differences.
+Section 3 compares within-country inequality trends in the two datasets, starting with the metrics where they fall furthest apart.
+For this we use the Gini of disposabale income (or consumption) and the top 1% share of pretax income as metrics emblematic of the *PIP* and *WID* datasets respectively.
+Section 4 takes a first simple step to bridge betweent he datasets, by aligning the coverage. Section 5 considers a range of inequality indices, comparing the datasets on a like-for-like basis.
+Section 5 zooms in subset of countries where there is most disagreement. the role of welfare metrics and specific data quality issues for 
+Section 6 concludes, drawing out the main overlaps and disagreements between the two datasets, accounting for the steps to
+
+# Two 'worlds' of inequality measurement
 

### 2023-04-28 PhD_pages/Paper_1/paper1.qmd
Paper 1: Drafting

@@
 Their use and impact, however, has taken place within largely separate arenas.
-Many methodological differences separate the two datasets, including the  underlying data sources they rely on, the methods and assumptions applied to those sources, the welfare concept, the summary indices of inequality favoured, and coverage they provide.
-This lack of alignment between the datasets greatly limits our ability to combine the insights provided by each into a coherent empirical view informed by both.
+Many methodological differences separate the two datasets, including the  underlying data sources they rely on, the methods and assumptions applied to those sources, the concept of welfare measured, the summary indices of inequality typically favoured, and coverage they provide.
 The two datasets represent distinct and, to a large extent, incommensurable 'worlds' of inequality measurement.
-
-Moreover, these two worlds of measurement yield different views concerning the nature and evolution of inequality.
-
-Firstly, while the *WID* data and its associated global reports suggest rising inequality was a near-ubiquitous phenomenon [@alvaredo_world_2018; @chancel_world_2021], in contrast the World Bank *PIP* data and its associated reports offer a considerably more mixed picture, in which inequality has been rising and falling in roughly equal numbers of countries [@world_bank_poverty_2016].[^WID_PIP_quotes]
-
-[^WID_PIP_quotes]:
-    > Income and wealth inequalities have been on the rise nearly everywhere since the 1980s, following a series of deregulation and liberalization programs which took different forms in different countries.
-[@chancel_world_2021, pp.
-xx]
-
-    > In recent decades, income inequality has increased in nearly all countries, but at different speeds... [@alvaredo_world_2018, pp. xx]
-
+This lack of alignment between the datasets limits our ability to combine the insights provided by each into a coherent empirical view informed by both.
+
+This disconnect is particularly problematic given the starkly contrasting perspectives on inequality the two worlds provide.
+
+Analysis of the *PIP* data suggests that within-country inequality trends have been decidedly mixed in recent decades.
+Inequality is found to have fallen in roughly as many countries as it has risen, with no substantial trend in either direction on average [@world_bank_poverty_2016].[^PIP_quotes] In contrast, summaries of the *WID* data emphasise that rising inequality within countries has been an almost universal phenomenon [@alvaredo_world_2018; @chancel_world_2021].[^WID_quotes]
+
+[^PIP_quotes]:
     > [Since c.1993] the number of countries with rising inequality was **slightly higher** than the number of countries with falling inequality.
@@
 
-Secondly, the datasets provide contrasting accounts as to how inequalities *within* countries compare to the differences in average incomes seen *between* countries.
-Most analyses of global income inequality to date have relied heavily on the *PIP* data [@Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @milanovic_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk].
-A consistent result from these studies is that, despite recent convergence [@patel_new_2021], differences in average incomes between countries continue to make up the lion's share of global inequality.
-On such accounts, a person's income is largely determined by the country in which they happen to have been born [@milanovic_global_2015].
+[^WID_quotes]:
+    > Income and wealth inequalities have been on the rise nearly everywhere since the 1980s, following a series of deregulation and liberalization programs which took different forms in different countries.
+[@chancel_world_2021, pp.
+xx]
+
+    > In recent decades, income inequality has increased in nearly all countries, but at different speeds... [@alvaredo_world_2018, pp. xx]
+
+The two datasets have also yielded starkly contrasting accounts of how big a contribution within-country inequalities make to overall global inequality.
+Most analyses of global income inequality to date have relied heavily on the *PIP* data [@kanbur_global_2022; @lakner_global_2016; @anand_global_2015; @milanovic_global_2015; @Nino-Zarazua2017-tv].
+A consistent result from these studies is that, despite some recent convergence in average incomes between countries [@patel_new_2021], such differences  continue to make up the lion's share of global inequality.
+On such accounts, a person's income is largely determined purely by the country in which they happen to have been born [@milanovic_global_2015].
 More recent analysis of the *WID* data, however, flips this established result on its head.
@@
 
-Thus despite the greatly improved range of distributional data at our disposal, considerable ambiguity continues to surround some of the most basic descriptive questions concerning inequality around the world.
+The two worlds of *measurement* thus depict two very different empirical *realities*.
+Knowing which of these worlds we live in matters not just for our understanding of inequality today, but also our understanding of how we should seek to combat inequality in the future.
+A near ubiquitous rise in inequality apparent in the *WID* data points to fundamental 'forces for divergence' within capitalism [@piketty_capital_2014] and the need for global coordination, particularly on the taxation of capital, to counter them [@saez_triumph_2019].
+If inequalities within countries now predominate within the global distribution of income, such avenues will also be central to tackling global inequality too [@piketty_brief_2022].
+Conversely, the much more mixed picture seen in the *PIP* data highlights more prominently the past role and future scope of national policies to shape the extent of inequality within countries [@ravallion_inequality_2018].
+From this viewpoint, such inequalities in any case remain a small share of overall global inequality such that the global distribution will continue to depend heavily on future aggregate growth rates [@kanbur_global_2022].
+
+More fundamentally, the two datasets support differing narratives concerning globalization. [TODO: elabourate].
+
+Despite laudable efforts to make transparent, in practice a black box.
+Treatmetn of the underlying data and assumptions.
+Whilst strengths and weaknesses studied for particular countries with good data, when applied globally we really don't know.
+Despite the great data, fundamental ambiguity surrounding the most basic aspects.
 
@@
 
-# Bridging the gap for particular countries
+# Bridging the gap for particular countries / welfare concept
+
+"Korinek, Mistiaen, and
+Ravallion (2006) estimate that correcting
+for such selective compliance adds about
+five percentage points to the Gini index for
+the United States."
 

### 2023-04-28 PhD_pages/older_structure/paper_pages/paper3.qmd
Restructuring folders so that everything is contained in one complete subfolder per paper
Paper 1: Drafting

--- 2022-11-30 PhD_pages/paper_pages/paper3.qmd
@@
 ---
+
+NB: I should look at the global estimates of factor shares mentioned in some UN reports (maybe ILO data?). e.g. 
+- "The share of wages in total GDP declined in a majority of countries
+(91 out of 133 with data) from 1995 to 2014 (ILO, 2016)." (https://www.un.org/development/desa/dspd/wp-content/uploads/sites/22/2020/02/World-Social-Report2020-FullReport.pdf).
+- "A decline in the labour share of income from 2014 to 2019 ‐ from 54.1 per cent to 52.6 per cent ‐ represents upward pressure on inequality." https://www.un.org/sustainabledevelopment/inequality/
 

### 2023-05-02 PhD_pages/Paper_1/paper1.qmd
Paper 1: Drafting

@@
 
+# On such accounts, a person's income is determined largely simply by the country in which they happen to have been born [@milanovic_global_2015].
+
+# A near ubiquitous rise in inequality apparent in the *WID* data points to fundamental 'forces for divergence' within capitalism [@piketty_capital_2014] and the need for global coordination, particularly on the taxation of capital, to counter them [@saez_triumph_2019].
+If inequalities within countries now predominate within the global distribution of income, such avenues will also be central to tackling global inequality too [@piketty_brief_2022].
+Conversely, the much more mixed picture seen in the *PIP* data highlights more prominently the past role and future scope of national policies to shape the extent of inequality within countries [@ravallion_inequality_2018].
+From this viewpoint, such inequalities in any case remain a small share of overall global inequality such that the global distribution will continue to depend heavily on future aggregate growth rates [@kanbur_global_2022].
+
 ```
@@
 Economic inequalities around the world are vast.
-The world's richest 1% enjoy an average income that is around 150 times that of the poorer half of the world's population [@chancel_world_2021].
-While the richest may afford private flights to space, roughly three billion people are unable to afford even just the cost of a healthy diet [@herforth_cost_2020].
+While the richest can afford private flights to space, roughly three billion people are unable to afford even just the cost of a healthy diet [@herforth_cost_2020].
+The world's richest 1% enjoy an average income that is around 150 times that of those in the poorer half of the world's population [@chancel_world_2021].
 
@@
 Their use and impact, however, has taken place within largely separate arenas.
-Many methodological differences separate the two datasets, including the  underlying data sources they rely on, the methods and assumptions applied to those sources, the concept of welfare measured, the summary indices of inequality typically favoured, and coverage they provide.
-The two datasets represent distinct and, to a large extent, incommensurable 'worlds' of inequality measurement.
-This lack of alignment between the datasets limits our ability to combine the insights provided by each into a coherent empirical view informed by both.
+Many methodological differences distinguish the *PIP* and *WID* datasets.
+This includes the underlying data sources they rely on; the methods and assumptions applied to those sources; the concept of welfare measured; the summary indices of inequality typically favoured; and coverage they provide.
+This lack of alignment makes it difficult to combine the insights each provide into a coherent empirical view informed by both.
+The two datasets represent distinct and, to a large extent, incommensurable 'worlds' of inequality measurement – used in isolation rather than in conjunction.
 
@@
 
-The two datasets have also yielded starkly contrasting accounts of how big a contribution within-country inequalities make to overall global inequality.
+The two datasets have also yielded opposing accounts of the contribution made by within-country inequalities to overall global inequality.
 Most analyses of global income inequality to date have relied heavily on the *PIP* data [@kanbur_global_2022; @lakner_global_2016; @anand_global_2015; @milanovic_global_2015; @Nino-Zarazua2017-tv].
-A consistent result from these studies is that, despite some recent convergence in average incomes between countries [@patel_new_2021], such differences  continue to make up the lion's share of global inequality.
-On such accounts, a person's income is largely determined purely by the country in which they happen to have been born [@milanovic_global_2015].
+A consistent result from these studies is that differences between countries' average incomes continue to make up the lion's share of global inequality, despite some recent convergence [@patel_new_2021].
 More recent analysis of the *WID* data, however, flips this established result on its head.
-@chancel_global_2021 find within-country inequalities instead 'dominate' inequality between countries, contributing 78% in a decompositon of the global Theil index.
-
-The two worlds of *measurement* thus depict two very different empirical *realities*.
-Knowing which of these worlds we live in matters not just for our understanding of inequality today, but also our understanding of how we should seek to combat inequality in the future.
-A near ubiquitous rise in inequality apparent in the *WID* data points to fundamental 'forces for divergence' within capitalism [@piketty_capital_2014] and the need for global coordination, particularly on the taxation of capital, to counter them [@saez_triumph_2019].
-If inequalities within countries now predominate within the global distribution of income, such avenues will also be central to tackling global inequality too [@piketty_brief_2022].
-Conversely, the much more mixed picture seen in the *PIP* data highlights more prominently the past role and future scope of national policies to shape the extent of inequality within countries [@ravallion_inequality_2018].
-From this viewpoint, such inequalities in any case remain a small share of overall global inequality such that the global distribution will continue to depend heavily on future aggregate growth rates [@kanbur_global_2022].
-
-More fundamentally, the two datasets support differing narratives concerning globalization. [TODO: elabourate].
-
-Despite laudable efforts to make transparent, in practice a black box.
-Treatmetn of the underlying data and assumptions.
-Whilst strengths and weaknesses studied for particular countries with good data, when applied globally we really don't know.
-Despite the great data, fundamental ambiguity surrounding the most basic aspects.
-
-This is the first of a pair of papers that try to address this problem by bridging between the *PIP* and *WID* datasets, and the two 'worlds' of measurement they manifest.
-The novelty of the papers lies preciesly in considering both datasets in conjunction, in contrast to the many studies that consider only one in isolation.
-In taking such a combined view, we are able to 'triangulate' a more robust picture of inequality around the world by highlighting where the datasets agree and reinforce one another, and pinpointing more precisely where and why they do not.
-
-The present paper the focuses on trends in within-country inequality.
-A companion paper considers overall global inequality, factoring in differences in average incomes between countries alongside the within-country inequalities analysed here.
+@chancel_global_2021 find within-country inequalities instead now 'dominate' inequality between countries, contributing 78% in a decompositon of the global Theil index.
+
+Alongside these divergent assesments of *how* inequality has changed around the world, the two approaches to measurement are frequently used to evidence contrasting accounts of *why* it has changed in this way and hence where to target efforts to combat inequality in the future.
+[TODO: Expand thought: i.e.
+marginal labour productivies/wage distribution vs capital incomes/fundamental force for divergence/top marginal taxes/top wages bargaining.].
+
+The two 'worlds' of inequality measurement adopted by the *PIP* and *WID* datasets have yielded two very different views of inequality around the world and consequently two very different sets of recommendations as to how to address it.
+
+Which of these realities should we give most credence to?
+A simplistic reading of the top incomes literature would see the *WID* data as unambiguously providing greater reliability, thanks to the incorporation of tax data and other sources in order to correct for the well-known issue of the 'missing rich' within survey data [@lustig_missing_2019].
+The need for such a correction is amply demonstrated in many cases.
+@korinek_survey_2006, for example, find richer households in the United States to be less likely to participate in surveys and estimate this to bias the Gini coefficient downwards by around five percentage points.
+[TODO: More examples, including underreporting].
+The *PIP* data, based purely on survey data, makes no such correction and is thus likely to understimate the extent of inequality in many cases.
+
+A more agnostic assessment of the two datasets is, however, warrented on a number of grounds.
+Firstly, the extent of the upward revision to inequality estimates resulting from top income adjustments varies from country to country.
+In some cases it may be quite modest, as demonstrated by @carranza_top_2021.
+They apply *WID*-style top income adjustments to survey data for 26 European countries and find an average increase in the Gini of around 2.4 points, with notable differences across countries and limited impact on trends.
+Secondly, for many countries/periods included in the *WID* estimates, the richer range of data sources that the dataset's methodology seeks to incorporate do not exist.
+In the place of data, various imputation and interpolation strategies must be employed.^[For example, see the methods outlined in @chancel_towards_2019 in the context of Africa, for which data quality and coverage is especially low] The terms of this tradeoff are not clear.
+While the incorporation of tax data is a proven asset in many individual cases, the value of top income adjustment methodologies when applied *globally* is harder to demonstrate.
+Lastly, deviations between survey and tax data appear to be restricted to the *very* top of the distribution.
+After aligning the income concept across survey and tax data for United States, Germany and France, @yonzan_drawing_2022 find that the sources correspond extremely well up until the top 1% of the distribution.
+The apparent 'accuracy' of survey data is thus likely to depend on the inequality index chosen, and in particular to its sensitity to changes at the top of the distribution.
+It is thus a question whose answer will depend also on normative judgements concerning the nature of inequality [@atkinson_measurement_1970].
+
+Overall, while for particular questions and cases, much can be said about the relative strengths and weaknesses of the two approaches to measurement employed in the *PIP* and *WID* datasets, there is no straightforward answer to the question of which offers us the most reliable or relevant global view on inequality.
+
+A different route, however, may be taken to try to reconcile the divergent views of inequality drawn from the two datasets.
+A limitation of the existing .tend to be made insolation.
+This means areas of agreement are overlooked and the reasons for disagreement are obfuscated.
+By using the two datasets in conjunction, rather than in isolation, we can 'triangulate' a more robust and coherent empirical view informed by both.
+[TODO: Add footnote citing Tony Atkinson on the idea of triangulation].
+This is the first of a pair of papers that seeks to do this.
+The present paper focuses on trends in within-country inequality – a companion paper takes a similar approach but considers overall global inequality.
+It triangulates between the *PIP* and *WID* data by separately aligning coverage, the inequality index, and where possible the welfare concept.
+
+The result of this is X, Y, Z.
 
@@
 Section 2 provides a more detailed description of the *PIP* and *WID* datasets and their main differences.
-Section 3 compares within-country inequality trends in the two datasets, starting with the metrics where they fall furthest apart.
-For this we use the Gini of disposabale income (or consumption) and the top 1% share of pretax income as metrics emblematic of the *PIP* and *WID* datasets respectively.
-Section 4 takes a first simple step to bridge betweent he datasets, by aligning the coverage. Section 5 considers a range of inequality indices, comparing the datasets on a like-for-like basis.
-Section 5 zooms in subset of countries where there is most disagreement. the role of welfare metrics and specific data quality issues for 
-Section 6 concludes, drawing out the main overlaps and disagreements between the two datasets, accounting for the steps to
+Section 3 compares within-country inequality trends in the two datasets, starting with the metrics they are most associated with and where they fall furthest apart: for *PIP* the Gini of disposabale income (or consumption); for *WID* the top 1% share of pretax national income.
+Section 4 takes a first simple step to bridge between the datasets, by restricting the analysis to only those countries available in both datasets.
+Section 5 considers a range of inequality indices, comparing the datasets on a like-for-like basis.
+Section 5 zooms in subset of countries where there is most disagreement and using Luxembourg Income Study considers the contribution of the different taget welfare concepts employed within the two datasets.
+Section 6 concludes.
 

### 2023-05-02 PhD_pages/Paper_1/paper1.qmd
Paper 1: Drafting

@@
 ---
-
-::: {.callout-note collapse="true" appearance="minimal"}
-### Load packages and data
 
@@
 
-:::
-
 ```{python}
@@
 # Bourguignon (2015)
-# The share of wages in total GDP declined in a majority of countries
-# (91 out of 133 with data) from 1995 to 2014 (ILO, 2016).
+# WID methodology papers:
+
+# - Brazil: https://wid.world/document/extreme-persistent-inequality-new-evidence-brazil-combining-national-accounts-surveys-fiscal-data-2001-2015-wid-world-working-paper-201712/
+#  - Assouad (2017) on Lebanon and Czajka (2017) on Côte d’Ivoire]
+
+# TEXT GRAVEYARD
 
@@
 
+# The two datasets differ along a range of dimensions, including the underlying sources relied on, the measures of welfare used, the summary indicators commonly favoured, and the temporal and geographic coverage available.
+
+# These differences come largely as a bundle.
+Choosing between the *WID* or *PIP* data implies not only a technical choice between data sources and methodolodies, but also a more fundamental choice as to which concept of inequality to consider, for which places and time periods.
+This has a led to a degree of specialisation in their use.
+
+# The *PIP* data has typically been the favoured choice for investigating the whole distribution of welfare across households, particularly at a global scale [@Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @milanovic_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk].
+This includes the World Bank's own research and monitoring activties concerning global poverty, inequality and growth across the distribution [see for example @world_bank_poverty_2016; @world_bank_poverty_2022].
+The dataset is also relied on by the United Nations to set and monitor many of the targets in this area that form part of its *Sustainable Development Goals*.
+
+# The *WID* data has more commonly been used to understand top dynamics before taxes.
+Notwithstanding substantial efforts to global, high income countries due to the data required.
+in which today's rich countries are overrepresented .
+In policy, advocacy and public debate, the data used to highlight the very large and, in many cases, growing advantage enjoyed by the richest individuals in these societies.
+In academia, top incomes data long-run processes fundamental – capital accumulation.[@atkinson_top_2007; @piketty_capital_2014]  On political economy channels, social conflict between capital and labour, wage bargaining norms and the interaction with marginal tax rates.
+More generally the data have led to a fundamental revaluation of the causes of inequality that move beyond orthodox accounts given purely in terms of marginal productivities of workers with heterogeneous skills or education [@alvaredo_top_2013; @piketty_capital_2014-1].
+
+# Thus whilst both the *PIP* and *WID* datasets have played substantial agenda-setting roles in academic research, policy and public debate, this has tended to occur in largely separate arenas.
+Moreoever, no way to join up the insights they offer.
+Very confusing terrain for people to understand.
+We are left with two distinct and, to a large extent, incommensurable 'worlds' of inequality measurement.
+
+# Recent extensions have brought down some of these barriers. PIP now includes high-income countries. WID now global coverage, and post-tax income concepts. In
+
+# At the same time, these developments in the two datasets have made this disconnect between them all the more apparent. Fundamental differences on the basics. Two in particular.
+
+# 1
+# 2
+
+# Whilst for individual countries, the connections may be well understood, globally the value of the different assumptions is not really known. We have to pick blindly.
+
+# ### Connecting up the worlds
+
+# This is the first of two papers that seeks to join up.
+
+# The dataset is used for the Bank's own research and monitoring activities which, whilst traditionally focussing on absolute poverty, have more recently encompassed trends in inequality, relative poverty and growth across the distribution.
+Notably, the data is alo used to set and monitor many of the targets, on both poverty and inequality, that form part of the UN's Sustainable Development Goals.
+More generally, the source has been indispensible for research taking a global view on inequality [@Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @milanovic_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk].
+
+# The *WID* database emerged out of the burgeoning literature on top incomes [@atkinson_top_2014] and has come to play a central and, to some extent, coordinating role for this research agenda.
+The key innovation of the 'top incomes' approach has been the incorporation of tax data, as either an alternative or complement to the more orthodox reliance on household survey data.
+The primary advantage is that tax data can provide a better view of the top of the income distribution, which is often poorly captured by survey data [@lustig_missing_2019].
+However, the choice of data source has typically brought with it a number of other differences that mark out the 'top incomes' approach.
+This includes a particular focus on the top of the distribution (for which tax data is more universally available), and on income measured before tax (in terms of which the underlying data is gathered).
+This contrasts with more orthodox approaches to inequality measurement in which disposable income has tended to be seen as the more relevant measure of welfare [@canberra_group_canberra_2011], and greater attention is given to measures of overall inequality, such as the Gini coefficient, or to relative poverty.
+
+# Another benefit of tax data is the much longer timeframes it is sometimes available for.
+The estimates gathered in *WID* provide a long-run perspective on the evolution of inequality over the twentieth century, albeit for a more limited geography in which today's rich countries are somewhat overrepresented [@atkinson_top_2007; @piketty_capital_2014].
+
+# Moreover, the top incomes methodology carried forward by *WID* developed with a distinctive set of questions concerning the *causes* of inequality in mind.
+Particular attention is given to the distribution of wealth and capital incomes as drivers of personal income inequality, in contrast to orthodox accounts focussed purely on the dispersion of wages [@alvaredo_top_2013; @piketty_capital_2014-1].
+This ...
+hidden from view given the concentration of capital incomes at the top, captured poorly by survey data.
+
+# ### Two 'worlds' of inequality measurement
+# Tw worlds – methods, conceptions of inequality, time periods and geographic coverage and research questions come bundled.
+
+# Always unsatisfying. Very difficult to unpick. Peple do it for individual countries..
+
+# A wholeslae comparison was generally avoided due to the very different coverage. Used in parallel.
+
+# As the databases have developed and expanded, however, this division of labour between the two datasets has become increasingly blurred.
+PIP includes high-income countries.
+WID coverage is global.
+DINA focus on comparability.
+Increasingly, the datasets are used to answer the same range of questions, speak to the same world.
+
+# Where they now meet, the result is very confusing, one that, despite the great strides forward, leaves uncertainty around the most basic questions.
+
+# This is most visible in two fundamental disagreements between the two 'worlds'.
+
+# Firstly, they give very different views of how ubiquitous a phenomenon rising inequality is within countries around the world.
+The view from *World Inequality Database* and its associated global report is that inequality has risen 'nearly everywhere' in recent decades [@alvaredo_world_2018; @chancel_world_2021].
+In contrast, the World Bank data and its associated biannual report offers a much more mixed picture – in which inequality is rising and falling in roughly equal numbers of countries.
+
+# Secondly, the datasets paint as to how such inequalities *within* countries compare to the differences in average incomes seen *between* richer and poorer countries.
+
+# Most analyses of global income inequality to date have relied heavily on the *PIP* data [@Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @milanovic_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk].^[TODO: Add footnote explaining transition from Povcal] *[TODO: Check and more refs.]*.
+A consistent result from this literature is that inequalities within countries remain small in comparison to the much larger differences seen between countries.
+This is despite the convergence seen in recent decades between countries, given higher economic growth rates across low- and middle-income countries.
+According to such research, a person's income remains largely determined by the country in which they happen to have been born [@milanovic_global_2015].
+
+# More recent analysis of the WID data flips this established result on its head. Chancel... estimate xx.
+
+# The datasets are enormous efforts, involving the work of hundreds of researchers.
+They are hugely valuable offering an unprecidented perspective on global inequality and how it is evolving.
+And yet the disconnect between the two devalues them both.
+Amidst the wealth of data, basic facts are not known.
+And the terrain is incredibly confusing to navigate for policimakers and the concerned public.
+
+# ### Connecting up the worlds
+
+# This paper, along with a companion paper, seeks to join up these two worlds. This first paper picks up the first question – within country. A companion paper considers global inequality.
+
+# Section 2 of this paper introduces the two worlds.
+# Section 3 compares within country trends between two periods.
+It begins with the inequlaity metrics for which these datasets are most associated, and in which the worlds fall furthest apart: Gini, Top 1% share.
+Already
+# Section 4 aligns coverage. It is not neutral. Already this goes some way to explaining the gap between the two datasets.
+# Section 5 aligns inequality indices. [Summary.]
+# Section 5 considers the role of welfare metrics and specific data quality issues for a subset of countries where there is most disagreement.
+
+# Summary of main conclusion!?
+
+# # 2. Two worlds of inequality measurement
+
+# As well as being the the two  not just two different databases. They in some sense represent *two different worlds* of inequality measurement.
+
+# Different methods. These yield different pictures of the extent of inequality and how it has changed over time. Motivated by different ideas about the salient factors shaping inequality.
+
+# ### PIP
+
+# ### WID
+
+# They differ as to how ubiquitous a phenomenon rising inequality is around the world.
+The view from *World Inequality Database* and its associated global report is that inequality is rising in 'nearly everywhere' [@alvaredo_world_2018; @chancel_world_2021].[^WID_quotes] In contrast, the World Bank data and its biannual report offers a much more mixed picture – in which inequality is rising and falling in roughly equal numbers of countries.
+
+# [^WID_quotes]:
+#     > Income and wealth inequalities have been on the rise nearly everywhere since the 1980s, following a series of deregulation and liberalization programs which took different forms in different countries.
+[@chancel_world_2021, pp.
+xx]
+
+#     > In recent decades, income inequality has increased in nearly all countries, but at different speeds... [@alvaredo_world_2018, pp. xx]
+
+# Decompositions of global inequality into between- and within-country components.
+
+# Capital vs labour share. .– including fundamental laws of capitalism with regard to capital accumulation (Piketty) –
+
+# ### The disconnect between these datasets devaules them both
+
+# These datasets are incredibly valuable – playing to motivate action.
+The example in the opening sentence above illustrates this well.
+It is only thanks to the WID data that we can say X.
+It is only thanks to the PIP data that we can say that three billion cannot afford a healthy diet.
+
+# Often it is presented that there is a division of labour between these dataset – WID is for inequality, especially at the top.PIP is for poverty.
+
+# But the point is that it is only one world that we live in. The disconnect is not something we can live with, but rather presents a challenge to both datasets.
+
+# There remains fundamental ambiguity surrounding which of two possible worlds we live in.
+
+# The '*WID* world' rising inequality within countries has been an almost universal phenomenon.
+suggests the winners from globalization and economic growth overwhelmingly have been a global elite.
+It sees the bulk of global inequalities as lying within nations, but emphasises global factors in terms of causes and hence the need for international coordination to combat inequality.
+
+# The '*PIP* world' – within-country inequality trends in recent decades have been mixed, and the differences in average incomes between countries remain the lion's share of the globally.As such the importance of national policy (Ravallion, 2018; Corlett, 2016).
+access to both produced and human capital.
+
+# Knowing which of these worlds we live in matters a great deal – for our understanding of inequality today, how it has changed, and also where efforts to combat inequality in the future are likely to be most successful.
+
+# Trouble is that these are all bundled up. You must chose. On what basis can you make a choice?
+
+# Known problems on both sides.
+
+# For PIP – top incomes, and lack of comparability of the survey data – notably mix of incme and consumption data.
+
+# For WID – tha many assumptions.
+An emphasis placed on the incorporation of tax data, but when taking a global view confonte by the fact that such tax data is often nonexistent.
+Top income corrections based on strong assumptions.
+
+# Given the presentation, the extent to which different assumptions, indices, methodological choices affects the data and – is not something that can be read off' the data.
+
+# In using the datasets a global view on inequality, requires choosing wholesale between methodological 'bundles' whose relative strengths and weaknesses, applied globally, are not well understood.
+
+# Also makes for a very confusing terrain for policymakers, media and the public.
+
+# The two datasets are incredible efforts, involving the work of hundreds of researchers.
+Both datasets have continued to improve adn expand.
+In terms of data availability we are thus in a better position than ever to take a global viw on inequlity.
+And yet, amidst this wealth of data, limited understanding of very basics concerning inequality around the world.
+
+# ### Outline: Connecting the two worlds
+# *(In this section I will explain the plan for the paper)*
+
+# In this paper I aim to bring these two worlds of inequality measurement closer. To Triangulate what they do agree on and identify where they disagree.
+
+# The first, most obvious, step is simply to compare the in a systemtatic way.
+This very basic step is curiously lacking in the literature.
+We begin by comparing trends in each world based on the inequality indices most typical in each: the Gini coefficient for the PIP data, and the top 1% share of pre-tax income for WID.
+
+# Given incomplete coverage in both datasets, we 'line up' estimates to reference years to see change over time.
+We use two periods, presenting two different tradeoffs with respect to data coverage.
+The first is 1993-2018??
+This maximises global coverage – 100 countries covering roughly 90% of the world poplation (Figure XX).
+The downside is that this period misses out – transition, liberalisation and structural adjustment.
+Question begging, by avoiding specifically the period where inequality thought to rise.
+
+# The second is 1980-2018. Here we lose coverage, especially in the PIP data (Figure XX).
+
+# The two measures selected – Gini coefficient of disposable income/consumption, and the top 1% share of pre-tax income – are where the PIP and WID worlds respectively are at their furthest apart.
+Yet in terms of broad global trends, there is much that even these very different measures can agree on:
+# - x
+# - y
+# - z.
+
+# There are big differences though:
+# - Somewhat more pronounced rise among advanced economies.
+# - Big divergence in trend some of the most populous countries
+
+# Subsequent sections attempt to bridge the gaps in order to learn where those differences emerge.
+
+# - Section 2: Aligning coverage. This reduces the gap already. Coverage of countries is not... with respect to inequality trends.
+
+# - Inequality indices chosen. We align. Compare Top 1%. And Gini.
+#     - Explain results.
+
+# - Welfare concept.
+We look at different income concepts for the small number of coutnries for which such data exists.
+To this end we also include data from the Luxembourg Income Study.
+Similar to PIP in that it relies on survey data.
+But provides access to microdata, allowing for the construction of different welfare aggregates.
+We focus on the countries with the biggest difference in trend and biggest population.
+I.e.
+those that contribute most to the difference in the weighted an unweigted global averages.
+
+# What I won't do is a detailed lining up.
+# Although both institutions continue to take many steps to provide transparency and reproducibility to their estimates, they remain a black box. Makes it impossible to do this lining up.
+
+# In many particular contexts – where detailed lining up is possible – there relative strengths and weaknesses may be clear.
+(This is what Morllei does.
+And Nolan).
+But at the global level the makes this impractical.
+
+#     - Summarise: really depends on the country.
+
+# ----
+
 ```
@@
 
-This disconnect is particularly problematic given the starkly contrasting perspectives on inequality the two worlds provide.
+This disconnect is particularly problematic given the starkly contrasting perspectives on inequality found in studies drawing on the two datasets.
 
@@
 
-Alongside these divergent assesments of *how* inequality has changed around the world, the two approaches to measurement are frequently used to evidence contrasting accounts of *why* it has changed in this way and hence where to target efforts to combat inequality in the future.
-[TODO: Expand thought: i.e.
-marginal labour productivies/wage distribution vs capital incomes/fundamental force for divergence/top marginal taxes/top wages bargaining.].
-
-The two 'worlds' of inequality measurement adopted by the *PIP* and *WID* datasets have yielded two very different views of inequality around the world and consequently two very different sets of recommendations as to how to address it.
+Alongside these divergent assesments of *how* inequality has changed around the world, the two approaches to measurement are frequently used to evidence contrasting accounts of *why* it has changed in this way, and hence where to target efforts to combat inequality in the future.
+[TODO: Expand this thought: i.e.
+marginal labour productivies vs capital incomes/top wages bargaining & top marginal tax rates.
+National vs global factors, policy vs more fundamental forces.].
+
+In sum, the two 'worlds' of inequality measurement adopted by the *PIP* and *WID* datasets have yielded two very different views of inequality around the world and consequently two very different sets of recommendations as to how to address it.
 
@@
 A simplistic reading of the top incomes literature would see the *WID* data as unambiguously providing greater reliability, thanks to the incorporation of tax data and other sources in order to correct for the well-known issue of the 'missing rich' within survey data [@lustig_missing_2019].
-The need for such a correction is amply demonstrated in many cases.
-@korinek_survey_2006, for example, find richer households in the United States to be less likely to participate in surveys and estimate this to bias the Gini coefficient downwards by around five percentage points.
+The case for such a correction is amply demonstrated in many instances.
+@korinek_survey_2006, for example, find richer households in the United States are less likely to participate in surveys, and they estimate that this biases the Gini coefficient downwards by around five percentage points.
 [TODO: More examples, including underreporting].
@@
 A more agnostic assessment of the two datasets is, however, warrented on a number of grounds.
-Firstly, the extent of the upward revision to inequality estimates resulting from top income adjustments varies from country to country.
-In some cases it may be quite modest, as demonstrated by @carranza_top_2021.
-They apply *WID*-style top income adjustments to survey data for 26 European countries and find an average increase in the Gini of around 2.4 points, with notable differences across countries and limited impact on trends.
+Firstly, the extent of the upward revision to inequality estimates resulting from top income adjustments varies considerably from country to country, and in some cases may be quite modest.
+This is demonstrated by @carranza_top_2021 who apply a *WID*-style top income adjustment to survey data for 26 European countries and find an average increase in the Gini of around 2.4 points, with notable differences across countries and limited impact on trends.
 Secondly, for many countries/periods included in the *WID* estimates, the richer range of data sources that the dataset's methodology seeks to incorporate do not exist.
-In the place of data, various imputation and interpolation strategies must be employed.^[For example, see the methods outlined in @chancel_towards_2019 in the context of Africa, for which data quality and coverage is especially low] The terms of this tradeoff are not clear.
-While the incorporation of tax data is a proven asset in many individual cases, the value of top income adjustment methodologies when applied *globally* is harder to demonstrate.
+In the place of data, various imputation and interpolation strategies must be employed.^[For example, see the methods outlined in @chancel_towards_2019 in the context of Africa, for which data quality and coverage is especially low] The terms of this tradeoff – between drawing on a greater set of information, and needing to fill in the gaps where this information does not exist – are not clear.
+While the incorporation of tax data is a proven asset in many individual cases, the value and reliability of top income adjustment methods when applied globally is harder to demonstrate.
 Lastly, deviations between survey and tax data appear to be restricted to the *very* top of the distribution.
@@
 The apparent 'accuracy' of survey data is thus likely to depend on the inequality index chosen, and in particular to its sensitity to changes at the top of the distribution.
-It is thus a question whose answer will depend also on normative judgements concerning the nature of inequality [@atkinson_measurement_1970].
-
-Overall, while for particular questions and cases, much can be said about the relative strengths and weaknesses of the two approaches to measurement employed in the *PIP* and *WID* datasets, there is no straightforward answer to the question of which offers us the most reliable or relevant global view on inequality.
-
-A different route, however, may be taken to try to reconcile the divergent views of inequality drawn from the two datasets.
-A limitation of the existing .tend to be made insolation.
-This means areas of agreement are overlooked and the reasons for disagreement are obfuscated.
+It is thus a question whose answer will depend also on normative judgements concerning the nature of inequality and how it should be measured [@atkinson_measurement_1970].
+
+Overall, while much can be said about the relative strengths and weaknesses of the two approaches to measurement employed in the datasets for particular questions and cases, there is no straightforward answer to the question of which offers us the most reliable or relevant *global* view on inequality.
+
+A different route, however, may be taken to try to reconcile the divergent views of inequality often drawn from the two datasets.
+A substantial limitation of existing work on inequality around the world is that consideration of the *PIP* and *WID* data tends to be given in insolation.[^isolated_examples] As a consequence, areas of *agreement* between the dataset are overlooked, and the reasons behind disagreements – of which, given the range of methodological differences, there may be many – are given little consideration.
 By using the two datasets in conjunction, rather than in isolation, we can 'triangulate' a more robust and coherent empirical view informed by both.
 [TODO: Add footnote citing Tony Atkinson on the idea of triangulation].
-This is the first of a pair of papers that seeks to do this.
-The present paper focuses on trends in within-country inequality – a companion paper takes a similar approach but considers overall global inequality.
-It triangulates between the *PIP* and *WID* data by separately aligning coverage, the inequality index, and where possible the welfare concept.
-
+This is the first of a pair of papers aiming to do this.
+The present paper focuses on trends in within-country inequality and a companion paper takes a similar approach but considers overall global inequality.
+
+[^isolated_examples]:
+    TODO: finish footnote.
+    Examples include the flagship reports associated with the two datasets – see for example @chancel_global_2021 and @world_bank_poverty_2022.
+Policy papers: @undp_beyond_2019 consider only WID data, with some minor cross-references to PIP data.
+Within academic research @kanbur_global_2022, @nino-zarazua_global_2017 and @ravallion_income_2014 rely on PIP data.
+Explain about @van_zanden_changing_2014.
+
+    Two notable counter-examples that have sought to incorporate both 'worlds' are @anand_global_2015 and @lakner_global_2016.
+Both studies recalculate global inequality measures after making a rough adjustment to the survey data so as to match the top income shares observed in *WTID*, a successor to *WID*.
+Their estimates extend only to 2005 and 2008 respectively and were made prior to the substantial increases in coverage now available in *WID*.
+Moreover, neither consider average within-country trends – the primary focus of the present paper.
+
+    @united_nations_world_2020 also draws on evidence from both datasets, but presents the results in parallel, without attempting to draw from them a coherent view.
+
+I triangulate between the *PIP* and *WID* data by separately aligning coverage and the inequality indices used.
 The result of this is X, Y, Z.
+For a subset of countries, the welfare concept, using Luxembourg Income Study, as well as *WID*'s own estiamtes of post-tax inequality where available.
 
@@
 
-The two datasets differ along a range of dimensions, including the underlying sources relied on, the measures of welfare used, the summary indicators commonly favoured, and the temporal and geographic coverage available.
-
-These differences come largely as a bundle.
-Choosing between the *WID* or *PIP* data implies not only a technical choice between data sources and methodolodies, but also a more fundamental choice as to which concept of inequality to consider, for which places and time periods.
-This has a led to a degree of specialisation in their use.
-
-The *PIP* data has typically been the favoured choice for investigating the whole distribution of welfare across households, particularly at a global scale [@Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @milanovic_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk].
-This includes the World Bank's own research and monitoring activties concerning global poverty, inequality and growth across the distribution [see for example @world_bank_poverty_2016; @world_bank_poverty_2022].
-The dataset is also relied on by the United Nations to set and monitor many of the targets in this area that form part of its *Sustainable Development Goals*.
-
-The *WID* data has more commonly been used to understand top dynamics before taxes.
-Notwithstanding substantial efforts to global, high income countries due to the data required.
-in which today's rich countries are overrepresented .
-In policy, advocacy and public debate, the data used to highlight the very large and, in many cases, growing advantage enjoyed by the richest individuals in these societies.
-In academia, top incomes data long-run processes fundamental – capital accumulation.[@atkinson_top_2007; @piketty_capital_2014]  On political economy channels, social conflict between capital and labour, wage bargaining norms and the interaction with marginal tax rates.
-More generally the data have led to a fundamental revaluation of the causes of inequality that move beyond orthodox accounts given purely in terms of marginal productivities of workers with heterogeneous skills or education [@alvaredo_top_2013; @piketty_capital_2014-1].
-
-Thus whilst both the *PIP* and *WID* datasets have played substantial agenda-setting roles in academic research, policy and public debate, this has tended to occur in largely separate arenas.
-Moreoever, no way to join up the insights they offer.
-Very confusing terrain for people to understand.
-We are left with two distinct and, to a large extent, incommensurable 'worlds' of inequality measurement.
-
-Recent extensions have brought down some of these barriers. PIP now includes high-income countries. WID now global coverage, and post-tax income concepts. In
-
-At the same time, these developments in the two datasets have made this disconnect between them all the more apparent. Fundamental differences on the basics. Two in particular.
-
-1
-2
-
-Whilst for individual countries, the connections may be well understood, globally the value of the different assumptions is not really known. We have to pick blindly.
-
-### Connecting up the worlds
-
-This is the first of two papers that seeks to join up.
-
-The dataset is used for the Bank's own research and monitoring activities which, whilst traditionally focussing on absolute poverty, have more recently encompassed trends in inequality, relative poverty and growth across the distribution.
-Notably, the data is alo used to set and monitor many of the targets, on both poverty and inequality, that form part of the UN's Sustainable Development Goals.
-More generally, the source has been indispensible for research taking a global view on inequality [@Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @milanovic_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk].
-
-The *WID* database emerged out of the burgeoning literature on top incomes [@atkinson_top_2014] and has come to play a central and, to some extent, coordinating role for this research agenda.
-The key innovation of the 'top incomes' approach has been the incorporation of tax data, as either an alternative or complement to the more orthodox reliance on household survey data.
-The primary advantage is that tax data can provide a better view of the top of the income distribution, which is often poorly captured by survey data [@lustig_missing_2019].
-However, the choice of data source has typically brought with it a number of other differences that mark out the 'top incomes' approach.
-This includes a particular focus on the top of the distribution (for which tax data is more universally available), and on income measured before tax (in terms of which the underlying data is gathered).
-This contrasts with more orthodox approaches to inequality measurement in which disposable income has tended to be seen as the more relevant measure of welfare [@canberra_group_canberra_2011], and greater attention is given to measures of overall inequality, such as the Gini coefficient, or to relative poverty.
-
-Another benefit of tax data is the much longer timeframes it is sometimes available for.
-The estimates gathered in *WID* provide a long-run perspective on the evolution of inequality over the twentieth century, albeit for a more limited geography in which today's rich countries are somewhat overrepresented [@atkinson_top_2007; @piketty_capital_2014].
-
-Moreover, the top incomes methodology carried forward by *WID* developed with a distinctive set of questions concerning the *causes* of inequality in mind.
-Particular attention is given to the distribution of wealth and capital incomes as drivers of personal income inequality, in contrast to orthodox accounts focussed purely on the dispersion of wages [@alvaredo_top_2013; @piketty_capital_2014-1].
-This ...
-hidden from view given the concentration of capital incomes at the top, captured poorly by survey data.
-
-### Two 'worlds' of inequality measurement
-Tw worlds – methods, conceptions of inequality, time periods and geographic coverage and research questions come bundled.
-
-Always unsatisfying. Very difficult to unpick. Peple do it for individual countries..
-
-A wholeslae comparison was generally avoided due to the very different coverage. Used in parallel.
-
-As the databases have developed and expanded, however, this division of labour between the two datasets has become increasingly blurred.
-PIP includes high-income countries.
-WID coverage is global.
-DINA focus on comparability.
-Increasingly, the datasets are used to answer the same range of questions, speak to the same world.
-
-Where they now meet, the result is very confusing, one that, despite the great strides forward, leaves uncertainty around the most basic questions.
-
-This is most visible in two fundamental disagreements between the two 'worlds'.
-
-Firstly, they give very different views of how ubiquitous a phenomenon rising inequality is within countries around the world.
-The view from *World Inequality Database* and its associated global report is that inequality has risen 'nearly everywhere' in recent decades [@alvaredo_world_2018; @chancel_world_2021].
-In contrast, the World Bank data and its associated biannual report offers a much more mixed picture – in which inequality is rising and falling in roughly equal numbers of countries.
-
-Secondly, the datasets paint as to how such inequalities *within* countries compare to the differences in average incomes seen *between* richer and poorer countries.
-
-Most analyses of global income inequality to date have relied heavily on the *PIP* data [@Kanbur2022-lk; @lakner_global_2016; @anand_global_2015; @milanovic_global_2015; @Nino-Zarazua2017-tv; @Ravallion2014-mk].^[TODO: Add footnote explaining transition from Povcal] *[TODO: Check and more refs.]*.
-A consistent result from this literature is that inequalities within countries remain small in comparison to the much larger differences seen between countries.
-This is despite the convergence seen in recent decades between countries, given higher economic growth rates across low- and middle-income countries.
-According to such research, a person's income remains largely determined by the country in which they happen to have been born [@milanovic_global_2015].
-
-More recent analysis of the WID data flips this established result on its head. Chancel... estimate xx.
-
-The datasets are enormous efforts, involving the work of hundreds of researchers.
-They are hugely valuable offering an unprecidented perspective on global inequality and how it is evolving.
-And yet the disconnect between the two devalues them both.
-Amidst the wealth of data, basic facts are not known.
-And the terrain is incredibly confusing to navigate for policimakers and the concerned public.
-
-### Connecting up the worlds
-
-This paper, along with a companion paper, seeks to join up these two worlds. This first paper picks up the first question – within country. A companion paper considers global inequality.
-
-Section 2 of this paper introduces the two worlds.
-Section 3 compares within country trends between two periods.
-It begins with the inequlaity metrics for which these datasets are most associated, and in which the worlds fall furthest apart: Gini, Top 1% share.
-Already
-Section 4 aligns coverage. It is not neutral. Already this goes some way to explaining the gap between the two datasets.
-Section 5 aligns inequality indices. [Summary.]
-Section 5 considers the role of welfare metrics and specific data quality issues for a subset of countries where there is most disagreement.
-
-Summary of main conclusion!?
-
-# 2. Two worlds of inequality measurement
-
-As well as being the the two  not just two different databases. They in some sense represent *two different worlds* of inequality measurement.
-
-Different methods. These yield different pictures of the extent of inequality and how it has changed over time. Motivated by different ideas about the salient factors shaping inequality.
-
-### PIP
-
-### WID
-
-They differ as to how ubiquitous a phenomenon rising inequality is around the world.
-The view from *World Inequality Database* and its associated global report is that inequality is rising in 'nearly everywhere' [@alvaredo_world_2018; @chancel_world_2021].[^WID_quotes] In contrast, the World Bank data and its biannual report offers a much more mixed picture – in which inequality is rising and falling in roughly equal numbers of countries.
-
-[^WID_quotes]:
-    > Income and wealth inequalities have been on the rise nearly everywhere since the 1980s, following a series of deregulation and liberalization programs which took different forms in different countries.
-[@chancel_world_2021, pp.
-xx]
-
-    > In recent decades, income inequality has increased in nearly all countries, but at different speeds... [@alvaredo_world_2018, pp. xx]
-
-Decompositions of global inequality into between- and within-country components.
-
-Capital vs labour share. .– including fundamental laws of capitalism with regard to capital accumulation (Piketty) –
-
-### The disconnect between these datasets devaules them both
-
-These datasets are incredibly valuable – playing to motivate action.
-The example in the opening sentence above illustrates this well.
-It is only thanks to the WID data that we can say X.
-It is only thanks to the PIP data that we can say that three billion cannot afford a healthy diet.
-
-Often it is presented that there is a division of labour between these dataset – WID is for inequality, especially at the top.PIP is for poverty.
-
-But the point is that it is only one world that we live in. The disconnect is not something we can live with, but rather presents a challenge to both datasets.
-
-There remains fundamental ambiguity surrounding which of two possible worlds we live in.
-
-The '*WID* world' rising inequality within countries has been an almost universal phenomenon.
-suggests the winners from globalization and economic growth overwhelmingly have been a global elite.
-It sees the bulk of global inequalities as lying within nations, but emphasises global factors in terms of causes and hence the need for international coordination to combat inequality.
-
-The '*PIP* world' – within-country inequality trends in recent decades have been mixed, and the differences in average incomes between countries remain the lion's share of the globally.As such the importance of national policy (Ravallion, 2018; Corlett, 2016).
-access to both produced and human capital.
-
-Knowing which of these worlds we live in matters a great deal – for our understanding of inequality today, how it has changed, and also where efforts to combat inequality in the future are likely to be most successful.
-
-Trouble is that these are all bundled up. You must chose. On what basis can you make a choice?
-
-Known problems on both sides.
-
-For PIP – top incomes, and lack of comparability of the survey data – notably mix of incme and consumption data.
-
-For WID – tha many assumptions.
-An emphasis placed on the incorporation of tax data, but when taking a global view confonte by the fact that such tax data is often nonexistent.
-Top income corrections based on strong assumptions.
-
-Given the presentation, the extent to which different assumptions, indices, methodological choices affects the data and – is not something that can be read off' the data.
-
-In using the datasets a global view on inequality, requires choosing wholesale between methodological 'bundles' whose relative strengths and weaknesses, applied globally, are not well understood.
-
-Also makes for a very confusing terrain for policymakers, media and the public.
-
-The two datasets are incredible efforts, involving the work of hundreds of researchers.
-Both datasets have continued to improve adn expand.
-In terms of data availability we are thus in a better position than ever to take a global viw on inequlity.
-And yet, amidst this wealth of data, limited understanding of very basics concerning inequality around the world.
-
-### Outline: Connecting the two worlds
-*(In this section I will explain the plan for the paper)*
-
-In this paper I aim to bring these two worlds of inequality measurement closer. To Triangulate what they do agree on and identify where they disagree.
-
-The first, most obvious, step is simply to compare the in a systemtatic way.
-This very basic step is curiously lacking in the literature.
-We begin by comparing trends in each world based on the inequality indices most typical in each: the Gini coefficient for the PIP data, and the top 1% share of pre-tax income for WID.
-
-Given incomplete coverage in both datasets, we 'line up' estimates to reference years to see change over time.
-We use two periods, presenting two different tradeoffs with respect to data coverage.
-The first is 1993-2018??
-This maximises global coverage – 100 countries covering roughly 90% of the world poplation (Figure XX).
-The downside is that this period misses out – transition, liberalisation and structural adjustment.
-Question begging, by avoiding specifically the period where inequality thought to rise.
-
-The second is 1980-2018. Here we lose coverage, especially in the PIP data (Figure XX).
-
-The two measures selected – Gini coefficient of disposable income/consumption, and the top 1% share of pre-tax income – are where the PIP and WID worlds respectively are at their furthest apart.
-Yet in terms of broad global trends, there is much that even these very different measures can agree on:
-- x
-- y
-- z.
-
-There are big differences though:
-- Somewhat more pronounced rise among advanced economies.
-- Big divergence in trend some of the most populous countries
-
-Subsequent sections attempt to bridge the gaps in order to learn where those differences emerge.
-
-- Section 2: Aligning coverage. This reduces the gap already. Coverage of countries is not... with respect to inequality trends.
-
-- Inequality indices chosen. We align. Compare Top 1%. And Gini.
-    - Explain results.
-
-- Welfare concept.
-We look at different income concepts for the small number of coutnries for which such data exists.
-To this end we also include data from the Luxembourg Income Study.
-Similar to PIP in that it relies on survey data.
-But provides access to microdata, allowing for the construction of different welfare aggregates.
-We focus on the countries with the biggest difference in trend and biggest population.
-I.e.
-those that contribute most to the difference in the weighted an unweigted global averages.
-
-What I won't do is a detailed lining up.
-Although both institutions continue to take many steps to provide transparency and reproducibility to their estimates, they remain a black box. Makes it impossible to do this lining up.
-
-In many particular contexts – where detailed lining up is possible – there relative strengths and weaknesses may be clear.
-(This is what Morllei does.
-And Nolan).
-But at the global level the makes this impractical.
-
-    - Summarise: really depends on the country.
-
-----
-
-# 3. Comparing the two worlds
-
-We begin by comparing the metrics where the worlds are furthest apart. PIP Gini. Wid Top 1% share.
+[TODO: More details on PIP and WID and how they differ in terms of concepts and methods.
+
+- Include discussion of inter/extrapolation in WID data.
+E.g.
+Rescaling factors from tax data for one year is often applied to many other years.
+In the case of Africa, the rescalling factors seen in the few countries with tax data (for at least some years) are applied to all other countries.]
+
+- Include discussion of comparability and quality of survey data in PIP.
+
+## Coverage
+
+# 3. An out-of-the-box comparison of the *PIP* and *WID* datasets
+
+[TODO: Describe the basic method used to compare the data:
+
+- I take pairs of estimates in or close to reference years
+    - Footnote: For PIP data, which is a mix of income and consumption points, I only take a pair of observations where the income concept does not match where matchin observations are not available.
+This is only the case for 3 out of 90 countries in one specification.
+    - I use two periods, chosen to hit two different points in the trade-off maximising coverage and the length/relevance of the period studied: 1993-2015 (+/- 3 years) and 1980-2018 (+/- 5 years)
+
+- I consider unweighted and population-weighted average trends, and also the number of countries seeing substantive rises or falls.
+
+]
+
+We begin by comparing trends in the inequality metrics most characteristic of the two 'worlds' of measurement and where the datasets fall the furthest apart conceptually: for *PIP* the Gini of disposabale income (or consumption); for *WID* the top 1% share of pretax national income.
 
@@
 
-@fig-pip-wid-1993-2015-all-obs comoares the level of inequality, as measured by the two metrics, in or around 1993 and 2015.
-A band is shown along the 45-degree line.
-Countries falling within this band saw little change in inequality.
-For countries above the band, inequality in 2015 was substantively higher than in 1993.
-Countries falling below the band instead saw a substantive fall.
-
-@tbl-pip-wid-1993-2015-all-obs-absolute-count provides a summary of the .
-A 'substantive' change here is defined as a movement of more than 2 percentage points in the case of the PIP Gini coefficient, and more than 1 percentage point in the case of the WID top 1% share.
-[TODO: add table in appendix showing that using relative thresholds (+/-5% across both metrics) makes little difference.]
-
-According to PIP estimates for the Gini coefficient, inequality was rising and falling in roughly an equal number of countries over the period (broadly replicating the analysis presented in @world_bank_poverty_2016).
-
-@tbl-pip-wid-1993-2015-all-obs-averages presents averages for the two reference years.
-The unweighted average Gini coefficient within the PIP data fell somewhat over the period, from 39.8 to 30.0 points.
-As we see from @fig-pip-wid-1993-2015-all-obs (top panel), contributing to this average decline are a number of countries showing very substantial falls – in some cases declines of 10 points or more.
-These are seen mostly in Sub-Saharan Africa, in addition to a number of Latin American and Eastern European and Central Asian countries.
-[TODO: Mention particular questions around survey comparability in SSA – this is a subject that will be taken up in section X.]
-
-WID top 1% share estimates show decidedly more cases of rising inequality – in roughly two-thirds of countries.The average, again, is pulled down by some particularly substantial falls, which in the WID data are more exclusively concentrated in Sub-Saharan Africa.
-This leaves the average top 1% share remarkably stable over the period, showing a modest rise from 14.5% to 14.8%.
-
-In both datasets, we see that inequality trends were less benign in particularly populous countries.
-This is apparent from a comparison of the unweighted and population-weighted averages (@tbl-pip-wid-1993-2015-all-obs-averages).
-The moderate fall seen in the unweigted average PIP Gini coefficient is no longer visible in the weighted average.
-In the WID data, in contrast to the unweighted average, the weighted average rose by more than a third, from 13% to 17%.
-
-As @fig-pip-vs-wid-1993-2015 illustrates, it is also in particularly populous countries where we see some of the starkest differences between the two datasets.
-The figure plots the change in the two metrics, along with the line of best fit.
-[TODO: 
-- Plot also for relative change (%, not %-pt).
-- But what should we expect? Plot the relationship between Gini change and top 1% share change assuming lognormality. Possibly this makes more sense as part of section 3 on inequality indices.]
-
-This includes India, United States, for which far more substantial rises are seen in the WID top 1% share.
-In Brazil and Nigeria, substantial falls present in the PIP Gini are not reflected in the WID data.
-Mexico and Russia stand out in particular in seeing large movements in opposite directions within the two datasets.
-
-Regionally [TODO: summarise regional trends].
-
-```{python}
-# Set-up to compare WID and PIP top 10% share, all observations
+```{python}
+# Set-up to compare PIP Gini and WID top 1% share, all observations, for shorter time period
 reference_vals = [1993, 2015]
@@
 
-output_dict = prep_data_and_change_summary(
+output_dict_1993 = prep_data_and_change_summary(
       source_dfs = source_dfs,
@@
 # Store data for OWID upload – prepared further in appendix below
-df_owid_1993_2015 = output_dict['owid_format']
+df_owid_1993_2015 = output_dict_1993['owid_format']
 
@@
 
-#### Figures 1-2
-
-::: {.panel-tabset}
-## All available observations
+@fig-pip-wid-1993-2015-all-obs compares the level of inequality, as measured by the two metrics, in or around 1993 and 2015.
+A band is shown along the 45-degree line.
+Countries falling within this band saw little change in inequality.
+For countries above the band, inequality in 2015 was substantively higher than in 1993.
+Countries falling below the band instead saw a substantive fall.
+The data is summarised in @tbl-pip-wid-1993-2015-all-obs-absolute-count, which provides a counts of the number of countries with rising, stable or falling inequality, and @tbl-pip-wid-1993-2015-all-obs-averages which presents unweighted and population-weighted averages for the two reference years.
+In both @fig-pip-wid-1993-2015-all-obs and @tbl-pip-wid-1993-2015-all-obs-absolute-count, a 'substantive change over the period is defined as a movement of more than 2 percentage points in the case of the PIP Gini coefficient, and more than 1 percentage point in the case of the WID top 1% share.
+[TODO: add table in appendix showing that using *relative* thresholds – i.e.
++/-5% (and not %-point) across both metrics – makes little difference.]
+
+#### Figure 1
 
 ![PIP Gini and WID Top 1% shahre in 1993 vs 21015 – All available observations](images/figures/PIPGini-vs-WIDTop1-1993-vs-2015-01.png){#fig-pip-wid-1993-2015-all-obs}
-
-## Only observations available in both datasets
-![PIP Gini and WID Top 1% shahre in 1993 vs 21015 – Matching observations only](images/figures/PIPGini-vs-WIDTop1-1993-vs-2015-02.png){#fig-pip-wid-1993-2015-matching-obs}
-
-:::
-
-#### Tables 1-2
-::: {.panel-tabset}
-## All available observations
 
@@
 #| tbl-cap: Change in inequality 1993 vs 2015 - All availalbe observations
-output_dict['all_sources_False']['absolute_counts']
-
-```
-
-## Only observations available in both datasets
-
-```{python}
-#| label: tbl-pip-wid-1993-2015-matching-obs-absolute-count
-#| tbl-cap: Change in inequality 1993 vs 2015 - Matching observations only
-output_dict['all_sources_True']['absolute_counts']
-
-```
-
-:::
-
-#### Tables 3-4
-
-::: {.panel-tabset}
-## All available observations
+output_dict_1993['all_sources_False']['absolute_counts']
+
+```
 
@@
 #| tbl-cap: Unweighted and weighted average in 1993 vs 2015 - All availalbe observations
-output_dict['all_sources_False']['averages']
-
-```
-
-## Only observations available in both datasets
-
-```{python}
-#| label: tbl-pip-wid-1993-2015-matching-obs-averages
-#| tbl-cap: Unweighted and weighted average in 1993 vs 2015 – Matching observations only
-output_dict['all_sources_True']['averages']
-
-```
-
-:::
-
-```{python}
-import statsmodels.formula.api as smf
-from matplotlib import pyplot as plt
-
-change_compare = output_dict['all_sources_True']['prepped_data'][['source_var', 'Entity', 'change', 'region']]
-
-change_compare = change_compare.pivot_table(index=['Entity', 'region'], columns='source_var', values='change').reset_index()
-
-change_compare = change_compare.rename(columns={"WID: Top 1pc share – pretax": 'change_WID', 'PIP: Gini': 'change_PIP'})
-
-# Save the data to upload to WID for plot
-change_compare_owid = change_compare
-change_compare_owid['year'] = 2000
-change_compare_owid = change_compare_owid.rename(columns={'Entity': 'country'})
-
-fp = 'data_appendix/figures_and_tables/Compare WID and PIP CHANGE in inequality 1993 vs 2015 (Joe temp).csv'
-
-change_compare_owid.to_csv(fp, index=False)
-
-# Initialise and fit linear regression model using `statsmodels`
-model = smf.ols('change_WID ~ change_PIP', data=change_compare)
-model = model.fit()
-# model.params
-
-```
-
-![Change in PIP Gini vs WID Top 1% share in 1993 vs 2015](images/figures/Change-1993-2015.svg){#fig-pip-vs-wid-1993-2015}
+output_dict_1993['all_sources_False']['averages']
+
+```
+
+According to *PIP* estimates for the Gini coefficient, inequality was rising and falling in roughly an equal number of countries over the period.
+This broadly replicates the analysis presented in @world_bank_poverty_2016.
+The unweighted average fell somewhat over the period, from 39.8 to 30.0 points.
+As we see from @fig-pip-wid-1993-2015-all-obs (top panel), contributing to this average decline are a number of countries showing very substantial falls – in some cases declines of 10 points or more.
+These are seen mostly in Sub-Saharan Africa, in addition to a number of Latin American and Eastern European and Central Asian countries.
+[TODO: Mention particular questions around survey comparability in SSA.]
+
+*WID* top 1% share estimates show decidedly more cases of rising inequality – in roughly two-thirds of countries.
+The average, again, is pulled down by some particularly substantial falls, which in the *WID* data are more exclusively concentrated in Sub-Saharan Africa.
+This leaves the average top 1% share remarkably stable over the period, showing a modest rise from 14.5% to 14.8%.
+
+In both datasets, we see that inequality trends were less benign in particularly populous countries.
+This is apparent from a comparison of the unweighted and population-weighted averages (@tbl-pip-wid-1993-2015-all-obs-averages).
+The moderate fall seen in the unweigted average *PIP* Gini coefficient is no longer visible in the weighted average.
+In the *WID* data, in contrast to the unweighted average, the weighted average rose by more than a third, from 13% to 17%.
+
+As @fig-pip-vs-wid-1993-2015 (section 4) illustrates, it is also in particularly populous countries where we see some of the starkest differences between the two datasets.
+The figure plots the change in the two metrics, along with the line of best fit.
+[TODO: 
+- Plot also for relative change (%, not %-pt).
+- But what should we expect? Plot the relationship between Gini change and top 1% share change assuming lognormality. Possibly this makes more sense as part of section 3 on inequality indices.]
+
+This includes India, United States, for which far more substantial rises are seen in the WID top 1% share.
+In Brazil and Nigeria, substantial falls present in the PIP Gini are not reflected in the WID data.
+Mexico and Russia stand out in particular in seeing large movements in opposite directions within the two datasets.
+
+Regionally [TODO: summarise regional trends].
 
@@
 ```{python}
-# Set-up to compare WID and PIP top 10% share, all observations
+# Set-up to compare PIP Gini and WID Top 1% share, all observations, for longer period
 reference_vals = [1980, 2018]
@@
 
-output_dict = prep_data_and_change_summary(
+output_dict_1980 = prep_data_and_change_summary(
       source_dfs = source_dfs,
@@
 # Store data for OWID upload – prepared further in appendix below
-df_owid_1980_2018 = output_dict['owid_format']
+df_owid_1980_2018 = output_dict_1980['owid_format']
 
@@
 
-#### Figures X-Y
-
-#### Tables 3-4
-
-::: {.panel-tabset}
-## All available observations
+[TODO: Make similar figure as figure 1]
+
+```{python}
+#| label: tbl-pip-wid-1980-2018-all-obs-absolute-count
+#| tbl-cap: Change in inequality 1980 vs 2018 - All availalbe observations
+output_dict_1980['all_sources_False']['absolute_counts']
+
+```
 
@@
 #| tbl-cap: Unweighted and weighted average in 1980 vs 2018 - All availalbe observations
-output_dict['all_sources_False']['averages']
-
-```
-
-## Only observations available in both datasets
-
-```{python}
-#| label: tbl-pip-wid-1980-2018-matching-obs-averages
-#| tbl-cap: Unweighted and weighted average in 1980 vs 2018– Matching observations only
-output_dict['all_sources_True']['averages']
-
-```
-
-:::
+output_dict_1980['all_sources_False']['averages']
+
+```
 
@@
 
+[TODO: Plot in appendix – Just like figure 1, but show only the data points missing from WID in the PIP pannel. And visa versa.]
+
+```{python}
+import statsmodels.formula.api as smf
+from matplotlib import pyplot as plt
+
+change_compare = output_dict_1993['all_sources_True']['prepped_data'][['source_var', 'Entity', 'change', 'region']]
+
+change_compare = change_compare.pivot_table(index=['Entity', 'region'], columns='source_var', values='change').reset_index()
+
+change_compare = change_compare.rename(columns={"WID: Top 1pc share – pretax": 'change_WID', 'PIP: Gini': 'change_PIP'})
+
+# Save the data to upload to WID for plot
+change_compare_owid = change_compare
+change_compare_owid['year'] = 2000
+change_compare_owid = change_compare_owid.rename(columns={'Entity': 'country'})
+
+fp = 'data_appendix/figures_and_tables/Compare WID and PIP CHANGE in inequality 1993 vs 2015 (Joe temp).csv'
+
+change_compare_owid.to_csv(fp, index=False)
+
+# Initialise and fit linear regression model using `statsmodels`
+model = smf.ols('change_WID ~ change_PIP', data=change_compare)
+model = model.fit()
+# model.params
+
+```
+
+![Change in PIP Gini vs WID Top 1% share in 1993 vs 2015](images/figures/Change-1993-2015.svg){#fig-pip-vs-wid-1993-2015}
+
+[TODO: same plot but for 1980 period]
+
 As such, differences in coverage already go some way to explaining the differences between the two datasets. 
@@
 
-Plot – Just like figure 1, but show only the data points missing from WID in the PIP pannel. And visa versa.
-
-# 3. Aligning indices
+# 4. Aligning indices
 
@@
 # Appendix
+
+## Comparison of the WID and PIP datasets after aligning coverage
+
+### Inequality since 1993
+
+#### Figure X
+![PIP Gini and WID Top 1% share in 1993 vs 2015 – Matching observations only](images/figures/PIPGini-vs-WIDTop1-1993-vs-2015-02.png){#fig-pip-wid-1993-2015-matching-obs}
+
+```{python}
+#| label: tbl-pip-wid-1993-2015-matching-obs-absolute-count
+#| tbl-cap: Change in inequality 1993 vs 2015 - Matching observations only
+output_dict_1993['all_sources_True']['absolute_counts']
+
+```
+
+```{python}
+#| label: tbl-pip-wid-1993-2015-matching-obs-averages
+#| tbl-cap: Unweighted and weighted average in 1993 vs 2015 – Matching observations only
+output_dict_1993['all_sources_True']['averages']
+
+```
+
+### Inequality since 1980
+
+[TODO: same as above for post-1993 period.]
+
+```{python}
+#| label: tbl-pip-wid-1980-2018-matching-obs-averages
+#| tbl-cap: Unweighted and weighted average in 1980 vs 2018– Matching observations only
+output_dict_1980['all_sources_True']['averages']
+
+```
+
 ```{python}

### 2023-05-03 PhD_pages/Paper_1/data_appendix/1_A_prepare_PIP.qmd
Paper 1: restructuring to prep data separate to paper

@@
 import pandas as pd
+import json
+import requests
 
-pd.set_option('display.max_rows', None)
 ```
@@
 ::: {.callout-note collapse="true" appearance="minimal"}
-### Load PIP data
+### Load PIP percentile data
+```{python}
+fp = "data/original/pip_percentiles.csv"
+
+df_pip_percentiles = pd.read_csv(fp)
+
+```
+
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Keep top 1% shares
+```{python}
+
+drop_cols = ['percentile', 'avg_welfare', 'pop_share', 'quantile']
+
+df_pip_percentiles = df_pip_percentiles[df_pip_percentiles['percentile'] == 100]\
+  .drop(drop_cols, axis=1)\
+  .rename(columns={
+    'welfare_share': 'Top 1pc share'
+  })
+
+# Convert shares to percent
+df_pip_percentiles['Top 1pc share'] = df_pip_percentiles['Top 1pc share'] * 100
+```
+
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Standardize country names in percentile data
+
+```{python}
+# # Initially I save a csv of country-codes included in the percentil data, to pass through the OWID country name standardization tool
+# unique_countries = df_pip_percentiles['country_code'].unique()
+# df_pip_percentiles_countries = pd.DataFrame({'country': unique_countries})
+
+# df_pip_percentiles_countries.to_csv('data/original/pip_percentiles_countries.csv', index=False)
+
+# Read in name mapping file
+fp = 'data/original/pip_percentiles_countries_country_standardized.csv'
+name_mapping = pd.read_csv(fp)
+
+# Merge into
+
+# Merge mapping into data
+
+df_pip_percentiles = pd.merge(df_pip_percentiles, name_mapping, left_on='country_code', right_on='country', how = 'left')
+
+# Drop columns used in the construction of the standardized country column
+df_pip_percentiles = df_pip_percentiles.drop(columns=['country','country_code'])
+
+df_pip_percentiles = df_pip_percentiles.rename(columns = {'Our World In Data Name': 'country'})
+```
+
+```{python}
+df_pip_percentiles.head()
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Load PIP main data (already prepared by OWID)
 We load the data from a file prepared from the PIP API by Our World in Data.
@@
 ```{python}
-# df_pip<- read.csv("https://raw.githubusercontent.com/owid/poverty-data/main/datasets/pip_dataset.csv")
+# df_pip = pd.read_csv("https://raw.githubusercontent.com/owid/poverty-data/main/datasets/pip_dataset.csv")
 
@@
 df_pip = df_pip.loc[df_pip["ppp_version"] == 2017]
+```
+:::
+
+::: {.callout-note collapse="true" appearance="minimal"}
+### Merge top 1% shares from percentile data into main PIP data
+
+```{python}
+df_pip = pd.merge(df_pip, df_pip_percentiles, on=['country', 'year', 'reporting_level', 'welfare_type'], how='left')
+
+df_pip.head()
 ```
@@
 
+# Top 10 / bottom 50% share
+df_pip["Ratio Top 10 Bottom 50 share"] = df_pip["decile10_share"]/df_pip["Bottom 50pc share"] * 100
+
 ```
@@
   "Gini",
+  "Top 1pc share",
   "Top 10pc share",
   "Bottom 50pc share",
+  "Ratio Top 10 Bottom 50 share",
   "MLD",

### 2023-05-03 PhD_pages/Paper_1/data_appendix/prepare_data_for_figures_and_tables.qmd
Paper 1: restructuring to prep data separate to paper

---
title: "Appendix 1.H: Preparing figures and tables used in the paper"
format: html
execute:
  echo: false
  warning: false
---

```{python}

from plotnine import *
import plotly.express as px
import pandas as pd
import numpy as np
# from IPython.display import display, HTML, IFrame, Markdown
# import dataframe_image as dfi
# import imgkit

from G_functions_for_figures_and_tables import *

# pd.set_option('display.max_rows', None)

```

```{python}

fp = 'data/clean/pip.csv'

df_pip = pd.read_csv(fp)

fp = 'data/clean/wid.csv'

df_wid = pd.read_csv(fp)

fp = 'data/clean/lis.csv'

df_lis = pd.read_csv(fp)

fp = 'data/clean/region_mapping.csv'

df_regions = pd.read_csv(fp)

fp = 'data/clean/population.csv'

df_pop = pd.read_csv(fp)

fp = 'data/clean/region_population.csv'

df_pop_regions = pd.read_csv(fp)

# Gather data into dictionary
source_dfs = {
    "PIP": df_pip,
    "WID": df_wid,
    "LIS": df_lis
    }

# This specifies the +/- for the 'stable' category in the tables below
tolerance_lookup =  {
                'var_search_term': [
                    'Gini',
                    'Top 1pc share', 
                    'Top 10pc share',
                    'Bottom 50pc share',
                    "Ratio Top 10 Bottom 50 share"],
                'var_tolerance': [
                    1,
                    1,
                    1,
                    1,
                    1
                ], 
                'relative_tolerance': [
                    0.05
                ]
            }

```

```{python}

selected_vars = [
        'PIP: Gini',
        "PIP: Top 10pc share",
        'PIP: Bottom 50pc share',
        'PIP: Ratio Top 10 Bottom 50 share',
        'WID: Gini – pretax',
        "WID: Top 1pc share – pretax",
        "WID: Top 10pc share – pretax",
        'WID: Bottom 50pc share – pretax',
        'WID: Ratio Top 10 Bottom 50 share – pretax'
        ]

```

Prep data for 1993-2015 period

```{python}
# Set-up to compare PIP Gini and WID top 1% share, all observations, for shorter time period
reference_vals = [1993, 2015]
max_dist_from_refs = 3

prepped_data_dict = prep_data(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_regions = df_regions,
      region_col = 'region',
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      reference_col = "Year",
      group_by_col = "Entity",
      tolerance_lookup =  tolerance_lookup,
      outlier_cut_off_upper = None
      )

owid_format = prepped_data_to_owid_format(
      prepped_data_all_obs = prepped_data_dict['all_obs'],
      prepped_data_matching_obs = prepped_data_dict['matching_obs'],
      reference_vals = reference_vals
      )
owid_format

# # 
# Save data to be uploaded to OWID to make charts
fp = 'figures_and_tables/Compare WID and PIP inequality data 1993 vs 2015 (Joe temp).csv'

owid_format.to_csv(fp, index=False)

```

```{python}
summary_tables_dict = prep_summary_tables(
    prepped_data_dict = prepped_data_dict,
    df_pop_regions = df_pop_regions,
    selected_vars = [ 'PIP: Gini',
                    'WID: Top 10pc share – pretax'],
    reference_vals = reference_vals)

summary_tables_dict['all_obs']['absolute_counts']

```

Prep data for 1980-2018 period

```{python}
# Set-up to compare PIP Gini and WID top 1% share, all observations, for shorter time period
reference_vals = [1980, 2018]
max_dist_from_refs = 5

prepped_data_dict = prep_data(
      source_dfs = source_dfs,
      df_pop = df_pop,
      df_regions = df_regions,
      region_col = 'region',
      selected_vars = selected_vars,
      reference_vals = reference_vals,
      max_dist_from_refs = max_dist_from_refs,
      min_dist_between = 1,
      reference_col = "Year",
      group_by_col = "Entity",
      tolerance_lookup =  tolerance_lookup,
      outlier_cut_off_upper = None
      )

owid_format = prepped_data_to_owid_format(
      prepped_data_all_obs = prepped_data_dict['all_obs'],
      prepped_data_matching_obs = prepped_data_dict['matching_obs'],
      reference_vals = reference_vals
      )

# # 
# Save data to be uploaded to OWID to make charts
fp = 'figures_and_tables/Compare WID and PIP inequality data 1980 vs 2018 (Joe temp).csv'

owid_format.to_csv(fp, index=False)

# prepped_data_dict['all_obs']
```

### 2023-05-03 PhD_pages/Paper_1/paper1.qmd
Paper 1: restructuring to prep data separate to paper

@@
                     'Top 10pc share',
-                    'Bottom 50pc share'],
+                    'Bottom 50pc share',
+                    "Ratio Top 10 Bottom 50 share"],
                 'var_tolerance': [
+                    1,
                     1,
@@
 - Include discussion of comparability and quality of survey data in PIP.
+
+- Plot forthe appendix: make a direct comparison for Africa region with the same inequality index.
+Necessarily tethered to the survey data (unlike say Mexico) The changes are amplified in WID (falls are larger, rises are larger).
 
@@
 
-Summary, across both periods. (Reflect this back in the intro.)
-
-# 3. Aligning coverage
-
-Whilst the number of countries available in the two datasets is very similar, they are not the same countries.
-In particular, LAC is largely absent from WID.
-(Footnote about LAC in WID??
-Are the global results based on linear interpolation?...
-An interpolation that contradicts the survey data?) Also fewer Sub-Saharan African countries.
-In PIP, fewer rich countries in Western Europe and its offshoots.
-
-Importantly, ...
-are not neutral with respect to.The regions where WID is lacking coverage – LAC and SSA – are regions in which inequality was generally falling, according to the PIP Gini coefficient.
-The reverse is also the case: Regions that PIP are lacking – high income countries in W.
-Europe and offshoots – were regions where inequality was generally rising, according to the WID top 1% shares data.
-
-[TODO: Plot in appendix – Just like figure 1, but show only the data points missing from WID in the PIP pannel. And visa versa.]
-
-```{python}
-import statsmodels.formula.api as smf
-from matplotlib import pyplot as plt
-
-change_compare = output_dict_1993['all_sources_True']['prepped_data'][['source_var', 'Entity', 'change', 'region']]
-
-change_compare = change_compare.pivot_table(index=['Entity', 'region'], columns='source_var', values='change').reset_index()
-
-change_compare = change_compare.rename(columns={"WID: Top 1pc share – pretax": 'change_WID', 'PIP: Gini': 'change_PIP'})
-
-# Save the data to upload to WID for plot
-change_compare_owid = change_compare
-change_compare_owid['year'] = 2000
-change_compare_owid = change_compare_owid.rename(columns={'Entity': 'country'})
-
-fp = 'data_appendix/figures_and_tables/Compare WID and PIP CHANGE in inequality 1993 vs 2015 (Joe temp).csv'
-
-change_compare_owid.to_csv(fp, index=False)
-
-# Initialise and fit linear regression model using `statsmodels`
-model = smf.ols('change_WID ~ change_PIP', data=change_compare)
-model = model.fit()
-# model.params
-
-```
-
-![Change in PIP Gini vs WID Top 1% share in 1993 vs 2015](images/figures/Change-1993-2015.svg){#fig-pip-vs-wid-1993-2015}
-
-[TODO: same plot but for 1980 period]
-
-As such, differences in coverage already go some way to explaining the differences between the two datasets. 
-@fig-pip-wid-1993-2015-matching-obs and @tbl-pip-wid-1993-2015-matching-obs-averages show the data, restricting the observations to those available in both datasets.
-This brings the two series closer.
-Explain.
-(Coverage falls)
-
-Summarise remaining differences.
-
-# 4. Aligning indices
-
-- Show that measure doesn't matter for WID until top 1 (check Palma ratio)
-
-  - A table, using relative threshold (+/-5%) Different measures: Gini, Palma, top 10, top 1.
-
-  - Regressions – all pairs of the four measures. See the r2 and note outliers for each.
-
-- Check out how much different metrics matter for PIP.
-
-  - Same table and regressions.
-
-- For each of the 4 measures, do a direct comparison across sources.
-
-  - For both 1993 and 1980: matching observations only and only global results.
-
-(Sub-Saharan Africa: example where the Gini fell in more cases than the top 10%.?)
-
-## Similarities and differences
-
-Summarise where we're at after aligning the inequality metric.
-
-# Bridging the gap for particular countries / welfare concept
-
-"Korinek, Mistiaen, and
-Ravallion (2006) estimate that correcting
-for such selective compliance adds about
-five percentage points to the Gini index for
-the United States."
-
-## Sub-Saharan Africa
-Notes on the WID methodology. Basically no tax data. Based on guesses about how you shift consumption to pre-tax income, and then top income adjustment.
-
-Plot: make a direct comparison for this region with the same index. NEcessarily tethered to the survey data (unlike say Mexico) The changes are amplified in WID (falls are larger, rises are larger).
-
-Also note the quality/comparability of the PIP survey data.
-
------
-
-### Aligning coverage
-
-## Global data on inequality
-
-### Global databases
-
-A global view faces the hard limitation of the underlying data. This has two sides: firstly is there any data? Secondly the comparability of the data, across countries and over time.
-
-Much improved – *harmonized* secondary databases. Three of note.
-
-- PIP
-
-- WID
-
-- LIS
-
-### Two worlds of inequality measurement
-
-#### Connecting up the worlds is valuable
-
-- for Policy
-- for academic research
-
-## Triangulating a global view
-
-### Since 1990
-
-### Since 1980
-
-### Comparing metrics
-
-### Since 1990
-
 Similarities:
@@
 
+# 3. Aligning coverage
+
+Whilst the number of countries available in the two datasets is very similar, they are not the same countries.
+In particular, LAC is largely absent from WID.
+(Footnote about LAC in WID??
+Are the global results based on linear interpolation?...
+An interpolation that contradicts the survey data?) Also fewer Sub-Saharan African countries.
+In PIP, fewer rich countries in Western Europe and its offshoots.
+
+Importantly, ...
+are not neutral with respect to.The regions where WID is lacking coverage – LAC and SSA – are regions in which inequality was generally falling, according to the PIP Gini coefficient.
+The reverse is also the case: Regions that PIP are lacking – high income countries in W.
+Europe and offshoots – were regions where inequality was generally rising, according to the WID top 1% shares data.
+
+[TODO: Plot in appendix – Just like figure 1, but show only the data points missing from WID in the PIP pannel. And visa versa.]
+
+```{python}
+import statsmodels.formula.api as smf
+from matplotlib import pyplot as plt
+
+change_compare = output_dict_1993['all_sources_True']['prepped_data'][['source_var', 'Entity', 'change', 'region']]
+
+change_compare = change_compare.pivot_table(index=['Entity', 'region'], columns='source_var', values='change').reset_index()
+
+change_compare = change_compare.rename(columns={"WID: Top 1pc share – pretax": 'change_WID', 'PIP: Gini': 'change_PIP'})
+
+# Save the data to upload to WID for plot
+change_compare_owid = change_compare
+change_compare_owid['year'] = 2000
+change_compare_owid = change_compare_owid.rename(columns={'Entity': 'country'})
+
+fp = 'data_appendix/figures_and_tables/Compare WID and PIP CHANGE in inequality 1993 vs 2015 (Joe temp).csv'
+
+change_compare_owid.to_csv(fp, index=False)
+
+# Initialise and fit linear regression model using `statsmodels`
+model = smf.ols('change_WID ~ change_PIP', data=change_compare)
+model = model.fit()
+# model.params
+
+```
+
+![Change in PIP Gini vs WID Top 1% share in 1993 vs 2015](images/figures/Change-1993-2015.svg){#fig-pip-vs-wid-1993-2015}
+
+[TODO: same plot but for 1980 period]
+
+As such, differences in coverage already go some way to explaining the differences between the two datasets. 
+@fig-pip-wid-1993-2015-matching-obs and @tbl-pip-wid-1993-2015-matching-obs-averages show the data, restricting the observations to those available in both datasets.
+This brings the two series closer.
+Explain.
+(Coverage falls)
+
+Summarise remaining differences.
+
 Countries where the sources paint a radically different picture that affect the aggregate results:
@@
 - India: Insane rise in WID 34->56
-- China:
 - Russia: Massive drop in PIP. Insane rise in WID.
@@
 
-What explains the different trends in these countries? (We can maybe look in more detail, at different welfare concepts etc.)
-
-It sometimes goes the other way: Indonesia – Rise in PIP, fall in WID.
-
-:::
-
-#### Method 2: Use a regression to obtain the time trend
-
-### Results for a range of inequality metrics
-
-::: {.callout-note collapse="true" appearance="minimal"} 
-## Calculations for comparisons
-
-## Explaining differences between PIP and WID data
-
-### For a subset of countries we can bridge between the PIP and WID data
-
-Besides the underlying data sources, there are many differences between what is being measured in the PIP and WID data - notably the welfare concept and unit of analysis/equivalization.
-
-In order to get a better understanding of where differences between the datasets arise, for a subset of countries we can bridge between the two somewhat. For example:
-
--   WID has post-tax income series for some countries
--   We can use LIS to look at how changing the welfare concept, unit of analysis, equivalisation affects things.
-(For equivalisation, we could also possibly use the newly available PIP harmonized microdata).
-
-### More in-depth discussion of particular outliers
-
--   Russia is a very notable case where the two datasets tell very different stories.
+It sometimes goes the other way: 
+- Indonesia: Rise in PIP, fall in WID.
+- Some SSA countries: Larger falls in WID.
+
+# 4. Aligning indices
+
+- For a range of metrics, check which countries disagree on change category using a relative definitnio (+/-5%). And report global averages.
+
+```{python}
+# Set-up to compare PIP
+reference_vals = [1993, 2015]
+max_dist_from_refs = 3
+
+selected_vars = [
+        'PIP: Gini',
+        'PIP: Ratio Top 10 Bottom 50 share'
+        ]
+
+output_dict_PIP_indices_1993 = prep_data_and_change_summary(
+      source_dfs = source_dfs,
+      df_pop = df_pop,
+      df_pop_regions = df_pop_regions,
+      df_regions = df_regions,
+      region_col = 'region',
+      selected_vars = selected_vars,
+      reference_vals = reference_vals,
+      max_dist_from_refs = max_dist_from_refs,
+      min_dist_between = 1,
+      reference_col = "Year",
+      group_by_col = "Entity",
+      outlier_cut_off_upper = None,
+      tolerance_lookup =  tolerance_lookup
+      )
+
+```
+
+```{python}
+#| label: tbl-pip-indices-counts
+#| tbl-cap: Change in inequality 1980 vs 2018 - All availalbe observations
+output_dict_PIP_indices_1993['all_sources_False']['absolute_counts']
+
+```
+
+```{python}
+#| label: tbl-pip-indices-avgs
+#| tbl-cap: Unweighted and weighted average in 1980 vs 2018 - All availalbe observations
+output_dict_PIP_indices_1993['all_sources_False']['averages']
+
+```
+
+- Show that measure doesn't matter for WID until top 1 (check Palma ratio)
+
+  - A table, using relative threshold (+/-5%) Different measures: Gini, Palma, top 10, top 1.
+
+  - Regressions – all pairs of the four measures. See the r2 and note outliers for each.
+
+- Check out how much different metrics matter for PIP.
+
+  - Same table and regressions.
+
+- For each of the 4 measures, do a direct comparison across sources.
+
+  - For both 1993 and 1980: matching observations only and only global results.
+
+(Sub-Saharan Africa: example where the Gini fell in more cases than the top 10%.?)
+
+## Similarities and differences
+
+Summarise where we're at after aligning the inequality metric.
+
+# 5. Bridging the gap for particular countries / welfare concept
 

### 2023-05-03 PhD_pages/Paper_1/data_appendix/G_functions_for_figures_and_tables.py

@@ -224,7 +224,6 @@ def prep_data(
       reference_vals,
       max_dist_from_refs,
       min_dist_between,
-      all_metrics_requirement,
       reference_col = "Year",
       group_by_col = "Entity",
       tolerance_lookup =  {
@@ -245,234 +244,286 @@ def prep_data(
 
     selected_source_var_dict = prep_selected_vars(selected_vars)
 
+    prepped_data_dict = {}
+
+    for obs_requirement in ['all_obs', 'matching_obs']:
+
+      if obs_requirement == 'all_obs':
+        all_metrics_requirement = False
+      if obs_requirement == 'matching_obs':
+        all_metrics_requirement = True
+
+      matches = []
+
+      for i in range(len(selected_source_var_dict['sources'])):
+          
+          source = selected_source_var_dict["sources"][i]
+          var = selected_source_var_dict["var_names"][i]
+
+          
+
+          df = source_dfs[source]
+
+          if source == "PIP":
+
+              matched = pip_welfare_routine(
+                  df = df,
+                  reference_vals = reference_vals,
+                  max_dist_from_refs = max_dist_from_refs,
+                  min_dist_between = min_dist_between,
+                  reference_col = reference_col,
+                  group_by_col = group_by_col,
+                  value_col = var
+                  )
+
+          else:
+
+              matched = merge_two_ref_matches(
+                  df = df,
+                  reference_vals = reference_vals,
+                  max_dist_from_refs = max_dist_from_refs,
+                  min_dist_between = min_dist_between,
+                  reference_col = reference_col,
+                  group_by_col = group_by_col,
+                  value_col = var
+                  )
+
+          matches.append(matched)
+
+
+      prepped_data = pd.concat(matches, keys=selected_vars)
+
+      # Tidy up indexes
+      prepped_data = prepped_data.reset_index()
+      
+      prepped_data = prepped_data.drop('level_1', axis=1)
+
+      prepped_data = prepped_data\
+          .rename(columns={"level_0": "source_var"})
+
+
+
+
+      # Store the names of the columns to be used onthe X and Y axis
+      x_axis = f'value{reference_vals[0]}'
+      y_axis = f'value{reference_vals[1]}'
+
+
+
+      # Apply outlier cut off, if specified
+      if not pd.isna(outlier_cut_off_upper):
[…]
